### 2015-02-27 manuscript/plos_template.tex
add paper

% Template for PLoS
% Version 2.0 July 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% Be advised that this is merely a template 
% designed to facilitate accurate translation of manuscript content 
% into our production files. 
%
% This template contains extensive comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template 
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication and enters production, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% DO NOT ADD EXTRA PACKAGES TO THIS TEMPLATE unless absolutely necessary.
% Packages included in this template are intentionally
% limited and basic in order to reduce the possibility
% of issues during our production process.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text. For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
%
% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
% For bolding characters in mathmode, please use $\mathbf{}$ 
%
% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
%
% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
%
% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
%
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\usepackage{bm}
\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
% \DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
% \usepackage{setspace} 
% \doublespacing

\usepackage{graphicx}
\usepackage{subfigure}

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}

%% Include all macros below. Please limit the use of macros.

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Extraction of Latent Probabilistic Mutational Signature in Cancer Genomes}
}
% Insert Author names, affiliations and corresponding author email.
\\
Yuichi Shiraishi$^{1,\ast}$, 
Georg Tremmel$^{1}$, 
Satoru Miyano$^{1}$, 
Matthew Stephens$^{2,3}$
\\
\bf{1} Laboratory of DNA Information Analysis, Human Genome Center, Institute of Medical Science, The University of Tokyo, Tokyo, Japan
\\
\bf{2} Dept. of Human Genetics, University of Chicago, Chicago, Illinois, United States of America
\\
\bf{3} Dept. of Statistics, University of Chicago, Chicago, Illinois, United States of America
% \bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: yshira@hgc.jp
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

Thanks to the advances in  recent high throughput sequencing technologies,
a massive amount of somatic mutations from cancer genome sequencing data become available.
Accordingly, it becomes possible to detect characteristic patterns of somatic mutations or ``mutation signatures'' 
at an unprecedented resolution with the expectations of revealing novel causes and mechanisms of tumorigenesis.

Several statistical approaches for extracting characteristic mutation signatures have been proposed.
However, in previous approaches,
since the number of parameters increases exponentially as more contextual factors are taken into account,
we could not treat many contextual factors all together because of instability of estimation results.
Furthermore, interpretation of mutations signatures of huge dimensional vectors is often troublesome.

In this paper, we propose a novel approach based on hierarchical probabilistic modeling.
The proposed approach reduces the number of parameters by the independence assumption on each factor 
so that we can obtain more robust and interpretable estimates.
Using synthetic and real data, we demonstrate that the proposed approach can not only give highly robust estimates,
but also capture novel characteristics such as base frequency at the two base 5' to the mutated sites.
In addition, we clarify the relationships between the proposed approach and the ``mixed-membership models'', 
that have been actively studied in statistical machine learning and statistical genetics community.
Recognizing these relationships will help us to develop a grasp of mutation signature extraction problems,
and will allow us to utilize a lot of techniques accumulated on other fields to further improve the statistical methods.

Finally, we have prepared an R package of the proposed approach (probabilistic mutation signature, pmsignature),
which is available at \url{https://github.com/friend1ws/pmsignature}.

% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}

It is known that the pattern of somatic mutations depends on cancer types and even individuals within the same caner type.
For example, C $>$ A mutations are frequent in lung caner
whereas C $>$ T and CC $>$ TT mutations are frequent in skin cancer,
and their patterns are usually associated to some kinds of carcinogens.
However, since the cost and throughput of sequencing technologies were limited, 
high-resolution analysis of mutation patterns was not possible.

With the coming of high-throughput sequencing technologies, we can now obtain a massive amount of somatic mutation data from cancer genomes,
and there is a great possibility of detecting novel mutation patterns leading to identification of novel carcinogens
and obtaining better classification of cancer genomes based on mutation patterns differences.
On the other hand, there is an increasing need for novel statistical method for mutation patterns analysis 
from vast amount of somatic mutation data.

In this paper, we provide a novel statistical tools for clarifying characteristic mutation patterns
from massive amounts of cancer somatic mutation data, 
which will give us more robust and interpretable mutation patterns compared to previous approaches.
We demonstrate the efficiency of our method using simulated and real data.

\section*{Introduction}

Cancer is a genomic disease. 
As we lead a life, DNAs within the cells throughout our body acquire a number of random somatic mutations
mainly caused by DNA replication errors and exposures to mutagens such as chemical substances, radioactivities and inflammatory reactions.  
Although most mutations are harmless (called ``passenger mutations''), 
a small portions of mutations at some specific sites in cancer genes (called ``driver mutations'') 
confer growth activities to the cells having them over other cells,
allowing such as autonomous proliferation and tissue invasion,
and contribute to oncogenesis \cite{stratton2009cancer}.
The main goal in cancer genome study has been to find the driver mutations to understand the mechanism of cancer development.
On the other hand,  even passenger mutations can also give us an important information, 
because they often show specific mutation patterns, or ``mutation signatures'', which reflect driving forces causing somatic mutations.
Classical analyses of mutation patterns have revealed a number of relationships between mutation patterns and carcinogens.
For example, C $>$ A mutations at CpG sites are abundant in lung cancers with smoking history,
and these are caused by benzo(a)pyrene included in tobacco smoke \cite{pmid12379884}.
Also, C $>$ T and CC $>$ TT  mutations are abundant in ultraviolet-light-associated skin cancers, 
and these are caused by pyrimidine dimers as a result of ultraviolet radiation \cite{pmid15748635}.

There were several problems in these classical studies \cite{pmid12379884, pmid15748635}.
Due to limited sequencing throughput, 
they mostly aggregated mutations collected from each individuals over the same cancer types focusing on a few cancer genes such as TP53
where high mutation frequencies could be expected,
and compared mutation patten profiles among different cancer types.
However, many of those mutations in those cancer genes are considered to be driver mutations adding selective activities to the cells,
which can give biased mutation profiles from those purely reflect driving forces of mutations.
Furthermore, the mutation pattern profile not for each cancer type but for each individual could not be explored in the above approach.

Nowadays, recent advancement in high-throughput sequencing technologies provides vast amounts of passenger mutations
and great chances for not only identification of novel driver mutations but also investigation of sample-by-sample mutation patterns in an unbiased way. 
A study using 21 breast cancer samples identified the association between C $>$ [GT] mutations at TpC sites, 
which is later proved to be caused by APOBEC proteins, and the novel phenomenon called {\it kataegis} \cite{pmid22608084}. 
Moreover, the recent landmark study revealed a landscape of mutation signatures using from 7,034 primary cancer samples of 30 different classes \cite{pmid23945592}. 
Furthermore, it is expected that detection of novel mutation signatures and associated mutagens can lead to identification of novel mutagens and prevention of cancer.

At the same time, for extracting prominent mutation signatures from vast amounts of somatic mutation data,
there is an increasing need for novel latent variable modeling approaches.
Currently, only a few approaches have been proposed. 
In \cite{pmid23318258}, nonnegative matrix factorization (NMF, \cite{pmid10548103}) is used for the matrix whose elements represent frequencies of mutation patterns for each sample. 
In \cite{pmid23628380}, the number of each mutation patterns is assumed to be generated by Poisson distribution whose parameters are linear combinations of latent mutational signatures.

One of the major problems of the previous approach is that 
the number of parameters in the model grows exponentially 
with the number of contextual factors to take into account.
Many of recent approaches take into account just immediate 5' and 3' bases as contextual factors in addition to substitution patterns.
However, when considering two 5' and 3' bases to the mutated sites,
the estimated mutation signatures often become unstable due to the curse of high dimensionality, 
despite there is a potential need to observe two base 5' and 3' positions to the mutated sites \cite{pmid9683596}.
Also, it is often very hard to interpret and extract important characteristics from the estimate with high-dimensional parameters.
Furthermore, considering that latent variable modeling have enormous number of researches in statistics and machine learning, 
there is a great possibility that we can further elaborate statistical methods for mutational signature extraction problems.

In this paper, we present a novel probabilistic approach.
To avoid the problem of the exponential increase of the number of parameters,
we assumed independence among contextual factors, which can give more interpretable and robust estimation.
We demonstrate that independence assumption can robustly capture known mutational signatures with additional contextual informations.

In addition, even though it has not been widely discussed, 
the problem of identifying mutation signatures is closely related to "mixed-membership models'' in other fields, 
such as ``population admixture'' models \cite{pmid10835412} in the analysis of population structure, 
and the ``latent dirichlet allocation'' (LDA) model for document clustering \cite{Blei:2003}.
In this paper, we explicitly show the relationships between the proposed model and the mixed-membership models and discuss the relationships,
which we believe will be highly helpful for future elaboration of the statistical method for mutation signature extraction problems.

The R package for the proposed method,  {\bf pmsignature} ({\bf p}robabilistic {\bf m}utational signature),
is available at \url{https://github.com/friend1ws/pmsignature}).
The core part of the estimation process is implemented in C++ by way of the Rcpp package \cite{eddelbuettel2011rcpp},
which enables us to handle millions of somatic mutations from thousands of cancer genomes using standard desktop computers.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Methods}

\subsection*{Independence assumption in mutation signatures}

The term ``mutation signature'' is used to describe a characteristic mutational pattern observed in cancer genomes, 
that is often related to some carcinogens
(e.g., significantly frequent C $>$ A mutations in lung cancers with smoking histories).
Mathematically, mutation signatures have been characterized as frequencies or probabilities over mutation pattern \cite{pmid23318258,pmid23628380}.

Typically, mutation patterns are categorized 
by 6 substitution patterns (C$>$A, C$>$G, C$>$T, T$>$A, T$>$C, T$>$G, the original base is usually fixed to C or T for removing the redundancy of taking complementary strands),
or 96 (6 $\times$ 4 $\times$ 4) patterns by considering immediate 5' and 3' flanking bases as well as 6 substitution patterns.
Furthermore, taking account of the transcription direction (plus or minus), 
categorization by 192 mutation patterns is sometimes used \cite{pmid23945592, pmid23318258}.
In this paper, we call the factors composing mutation patterns (such as substitution patterns, adjacent bases and transcriptional strand) 
as ``mutation features.''

One of the major problems is that 
total number of mutation patterns exponentially increases
as we increase the number of mutation features to take into account.
For example, if we want to take account of up to $n$ bases 5' and 3' to the mutated site 
(which we call $-n$ site and $+n$ site, respectively, in this paper) as well as 6 substitution patterns, 
the number of parameters becomes $6 \times 4^{2n} - 1$. 
On the other hand, the mutation rates are mostly below 10 per Mb \cite{pmid23945592},
(where about 30,000 mutations are expected in entire genomic regions).
Therefore, considering even $\pm 2$ sites and substitution patterns (1535 parameters)
is often difficult because of instability of parameter estimation.

On the other hand, observing the mutation signatures observed in the previous studies (see Figure \ref{APOBEC_example}), 
many mutation signatures had clear characteristics such as:
\begin{itemize}
\item 
A certain value of mutation features is often dominant in many known mutation signatures. 
For example, the base T is highly dominant at the $-1$ site in the APOBEC signature, 
and C $>$ T substitution is highly dominant in the ultraviolet signature \cite{pmid23945592}.

\item
Proportional relationships among mutation features are often observed in many signatures. 
For example, in the APOBEC signature, 
the frequencies of bases at the $+1$ site are largely proportional across 3 major substitution patterns C$>$T, C$>$G and C$>$A,
and show consistent tendencies (A and T bases are more frequent than C and G bases).

\end{itemize}

These characteristics imply that the parameter spaces of mutation signatures can be degenerated to lower dimensional spaces in many cases, 
and we can potentially reduce the number of parameters by imposing additional restrictions.

In this paper, we propose to represent mutation signatures with independent multiple multi-nominal distributions
on each mutation feature. 
By doing this, we can reduce the number of parameters from exponential to linear with respect to the number of mutation features. 
When we consider up to $\pm n$ sites as well as 6 substitution patterns, 
the required number of parameters becomes $5 + 6n$. 
For example, when considering $\pm 2$ sites and substitution patterns,
the number of parameters becomes 17, which is far less compared to 1535 in the previous approach.
Furthermore, independence assumption enables us to come up with fairly interpretable representation for mutation signatures.

It may be argued that the independence assumption does not conform to real mechanisms of mutational processes 
because it is highly unlikely that mutagens independently choose each mutation feature such as flanking bases and substitution patterns. 
However, we would like to present the following justifications for the independence assumption:

\begin{enumerate}
\item
For representing the transcription factor binding motifs, the position specific weight matrix (PSWM), 
which is equivalent to independent multiple multinomial distributions, 
has been quite successfully utilized 
even though it is never likely that transcription factors independently choose the bases of their binding sites. 
This is probably because PSWM can capture important characteristics of transcription factor binding sites 
in spite of the independence assumption,
and can be represented in an intuitively interpretable way via sequencing logos \cite{pmid2172928}.

\item
As we demonstrate in later sections, 
the proposed approach can robustly extract most of previously-collected mutation signatures even independence assumption is imposed.

\item
Although most mutation signatures show proportional relationships among mutation features, 
a few signatures seem to violate proportional relationships and cannot be explained by independent models. 
For example, the pol $\epsilon$ mutation signature \cite{pmid23945592} puts strong probability masses 
on just two pattern TpCpT $>$ TpApT and TpCpG $>$ TpTpG, 
and the frequencies of substitution patterns and $+1$ bases are not proportional unlike the APOBEC signature.
However,  by using multiple mutation signatures, we can represent this phenomenon even with independence assumption.
\end{enumerate}

\subsection*{Mathematical representation of mutation signatures}

In this subsection, 
we show how to mathematically describe mutations collected from genome sequencing studies and mutational signatures.
First, let $\Delta^S = \{ (t_1, \cdots, t_S) |\ t_s \geq 0 (\forall s = 1, \cdots, S), \sum_{s=1}^S t_s = 1 \}$ denote S-dimensional simplex,
which is used to represent nonnegative vectors summing to $1$ throughout the paper.

Suppose each somatic mutation has $L$ mutation features, $\bm{m} = (m_1, m_2, \cdots, m_L)$, 
where each $m_l$ can take $M_l$ discrete values. Also, we set $\bm{M} = (M_1, \cdots, M_L)$.
Let $\bm{x}_{i, j} = (x_{i, j, 1}, \cdots, x_{i, j, L}), (i = 1, \cdots, I, j = 1, \cdots, J_i)$ denote the mutation feature vector for the $j$-th mutation of $i$-th cancer genome,
where $x_{i, j, l} \in \{1, \cdots, M_l \}$, $I$ is the number of available cancer genomes and $J_i$ is the number of mutations in the $i$-th cancer genome.
When taking account of 6 substitution patterns and $\pm 2$ sites,
$\bm{M} = (6, 4, 4, 4, 4)$.
See Table \ref{tab_rep} for other representation example.

Suppose that there are $K$ mutation signatures,
Let  $\bm{f}_{k, l} = (f_{k, l, 1}, \cdots, f_{k, l, M_l }) \in \Delta^{M_l}$ denote 
the multinomial distribution parameter of $k$-th mutation signature and the $l$-th mutation feature, 
then the probabilistic distribution of $k$-th signature corresponds to $\{ \bm{f}_{k, l} \}_{l=1, \cdots, L}$.

Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
We give a way of visualizing probabilistic mutation signature (see Fig. \ref{mutSig_example}), 
which is reminiscent of sequencing logos \cite{pmid2172928}.

\subsection*{Non-independent model as a special case of the proposed method}

The previous approaches, 
where each mutation signature is not a set multiple multinomial distributions but one high-dimensional multinomial distribution,
can be considered as a special case of the above framework.

Suppose each mutation feature vector (such as substitution patterns, adjacent bases, and so on) 
is re-labeled as a single-valued mutation feature by some lexicographical order (see Table \ref{tab_rep}),
then each mutation signature becomes one multinomial distribution. 
We call this representation and mutational process modeling as ``full representation,'' and ``full model.''
In this case, when taking account of 6 substitution patterns and $\pm 2$ sites,
$\bm{M} = (1536)$.
On the other hand, the usual representation, where each mutation feature has its corresponding value and individual multinomial distribution,
is called as ``independent representation,'' and ``independent model.''

The full representation model potentially represent complicated mutational processes
(e.g., a situation where C $>$ A is frequent at ApCpG sites and C $>$ T is frequent at TpCpA sites) with one signature.
However, when many mutation contextual factors are taken into account and the number of free parameters get huge,
estimated results tend to be unstable and unreliable.
Furthermore, there is a great risk that we over-interpreting the seemingly complex estimated results.

\subsection*{Generative model of mutation features for each cancer genome}

Each cancer genome naturally has multiple mutation signatures,
because we are living exposed to a variety of carcinogens. 
Also, the strength of each mutation signature varies among cancer samples, 
depending on lifestyles, genetic difference and so on.
We represent the distribution of mutation signatures for the $i$-the cancer genome 
by $\bm{q}_i = (q_{i,1}, q_{i,2}, \cdots, q_{i, K}) \in \Delta^K,  (i = 1, \cdots, I)$.

We adopt a two step model for a generative model of mutations.
First, one of the contributing mutation signature is chosen for each mutation 
depending on the signature distribution parameter of the corresponding cancer genome $\{ \bm{q}_i \}$.
Then, according to the probability distribution of the selected mutation signature, 
mutation features such as substitution patterns and flanking base pairs are generated.

The detailed description of generative process of $\{ \bm{x}_{i,j} \}$ is as follows:
For the $j$-th mutation in the $i$-th cancer genome, 
\begin{enumerate}
\item
Generate $z_{i,j} \sim \text{Multinomial} (\bm{q}_m)$, where $z_{i,j} \in \{1,\cdots,K \}$ is the underlying mutation signature causing that mutation.
\item 
For each $l (= 1, \cdots, L)$, generate the values of each mutation feature $x_{i,j,l} \sim \text{Multinomial}(\bm{f}_{z_{i,j},l})$.
\end{enumerate}

\subsection*{Relationship with mixed-membership models}

The two step generative model described in the previous subsection has close relationships with
the mixed-membership models that have been adopted in many other applications, 
such as document classification and population structure inference problems. 
In this subsection, we show the relationships between the proposed method and mixed-membership models,
slightly abusing notations to contrast the relationships with the proposed method.

In topic models \cite{Hofmann:1999,Blei:2003},
which is a form of mixed-membership models frequently used in document classification problems,
each document is assumed to have $K$ different ``topics'' in varying proportions ($\bm{q}_i \in \Delta^K$),
where each topic is characterized by a word frequency (a multinomial distribution on a set of words $W$ ($\bm{f}_k \in \Delta^W$).
And each word is assumed to be generated by one of $K$ multinomial distributions (topics).
The detailed generative process of the $j$-th word in the $i$-th document $x_{i,j}$ is:  
\begin{enumerate}
\item
Generate the underlying topic for the $j$-th word, $z_{i,j} \sim \text{Multinomial} (\bm{q}_i)$., where $z_{i,j} \in \{1,\cdots,K \}$.
\item
Generate $x_{i,j} \sim \text{Multinomial} (\bm{f}_{z_{i,j}})$, where $x_{i,j} \in \{1, \cdots, W \}$.
\end{enumerate}
Actually, the proposed model in case of $L = 1$ or ``full representation''  is mostly the same as topic models.

On the other hand, in population structure inference problems \cite{pmid10835412, pmid19648217}, 
each individual is assumed to be an admixture of $K$ ancestries in varying proportions, 
where each ancestry is characterized by the allele frequency at each SNP locus,
and each SNP genotype of an individual are assumed to be generated by the two step model:
first, one of the ancestries are chosen by the admixture ratio of ancestries given for each individual,
and then the SNP genotype is generated according to the allele frequency of the selected ancestry at that locus.
The relationships among the mutation signature model, topic models and population structure models are summarized in the Table \ref{tab_pop}.

Furthermore, close relationships between mixed-membership models and nonnegative matrix factorization,
which has been successfully used in the previous studies for mutational signature problems \cite{pmid22608084, pmid23318258, pmid23945592},
have been pointed out \cite{ding2008equivalence}.
In fact, the proposed method can be seen as non-negative matrix factorization with additional restrictions.
See Supplementary Materials for the detail of the relationships between the proposed approach and nonnegative matrix factorization.

\subsection*{Estimating parameters}

The parameters $\{ \bm{f}_{k, l} \}$ and $\{ \bm{q}_i \}$
are not given and have to be estimated from the available mutation data $\{ \bm{x}_{i,j} \}$.
On estimating the parameters for mixed-membership models, a number of approaches have been proposed.

After realizing the relationships with mixed-membership models, 
a number of  past parameter estimation techniques proposed for mixed-membership models can be tailored for the proposed model.
EM-algorithm (or its variant, called tempered EM algorithm) have been adopted in classical topic models 
for document classification (\cite{Hofmann:1999}) and population structure estimation (\cite{tang2005estimation}).
In the population structure estimation problem, \cite{pmid19648217} has proposed 
a fast block relaxation scheme using sequential quadratic programming for block updates
with a quasi-Newton acceleration of convergence \cite{zhou2011quasi}, demonstrating a great improvement over EM-algorithm.
Similar techniques are used in \cite{taddy2011estimation} for the document classification.
For the estimation of Bayesian mixed-membership models, 
(collapsed) Gibbs sampling \cite{pmid10835412,pmid14872004} and 
variational method \cite{Blei:2003,teh2006collapsed,raj2014variational} have been proposed.

In this paper, we adopt a relatively simple approach that uses EM-algorithm.
However, there is a great possibility that we can devise far more efficient approach based on past experiences.
Let $g_{i, \bm{m}}$ denote the number of the $i$-th sample's mutations with the mutation feature vector $\bm{m}$.

Introducing the auxiliary variables $\theta_{i, k, \bm{m}}$, 
we update these auxiliary variables in the E-step as

\begin{equation*}
\theta_{i, k, \bm{m}} = \frac{  q_{i,k} \prod_{l=1}^L f_{k,l,m_l} }{ \sum_{k^{\prime} = 1}^K q_{i, k^{\prime} } \prod_{l=1}^L f_{k^{\prime}, l, m_l } }.
\end{equation*}

Then, in the M-step, we update the parameters $\{ f_{k, \bm{m} } \}$ and $\{ q_{i, k} \}$ as

\begin{equation*}
f_{k, l, p} = \frac{ \sum_{\bm{m} : m_l = p} g_{i, \bm{m}} \theta_{i, k, \bm{m}} }{ \sum_{p^{\prime} } 
\sum_{\bm{m} : m_l = p^{\prime}}  g_{i, \bm{m}}\theta_{i, k, \bm{m}} },
\end{equation*}
\begin{equation*}
q_{i, k} = \frac{ \sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k, \bm{m} } }{  \sum_{k^{\prime} }\sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k^{\prime}, \bm{m} } }.
\end{equation*}

In addition, we use SQUAREM \cite{varadhan2008simple},
which is a general framework for accelerating the convergence of any fixed-point iteration such as EM algorithm. 
Furthermore, to reduce problems with convergence to local minima, 
we performe EM-algorithm several times (10 times in this paper) changing the initial points,
and adopt the estimate with the maximum log-likelihood.
For the derivation of EM-algorithm, please see Supplementary Material.

\subsection*{Adding background signatures}

There may be a possibility that the intrinsic composition of the genome sequence influences the estimated mutation signatures.
For example, the number of observed C $>$ T transitions at CpG sites may increase at promoter regions,
just because CpG dinucleotides are more frequent in those regions.
In the previous research \cite{pmid23628380}, this background problem was dealt 
by explicitly incorporating mutation ``opportunity'' coefficients into the model.

Here, to offset the influences of intrinsic sequence composition, we add background signatures 
$\{ \bm{f}_{0, \bm{m}} \} \in \Delta^{M_1 \times  \cdots \times M_L}$.
For example, when obtaining the background signatures in case of considering substitution patterns and up to $\pm 2$ bases, 
we first calculate the frequencies of 5-mers where complement sequences are taken when the central bases are A or G,
and then the frequencies divided by 3 was give on each mutation feature vectors $\bm{m}$
considering the alternated bases are equally likely.
Since we mainly deal with mutation data from exome sequencing,
background signatures were calculated on entire exonic regions.

\subsection*{Estimating standard errors}

The standard errors for the parameter estimates are calculated using bootstrap, 
where somatic mutations are resampled according to the empirical distribution of the original data $\{ \bm{x}_{i,j} \}$ for each cancer genome.
For each bootstrap resample, we performed the re-estimation using parameters obtained for the original data as initial points,
and calculated sample standard errors of the inferred mutational signatures as estimates of parameter standard errors.
In this paper, we performed 100 bootstrap resampling.

\subsection*{Selecting the number of signatures}

Determining the number of mutation signatures $K$ is an important and challenging task. 
One approach is to utilize some statistical information criteria such as AIC \cite{akaike1974new}, BIC \cite{schwarz1978estimating}.
In the population structure problems, for example, 
the Bayesian deviance \cite{pmid10835412}, 
and cross-validation \cite{alexander2011enhancements} have been suggested.
One previous study on mutation signature problems \cite{pmid23628380} utilized BIC.
The problem of using these statistical information criteria is that most of them are based on the likelihood,
where slight deviations between the specified probabilistic models and the reality sometimes lead to 
larger number of mutation signatures for compensating those deviations,
and produce results with a risk of over-interpretation.

In this paper, instead of utilizing statistical information criteria,
we adopt following strategies:

\begin{itemize}
\item
After calculating the likelihood and standard errors of parameters for a range of $K$,
the value of $K$ is determined at the point where the likelihood is sufficiently high, 
and the standard errors are sufficiently low \cite{pmid23318258}.

\item
When, for $k_1$-th and $k_2$-th mutation signatures, 
we could detect strong correlations between the estimated membership parameters 
for each cancer genome
($(q_{1,k_1}, q_{2,k_1}, \cdots, q_{I,k_1})$ and $(q_{1,k_2}, q_{2,k_2}, \cdots, q_{I,k_2})$),
and the two mutation signatures ($\{ f_{k_1, \bm{m} } \}$ and $\{ f_{k_2, \bm{m} } \}$ ) show similar patterns,
then it is likely that an excess amount of $K$ forced to split one mutation signature into two.
We stop increasing $K$ before these pairs of mutation signatures are observed.

\end{itemize}

The strategies listed above are probably not exhaustive, 
and we should add other practical strategies observing various situations.
In addition, it would be nice is we could devise automated and practical approaches for choosing $K$,
which is a possible future challenges.

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Experiments on synthetic data}

First, we would like to investigate whether the proposed approach can extract ``true'' mutation signatures or not.
Since we can not know the true mutation signatures in real biological data, we resorted to simulation studies.

Here, we generated a set of mutations changing the number of cancer genomes (10, 25, 50, 100), 
and the number of mutations for each cancer genome (10, 25, 50, 100, 250, 500, 1000).
The number of mutation signatures were set to 5 including background mutation ratio.
The mutation feature parameters and signature distribution parameters were generated by Dirichlet distribution,

\begin{equation*}
 f_{k,l,m_l} \sim \text{Dir} (\alpha \bm{1} ),\ k = 1, \cdots, K,\  l = 1, \cdots, L.
\end{equation*}
 \begin{equation*}
 q_{i, k} \sim \text{Dir} (\gamma \bm{1} ),\ i = 1, \cdots, I,
 \end{equation*}
where the $\alpha$ and $\gamma$ represent the amounts of dispersion 
for the mutation features parameters and signature distribution parameters, respectively.
When these values are smaller, only a fewer components can have larger probability masses.
and the rest will have much smaller masses.  
When these are large, all the components tend to have evenly-distributed probability masses.

Overall, we could estimate the mutation signatures very accurately.
For example, when the number of cancer genomes and mutations for each cancer genome is 25 and 100, respectively,
we could recover the true signatures very accurately in most cases (see Supplementary Figure 1 as an example).
Furthermore, in most cases, the log-likelihood stopped increasing at five mutation signatures, 
whereas the standard error of the estimated parameters started increasing past five mutation signatures (see Supplementary Figure 2),
indicating that true number of mutation signatures can be recovered by observing the trade-off between likelihood and standard-errors.

As expected, as we increase the number of either cancer genomes or mutations, the accuracy of the estimates increased.
Also, as we increase the value of $\alpha$, then the accuracy of the estimated signatures got worse.
However, even we increase the value of $\gamma$, the accuracy of the estimate signatures did not changed much.
Therefore, the dispersion of mutation feature parameters influences the accuracy of mutation signatures
more sensitively than signature distribution parameters.

\subsection*{Experiments using cancer genomes from urothelial carcinoma of the upper urinary tract}

In this section, we compare the ``full representation'' and the ``independent representation'' by examining mutation signatures obtained by the two approaches,
and investigate the robustness by downsampling experiments.
The dataset used here is a list of 14717 somatic substitutions 
collected from the study of 26 urothelial carcinomas of the upper urinary tract (UTUC) \cite{pmid23926200},
where they found a novel mutation signature: 
T $>$ A substitutions at CpTpG sites with a strong strand specificity caused by aristolochic acids (AAs).

First, we performed the proposed method,
considering substitution patterns, up to $\pm 2$ sites and strand directions as mutation features.
For the full representations, we assign one integer for each combination of these features, and thus $L = 1, \bm{M} = (3072)$.
For the independent representation, $L = 6, \bm{M} = (6, 4, 4, 4, 4, 2)$.

First, performed on various number of mutation signatures $K$, we observed how the detected mutation signatures changes (See Supplementary Figure 3).
See Supplementary Figure 4 for the values of likelihood and bootstrap-errors.
For $K = 2$ including a back-ground signatures., a mutation signature which seem to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription direction) was observed.
For $K = 3$, an additional mutation signature corresponding APOBEC enzyme (C $>$ Y at TpCpN sites) was observed
in addition to the AA mutation signature.
For $K = 4$, an additional signature (T $>$ A at NpTpN sites with strong transcription direction) 
that is similar to the AA signature was observed. 
When we check the correlation of estimated membership parameter for each cancer genome, 
strong correlation between the AA signature (CpTpG $>$ CpApG) can be observed ($R = 0.77$, see Supplementary Figure 5). 
This additional signature may be just making up for the residual of the AA signature 
which the original AA signature could not explain due to a slight deviance of probabilistic model.
Similar results were obtained when using full the representation model ($L = 1, \bm{M} = (3072)$).
Therefore, $K = 3$ seems to be one reasonable choice in terms of the interpretability
(Fig. \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}).

For the independent representation model, we could observed the depletion of G base at the $-2$ sites,
which is consistent to the previous study \cite{pmid23318258} and the result in the next subsection.
On the other hand, for the full representation model, this tendency was rather mild.
Inferred AA mutation signature had no clear characteristics at two bases 5' and 3' to the mutated site 
compared to the APOBEC mutation signature.

Next, setting the number of mutation signatures $K = 3$, assuming that the mutation signatures obtained using whole 14717 substitutions as a gold standard, 
we performed the proposed method on down-sampled data (1\%, 2.5\%, 5\%, 10\%, 25\%, 50\%), 
and compared obtained signatures with the gold standard.
For measuring the deviations of the mutation signature, the cosine distance is used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
so that the comparison between the full representation and the independent representation become possible for the same number of adjacent bases.
Trials for downsampling experiment for each ratio and model were repeated for 100 times.

As the Fig. \ref{UTUC:APOBEC_downsampling} shows,  the independent representation could recover the original signatures even when about 90 \% of the data was removed.
In this experiment, the AA signature was more robust than the APOBEC signature.
This is because the number of T $>$ A substitutions at GpTpC sites are more frequent in this dataset.
These results indicate that while both the estimation under the full representation and the independent representation could efficiently extract known mutation signatures,
and the estimation under the independent representation is more robust than the full representation.

\subsection*{Application to pan-cancer somatic mutation data}

Finally, we have applied the proposed method for the pan-cancer somatic mutation data \cite{pmid23945592}.
For each cancer type, the proposed method was applied separately.
By changing the number of mutation signatures and calculating values of the log-likelihood and bootstrap errors,
we have chosen the number of mutation signatures for each cancer type.
The estimated mutation signatures across cancer types were clustered by Frobenius Distance.
 % (see Supplementary Material for the values of the log-likelihoods and bootstrap errors for each cancer type).

The Figure \ref{nature2013_sig_summary},  \ref{nature2013_sig_member} shows the summary of the obtained mutation signatures.
In total, 27 mutation signatures were extracted.
The signature 1 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
The signature 2 (T $>$ [GT] at TpCpN sites) was observed in 12 cancer types, and related to the activity of APOBEC family.
The signature 6 and the signature 7 (C $>$ T at TpCpG and C $>$ A at TpCpT, respectively) were observed in colorectal and uterine cancers,
and associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
Interestingly, for representing the signature for the Pol $\epsilon$ dysfunction, which was represented by one signature in the previous study,
the proposed method utilized two mutation signatures for compensating simplified representation powers.
The signature 10 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
and seems to be associated with temozolomide, an anticancer drug.
The signature 13 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
The signature 19 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear.

Furthermore, a number of prominent mutation signatures observed in one cancer type 
had common characteristics with those observed in the previous study:
the signature 9 (T $>$ G at CpTpT) observed in oesophagus cancers, 
the signature 11 (C $>$ A at [CT]pCpT) observed in low grade gliomas,
the signature 16 (T $>$ C at NpTpA) observed in liver cancers, 
the signature 22 (C $>$ A at NpCp[AT]) observed in neuroblastomas,
the signature 24 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas, 
and the signature 26 (T $>$ C at GpTpN) in stomach cancers.
These results indicate that the proposed method can capture 
many of prominent mutation signatures detected in the previous study.

Many of the signatures captured in this study 
had prominent features at the two bases 5' and 3' to the mutated sites.
For the signature 2, the signatures captured different cancer types are highly consistent (Supplementary Figrue 6),
and they consistently showed depletion of G base at two bases 5' to the mutated site,
backing up the result obtained from UCUT data in the previous subsection and the previous study  \cite{pmid23318258}.
For the signature 6 and signature 7 (Supplementary Figrue 7, 8), 
signatures for the two different cancer types consistently showed abundance of T base at two bases 5' to the mutated site,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively.
For the signature 19, depletion of C base at two bases 5' to the mutated site is consistently observed (Supplementary Figrue 9).
Therefore, the proposed method can capture the characteristic features at the two bases 5' and 3' to the mutated sites.

\section*{Discussion}

In this paper, we proposed a novel framework for extracting mutation signatures from a set of somatic point substitutions.
We have shown close relationships with mixed membership models, 
that have been deeply investigated in statistical genetics and machine learning fields, as well as nonnegative matrix factorization.
We have demonstrated that the proposed method could capture meaningful characteristics of mutational patterns.

There is a lot of room for improvement by learning from past experiences and knowledge.
First, introducing certain prior distributions or penalty terms can lead more efficiency in terms of both accuracy and interpretation,
considering the success in machine learning and statistical genetics communities.
\cite{hoyer2004non, engelhardt2010analysis}.
In the documentation classification problem, 
adopting determinantal point process priors \cite{kulesza2012determinantal, kwok2012priors}
is demonstrated to be helpful for obtaining more diverse and intuitively interpretable topics,
avoiding over fine-tuning for frequent groups of words with multiple similar topics.
Second, utilizing other metrics than Euclidean distance and Kullback-Leibler Divergence might be worth investigating.
For nonnegative matrix factorization, 
the metric called $\beta$ divergence \cite{basu1998robust, eguchi2001robustifying}, 
has been successfully utilized in many research fields (reviewed in \cite{fevotte2011algorithms}).
Third, for avoiding the problem of determining the number of latent variables, 
introducing Hierarchical dirichlet processes \cite{teh2006hierarchical}, 
which is a natural extension of topic models to nonparametric Bayesian frameworks, might be helpful.

In this paper, we have demonstrated that the independent representations could obtain more robust estimates than the full representations.
Also, even in the independent representations, more complex mutational process could be represented by utilizing multiple mutation signatures.
However, we still do not know about the exact mechanisms of mutational processes and 
there might be those that should be represented by the full representations or somewhere between the full representations and the independent representations.
We should keep exploring appropriate representations of mutation signature with expertise from biology and chemistry.

We have just focused on somatic substitutions in this paper.
However, there are a variety of mutations in cancer genomes, 
such as insertions, deletions, double nucleotides substitutions, structural variations and copy number alterations.
Our framework can potentially treat those mutations by considering appropriate mutation features.
For example, for deletions, potential mutation features can be the lengths of deletions, adjacent bases, and so on.
However, detailed investigation on what mutation features are practical is a future problem.

% Do NOT remove this, even if you are not including acknowledgments.

\section*{Acknowledgments}

The first author would like to thank to Dr. Daichi Mochihashi for helpful discussion and comments on earlier version on the proposed method.
Many of the contents in this paper is deeply influenced by what was going on at Matthew Stephens Laboratory,
when the first author stayed at the University of Chicago as a visiting scholar.
The first author would like to thank to the members in Matthew Stephens and John Novembre Laboratory, especially Dr. John Novembre,
Dr. Jacob Degner for helpful discussion and comments.

% \section*{References}

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2009.bst
% style file and paste the contents of your .bbl file
% here.
%

% \bibliographystyle{plos2009}

\bibliography{plos_template}

% \section*{Figure Legends}
% This section is for figure legends only, do not include
% graphics in your manuscript file.
%
%\begin{figure}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure caption.  
%}
%\label{Figure_label}
%\end{figure}

\newpage

 \section*{Tables}
% 
% See introductory notes if you wish to include sideways tables.
%
% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements. Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
%
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\begin{table}[!ht]
\caption{
{\bf Example of representation for mutation patterns (substitution patterns and one 5' and 3' bases)}
In the independent representation, the elements of vector show substitution patterns, 5' adjacent bases and 3' adjacent bases, respectively.
For the substitution pattens, 1 to 6 values are assigned to C$>$A, C$>$G, C$>$T, T$>$A, T$>$C and T$>$G in this order.
For 5' and 3' adjacent bases, 1to 4 values are assigned to A, C, G and T.
Note that the original base is fixed to C or T to remove the redundancy of complement sequences.
}
\begin{center}
\begin{tabular}{|c|c|c|} \hline
mutation pattern & full representation & independent representation  \\ \hline
$L$ & 1  & 3  \\
$\bm{M}$ & (96)  & (6, 4, 4) \\  \hline
ApCpA $\to$ ApCpA & (1)  & (1, 1, 1)  \\ 
ApCpC $\to$ ApApC & (2)  & (1, 1, 2)  \\ 
ApCpG $\to$ ApApG & (3)  & (1, 1, 3)  \\
ApCpT $\to$ ApApT & (4)  & (1, 1, 4)  \\
CpCpA $\to$ CpApA & (5)  & (1, 2, 1) \\
$\cdots$ & $\cdots$ & $\cdots$ \\
ApCpA $\to$ ApGpA & (17)  & (2, 1, 1) \\
$\cdots$ & $\cdots$ & $\cdots$ \\
TpTpT $\to$ TpGpT & (96)  & (6, 4, 4) \\ \hline
\end{tabular}
\end{center}
\label{tab_rep}
\end{table}

\clearpage

\begin{table}[!ht]
\caption{
{\bf Relationships among mutation signature model, topic models, and population structure models.}
}
\begin{center}
\begin{tabular}{|l|l|l|l|} \hline
problem & $\bm{x}_{i,j}$ & $\bm{f}_{k}$ & $\bm{q}_{i}$  \\ \hline
mutation signature model & the $j$-th mutation  & the feature dist.  & the signature dist.  \\
& in the $i$-th cancer genome & for the $k$-th signature & for the $i$-th cancer genome \\ \hline
topic model &  the $j$-th word & the word dist. & the topic dist. \\ 
& in the $i$-th document & for the $k$-th topic & for the $i$-th document \\ \hline
population structure model & the $j$-th locus genotype & the allele freq.  & the admixture dist. \\
& of the individual $i$  & for the ancestry $k$ & for the individual $i$ \\
\hline 
\end{tabular}
\end{center}
\label{tab_pop}
\end{table}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=15cm,height=4cm]{APOBEC_example.eps}
\caption{The APOBEC signature extracted in the previous study (Signature 2 in \cite{pmid23945592}).
The barplot is divided by 6 substitution patterns.
In each division, 16 bars shows joint probability of 16 patterns of immediate 5' and 3' bases combinations 
(ApNpA, ApNpC, ApNpG, ApNpT, CpNpA, $\cdots$, TpTpT) and substitution patterns.
The strong intensities of the last four bars for the substitution patterns with original base C 
indicates that the immediate 5' base is mostly confined to T in this signature.
}
\label{APOBEC_example}
\end{figure*}

\clearpage

\begin{figure*}

\begin{minipage}{0.45\textwidth}
\begin{center}
\includegraphics[width=8cm,height=5cm]{pancreas_4_2.eps}

\end{center}
\end{minipage}
\begin{minipage}{0.55\textwidth}

\begin{center}
\makeatletter
 \def\@captype{table}
\makeatother

\subtable[substitution pattern]{\label{tab:tab11}
\small
\begin{tabular}{|c|c|c|c|c|c|}
\hline
C$>$A & C$>$G & C$>$T & T$>$A & T$>$C & T$>$G \\
\hline
0.061 & 0.020 & 0.514 & 0.000 & 0.405 & 0.000 \\
\hline
\end{tabular}
}

\subtable[adjacent bases]{\label{tab:tab12}
\small
\begin{tabular}{|c|c|c|c|c|}
\hline
positions to the mutated site & A & C &  G & T \\
\hline
-2 & 0.146 & 0.257 & 0.128 & 0.469 \\
-1 & 0.183 & 0.389 & 0.331 & 0.096 \\
+1 & 0.161 & 0.107 & 0.013 & 0.719 \\
+2 & 0.019 & 0.265 & 0.367 & 0.350 \\
\hline
\end{tabular}
}

\subtable[strand direction]{\label{tab:tab13}
\small
\begin{tabular}{|c|c|}
\hline
plus strand & minus strand  \\
\hline
0.375 & 0.625 \\
\hline
\end{tabular}
}

% \caption{sample}
% \label{label 2}
\end{center}
\end{minipage}
\caption{An example of a mutation signature and its visualization in the proposed approach.
Here, mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).}
\label{mutSig_example}

\end{figure*}

\begin{figure*}[ht]
\centering

\subfigure[APOBEC signature for the full model]{%
  \includegraphics[height=4cm,width=7.5cm,clip]{APOBEC_full5_dir_sig.eps}
  \label{UTUC:APOBEC_full5_sig}}
\quad
\subfigure[AA signature for the full model]{%
  \includegraphics[height=4cm,width=7.5cm,clip]{AA_full5_dir_sig.eps}
  \label{UTUC:AA_full5_sig}}

\subfigure[APOBEC signature for the independent model]{%
  \includegraphics[height=4cm,width=7.5cm,clip]{APOBEC_ind5_dir_sig.eps}
  \label{UTUC:APOBEC_ind5_sig}}
\quad
\subfigure[AA signature for the independent model]{%
  \includegraphics[height=4cm,width=7.5cm,clip]{AA_ind5_dir_sig.eps}
  \label{UTUC:AA_ind5_sig}}

\subfigure[APOBEC signature stability]{%
  \includegraphics[height=5.5cm,width=7.5cm,clip]{APOBEC_downsampling.eps}
  \label{UTUC:APOBEC_downsampling}}
\quad
\subfigure[AA signature stability]{%
  \includegraphics[height=5.5cm,width=7.5cm,clip]{AA_downsampling.eps}
  \label{UTUC:AA_downsampling}}
%
\caption{The mutation signatures for UTUC data, and the results of downsampling experiments. 
A total of 3072 elements in the full model mutation signatures were shown divided by substitution patterns and strand directions.}
\label{UTUC}
\end{figure*}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=15cm,height=6cm]{facet_result.eps}
\caption{The accuracy of the proposed approach for the simulated data when changing
the number of samples and mutations, 
and the amounts of dispersion parameters for the mutation features and signature distribution parameters.
}
\label{sim_facet}
\end{figure*}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=16cm,height=16cm, angle=270]{signature_summary.eps}
\caption{The summary of mutation signatures obtained using the proposed method,
where the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
}
\label{nature2013_sig_summary}
\end{figure*}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=15cm,height=15cm]{membership_summary.eps}
\caption{The summary of membership of each mutation signature across cancer types.}
\label{nature2013_sig_member}
\end{figure*}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=15cm,height=15cm]{APOBEC.eps}
\caption{The intensities at two 5' base to the mutated site for 14 mutation signatures categorized as signature 2}
\label{nature2013_sig_apobec}
\end{figure*}

% \section*{Supporting Information Legends}
%
% Please enter your Supporting Information captions below in the following format:
%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
% 
%\begin{description}
%\item {\bf}
%\item {\bf}
%\end{description}

\end{document}

### 2015-02-27 figure/script/AlexandrovEtAl_figures_temp.R

@@ -1,0 +1,431 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+##########
+# functions
+
+
+visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2) {
+  
+  if (is.na(baseCol)) {
+    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
+    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
+  }
+  
+  centerBase <- (1 + numBases) / 2;
+  
+  v1 <- vF[1,1:6];
+  V2 <- vF[2:(numBases),1:4];
+  A <- matrix(0, numBases, 4);
+  B <- matrix(0, 4, 4);
+  
+  if (trDir == TRUE) {
+    v3 <- vF[(numBases + 1),1:2];
+  }
+  
+  for (l in 1:numBases) {
+    if (l < centerBase) {
+      A[l, ] <- V2[l, ];
+    } else if (l > centerBase) {
+      A[l, ] <- V2[l - 1, ];
+    }
+  }
+  A[centerBase,2] <- sum(v1[1:3]);
+  A[centerBase,4] <- sum(v1[4:6]);
+  
+  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
+  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
+  
+  num2base <- c("A", "C", "G", "T");
+  
+  # frame();
+  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
+  
+  startx <- 0;
+  for(l in 1:numBases) {
+    
+    for(w in 1:4) {
+      endx <- startx + A[l,w]
+      polygon(c(startx, endx, endx, startx), c(0, 0, 1, 1), col = baseCol[w], border=F);
+      if (endx - startx > 1 / 4) {
+        text(0.5 * (endx + startx), 0.5, num2base[w], col="white", cex=charSize)
+      }
+      startx <- endx;
+    }
+    startx <- startx + 0.25;
+  }
+  
+  startx <- (centerBase - 1) * 1.25;
+  for (w in 1:4) {
[…]

### 2015-03-01 manuscript/plos_template.tex
paper

@@
 \centering
-\includegraphics[width=15cm,height=4cm]{APOBEC_example.eps}
+\subfigure[APOBEC signature in the previous study]{%
+  \includegraphics[width=15cm,height=4cm]{example96_APOBEC.eps}
+  \label{UTUC:APOBEC_full5_sig}}
+
+\subfigure[POLE signature in the previous study]{%
+  \includegraphics[width=15cm,height=4cm]{example96_POLE.eps}
+  \label{UTUC:AA_full5_sig}}
+
 \caption{The APOBEC signature extracted in the previous study (Signature 2 in \cite{pmid23945592}).
@@
 
-\begin{minipage}{0.45\textwidth}
+\begin{minipage}{0.6\textwidth}
 \begin{center}
-\includegraphics[width=8cm,height=5cm]{pancreas_4_2.eps}
+\includegraphics[width=10cm,height=6.66cm]{example_signature.eps}
 
@@
 \end{minipage}
-\begin{minipage}{0.55\textwidth}
+\begin{minipage}{0.4\textwidth}
 
@@
 \subtable[substitution pattern]{\label{tab:tab11}
-\small
+\footnotesize
 \begin{tabular}{|c|c|c|c|c|c|}
@@
 \hline
-0.061 & 0.020 & 0.514 & 0.000 & 0.405 & 0.000 \\
+0.201 & 0.080 & 0.314 & 0.000 & 0.405 & 0.000 \\
 \hline
@@
 \subtable[adjacent bases]{\label{tab:tab12}
-\small
+\footnotesize
 \begin{tabular}{|c|c|c|c|c|}
 \hline
-positions to the mutated site & A & C &  G & T \\
+position  & A & C &  G & T \\
 \hline
 -2 & 0.146 & 0.257 & 0.128 & 0.469 \\
--1 & 0.183 & 0.389 & 0.331 & 0.096 \\
+-1 & 0.183 & 0.389 & 0.331 & 0.097 \\
 +1 & 0.161 & 0.107 & 0.013 & 0.719 \\
-+2 & 0.019 & 0.265 & 0.367 & 0.350 \\
++2 & 0.019 & 0.265 & 0.366 & 0.350 \\
 \hline
@@
 \subtable[strand direction]{\label{tab:tab13}
-\small
+\footnotesize
 \begin{tabular}{|c|c|}
@@
 \subfigure[APOBEC signature for the full model]{%
-  \includegraphics[height=4cm,width=7.5cm,clip]{APOBEC_full5_dir_sig.eps}
+  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_APOBEC_full.eps}
   \label{UTUC:APOBEC_full5_sig}}
@@
 \subfigure[AA signature for the full model]{%
-  \includegraphics[height=4cm,width=7.5cm,clip]{AA_full5_dir_sig.eps}
+  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_AA_full.eps}
   \label{UTUC:AA_full5_sig}}
@@
 \subfigure[APOBEC signature for the independent model]{%
-  \includegraphics[height=4cm,width=7.5cm,clip]{APOBEC_ind5_dir_sig.eps}
+  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_APOBEC_ind.eps}
   \label{UTUC:APOBEC_ind5_sig}}
@@
 \subfigure[AA signature for the independent model]{%
-  \includegraphics[height=4cm,width=7.5cm,clip]{AA_ind5_dir_sig.eps}
+  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_AA_ind.eps}
   \label{UTUC:AA_ind5_sig}}
@@
 \subfigure[APOBEC signature stability]{%
-  \includegraphics[height=5.5cm,width=7.5cm,clip]{APOBEC_downsampling.eps}
+  \includegraphics[height=5.5cm,width=7.5cm,clip]{UTUC_downsampling_APOBEC.eps}
   \label{UTUC:APOBEC_downsampling}}
@@
 \subfigure[AA signature stability]{%
-  \includegraphics[height=5.5cm,width=7.5cm,clip]{AA_downsampling.eps}
+  \includegraphics[height=5.5cm,width=7.5cm,clip]{UTUC_downsampling_AA.eps}
   \label{UTUC:AA_downsampling}}
@@
 \centering
-\includegraphics[width=15cm,height=6cm]{facet_result.eps}
+\includegraphics[width=16cm,height=6cm]{simulation_result.eps}
 \caption{The accuracy of the proposed approach for the simulated data when changing
@@
 \centering
-\includegraphics[width=16cm,height=16cm, angle=270]{signature_summary.eps}
+\includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
 \caption{The summary of mutation signatures obtained using the proposed method,
@@
 \centering
-\includegraphics[width=15cm,height=15cm]{membership_summary.eps}
+\includegraphics[width=15cm,height=15cm]{corrplot.eps}
 \caption{The summary of membership of each mutation signature across cancer types.}
@@
 \centering
-\includegraphics[width=15cm,height=15cm]{APOBEC.eps}
+\includegraphics[width=10cm,height=10cm]{APOBEC_two5prime.eps}
 \caption{The intensities at two 5' base to the mutated site for 14 mutation signatures categorized as signature 2}

### 2015-03-01 figure/script/AlexandrovEtAl_figures_temp.R

@@ -152,7 +152,7 @@ convertFMatrixToVector <- function(Fmat, fdim) {
 # get the mutation signature of the specified cancer type and index
 getF <- function(type, K, fdim, ind) {
   
-  inputName <- paste("../../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
+  inputName <- paste("../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
   
   load(inputName);
   F <- Param@signatureFeatureDistribution;
@@ -165,7 +165,7 @@ getF <- function(type, K, fdim, ind) {
 # get the mutation signatures of the specified cancer type
 getSigMat <- function(type, K, fdim) {
   
-  inputName <- paste("../../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
+  inputName <- paste("../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
   
   load(inputName);
   F <- Param@signatureFeatureDistribution;
@@ -193,7 +193,7 @@ eps <- 0.6
 
 #' create the matrix showing the relationships between cancer type 
 #' and the specified numer of mutation signatures
-type2sigNum <- read.table("../../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
 
 #' reading and converting the mutation signatures for each cancer type
 for (i in 1:nrow(type2sigNum)) {
@@ -283,17 +283,17 @@ par(mar=c(0, 0, 0, 0));
 par(bg = rgb(0.9, 0.9, 0.9));
 par(xaxs = "i", yaxs = "i");
 par(mfrow=c(ceiling(length(Fs) / 4), 4));
-myOrder <- c(7, 11, 12, 15, 21,  # C > A
+# myOrder <- c(7, 11, 12, 15, 21,  # C > A
              1, 6, 10, 13, 18, 20, 23, # C > T
              2, 8, 14, 22, #C > any
              9, 16, 19, 24, #T > any
              3, 4, 5, 17, 25, 26)
 
-for (i in myOrder) {
[…]

### 2015-03-01 figure/script/UTUC_downsampling.R

@@ -13,7 +13,7 @@ Sig <- rep(0, 200 * 6 * 2);
 Type[1:1200] <- "full";
 for (i in 1:6) {
   Inds <- 1:100 + 200 * (i - 1);
-  a <- read.table(paste("../../../UTUC/result/cosineDist_full/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
+  a <- read.table(paste("../../UTUC/result/cosineDist_full/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
   Ratio[Inds] <- ratio[i];
   Ratio[Inds + 100] <-ratio[i];                       
   Dist[Inds] <- a[,"AA"];
@@ -26,7 +26,7 @@ for (i in 1:6) {
 Type[1201:2400] <- "ind";
 for (i in 1:6) {
   Inds <- 1:100 + 200 * (i - 1) + 1200;
-  a <- read.table(paste("../../../UTUC/result/cosineDist_ind/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
+  a <- read.table(paste("../../UTUC/result/cosineDist_ind/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
   Ratio[Inds] <- ratio[i];
   Ratio[Inds + 100] <-ratio[i];                       
   Dist[Inds] <- a[,"AA"];
@@ -60,7 +60,7 @@ ggplot(summary.AA, aes(x=ratio, y=cosine_similality, colour=type, group=type, fi
         legend.text = element_text(size = rel(1.5)),
         legend.title = element_text(size = rel(1.5)));
 
-ggsave("../../result/UTUC_downsampling_AA.eps", width=7.5, height=5.5, units = "in");
+ggsave("../../manuscript/UTUC_downsampling_AA.eps", width=7.5, height=5.5, units = "in");
 
 
 summary.APOBEC <- summary_UTUC.downsampling %>% filter(signature=="APOBEC");
@@ -85,7 +85,7 @@ ggplot(summary.APOBEC, aes(x=ratio, y=cosine_similality, colour=type, group=type
         legend.text = element_text(size = rel(1.5)),
         legend.title = element_text(size = rel(1.5)));
 
-ggsave("../../result/UTUC_downsampling_APOBEC.eps", width=7.5, height=5.5, units = "in");
+ggsave("../../manuscript/UTUC_downsampling_APOBEC.eps", width=7.5, height=5.5, units = "in");
 
 
 

### 2015-03-01 figure/script/UTUC_signature.R

@@ -1,31 +1,31 @@
 library(pmsignature);
 
-load("../../../UTUC/result/Param_full/3.Rdata");
+load("../../UTUC/result/Param_full/3.Rdata");
 Param <- resultForSave[[1]];
 
 if (sum(Param@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
   visPMSignature(Param, 1);
-  ggsave("../../result/UTUC_APOBEC_full.eps", height = 4, width = 6);
+  ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
   visPMSignature(Param, 2);
-  ggsave("../../result/UTUC_AA_full.eps", height = 4, width = 6); 
+  ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5); 
 } else {
   visPMSignature(Param, 2);
-  ggsave("../../result/UTUC_APOBEC_full.eps", height = 4, width = 6);
+  ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
   visPMSignature(Param, 1);
-  ggsave("../../result/UTUC_AA_full.eps", height = 4, width = 6);   
+  ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5);   
 }
   
-load("../../../UTUC/result/Param_ind/3.Rdata");
+load("../../UTUC/result/Param_ind/3.Rdata");
 Param <- resultForSave[[1]];
 
 if (sum(Param@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param@signatureFeatureDistribution[2,1,4] > 0.75) {
-  visPMSignature(Param, 1, charSize = 0.8);
-  dev.copy2eps(file="../../result/UTUC_APOBEC_ind.eps", height = 4, width = 6);
-  visPMSignature(Param, 2, charSize = 0.8);
-  dev.copy2eps(file="../../result/UTUC_AA_ind.eps", height = 4, width = 6);
+  visPMSignature(Param, 1, charSize = 1);
[…]

### 2015-03-01 figure/script/example_figures.R

@@ -8,7 +8,7 @@ G <- readMPFile(inputFile, numBases = 3, type = "full");
 Param <- getPMSignature(G, K = 2);
 
 # reading the signature from Alexandrov et al. Nature 2013
-nature2013_sig_raw <- read.table("../../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
+nature2013_sig_raw <- read.table("../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
 nature2013_sig <- t(nature2013_sig_raw[,4:30]);
 colnames(nature2013_sig) <- nature2013_sig_raw[,3];
 
@@ -16,10 +16,10 @@ Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[3,];
 Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[11,];
 
 visPMSignature(Param, 1)
-ggsave("../../result/example96_APOBEC.eps", width=12, height=3, units="in");
+ggsave("../../manuscript/example96_APOBEC.eps", width=12, height=3, units="in");
 
 visPMSignature(Param, 2)
-ggsave("../../result/example96_POLE.eps", width=12, height=3, units="in");
+ggsave("../../manuscript/example96_POLE.eps", width=12, height=3, units="in");
 ####################
 
 
@@ -37,10 +37,10 @@ Param@signatureFeatureDistribution[1,4,1:4] <- c(0.161, 0.107, 0.013, 0.719);
 Param@signatureFeatureDistribution[1,5,1:4] <- c(0.019, 0.265, 0.366, 0.350);
 Param@signatureFeatureDistribution[1,6,1:2] <- c(0.375, 0.625);
   
-visPMSignature(Param, 1);
+visPMSignature(Param, 1, charSize = 1.2);
 
-outputName <- "../../result/example_signature.eps";
-dev.copy2eps(file=outputName, height = 4, width = 8, pointsize = 18);
+outputName <- "../../manuscript/example_signature.eps";
+dev.copy2eps(file=outputName, height = 6.66, width = 10);
 par(.pardefault);
 
 

### 2015-03-01 figure/script/simulation_result.R

@@ -19,7 +19,7 @@ for (i1 in 1:7) {
   for (i2 in 1:4) {
     for (i3 in 1:3) {
       for (i4 in 1:3) {
-        a <- read.table(paste("../../../simulation/result/cosineDist/", mutNum[i1], "_", sampleNum[i2], "_", alpha[i3], "_", gamma[i4], ".txt", sep=""), sep="\t", header=FALSE);
+        a <- read.table(paste("../../simulation/result/cosineDist/", mutNum[i1], "_", sampleNum[i2], "_", alpha[i3], "_", gamma[i4], ".txt", sep=""), sep="\t", header=FALSE);
 
         Dist[Ind] <- rowSums(a) / 4;
         MutNum[Ind] <- mutNum[i1];
@@ -64,5 +64,5 @@ ggplot(summary_simulation.result, aes(x=factor(mutNum), y=cosine_similality, yma
         panel.grid.major.x = element_line(colour="grey60", linetype="dashed"),
         strip.text= element_text(face="bold", size=rel(1.2)));
 
-ggsave("../../result/simulation_result.eps", width=15, height=6);
+ggsave("../../manuscript/simulation_result.eps", width=15, height=6);
 

### 2015-03-01 supp/supp.tex
paper figures updated

\documentclass{article}
\textheight24cm\textwidth16cm\topmargin-1cm
\oddsidemargin0cm\evensidemargin0cm

\usepackage{amsmath,amssymb}
% \usepackage[dvipdfmx]{graphicx}
\usepackage{graphicx}
\usepackage{subfigure}

\usepackage{url}
\usepackage{bm}
\usepackage{setspace}

% \usepackage{algorithmic}
% \usepackage{algorithm}

\newtheorem{thm}{Theorem}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newcommand{\proof}{\noindent Proof.\ \ }
% \renewcommand{\algorithmiccomment}[1]{// #1}

%\pagestyle{empty}

\begin{document}

% \baselineskip 6.5mm

\vspace*{1.0cm}
{\LARGE \bf Supplementary material}
\vspace*{0.25cm}

\section*{Derivation of EM algorithm}

The complete log-likelihood including missing data $\{ \bm{z}_i \}$ for the proposed model is
\begin{equation*}
\sum_{i = 1}^I \sum_{j=1}^J \sum_{k=1}^K I(z_{i,j} = k) \Bigl( \sum_{l = 1}^L \log f_{k, l, x_{i,j,l}} + \log q_{i,k} \Bigr).
\end{equation*}
Here, we introduce the variable for conditional probability for $z_{i,j}$ given the parameters and the mutation features $\bm{x}_{i,j}$,
\begin{equation*}
\theta_{i,k, \bm{m}} = \Pr \bigl( z_{i,j} = k \big| \bm{x}_{i, j} = \bm{m}, \{ \bm{f}_{k, l} \}, \{  \bm{q}_i \} \bigr)
\end{equation*}
Note that this conditional probability just depends on the value of mutation feature $\bm{m} = (m_1, \cdots, m_L)$, not on the index $j$.
Then, the expected complete log-likelihood augmented by Lagrange multipliers is calculated as
\begin{equation*}
\sum_{i=1}^I \sum_{\bm{m}} g_{i, \bm{m} } \sum_{k=1}^K \theta_{i, k, \bm{m}}   \Bigl( \sum_{l = 1}^L \log f_{k, l, m_l} + \log q_{i,k} \Bigr)
+ \sum_{k=1}^K \sum_{l=1}^L \tau_{k,l} \bigl( 1 - \sum_{p=1}^{M_l} f_{k,l,p} \bigr)
+ \sum_{i=1}^I \rho_i \bigl( 1 - \sum_{k=1}^K q_{i,k} \bigr).
\end{equation*}
Differentiating it leads to following stationary equations: 
\begin{align*}
\sum_{i=1}^I \sum_{\bm{m}: m_l = p} g_{i, \bm{m}} \theta_{i,k,\bm{m}} - \tau_{k,l} f_{k,l,p} & = 0,\ (p=1, \cdots, M_l, k=1, \cdots, K,\ l=1, \cdots, L)., \\
\sum_{\bm{m}} g_{i,\bm{m}} \theta_{i,k,\bm{m}} - \rho_i q_{i, k} & = 0,\ (k = 1, \cdots, K, i = 1, \cdots, I).
\end{align*}
Then, by eliminating Lagrange multipliers, updating rules can be obtained.

\section*{Relationship with nonnegative matrix factorization}

First, for ease of explanation, let assume that the full representation'' representation ($L = 1$) is used.
Suppose that each $\bm{m}$ has unique appropriate index from 1 to  $| \bm{M} | = \prod_{l=1}^L M_l$ (the number of possible mutation patterns),
so that $\bm{m}$ can be indices of matrices.

Let $G = \{ g_{i, \bm{m}} \}$ denote the $I \times | \bm{M} | $ matrix, 
where $g_{i, \bm{m}}$ is the number of mutations whose mutation patters are equal to $\bm{m}$ in the $i$-th cancer genome.
Nonnegative matrix factorization aims to find low rank decomposition, 
$G \sim \tilde{Q} F$,
where $\tilde{Q} = \{ \tilde{q}_{i,k} \}$ and $F = \{ f_{k, \bm{m}} \}$ are nonnegative matrix, 
and row vectors of $F$ are often restricted to be sum to one. 
We used the notation $\tilde{Q}$ instead of $Q$
to represent that the row vectors of $\tilde{Q}$ are not normalized to sum to one in general.

For solving NMF, 
the previous study \cite{pmid23318258} used the following updating rule \cite{Lee00algorithmsfor}:
\begin{equation*}
f_{k,m} \gets f_{k,m} \frac{ ( \tilde{Q}^T G)_{k,m} }{ ( \tilde{Q}^T \tilde{Q} F  )_{k,m} },\  \
\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ (G F^T)_{i,k} }{ ( \tilde{Q} F F^T )_{i,k} },
\end{equation*}
that reduces the {\it Euclidean distance} $|| G - \tilde{Q} F ||$. 
Therefore, the optimization problem for the existing approach is
\begin{equation}
\begin{aligned}
& \text{minimize}
& & ||G - \tilde{Q} F || \\
& \text{subject to}
& & \sum_{ \bm{m} } f_{k, \bm{m}} =  1, \; k = 1, \cdots, K \\ &
& & f_{k, \bm{m}} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
\end{aligned}
\end{equation}

On the other hand, there is another type of updating rule \cite{Lee00algorithmsfor}:
\begin{equation*}
f_{k,m} \gets f_{k,m} \frac{ \sum_i \tilde{q}_{i,k} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_i \tilde{q}_{i,k} },
\end{equation*}
\begin{equation*}
\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ \sum_m f_{k,m} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_m f_{k,m} }.
\end{equation*}
that reduces the Kullback-Liebler Divergence:
\begin{equation*}
KL(G ||  \tilde{Q} F) = \sum_{i,m} \Bigl( g_{i,m} \log  \frac{ g_{i,m} }{ (\tilde{Q} F)_{i,m} } - g_{i,m} + ( \tilde{Q} f )_{i,m} \Bigr).
\end{equation*}

In general cases including the independent representation, 
there is restrictions $f_{k, \bm{m}} = \prod_l f_{k, l, m_l}$ by smaller set of parameters.
Let us consider the following optimization problem with the Kullback-Liebler Divergence and the restrictions on $F$:
\begin{equation}
\begin{aligned}
& \text{minimize}
& & KL(G ||  \tilde{Q} F) \\
& \text{subject to}
& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
\end{aligned}
\label{KL}
\end{equation}

In fact, this is equivalent to the proposed method, whose optimization problem can be written as:
\begin{equation}
\begin{aligned}
& \text{maximize}
& & L(Q, F | G) \bigl(= \sum_{i,m}  g_{i,m} \log (Q F)_{i,m} \bigr)  \\
& \text{subject to}
& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
& & \sum_k  q_{i, k} =  1, \; i = 1, \cdots, I \\ &
& & q_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
\end{aligned}
\label{likelihood}
\end{equation}

\begin{prop}
When $(Q, F) = (Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}), 
then $(\tilde{Q}, F) = (R^* Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}).
On the other hand, when $(\tilde{Q}, F) = (\tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}),
then $(Q, F) = (R^{*-1} \tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}),
where $R^* = \text{diag} (r_1^*, \cdots, r_I^*), r_i^* = \sum_{\bm{m}} g_{i,\bm{m}}, i = 1, \cdots, I$.
\end{prop}
\proof
This is because
\begin{equation*}
KL(G ||  \tilde{Q} F) = -  \sum_i \Bigl( ( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i \Bigr)  - L(Q, F | G) + \text{(constant value)},
\end{equation*}
where $Q$ is row-normalized matrix for $\tilde{Q}$, $\tilde{r}_i = \sum_k q_{i, k}$ for each $i$, 
and $( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i$ takes its maximum at $\tilde{r}_i = r_i^*$.
\hfill $\square$

% \bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
\bibliographystyle{plain}
\bibliography{document}

\clearpage

\begin{figure}
\centering
\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
\caption{}
\label{example_signature.pd}
\end{figure}

\clearpage

\begin{figure}
\centering
\includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
\caption{}
\label{example_summary.pd}
\end{figure}

\clearpage

\begin{figure}
\centering

\subfigure[Estimated mutation signatures ($K = 2$)]{%
  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K2.eps}
  \label{UTUC:sigK2_sig}}

\subfigure[Estimated mutation signatures ($K = 3$)]{%
  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K3.eps}
  \label{UTUC:sigK3_sig}}

\subfigure[Estimated mutation signatures ($K = 4$)]{%
  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K4.eps}
  \label{UTUC:sigK4_sig}}

\caption{The result of estimated mutation signatures and membership parameters for UTUC data 
when changing the number of mutation signatures ($K$).}
\label{UTUC_multiK}
\end{figure}

\clearpage

\begin{figure}
\centering
\includegraphics[width=15cm,height=5cm]{UTUC_stat.eps}
\caption{}
\label{example_summary.pd}
\end{figure}

\clearpage

\begin{figure}
\centering
\includegraphics[width=10cm,height=10cm]{UTUC_AA_AAlike_cor.eps}
\caption{}
\label{example_signature.pd}
\end{figure}

\clearpage

% \begin{figure}
% \centering
% \includegraphics[width=15cm,height=15cm]{sig2_group.eps}
% \caption{Mutation signatures obtained for several cancer types categorized as Signature 2}
% \label{example_summary.pd}
% \end{figure}

% \clearpage

% \begin{figure}
% \centering
% \includegraphics[width=15cm,height=3.5cm]{sig6_group.eps}
% \caption{Mutation signatures obtained for several cancer types categorized as Signature 6}
% \label{example_summary.pd}
% \end{figure}

% \clearpage

% \begin{figure}
% \centering
% \includegraphics[width=15cm,height=3.5cm]{sig7_group.eps}
% \caption{Mutation signatures obtained for several cancer types categorized as Signature 7}
% \label{example_summary.pd}
% \end{figure}

% \clearpage

% \begin{figure}
% \centering
% \includegraphics[width=15cm,height=3.5cm]{sig19_group.eps}
% \caption{Mutation signatures obtained for several cancer types categorized as Signature 19}
% \label{example_summary.pd}
% \end{figure}

\end{document}

### 2015-03-01 UTUC/stat_summary.R

@@ -1,0 +1,122 @@
+
+# strand direction information
+## ind model
+
+load("summary/hoang_ind5_dir_1_1.Rdata");
+
+strandInfo <- data.frame(sig=c("AA", "APOBEC"), type=c("ind", "ind"), strand=c("plus", "plus"), prob=c(lret[[1]][[1]][1,6,1], lret[[1]][[1]][2,6,1]));
+
+temp <- data.frame(sig=c("AA", "APOBEC"), type=c("ind", "ind"), strand=c("minus", "minus"), prob=c(lret[[1]][[1]][1,6,2], lret[[1]][[1]][2,6,2]));
+strandInfo <- rbind(strandInfo, temp);
+
+
+
+load("summary/hoang_full5_dir_1_1.Rdata");
+
+plus_AA_full <- sum(lret[[1]][[1]][1,,][seq(1, 3072, 2)]);
+plus_APOBEC_full <- sum(lret[[1]][[1]][2,,][seq(1, 3072, 2)]);
+
+temp <- data.frame(sig=c("AA", "APOBEC"), type=c("full", "full"), strand=c("plus", "plus"), prob=c(plus_AA_full, plus_APOBEC_full));
+strandInfo <- rbind(strandInfo, temp);
+
+minus_AA_full <- sum(lret[[1]][[1]][1,,][seq(2, 3072, 2)]);
+minus_APOBEC_full <- sum(lret[[1]][[1]][2,,][seq(2, 3072, 2)]);
+
+temp <- data.frame(sig=c("AA", "APOBEC"), type=c("full", "full"), strand=c("minus", "minus"), prob=c(minus_AA_full, minus_APOBEC_full));
+strandInfo <- rbind(strandInfo, temp);
+
+
+ggplot(strandInfo %.% filter(sig=="AA"), aes(x=type, y=prob, fill=strand)) + geom_bar(stat="identity");
+> ggplot(strandInfo %.% filter(sig=="APOBEC"), aes(x=type, y=prob, fill=strand)) + geom_bar(stat="identity");
+
+# 2 base 5' to the mutated site
+## ind model 
+
+load("summary/hoang_ind5_dir_1_1.Rdata");
+
+baseTwoFivePrime <- data.frame(sig=rep("AA", 4), type=rep("ind", 4), base=c("A", "C", "G", "T"), prob=lret[[1]][[1]][1,2,1:4]);
+
[…]

### 2015-03-01 UTUC/summary_plot.R

@@ -1,0 +1,38 @@
+
+PMSSummaryPlot <- function(type) {
+  
+  Ls <- rep(0, 5);
+  Bs <- rep(0, 5);
+  Cs <- rep(0, 5);
+  
+  for(K in 2:6) {
+    fileName <- paste("example/", type, "/", K, "_1_1.Rdata", sep="");
+    load(fileName);
+    Ls[(K - 1)] <- lret[[1]][[3]];
+    Cs[(K - 1)] <- max(cor(lret[[1]][[2]]) - 2 * diag(rep(1, K)));
+    
+    bF <- lret[[2]][[1]];
+    bFs <- matrix(0, 100, K - 1);
+    for(b in 1:100) {
+      for(k in 1:(K - 1)) {
+        bFs[b, k] <- sum(bF[b,k,,]);
+      }
+      Bs[(K - 1)] <- sum(bFs) / ((K - 1) * 100);
+    } 
+    
+  }
+  
+  par(mfrow=c(1, 3));
+  plot(2:6, Ls, xlab="#signature", ylab="log-likelihood", type="l");
+  grid();
+  plot(2:6, Bs, xlab="#signature", ylab="bootstrap-error", type="l");
+  grid();
+  plot(2:6, Cs, xlab="#signature", ylab="maximum cor. efficient", type="l");
+  grid(); 
+  
+}
+
+#' export eps width=1000, height=400
+PMSSummaryPlot("hoang_ind5_dir"); # 8 * 4 inches
+
+

### 2015-03-01 figure/script/AlexandrovEtAl_figures_temp.R

@@ -306,6 +306,16 @@ for (i in 1:length(Fs)) {
 }
 
 
+if (i < ceiling(length(Fs) / 4) * 4) {
+  i <- i + 1;
+  for (j in i:(ceiling(length(Fs) / 4) * 4)) {
+    plot.new();
+    plot.window(xlim=c(-0.3, 6.3), ylim=c(-0.3, 3.8));
+    polygon(c(-0.25, 6.6, 6.6, -0.25), c(-0.3, -0.3, 3.75, 3.75), col = "white", border = FALSE);
+  }
+}
+
+
 outputName <- "../../manuscript/AlexandrovEtAl_mergedSignature.eps";
 dev.copy2eps(file=outputName, height = ceiling(length(Fs) / 4) * 2.5, width = 16, pointsize = 18);
 par(.pardefault);

### 2015-03-01 figure/script/UTUC_stat.R

@@ -7,7 +7,7 @@ type <- c();
 
 
 for(K in 2:6) {
-  fileName <- paste("../../../UTUC/result/Param_ind/", K, ".Rdata", sep="");
+  fileName <- paste("../../UTUC/result/Param_ind/", K, ".Rdata", sep="");
   load(fileName);
   
   sigNum <- c(sigNum, K, K, K);
@@ -53,6 +53,6 @@ ggplot(UTUC_stat, aes(x = sigNum, y = value)) +
         panel.grid.major.x = element_line(colour="grey60", linetype="dashed"),
         strip.text = element_text(face="bold", size=rel(1.2)));
 
-ggsave("../../result/UTUC_stat.eps", width=15, height=5, units = "in");
+ggsave("../../supp/UTUC_stat.eps", width=15, height=5, units = "in");
 
 

### 2015-03-01 figure/script/UTUC_variousK.R

@@ -1,5 +1,9 @@
 # goal: generating several figures from the results of PMSignature on Hoang et al. (2013) data.
 
+# 2015 3/1 Y.S
+# The size of legend is influenced by the plot device size in RStudio.
+# I trid to fix this probelm, but I couldn't. 
+# When make the figures, please enlarge the device so that the legend sizes in the figures are natural...
 
 .pardefault <- par(no.readonly = TRUE);
 ##########
@@ -10,29 +14,92 @@ library(RColorBrewer);
 
 PMSSigPlot <- function(K) {
   
-  fileName <- paste("../../../UTUC/result/Param_ind/", as.character(K), ".Rdata", sep="");
+  fileName <- paste("../..//UTUC/result/Param_ind/", as.character(K), ".Rdata", sep="");
   load(fileName);
 
-  mat <- matrix(c(1:4, rep(K, 4)), 2, 4, byrow = TRUE);
-  
-  if (K <= 4) {
-    for (k in K:4) {
+  # prepare matrix for drawing multiple figures
+  mat <- matrix(c(1:3, rep(K, 3)), 2, 3, byrow = TRUE);
+  if (K <= 3) {
+    for (k in K:3) {
       mat[1, k] <- 0;
     }
   }
-  
   layout(mat);
+  
+  # memorize the current parameter and change the margin slightly
   tempMar <- par("mar");
-  par(mar= 0.4 * tempMar);
-  for (k in 1:(K-1)) {
-    visPMSignature(resultForSave[[1]], k);
+  par(mar = 0.4 * tempMar);
+ 
+  
+  F <- resultForSave[[1]]@signatureFeatureDistribution;
+  
+  # check the index corresponding each signature types
+  AA_ind <- NULL;
+  AA_like_ind <- NULL;
+  APOBEC_ind <- NULL;
+  for (k in 1:(K - 1)) {
+    # AA
+    if (F[k,1,4] > 0.8 & F[k,3,2] > 0.5 & F[k,4,3] > 0.5 & F[k,6,2] > 0.6) {
+      AA_ind <- k;
+    } else if (sum(F[k,1,1:3]) > 0.8 & F[k,3,4] > 0.7) {
+      APOBEC_ind <- k;
+    } else if (F[k,1,4] > 0.6 & F[k,6,2] > 0.6) {
+      AA_like_ind <- k;
+    }
   }
   
-  par(mar= 1.3 * tempMar);
-  Q <- resultForSave[[1]]@sampleSignatureDistribution;
[…]

### 2015-03-01 figure/script/simulation_example_stat_signature.R

@@ -66,7 +66,7 @@ ggplot(simulation_stat, aes(x = sigNum, y = value)) +
         panel.grid.minor.x = element_line(colour="grey60", linetype="dashed"),
         strip.text = element_text(face="bold", size=rel(1.2)));
 
-ggsave("../../result/simulation_stat.eps", width=15, height=5, units = "in");
+ggsave("../../supp/simulation_stat.eps", width=15, height=5, units = "in");
 
 
 
@@ -116,6 +116,6 @@ for (i in 1:length(relvect)) {
 }
 
 
-dev.copy2eps(file="../../result/simulation_signature_example.eps", height = 5.5, width = 15);
+dev.copy2eps(file="../../supp/simulation_signature_example.eps", height = 5.5, width = 15);
 
 

### 2015-03-01 manuscript/plos_template.tex
prepared signature list figures

@@
 \begin{figure*}[b]
-\centering
-\includegraphics[width=10cm,height=10cm]{APOBEC_two5prime.eps}
+
+\subfigure[APOBEC signature list]{%
+  \includegraphics[width=16cm,height=7.5cm]{signatureList_APOBEC.eps}
+  \label{APOBEC_sig_list}}
+
+\subfigure[POLE signature in the previous study]{%
+  \includegraphics[width=7.5cm,height=7.5cm]{APOBEC_two5prime.eps}
+  \label{APOBEC_twofive}}
+
+\caption{The intensities at two 5' base to the mutated site for 14 mutation signatures categorized as signature 2}
+\label{nature2013_sig_apobec}
+\end{figure*}
+
+\begin{figure*}[b]
+
+\subfigure[POLE1 signature list]{%
+  \includegraphics[width=8cm,height=2.5cm]{signatureList_POLE1.eps}
+  \label{POLE1_sig_list}}
+
+\subfigure[POLE2 signature list]{%
+  \includegraphics[width=8cm,height=2.5cm]{signatureList_POLE2.eps}
+  \label{POLE2_sig_list}}
+
+\subfigure[UV signature list]{%
+  \includegraphics[width=8cm,height=2.5cm]{signatureList_UV.eps}
+  \label{UV_sig_list}}
+
+\subfigure[LMST signature list]{%
+  \includegraphics[width=8cm,height=2.5cm]{signatureList_LMST.eps}
+  \label{LMST_sig_list}}
+
 \caption{The intensities at two 5' base to the mutated site for 14 mutation signatures categorized as signature 2}

### 2015-03-01 figure/script/AlexandrovEtAl_figures_temp.R

@@ -277,42 +277,112 @@ for (i in 1:length(removeList)) {
 }
 
 
+####################
+# change the order of the signature
+remainInd <- 1:length(Fs);
+newOrder <- c();
+
+for (k in remainInd) {
+  if (Fs[[k]][1,1] > 0.6) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+for (k in remainInd) {
+  if (Fs[[k]][1,2] > 0.6) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+for (k in remainInd) {
+  if (Fs[[k]][1,3] > 0.6) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+for (k in remainInd) {
+  if (sum(Fs[[k]][1,1:3]) > 0.75) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+
+for (k in remainInd) {
+  if (sum(Fs[[k]][1,4]) > 0.6) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+for (k in remainInd) {
+  if (sum(Fs[[k]][1,5]) > 0.6) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+for (k in remainInd) {
+  if (sum(Fs[[k]][1,6]) > 0.6) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+
+for (k in remainInd) {
+  if (sum(Fs[[k]][1,4:6]) > 0.75) {
+    newOrder <- c(newOrder, k);
+    remainInd <- remainInd[-which(remainInd == k)];
+  }
+}
+
+newOrder <- c(newOrder, remainInd)
+##########
+
+
 #' generate the figure of mutation signatures
 
 par(mar=c(0, 0, 0, 0));
 par(bg = rgb(0.9, 0.9, 0.9));
 par(xaxs = "i", yaxs = "i");
 par(mfrow=c(ceiling(length(Fs) / 4), 4));
-# myOrder <- c(7, 11, 12, 15, 21,  # C > A
[…]

### 2015-03-03 manuscript/plos_template.tex
added script for making two 5’ bases of UTUC APOBEC signature

--- 2015-03-01 manuscript/plos_template.tex
@@
 
-On the other hand, observing the mutation signatures observed in the previous studies (see Figure \ref{APOBEC_example}), 
+On the other hand, observing the mutation signatures observed in the previous studies, 
 many mutation signatures had clear characteristics such as:
@@
 For example, the base T is highly dominant at the $-1$ site in the APOBEC signature, 
-and C $>$ T substitution is highly dominant in the ultraviolet signature \cite{pmid23945592}.
+and C $>$ T substitution is highly dominant in the ultraviolet signature \cite{pmid23945592} (see Figure \ref{APOBEC_nature2013_sig}).
 
 \item
-Proportional relationships among mutation features are often observed in many signatures. 
-For example, in the APOBEC signature, 
-the frequencies of bases at the $+1$ site are largely proportional across 3 major substitution patterns C$>$T, C$>$G and C$>$A,
+Proportional relationships among mutation features are often observed in many signatures as shown in the APOBEC signature, 
+where the frequencies of bases at the $+1$ site are largely proportional across 3 major substitution patterns C$>$T, C$>$G and C$>$A,
 and show consistent tendencies (A and T bases are more frequent than C and G bases).
@@
 a few signatures seem to violate proportional relationships and cannot be explained by independent models. 
-For example, the pol $\epsilon$ mutation signature \cite{pmid23945592} puts strong probability masses 
+For example, the pol $\epsilon$ mutation signature \cite{pmid23945592} (see Figure \ref{POLE_nature2013_sig}) puts strong probability masses 
 on just two pattern TpCpT $>$ TpApT and TpCpG $>$ TpTpG, 
@@
 Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
-We give a way of visualizing probabilistic mutation signature (see Fig. \ref{mutSig_example}), 
+We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
 which is reminiscent of sequencing logos \cite{pmid2172928}.
@@
 
-Overall, we could estimate the mutation signatures very accurately.
-For example, when the number of cancer genomes and mutations for each cancer genome is 25 and 100, respectively,
-we could recover the true signatures very accurately in most cases (see Supplementary Figure 1 as an example).
-Furthermore, in most cases, the log-likelihood stopped increasing at five mutation signatures, 
+As Figure \ref{sim_facet} shows, we could estimate the mutation signatures very accurately overall.
+(see Supplementary Figure 1 for an example where the number of cancer genomes and mutations for each cancer genome is 25 and 100, respectively).
+In most cases, the log-likelihood stopped increasing at five mutation signatures, 
 whereas the standard error of the estimated parameters started increasing past five mutation signatures (see Supplementary Figure 2),
-indicating that true number of mutation signatures can be recovered by observing the trade-off between likelihood and standard-errors.
+indicating that true number of mutation signatures can be recovered by observing the trade-off between likelihood and standard-errors at least in an ideal case.
 
@@
 
-In this section, we compare the ``full representation'' and the ``independent representation'' by examining mutation signatures obtained by the two approaches,
+In this section, we compare the ``full model'' and the ``independent model'' by examining mutation signatures obtained by the two approaches,
 and investigate the robustness by downsampling experiments.
@@
 
-First, performed on various number of mutation signatures $K$, we observed how the detected mutation signatures changes (See Supplementary Figure 3).
-See Supplementary Figure 4 for the values of likelihood and bootstrap-errors.
-For $K = 2$ including a back-ground signatures., a mutation signature which seem to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription direction) was observed.
-For $K = 3$, an additional mutation signature corresponding APOBEC enzyme (C $>$ Y at TpCpN sites) was observed
-in addition to the AA mutation signature.
+First, performed on various number of mutation signatures $K$ (including a background signature), we observed how the detected mutation signatures changes (See Supplementary Figure 3).
+For $K = 2$, a mutation signature which seem to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription direction) was observed (Supplementary Figure 3 (a)).
+For $K = 3$, an additional mutation signature corresponding APOBEC enzyme (C $>$ [AGT] at TpCpN sites) was observed
+in addition to the AA mutation signature (Figure \ref{UTUC:APOBEC_ind5_sig}, \ref{UTUC:AA_ind5_sig}, Supplementary Figure 3 (b)).
 For $K = 4$, an additional signature (T $>$ A at NpTpN sites with strong transcription direction) 
-that is similar to the AA signature was observed. 
+that is somewhat similar to the AA signature was observed (Supplementary Figure 3 (c)). 
 When we check the correlation of estimated membership parameter for each cancer genome, 
-strong correlation between the AA signature (CpTpG $>$ CpApG) can be observed ($R = 0.77$, see Supplementary Figure 5). 
+strong correlation between the AA signature (CpTpG $>$ CpApG) and that additional AA-like signature can be observed ($R = 0.77$, see Supplementary Figure 5). 
 This additional signature may be just making up for the residual of the AA signature 
-which the original AA signature could not explain due to a slight deviance of probabilistic model.
-Similar results were obtained when using full the representation model ($L = 1, \bm{M} = (3072)$).
-Therefore, $K = 3$ seems to be one reasonable choice in terms of the interpretability
-(Fig. \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}).
+which the original AA signature could not explain due to a slight deviance of the probabilistic model.
+The values of likelihood continued increasing as the $K$ increase, while the bootstrap-errors started to increase at $K = 5$ (See Supplementary Figure 4).
+The strong correlation among estimated membership parameters started to be shown at $K = 4$
+Considering all these factors together, $K = 3$ seems to be one reasonable choice in terms of the interpretability,
+and we adopt $K = 3$ in the following.
 
@@
 which is consistent to the previous study \cite{pmid23318258} and the result in the next subsection.
-On the other hand, for the full representation model, this tendency was rather mild.
+On the other hand, for the full representation model, this tendency was rather mild (Figure \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}).
 Inferred AA mutation signature had no clear characteristics at two bases 5' and 3' to the mutated site 
@@
 
-Next, setting the number of mutation signatures $K = 3$, assuming that the mutation signatures obtained using whole 14717 substitutions as a gold standard, 
+Assuming that the mutation signatures obtained using whole 14717 substitutions as a gold standard, 
 we performed the proposed method on down-sampled data (1\%, 2.5\%, 5\%, 10\%, 25\%, 50\%), 
 and compared obtained signatures with the gold standard.
-For measuring the deviations of the mutation signature, the cosine distance is used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
-so that the comparison between the full representation and the independent representation become possible for the same number of adjacent bases.
+For measuring the deviations of the mutation signature, the cosine similarity is used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
+so that the comparison between the full representation and the independent representation become possible.
 Trials for downsampling experiment for each ratio and model were repeated for 100 times.
 
-As the Fig. \ref{UTUC:APOBEC_downsampling} shows,  the independent representation could recover the original signatures even when about 90 \% of the data was removed.
+As the Figure \ref{UTUC:APOBEC_downsampling} and \ref{UTUC:AA_downsampling} shows, the independent representation could recover the original signatures even when about 90 \% of the data was removed.
 In this experiment, the AA signature was more robust than the APOBEC signature.
@@
 In total, 27 mutation signatures were extracted.
-The signature 1 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
-The signature 2 (T $>$ [GT] at TpCpN sites) was observed in 12 cancer types, and related to the activity of APOBEC family.
-The signature 6 and the signature 7 (C $>$ T at TpCpG and C $>$ A at TpCpT, respectively) were observed in colorectal and uterine cancers,
+The signature 6 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
+The signature 13 (T $>$ [GT] at TpCpN sites) was observed in 11 cancer types, and related to the activity of APOBEC family.
+The signature 7 and the signature 1 (C $>$ T at TpCpG and C $>$ A at TpCpT, respectively) were observed in colorectal and uterine cancers,
 and associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
@@
 the proposed method utilized two mutation signatures for compensating simplified representation powers.
-The signature 10 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
+The signature 8 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
 and seems to be associated with temozolomide, an anticancer drug.
-The signature 13 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
-The signature 19 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
+The signature 9 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
+The signature 10 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
 and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear.
@@
 had common characteristics with those observed in the previous study:
-the signature 9 (T $>$ G at CpTpT) observed in oesophagus cancers, 
-the signature 11 (C $>$ A at [CT]pCpT) observed in low grade gliomas,
-the signature 16 (T $>$ C at NpTpA) observed in liver cancers, 
-the signature 22 (C $>$ A at NpCp[AT]) observed in neuroblastomas,
-the signature 24 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas, 
-and the signature 26 (T $>$ C at GpTpN) in stomach cancers.
+the signature 20 (T $>$ [CG] at CpTpT) observed in oesophagus cancers, 
+the signature 2 (C $>$ A at [CT]pCpT) observed in low grade gliomas,
+the signature 5 (C $>$ A at NpCp[AT]) observed in neuroblastomas,
+the signature 12 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas, 
+and the signature 18 (T $>$ C at GpTpN) in stomach cancers.
 These results indicate that the proposed method can capture 
@@
 had prominent features at the two bases 5' and 3' to the mutated sites.
-For the signature 2, the signatures captured different cancer types are highly consistent (Supplementary Figrue 6),
-and they consistently showed depletion of G base at two bases 5' to the mutated site,
+For the signature 13, the signatures captured different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}),
+and they consistently showed depletion of G base at two bases 5' to the mutated site (Figure \ref{APOBEC_twofive}),
 backing up the result obtained from UCUT data in the previous subsection and the previous study  \cite{pmid23318258}.
-For the signature 6 and signature 7 (Supplementary Figrue 7, 8), 
+For the signature 7 and signature 1, 
 signatures for the two different cancer types consistently showed abundance of T base at two bases 5' to the mutated site,
-and abundance of T bases at two bases 5' and 3' to the mutated site, respectively.
-For the signature 19, depletion of C base at two bases 5' to the mutated site is consistently observed (Supplementary Figrue 9).
+and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
+For the signature 19, depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).
 Therefore, the proposed method can capture the characteristic features at the two bases 5' and 3' to the mutated sites.
@@
   \includegraphics[width=15cm,height=4cm]{example96_APOBEC.eps}
-  \label{UTUC:APOBEC_full5_sig}}
+  \label{APOBEC_nature2013_sig}}
 
@@
   \includegraphics[width=15cm,height=4cm]{example96_POLE.eps}
-  \label{UTUC:AA_full5_sig}}
-
-\caption{The APOBEC signature extracted in the previous study (Signature 2 in \cite{pmid23945592}).
-The barplot is divided by 6 substitution patterns.
-In each division, 16 bars shows joint probability of 16 patterns of immediate 5' and 3' bases combinations 
-(ApNpA, ApNpC, ApNpG, ApNpT, CpNpA, $\cdots$, TpTpT) and substitution patterns.
-The strong intensities of the last four bars for the substitution patterns with original base C 
+  \label{POLE_nature2013_sig}}
+
+\caption{The APOBEC and POLE signatures extracted in the previous study (Signature 2 and 10 in \cite{pmid23945592}, respectively).
+The barplots are divided by 6 substitution patterns.
+In each division, 16 bars show joint probabilities of 16 patterns of immediate 5' and 3' bases combinations 
+(ApNpA, ApNpC, ApNpG, ApNpT, CpNpA, $\cdots$, TpTpT).
+(a) The strong intensities of the last four bars for the three substitution patterns with original bases C 
 indicates that the immediate 5' base is mostly confined to T in this signature.
+Also, the frequency of the immediate 3' bases for the mutation patterns at TpCpN sites are mostly proportional across three major substitution patterns (C $>$ A, C $>$ G and C $>$ T),
+showing that this signature might have more compact representations.
+(b) The strong intensities on TpCpT $>$ TpApT and TpCpG $>$ TpTpG are observed. This is a little complex in the sense that 
+the immediate 3' base and the substitution pattern show strong correlation.
 }
-\label{APOBEC_example}
+\label{nature2013_example}
 \end{figure*}
@@
 
-\begin{center}
+\begin{flushleft}
 \makeatletter
@@
 % \label{label 2}
-\end{center}
+\end{flushleft}
 \end{minipage}
@@
 
+\clearpage
+
+\begin{figure*}[b]
+\centering
+\includegraphics[width=16cm,height=6cm]{simulation_result.eps}
+\caption{The accuracy of the proposed approach for the simulated data 
+when changing the number of samples, mutations, 
+and the amounts of dispersion parameters ($\alpha$ and $\gamma$) for the mutation features and signature distribution parameters.
+}
+\label{sim_facet}
+\end{figure*}
+
 \begin{figure*}[ht]
 \centering
+
+\subfigure[APOBEC signature for the independent model]{%
+  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_APOBEC_ind.eps}
+  \label{UTUC:APOBEC_ind5_sig}}
+\quad
+\subfigure[AA signature for the independent model]{%
+  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_AA_ind.eps}
+  \label{UTUC:AA_ind5_sig}}
 
@@
 
-\subfigure[APOBEC signature for the independent model]{%
-  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_APOBEC_ind.eps}
-  \label{UTUC:APOBEC_ind5_sig}}
-\quad
-\subfigure[AA signature for the independent model]{%
-  \includegraphics[height=5cm,width=7.5cm,clip]{UTUC_AA_ind.eps}
-  \label{UTUC:AA_ind5_sig}}
-
 \subfigure[APOBEC signature stability]{%
@@
 \caption{The mutation signatures for UTUC data, and the results of downsampling experiments. 
-A total of 3072 elements in the full model mutation signatures were shown divided by substitution patterns and strand directions.}
+3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.}
 \label{UTUC}
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}[b]
-\centering
-\includegraphics[width=16cm,height=6cm]{simulation_result.eps}
-\caption{The accuracy of the proposed approach for the simulated data when changing
-the number of samples and mutations, 
-and the amounts of dispersion parameters for the mutation features and signature distribution parameters.
-}
-\label{sim_facet}
 \end{figure*}
@@
 \includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
-\caption{The summary of mutation signatures obtained using the proposed method,
+\caption{The summary of mutation signatures obtained in the reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method,
 where the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
@@
 \includegraphics[width=15cm,height=15cm]{corrplot.eps}
-\caption{The summary of membership of each mutation signature across cancer types.}
+\caption{The summary of membership of each mutation signature across 30 cancer types}
 \label{nature2013_sig_member}
@@
 \begin{figure*}[b]
-\centering
-\includegraphics[width=10cm,height=10cm]{APOBEC_two5prime.eps}
-\caption{The intensities at two 5' base to the mutated site for 14 mutation signatures categorized as signature 2}
-\label{nature2013_sig_apobec}
+
+\subfigure[APOBEC signatures obtained in each cancer type]{%
+  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
+  \label{APOBEC_sig_list}}
+
+\subfigure[Signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=7.5cm,height=7.5cm]{APOBEC_two5prime.eps}
+  \label{APOBEC_twofive}}
+
+\caption{APOBEC signatures obtained in each cancer type and the frequency of bases at two 5' to the mutated site}
+\label{nature2013_apobec}
+\end{figure*}
+
+\begin{figure*}[b]
+
+\subfigure[POLE1 signatures in each cancer type]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
+  \label{POLE1_sig_list}}
+
+\subfigure[POLE2 signature in each cancer type]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
+  \label{POLE2_sig_list}}
+
+\subfigure[UV signature in each cancer type]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
+  \label{UV_sig_list}}
+
+\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
+  \label{LMST_sig_list}}
+
+\caption{Several signatures obtained in multiple cancer types}
+\label{nature2013_sig_others}
 \end{figure*}

### 2015-03-03 figure/script/AlexandrovEtAl_figures_temp.R

@@ -535,9 +535,9 @@ for (i in 1:length(importantSigs)) {
   outputName <- paste("../../manuscript/signatureList_", names(importantSigs)[i], ".eps", sep="");
   
   if (length(types) > 4) {
-    dev.copy2eps(file=outputName, height = ceiling(length(types) / 4) * 2.5, width = 16, pointsize = 18);
+    dev.copy2eps(file=outputName, height = ceiling(length(types) / 4) * 2.4, width = 15);
   } else {
-    dev.copy2eps(file=outputName, height = 2.5, width = 4 * length(types), pointsize = 18);    
+    dev.copy2eps(file=outputName, height = 2.4, width = 3.75 * length(types));    
   }
   
   

### 2015-03-03 figure/script/UTUC_signature.R

@@ -1,31 +1,76 @@
 library(pmsignature);
 
 load("../../UTUC/result/Param_full/3.Rdata");
-Param <- resultForSave[[1]];
+Param_full <- resultForSave[[1]];
 
-if (sum(Param@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
-  visPMSignature(Param, 1);
+if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
+  visPMSignature(Param_full, 1);
   ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
-  visPMSignature(Param, 2);
+  visPMSignature(Param_full, 2);
   ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5); 
 } else {
-  visPMSignature(Param, 2);
+  visPMSignature(Param_full, 2);
   ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
-  visPMSignature(Param, 1);
+  visPMSignature(Param_full, 1);
   ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5);   
 }
   
 load("../../UTUC/result/Param_ind/3.Rdata");
-Param <- resultForSave[[1]];
+Param_ind <- resultForSave[[1]];
 
-if (sum(Param@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param@signatureFeatureDistribution[2,1,4] > 0.75) {
-  visPMSignature(Param, 1, charSize = 1);
[…]

### 2015-03-03 figure/script/UTUC_signature.R
slightly modified the script for UTUC APOBEC

@@ -64,7 +64,7 @@ if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + 
 
 
 
-ggplot(baseTwoFivePrime, aes(x=type, y=prob, fill=base)) +
+ggplot(baseTwoFivePrime, aes(x=model, y=prob, fill=base)) +
   geom_bar(stat="identity") +
   theme(legend.title = element_text(size = rel(1.5)),
         legend.text = element_text(size = rel(1.5)),
@@ -73,4 +73,4 @@ ggplot(baseTwoFivePrime, aes(x=type, y=prob, fill=base)) +
         axis.text.y = element_text(size=rel(1.5)),
         axis.title.y = element_text(size=rel(1.5)))
 
-ggsave("../../supp/UTUC_APOBEC_twoFivePrime.eps", width=2, height=8);
+ggsave("../../supp/UTUC_APOBEC_twoFivePrime.eps", width=4, height=8);

### 2015-03-05 AlexandrovEtAl/script/runall.sh
modified merge analysis script for Alexandrov et al

@@ -28,28 +28,34 @@ _COMMENT_OUT_
 
 # :<<_COMMENT_OUT_
 
+if [ ! -d log ]
+then
+    mkdir log
+fi
+
+
+SCRIPTDIR=`pwd`
 cd ../result
 RESULTDIR=`pwd`
 cd ../data
 DATADIR=`pwd`
 cd ../../common_script
-SCRIPTDIR=`pwd`
 
 while read type
 do
 
     for K in `seq 2 6`
     do
-        echo "qsub -l s_vmem=2G,mem_req=2 perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10"
-        qsub -l s_vmem=2G,mem_req=2 perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10
+        echo "qsub -l s_vmem=2G,mem_req=2 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10"
+        qsub -l s_vmem=2G,mem_req=2 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10
 
         if [ ${type} == Lung-Adeno ]
         then 
-            echo "qsub -l s_vmem=4G,mem_req=4 perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt ${RESULTDIR}/Param_ind5_dir/${type}.${K}.Rdata ${K} TRUE 10"
-            # qsub -l s_vmem=4G,mem_req=4 perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt ${RESULTDIR}/Param_ind5_dir/${type}.${K}.Rdata ${K} TRUE 10
[…]

### 2015-03-05 AlexandrovEtAl/script/subscript/getSample2type.sh
made the file for sample 2 cancer types

@@ -1,0 +1,5 @@
+#! /bin/sh
+
+# command for obtaining the relationship between sample and cancer types
+while read type; do for sample in `zcat ../../result/MPFormat/${type}.mp.txt.gz | cut -f 1 | sort -u`; do echo -e "${sample}\t${type}"; done; done < ../../data/cancer_types.txt > ../../../figure/data/AlexandrovEtAl_sample2type.txt
+

### 2015-03-06 figure/script/AlexandrovEtAl_figures.R
add scale figure

@@ -9,7 +9,7 @@ library(pmsignature);
 # functions
 
 
-visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2) {
+visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale = 0) {
   
   if (is.na(baseCol)) {
     gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
@@ -42,17 +42,20 @@ visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2) {
   
   num2base <- c("A", "C", "G", "T");
   
-  frame();
-  plot.window(xlim=c(-0.25, 1.25 * numBases + 0.25), ylim=c(-0.25, 3.25));
+  # frame();
+  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
+  
+  sizes <- 0.5 * (2 - apply(A, MARGIN = 1, FUN = function(x) {-sum(x * log2(x), na.rm = TRUE)}));
+  sizes <- sizes ** scale
   
   startx <- 0;
   for(l in 1:numBases) {
     
     for(w in 1:4) {
       endx <- startx + A[l,w]
-      polygon(c(startx, endx, endx, startx), c(0, 0, 1, 1), col = baseCol[w], border=F);
-      if (endx - startx > 1 / 4) {
-        text(0.5 * (endx + startx), 0.5, num2base[w], col="white", cex=charSize)
+      polygon(c(startx, endx, endx, startx), c(0, 0, sizes[l], sizes[l]), col = baseCol[w], border=F);
+      if (endx - startx > 1 / 4 & sizes[l] > 0.5) {
+        text(0.5 * (endx + startx), 0.5 * sizes[l], num2base[w], col="white", cex=charSize)
       }
       startx <- endx;
     }
@@ -152,7 +155,7 @@ convertFMatrixToVector <- function(Fmat, fdim) {
 # get the mutation signature of the specified cancer type and index
 getF <- function(type, K, fdim, ind) {
   
-  inputName <- paste("../../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
[…]

### 2015-03-06 figure/script/AlexandrovEtAl_figures.R.bk.R

@@ -42,8 +42,8 @@ visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2) {
   
   num2base <- c("A", "C", "G", "T");
   
-  # frame();
-  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
+  frame();
+  plot.window(xlim=c(-0.25, 1.25 * numBases + 0.25), ylim=c(-0.25, 3.25));
   
   startx <- 0;
   for(l in 1:numBases) {
@@ -152,7 +152,7 @@ convertFMatrixToVector <- function(Fmat, fdim) {
 # get the mutation signature of the specified cancer type and index
 getF <- function(type, K, fdim, ind) {
   
-  inputName <- paste("../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
+  inputName <- paste("../../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
   
   load(inputName);
   F <- Param@signatureFeatureDistribution;
@@ -165,7 +165,7 @@ getF <- function(type, K, fdim, ind) {
 # get the mutation signatures of the specified cancer type
 getSigMat <- function(type, K, fdim) {
   
-  inputName <- paste("../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
+  inputName <- paste("../../../AlexandrovEtAl/result/Param_ind5/", type, ".", as.character(K), ".Rdata", sep="");
   
   load(inputName);
   F <- Param@signatureFeatureDistribution;
@@ -193,7 +193,7 @@ eps <- 0.6
 
 #' create the matrix showing the relationships between cancer type 
 #' and the specified numer of mutation signatures
-type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+type2sigNum <- read.table("../../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
 
 #' reading and converting the mutation signatures for each cancer type
 for (i in 1:nrow(type2sigNum)) {
[…]

### 2015-03-06 supp/supp.tex
modified figure

@@
 \includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
-\caption{}
+\caption{An example of the relationship between true and estimated mutation signatures in the simulation.
+In this example, the numbers of cancer genomes and mutations for each cancer genome are 25 and 100, respectively,
+and the parameters $\alpha$ and $\gamma$ are both set to 1. 
+From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations}
 \label{example_signature.pd}
@@
 \includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
-\caption{}
+\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in the trial of the Supplementary Figure 1.}
 \label{example_summary.pd}
@@
 
-\subfigure[Estimated mutation signatures ($K = 2$)]{%
+\subfigure[$K = 2$]{%
   \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K2.eps}
@@
 
-\subfigure[Estimated mutation signatures ($K = 3$)]{%
+\subfigure[$K = 3$]{%
   \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K3.eps}
@@
 
-\subfigure[Estimated mutation signatures ($K = 4$)]{%
+\subfigure[$K = 4$]{%
   \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K4.eps}
@@
 \caption{The result of estimated mutation signatures and membership parameters for UTUC data 
-when changing the number of mutation signatures ($K$).}
+when changing the number of mutation signatures $K$.}
 \label{UTUC_multiK}
@@
 \includegraphics[width=15cm,height=5cm]{UTUC_stat.eps}
-\caption{}
-\label{example_summary.pd}
+\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in UTUC data.}
+\label{UTUC_stat}
 \end{figure}
@@
 \includegraphics[width=10cm,height=10cm]{UTUC_AA_AAlike_cor.eps}
-\caption{}
-\label{example_signature.pd}
+\caption{The relationships between the two membership parameters, AA (the first signature in the Supplementary Figure 3) and AA\_like (the second signature in the Supplementary Figure 3) signatures.}
+\label{UTUT_AA_AAlike}
 \end{figure}
@@
 
-% \begin{figure}
-% \centering
-% \includegraphics[width=15cm,height=15cm]{sig2_group.eps}
-% \caption{Mutation signatures obtained for several cancer types categorized as Signature 2}
-% \label{example_summary.pd}
-% \end{figure}
-
-% \clearpage
-
-% \begin{figure}
-% \centering
-% \includegraphics[width=15cm,height=3.5cm]{sig6_group.eps}
-% \caption{Mutation signatures obtained for several cancer types categorized as Signature 6}
-% \label{example_summary.pd}
-% \end{figure}
-
-% \clearpage
-
-% \begin{figure}
-% \centering
-% \includegraphics[width=15cm,height=3.5cm]{sig7_group.eps}
-% \caption{Mutation signatures obtained for several cancer types categorized as Signature 7}
-% \label{example_summary.pd}
-% \end{figure}
-
-% \clearpage
-
-% \begin{figure}
-% \centering
-% \includegraphics[width=15cm,height=3.5cm]{sig19_group.eps}
-% \caption{Mutation signatures obtained for several cancer types categorized as Signature 19}
-% \label{example_summary.pd}
-% \end{figure}
+\begin{figure}
+\centering
+\includegraphics[width=4cm,height=8cm]{UTUC_APOBEC_twoFivePrime.eps}
+\caption{The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
+\label{UTUC_APOBEC_twoFivePrime}
+\end{figure}
 

### 2015-03-06 figure/script/AlexandrovEtAl_figures.R

@@ -507,7 +507,8 @@ for (i in 1:length(importantSigs)) {
     plot.window(xlim=c(-0.3, 6.3), ylim=c(-0.3, 3.8));
     polygon(c(-0.25, 6.25, 6.25, -0.25), c(-0.25, -0.25, 3.75, 3.75), col = "white", border = FALSE);
     
-    visPMS_ind(F, numBases = 5, trDir = FALSE, charSize = 1);
+    # the setting of charSize is temporary... why the size changes by the number of rows.....
+    visPMS_ind(F, numBases = 5, trDir = FALSE, charSize = 0.6 + 0.1333 * ceiling(length(types) / 4));
     mtext(types[j],
           outer = FALSE,      # 作図領域の外の余白に書く
           side = 3,          # 上の余白に書く
@@ -579,7 +580,7 @@ ggplot(twoFivePrimeSig, aes(x=type, y=intensity, fill=base)) +
   theme(text = element_text(size=20)) +
   theme(axis.text.x= element_text(angle=60,hjust=1));
 
-ggsave("../../manuscript/APOBEC_two5prime.eps", width=6, height=6, units="in");
+ggsave("../../manuscript/APOBEC_two5prime.eps", width=8, height=5, units="in");
 
 
 
@@ -634,8 +635,9 @@ par(.pardefault);
 
 
 
-##########
+####################
 # change the scale
+# This might be not necessary for the paper
 ##########
 
 

### 2015-03-06 AlexandrovEtAl/script/runall.sh
performing transcription  direction analysis on Alexandrov data

@@ -46,16 +46,16 @@ do
 
     for K in `seq 2 6`
     do
-        echo "qsub -l s_vmem=2G,mem_req=2 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10"
-        qsub -l s_vmem=2G,mem_req=2 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10
+        # echo "qsub -l s_vmem=2G,mem_req=2 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10"
+        # qsub -l s_vmem=2G,mem_req=2 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt.gz ${RESULTDIR}/Param_ind5/${type}.${K}.Rdata ${K} FALSE 10
 
         if [ ${type} == Lung-Adeno ]
         then 
-            echo "qsub -l s_vmem=4G,mem_req=4 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt ${RESULTDIR}/Param_ind5_dir/${type}.${K}.Rdata ${K} TRUE 10"
-            # qsub -l s_vmem=4G,mem_req=4 -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log perform_pmsignature.sh ${RESULTDIR}/MPFormat/${type}.mp.txt ${RESULTDIR}/Param_ind5_dir/${type}.${K}.Rdata ${K} TRUE 10
[…]

### 2015-03-06 manuscript/plos_template.tex

@@
 and great chances for not only identification of novel driver mutations but also investigation of sample-by-sample mutation patterns in an unbiased way. 
-A study using 21 breast cancer samples identified the association between C $>$ [GT] mutations at TpC sites, 
+A study using 21 breast cancer samples identified the association between C $>$ [AGT] mutations at TpC sites, 
 which is later proved to be caused by APOBEC proteins, and the novel phenomenon called {\it kataegis} \cite{pmid22608084}. 
@@
 As Figure \ref{sim_facet} shows, we could estimate the mutation signatures very accurately overall.
-(see Supplementary Figure 1 for an example where the number of cancer genomes and mutations for each cancer genome is 25 and 100, respectively).
+(see Supplementary Figure 1 for an example).
 In most cases, the log-likelihood stopped increasing at five mutation signatures, 
@@
 where they found a novel mutation signature: 
-T $>$ A substitutions at CpTpG sites with a strong strand specificity caused by aristolochic acids (AAs).
-
-First, we performed the proposed method,
-considering substitution patterns, up to $\pm 2$ sites and strand directions as mutation features.
-For the full representations, we assign one integer for each combination of these features, and thus $L = 1, \bm{M} = (3072)$.
-For the independent representation, $L = 6, \bm{M} = (6, 4, 4, 4, 4, 2)$.
-
-First, performed on various number of mutation signatures $K$ (including a background signature), we observed how the detected mutation signatures changes (See Supplementary Figure 3).
+T $>$ A substitutions at CpTpG sites with a strong strand specificity caused by aristolochic acids (AA).
+
+We performed the proposed method considering substitution patterns, up to $\pm 2$ sites and strand directions as mutation features.
+For the full model, we assign one integer for each combination of these features, and thus $L = 1, \bm{M} = (3072)$.
+For the independent model, $L = 6, \bm{M} = (6, 4, 4, 4, 4, 2)$.
+
+First, performing on various number of mutation signatures $K$ (including a background signature) on the independent model, 
+we observed how the detected mutation signatures change (See Supplementary Figure 3).
 For $K = 2$, a mutation signature which seem to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription direction) was observed (Supplementary Figure 3 (a)).
@@
 that is somewhat similar to the AA signature was observed (Supplementary Figure 3 (c)). 
-When we check the correlation of estimated membership parameter for each cancer genome, 
-strong correlation between the AA signature (CpTpG $>$ CpApG) and that additional AA-like signature can be observed ($R = 0.77$, see Supplementary Figure 5). 
+When we checked the correlation of estimated membership parameter for each cancer genome, 
+strong correlation between the AA signature (CpTpG $>$ CpApG) and that additional AA-like signature could be observed ($R = 0.77$, see Supplementary Figure 5). 
 This additional signature may be just making up for the residual of the AA signature 
@@
 The values of likelihood continued increasing as the $K$ increase, while the bootstrap-errors started to increase at $K = 5$ (See Supplementary Figure 4).
-The strong correlation among estimated membership parameters started to be shown at $K = 4$
-Considering all these factors together, $K = 3$ seems to be one reasonable choice in terms of the interpretability,
-and we adopt $K = 3$ in the following.
-
-For the independent representation model, we could observed the depletion of G base at the $-2$ sites,
+The strong correlation among estimated membership parameters started to be shown at $K = 4$.
+Considering all these factors together, $K = 3$ seems to be a reasonable choice in terms of the interpretability,
+and we adopted $K = 3$ in the following.
+
+For the independent model, we could observed the depletion of G base at the $-2$ site,
 which is consistent to the previous study \cite{pmid23318258} and the result in the next subsection.
-On the other hand, for the full representation model, this tendency was rather mild (Figure \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}).
-Inferred AA mutation signature had no clear characteristics at two bases 5' and 3' to the mutated site 
-compared to the APOBEC mutation signature.
+On the other hand, for the full representation model, this tendency was rather mild (Figure \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}, Supplementary Figure 6).
+Inferred AA mutation signature had no clear characteristics at two bases 5' and 3' to the mutated site compared to the APOBEC mutation signature.
 
@@
 and compared obtained signatures with the gold standard.
-For measuring the deviations of the mutation signature, the cosine similarity is used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
-so that the comparison between the full representation and the independent representation become possible.
+For measuring the deviations of the mutation signature, the cosine similarity was used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
+so that the comparison between the full model and the independent model become possible.
 Trials for downsampling experiment for each ratio and model were repeated for 100 times.
 
-As the Figure \ref{UTUC:APOBEC_downsampling} and \ref{UTUC:AA_downsampling} shows, the independent representation could recover the original signatures even when about 90 \% of the data was removed.
+As the Figure \ref{UTUC:APOBEC_downsampling} and \ref{UTUC:AA_downsampling} shows, 
+the independent representation could recover the original signatures even when, e.g, about 90 \% of the data was removed.
 In this experiment, the AA signature was more robust than the APOBEC signature.
-This is because the number of T $>$ A substitutions at GpTpC sites are more frequent in this dataset.
-These results indicate that while both the estimation under the full representation and the independent representation could efficiently extract known mutation signatures,
-and the estimation under the independent representation is more robust than the full representation.
+We believe that this is because the number of T $>$ A substitutions at GpTpC sites are far more frequent in this dataset.
+These results indicate that the independent model can give more robust estimates than the full model.
 
@@
 The signature 6 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
-The signature 13 (T $>$ [GT] at TpCpN sites) was observed in 11 cancer types, and related to the activity of APOBEC family.
+The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 11 cancer types, and related to the activity of APOBEC family.
 The signature 7 and the signature 1 (C $>$ T at TpCpG and C $>$ A at TpCpT, respectively) were observed in colorectal and uterine cancers,
@@
 
-Furthermore, a number of prominent mutation signatures observed in one cancer type 
+In addition to the above signatues, a number of prominent mutation signatures observed in one cancer type 
 had common characteristics with those observed in the previous study:
@@
 and the signature 18 (T $>$ C at GpTpN) in stomach cancers.
-These results indicate that the proposed method can capture 
-many of prominent mutation signatures detected in the previous study.
 
@@
 had prominent features at the two bases 5' and 3' to the mutated sites.
-For the signature 13, the signatures captured different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}),
+For the signature 13, the signatures captured in 11 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}),
 and they consistently showed depletion of G base at two bases 5' to the mutated site (Figure \ref{APOBEC_twofive}),
 backing up the result obtained from UCUT data in the previous subsection and the previous study  \cite{pmid23318258}.
-For the signature 7 and signature 1, 
+For the Pol $\epsilon$ signatures (signature 7 and signature 1), 
 signatures for the two different cancer types consistently showed abundance of T base at two bases 5' to the mutated site,
@@
 For the signature 19, depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).
-Therefore, the proposed method can capture the characteristic features at the two bases 5' and 3' to the mutated sites.
+Although, as far as we know, there was no literature showing the characteristics of mutation signatures at $\pm 2$ sites,
+Considering from the robustness shown in the APOBEC signature, many of  the characteristics inferred here might be true.
+
+There results indicate that the proposed method can capture not only many of the prominent mutation signatures observed in the previous study,
+but also novel characteristics for $\pm 2$ sites.
 
@@
 The barplots are divided by 6 substitution patterns.
-In each division, 16 bars show joint probabilities of 16 patterns of immediate 5' and 3' bases combinations 
+In each division, 16 bars show joint probabilities of 16 combinations of the immediate 5' and 3' bases  
 (ApNpA, ApNpC, ApNpG, ApNpT, CpNpA, $\cdots$, TpTpT).
@@
 indicates that the immediate 5' base is mostly confined to T in this signature.
-Also, the frequency of the immediate 3' bases for the mutation patterns at TpCpN sites are mostly proportional across three major substitution patterns (C $>$ A, C $>$ G and C $>$ T),
-showing that this signature might have more compact representations.
+Also, the frequency of the immediate 3' bases for the mutation patterns at TpCpN sites are mostly proportional across three major substitution patterns (C $>$ A, C $>$ G and C $>$ T).
 (b) The strong intensities on TpCpT $>$ TpApT and TpCpG $>$ TpTpG are observed. This is a little complex in the sense that 
-the immediate 3' base and the substitution pattern show strong correlation.
+the immediate 3' base and the substitution pattern are not independent.
 }
@@
 Here, mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
-are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).}
+are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).
+In the bottom, the size of each box represents the frequencies of bases (A, C, G and T) at the flanking sites.
+In the top, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
+In the upper right, the height of the $+$ (non-coding strand) and $-$ (coding strand) boxes represents the frequencies of  stranded direction of mutations.
+}
 \label{mutSig_example}
@@
 %
-\caption{The mutation signatures for UTUC data, and the results of downsampling experiments. 
+\caption{The mutation signatures for the UTUC data, and the results of downsampling experiments. 
 3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.}
@@
 \includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
-\caption{The summary of mutation signatures obtained in the reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method,
+\caption{The summary of mutation signatures obtained in a reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method,
 where the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
@@
 \includegraphics[width=15cm,height=15cm]{corrplot.eps}
-\caption{The summary of membership of each mutation signature across 30 cancer types}
+\caption{The summary of membership of each mutation signature across 30 cancer types obtained using the proposed method.}
 \label{nature2013_sig_member}
@@
 
-\subfigure[Signature intensities at two 5' to the mutated site]{%
-  \includegraphics[width=7.5cm,height=7.5cm]{APOBEC_two5prime.eps}
-  \label{APOBEC_twofive}}
-
-\caption{APOBEC signatures obtained in each cancer type and the frequency of bases at two 5' to the mutated site}
-\label{nature2013_apobec}
-\end{figure*}
-
-\begin{figure*}[b]
-
 \subfigure[POLE1 signatures in each cancer type]{%
@@
   \label{POLE1_sig_list}}
-
+\quad
 \subfigure[POLE2 signature in each cancer type]{%
@@
   \label{UV_sig_list}}
-
+\quad
 \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
@@
 
-\caption{Several signatures obtained in multiple cancer types}
-\label{nature2013_sig_others}
+\subfigure[Signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
+  \label{APOBEC_twofive}}
+
+\caption{(a) APOBEC signatures obtained in 11 cancer types, 
+(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
+(f) the frequencies of bases at two 5' to the mutated site. 
+}
+\label{nature2013_apobec}
 \end{figure*}
+
+% \begin{figure*}[b]
+
+% \subfigure[POLE1 signatures in each cancer type]{%
+%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
+%   \label{POLE1_sig_list}}
+
+% \subfigure[POLE2 signature in each cancer type]{%
+%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
+%   \label{POLE2_sig_list}}
+
+% \subfigure[UV signature in each cancer type]{%
+%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
+%   \label{UV_sig_list}}
+
+% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
+%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
+%   \label{LMST_sig_list}}
+
+% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
+% \label{nature2013_sig_others}
+% \end{figure*}
 

### 2015-03-09 figure/script/AlexandrovEtAl_figures.R
add error bar plot

@@ -1,691 +1,0 @@
-
-library(Matrix);
-library(corrplot);
-library(pmsignature);
-
-.pardefault <- par(no.readonly = TRUE);
-
-##########
-# functions
-
-
-visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale = 0) {
-  
-  if (is.na(baseCol)) {
-    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
-    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
-  }
-  
-  centerBase <- (1 + numBases) / 2;
-  
-  v1 <- vF[1,1:6];
-  V2 <- vF[2:(numBases),1:4];
-  A <- matrix(0, numBases, 4);
-  B <- matrix(0, 4, 4);
-  
-  if (trDir == TRUE) {
-    v3 <- vF[(numBases + 1),1:2];
-  }
-  
-  for (l in 1:numBases) {
-    if (l < centerBase) {
-      A[l, ] <- V2[l, ];
-    } else if (l > centerBase) {
-      A[l, ] <- V2[l - 1, ];
-    }
-  }
-  A[centerBase,2] <- sum(v1[1:3]);
-  A[centerBase,4] <- sum(v1[4:6]);
-  
-  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
-  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
-  
-  num2base <- c("A", "C", "G", "T");
-  
-  # frame();
-  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
-  
-  sizes <- 0.5 * (2 - apply(A, MARGIN = 1, FUN = function(x) {-sum(x * log2(x), na.rm = TRUE)}));
-  sizes <- sizes ** scale
-  
-  startx <- 0;
-  for(l in 1:numBases) {
-    
-    for(w in 1:4) {
-      endx <- startx + A[l,w]
-      polygon(c(startx, endx, endx, startx), c(0, 0, sizes[l], sizes[l]), col = baseCol[w], border=F);
-      if (endx - startx > 1 / 4 & sizes[l] > 0.5) {
[…]

### 2015-03-09 figure/script/AlexandrovEtAl_figures_dir.R

@@ -1,0 +1,366 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("AlexandrovEtAl_function.R")
+
+
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4, 2);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
+  # indList <- which(closeSig[i,] > 0);
[…]

### 2015-03-09 figure/script/AlexandrovEtAl_function.R

@@ -1,0 +1,292 @@
+
+##########
+# functions for drawing figures for the Alexandrov et al analysis data
+
+
+visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale = 0) {
+  
+  if (is.na(baseCol)) {
+    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
+    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
+  }
+  
+  centerBase <- (1 + numBases) / 2;
+  
+  v1 <- vF[1,1:6];
+  V2 <- vF[2:(numBases),1:4];
+  A <- matrix(0, numBases, 4);
+  B <- matrix(0, 4, 4);
+  
+  if (trDir == TRUE) {
+    v3 <- vF[(numBases + 1),1:2];
+  }
+  
+  for (l in 1:numBases) {
+    if (l < centerBase) {
+      A[l, ] <- V2[l, ];
+    } else if (l > centerBase) {
+      A[l, ] <- V2[l - 1, ];
+    }
+  }
+  A[centerBase,2] <- sum(v1[1:3]);
+  A[centerBase,4] <- sum(v1[4:6]);
+  
+  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
+  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
+  
+  num2base <- c("A", "C", "G", "T");
+  
+  # frame();
+  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
+  
+  sizes <- 0.5 * (2 - apply(A, MARGIN = 1, FUN = function(x) {-sum(x * log2(x), na.rm = TRUE)}));
+  sizes <- sizes ** scale
+  
+  startx <- 0;
+  for(l in 1:numBases) {
+    
+    for(w in 1:4) {
+      endx <- startx + A[l,w]
+      polygon(c(startx, endx, endx, startx), c(0, 0, sizes[l], sizes[l]), col = baseCol[w], border=F);
+      if (endx - startx > 1 / 4 & sizes[l] > 0.5) {
+        text(0.5 * (endx + startx), 0.5 * sizes[l], num2base[w], col="white", cex=charSize)
[…]

### 2015-03-11 manuscript/plos_template.tex
various change

@@
 In total, 27 mutation signatures were extracted.
-The signature 6 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
-The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 11 cancer types, and related to the activity of APOBEC family.
-The signature 7 and the signature 1 (C $>$ T at TpCpG and C $>$ A at TpCpT, respectively) were observed in colorectal and uterine cancers,
-and associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
-Interestingly, for representing the signature for the Pol $\epsilon$ dysfunction, which was represented by one signature in the previous study,
-the proposed method utilized two mutation signatures for compensating simplified representation powers.
-The signature 8 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
-and seems to be associated with temozolomide, an anticancer drug.
+By comparing the composition of nucleotides and memberships of signatures with the result of the previous study,
+we could associate many of the detected signatures with known mutational processes.
+
+The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and related to the activity of APOBEC family.
+The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
+and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at $-2$ positions (Figure \ref{APOBEC_twofive}).
+This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
+and implies the robustness of the proposed approach.
+In terms of transcribed strand-specificity, the tendency is not consistent, implying that there is no significant strand-specificity in APOBEC signatures.
+
+The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) were observed in colorectal and uterine cancers,
+are probably associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
+The proposed method used two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
+while was represented by one signature in the previous study.
+Interestingly, signature 1 showed transcription biases whereas signature 8 did not.
+This phenomena are consistent for both colorectal and uterus cancers (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
+C $>$A at TpCpT mutations were shown to be enriched in leading strands of replication forks.
+whereas, there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
+Although replication strand biases is different from the transcription strand biases, these two biases may have strong corelation,
+considering the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
+Therefore, the transcription strand bias observed in the signature 1 may actually come from replication strand biases.
+
+Furthermore, both the signature 1 and 8 signatures showed strong abundance of T base at $-2$ positions,
+and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
+In fact, nonsense mutations in TP53 at R231X, which are shown to be enriched in cancers having Pol $\epsilon$,
+and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
+In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
+In fact, this mutation is C $>$ T at TpTpCpGpA, the most likely patten with the signature 8.
+This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.
+
+The signature 18 (T $>$ C at ApTp[AG]), observed in liver cancer, shows very strong strains specificity,
+were shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, though the source of this signature is still not clear.
+
+The signature 3 and 5 (C $>$ A at NpCpN), which were observed in head-and-neck cancers, three types of lung cancers,
+are probably associated with tobacco smoking.
+Furthermore, the fact that the estimated signature for each cancer shows higher prevalence on coding strands is consistent with the previous study.
+
+The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
+The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
+and seems to be associated with a chemotherapy drug, temozolomide.
+
 The signature 9 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
@@
 
-In addition to the above signatues, a number of prominent mutation signatures observed in one cancer type 
+The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, can be related to the one detected in the same cancer type in the previous study, but slightly different. 
+This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
+Actually, the mutation related to this signature is mostly coming from one cancer with a very high mutation rate.
+When we re-analyze the data removing this sample, the signature 2 disappeared.
+We suspect that the low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
+but cannot be decomposed due to the lack of samples having these signatures.
+Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.
+
+The signature 3 (C $>$ A at CpCpG), observed in several cancers,
+Actually, this signature correspond to the one that could not be validated in the previous study,
+and the pattern
+
+In addition to the above signatures, a number of prominent mutation signatures observed in one cancer type 
 had common characteristics with those observed in the previous study:
-the signature 20 (T $>$ [CG] at CpTpT) observed in oesophagus cancers, 
+the signature 21 (T $>$ [CG] at CpTpT) observed in oesophagus cancers, 
 the signature 2 (C $>$ A at [CT]pCpT) observed in low grade gliomas,
-the signature 5 (C $>$ A at NpCp[AT]) observed in neuroblastomas,
+the signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas,
 the signature 12 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas, 
-and the signature 18 (T $>$ C at GpTpN) in stomach cancers.
+and the signature 19 (T $>$ C at GpTpN) in stomach cancers.
+
+Although, our method detected most of the
 
@@
 
+\subfigure[Smoking signature  in each cancer type]{%
+  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
+  \label{LUNG_sig_list}}
+
 \subfigure[POLE1 signatures in each cancer type]{%
@@
 
-\subfigure[Signature intensities at two 5' to the mutated site]{%
-  \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
-  \label{APOBEC_twofive}}
+% \subfigure[Signature intensities at two 5' to the mutated site]{%
+% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
+ % \label{APOBEC_twofive}}
 

### 2015-03-11 figure/script/AlexandrovEtAl_figures.R

@@ -1,0 +1,505 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("AlexandrovEtAl_function.R")
+
+
+##########
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], fdim));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
+  # indList <- which(closeSig[i,] > 0);
[…]

### 2015-03-13 figure/script/AlexandrovEtAl_Glioma-Low-Grade_removed.R
script for generating figures about low grade glioma

@@ -1,0 +1,49 @@
+library(pmsignature)
+
+G <- readMPFile("../../AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
+
+##########
+# membership plot
+load("../../AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
+visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
+ggsave("../../supp/Glioma-Low-Grade.membership.eps", height = 4, width = 15, pointsize = 9);
+##########
+
+
+##########
+# remove the sample with maximum number of mutation and check the signatures
+mutNum <- getMutNum(G);
+mutMaxInd <- which(mutNum[,2] == max(mutNum[,2]));
+G@countData[3, G@countData[2,] == mutMaxInd] <- 0;
+BG <- readBGFile(G);
+
+Param2 <- getPMSignature(G, K = 2, BG = BG);
+Param3 <- getPMSignature(G, K = 3, BG = BG);
+Param4 <- getPMSignature(G, K = 4, BG = BG);
+Param5 <- getPMSignature(G, K = 5, BG = BG);
+Params <- c(Param2, Param3, Param4, Param5);
+
+alignOrder <- alignmentSignature(Params);
+
+
+tempMar <- par("mar");
+par(mar = 0.4 * tempMar);
+
+for (l1 in 1:length(Params)) {
+  
+  mat <- matrix(c(1:l1, rep(0, 4 - l1)), 1, 4, byrow = TRUE);
+  layout(mat);
+   
+  for (l2 in 1:l1) {
+    aInd <- which(alignOrder[[l1]] == l2);
+    visPMSignature(Params[[l1]], aInd);
+  }
+  
+  outputName <- paste("../../supp/Glioma-Low-Grade.removed_signature_K", (l1 + 1), ".eps", sep="");
+  dev.copy2eps(file=outputName, height = 3, width = 15, pointsize = 9);
+}
+
+
+
+
+

### 2015-03-13 figure/script/AlexandrovEtAl_Glioma-Low-Grade_removed.R
figure for oxidation artifact signatures

@@ -5,8 +5,11 @@ G <- readMPFile("../../AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz
 ##########
 # membership plot
 load("../../AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
-visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
+visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
 ggsave("../../supp/Glioma-Low-Grade.membership.eps", height = 4, width = 15, pointsize = 9);
+
+visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
+ggsave("../../supp/Glioma-Low-Grade.membership_log.eps", height = 4, width = 15, pointsize = 9);
 ##########
 
 

### 2015-03-13 figure/script/AlexandrovEtAl_oxidation.R

@@ -1,0 +1,56 @@
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+##########
+# function for checking the index of the oxidative signature
+checkOxidationInd <- function(Param) {
+  
+  maxK <- slot(Param, "signatureNum") - as.integer(slot(Param, "isBackGround"));
+  
+  sigInd <- 0;
+  for (k in 1:maxK) {
+    
+    F <- Param@signatureFeatureDistribution;
+    if (F[k,1,1] > 0.6 & F[k,2,2] > 0.5 & F[k,3,2] > 0.5) {
+      sigInd <- k;
+    }
+    
+  }
+  
+  return(sigInd);
+  
+}
+##########
+
+
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+oxidation_types <- c("Kidney-Clear-Cell", "Lung-Adeno", "Melanoma");
+
+for (i in 1:length(oxidation_types)) {
+  
+  K <- type2sigNum[type2sigNum[,1] == oxidation_types[i], 2];
+  inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
+  inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
+        
+  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
+  load(inputRdata);
+  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
+  ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 15, pointsize = 9);
+
+  oxInd <- checkOxidationInd(resultForSave[[1]])
+  
+  tempMar <- par("mar");
+  par(mar = 0.4 * tempMar);
+  
+  visPMSignature(resultForSave[[1]], oxInd, charSize = 0.7);
+  dev.copy2eps(file=paste("../../supp/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 2.4, width = 3.75);    
+
+  par(.pardefault);
+  
+}
+

### 2015-03-13 figure/script/oxidation.R
oxidation

@@ -1,0 +1,44 @@
+library(pmsignature);
+
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+oxidation_types <- c("Kidney-Clear-Cell", "Lung-Adeno", "Melanoma");
+
+for (i in 1:length(oxidation_types)) {
+  
+  K <- type2sigNum[type2sigNum[,1] == oxidation_types[i], 2];
+  inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
+  inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
+        
+  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
+  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
+  ggsave("../../supp/", oxidation_types[i], "_oxidation_membership.eps", height = 4, width = 15, pointsize = 9);
+
+  oxInd <- checkOxidationInd(resultForSave[[1]])
+  visPMSignature(resultForSave[[1]], oxInd);
+  dev.copy2eps(file="../../supp/", oxidation_types[i], "_oxidation_signature.eps", height = 2.4, width = 3.75);    
+
+}
+
+
+checkOxidationInd <- function(Param) {
+  
+  maxK <- slot(Param, "signatureNum") - as.integer(slot(Param, "isBackGround"));
+  
+  sigInd <- 0;
+  for (k in 1:maxK) {
+    
+    F <- Param@signatureFeatureDistribution;
+    if (F[k,1,1] > 0.6 & F[k,2,2] > 0.5 & F[k,3,2] > 0.5) {
+      sigInd <- k;
+    }
+    
+  }
+  
+  return(sigInd);
+  
+}
+

### 2015-03-13 manuscript/plos_template.tex
work on manuscript

@@
 
-\subsection*{Application to pan-cancer somatic mutation data}
-
-Finally, we have applied the proposed method for the pan-cancer somatic mutation data \cite{pmid23945592}.
-For each cancer type, the proposed method was applied separately.
+\subsection*{Application to somatic mutation data of 30 cancer types}
+
+Finally, we have applied the proposed method for the somatic mutation data of 30 cancer types \cite{pmid23945592}.
+For each cancer type, the proposed method was applied separately to see the robustness of the estimated signatures among different cancer types.
 By changing the number of mutation signatures and calculating values of the log-likelihood and bootstrap errors,
@@
 
-The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and related to the activity of APOBEC family.
+The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
 The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
@@
 and implies the robustness of the proposed approach.
-In terms of transcribed strand-specificity, the tendency is not consistent, implying that there is no significant strand-specificity in APOBEC signatures.
+The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
 
 The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) were observed in colorectal and uterine cancers,
-are probably associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
+are associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
 The proposed method used two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
@@
 whereas, there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
-Although replication strand biases is different from the transcription strand biases, these two biases may have strong corelation,
+Although replication strand biases is different from the transcription strand biases, these two biases may have strong correlation,
 considering the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
@@
 
-Furthermore, both the signature 1 and 8 signatures showed strong abundance of T base at $-2$ positions,
+Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at $-2$ positions,
 and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
@@
 
-The signature 18 (T $>$ C at ApTp[AG]), observed in liver cancer, shows very strong strains specificity,
-were shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, though the source of this signature is still not clear.
-
-The signature 3 and 5 (C $>$ A at NpCpN), which were observed in head-and-neck cancers, three types of lung cancers,
-are probably associated with tobacco smoking.
-Furthermore, the fact that the estimated signature for each cancer shows higher prevalence on coding strands is consistent with the previous study.
+The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, can be related to the one detected in the same cancer type in the previous study, but slightly different. 
+This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
+Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate, 
+and the signature 2 disappeared when we re-analyze the data removing this sample.
+We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
+but cannot be decomposed due to the lack of the number of samples having these signatures.
+Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.
+
+The signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
+The estimated signature in each cancer type shows higher mutation prevalence on the template strand, which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
+
+The signature 4 (C $>$ A at CpCpG) observed in several cancers shows remarkably high ratio of base C at the $-2$ position.
+This seems to correspond to the ``signatures R2''  (detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas) 
+that could not be validated in the previous study (see supplementary figures of the previous study \cite{pmid23318258}),
+meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
+Actually, the pattern of this signature matches to that of artifact by oxidation of DNA during acoustic shearing \cite{pmid23303777}.
+Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.
+
+The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer shows very strong strains specificity, implying the evidence of transcription-coupled nucleotide excision repair.
+This signature has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, though the source of this signature is still not clear.
 
@@
 The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
-and seems to be associated with a chemotherapy drug, temozolomide.
-
-The signature 9 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
-The signature 10 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
+and are probably associated with a chemotherapy drug, temozolomide.
+The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
+
+The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
 and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear.
-
-The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, can be related to the one detected in the same cancer type in the previous study, but slightly different. 
-This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
-Actually, the mutation related to this signature is mostly coming from one cancer with a very high mutation rate.
-When we re-analyze the data removing this sample, the signature 2 disappeared.
-We suspect that the low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
-but cannot be decomposed due to the lack of samples having these signatures.
-Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.
-
-The signature 3 (C $>$ A at CpCpG), observed in several cancers,
-Actually, this signature correspond to the one that could not be validated in the previous study,
-and the pattern
+When we investigate $\pm 2$ positions, the depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).
 
@@
 and the signature 19 (T $>$ C at GpTpN) in stomach cancers.
-
-Although, our method detected most of the
-
-Many of the signatures captured in this study 
-had prominent features at the two bases 5' and 3' to the mutated sites.
-For the signature 13, the signatures captured in 11 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}),
-and they consistently showed depletion of G base at two bases 5' to the mutated site (Figure \ref{APOBEC_twofive}),
-backing up the result obtained from UCUT data in the previous subsection and the previous study  \cite{pmid23318258}.
-For the Pol $\epsilon$ signatures (signature 7 and signature 1), 
-signatures for the two different cancer types consistently showed abundance of T base at two bases 5' to the mutated site,
-and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
-For the signature 19, depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).
-Although, as far as we know, there was no literature showing the characteristics of mutation signatures at $\pm 2$ sites,
-Considering from the robustness shown in the APOBEC signature, many of  the characteristics inferred here might be true.
 
@@
 In the top, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
-In the upper right, the height of the $+$ (non-coding strand) and $-$ (coding strand) boxes represents the frequencies of  stranded direction of mutations.
+In the upper right, the height of the $+$ box represents the frequencies of mutations in the coding strand (or the plus strand, the sense strand and the untranscribed strand)
+whose nucleotide sequences directly corresponds to mRNA, 
+whereas the height of $-$ box represents those in the template strand (or the minus strand, the antisense strand, the transcribed strand and the noncoding strand)
+whose sequences are copied during the synthesis of mRNA.
 }

### 2015-03-14 manuscript/plos_template.tex
modify several staffs for supplementary material figures

@@
 
-The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
-The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
-and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at $-2$ positions (Figure \ref{APOBEC_twofive}).
-This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
-and implies the robustness of the proposed approach.
-The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
-
 The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) were observed in colorectal and uterine cancers,
@@
 
-Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at $-2$ positions,
+Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at the $-2$ position,
 and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
@@
 
-The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, can be related to the one detected in the same cancer type in the previous study, but slightly different. 
+The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, 
+can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 7 (a)). 
 This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
-Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate, 
-and the signature 2 disappeared when we re-analyze the data removing this sample.
+Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate (Supplementary Figure 7 (b,c)), 
+and the signature 2 disappeared when we re-analyze the data removing this sample (Supplementary Figure 8).
 We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
@@
 
+The signature 4 (C $>$ A at CpCpG) observed in kidney clear cell carcinomas, lung adenocarcinomas and melanomas 
+shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 9 (a,b,c)).
+However, this signature is confide to only a few samples for every cancer type (see Supplementary Figure 9 (d, e, f)).
+This seems to correspond to the ``signatures R2''  detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas
+(see supplementary figures of the previous study \cite{pmid23318258})
+that could not be validated in the previous study,
+meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
+Actually, the pattern of this signature largely matches to that of the putative artifact by oxidation of DNA during acoustic shearing suggested in \cite{pmid23303777},
+we guess that this signature is the same as what is observed in that study.
+Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.
+
+The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
+The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
+and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}).
+This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
+and implies the robustness of the proposed approach.
+The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
+The Signature 15 and 16 may also be delated to the APOBEC signatures although the estimated forms are slightly different,
+and are not merged to the cluster of signature 13 in the current criteria.
+
 The signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
 The estimated signature in each cancer type shows higher mutation prevalence on the template strand, which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
-
-The signature 4 (C $>$ A at CpCpG) observed in several cancers shows remarkably high ratio of base C at the $-2$ position.
-This seems to correspond to the ``signatures R2''  (detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas) 
-that could not be validated in the previous study (see supplementary figures of the previous study \cite{pmid23318258}),
-meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
-Actually, the pattern of this signature matches to that of artifact by oxidation of DNA during acoustic shearing \cite{pmid23303777}.
-Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.
-
-The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer shows very strong strains specificity, implying the evidence of transcription-coupled nucleotide excision repair.
-This signature has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, though the source of this signature is still not clear.
-
+The signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas matches the pattern of that detected in the same cancer type in the previous study.
 The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
-The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, 
-and are probably associated with a chemotherapy drug, temozolomide.
+The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, and are probably associated with a chemotherapy drug, temozolomide.
 The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
-
 The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
-and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear.
-When we investigate $\pm 2$ positions, the depletion of C base at two bases 5' to the mutated site is consistently observed (Figure \ref{LMST_sig_list}).
+and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear,
+and the depletion of C base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_sig_list}).
+the signature 12 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas,
+
+The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, 
+though the source of this signature is still not clear.
+The fact that very strong strains specificity is shown and 
+the evidence of transcription-coupled nucleotide excision repair is implied is consistent to the previous studies \cite{pmid25362482, pmid23318258}.
 
@@
 the signature 21 (T $>$ [CG] at CpTpT) observed in oesophagus cancers,
-the signature 2 (C $>$ A at [CT]pCpT) observed in low grade gliomas,
-the signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas,
-the signature 12 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas, 
+
 and the signature 19 (T $>$ C at GpTpN) in stomach cancers.

### 2015-03-14 supp/supp.tex
modify several staffs for supplementary material figures

@@
 
+\clearpage
+
+\begin{figure*}
+
+% \subfigure[POLE1 signatures in each cancer type]{%
+%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
+%   \label{POLE1_sig_list}}
+
+% \subfigure[POLE2 signature in each cancer type]{%
+%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
+%   \label{POLE2_sig_list}}
+
+% \subfigure[UV signature in each cancer type]{%
+%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
+%   \label{UV_sig_list}}
+
+% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
+%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
+%   \label{LMST_sig_list}}
+
+% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
+% \label{nature2013_sig_others}
+% \end{figure*}
+
+\clearpage
+
+\begin{figure}
+\centering
+
+\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in \cite{pmid23945592})]{%
+  \includegraphics[height=3cm,width=12cm,clip]{example96_Glioma-Low-Grade.eps}
+  \label{LGG_nature2013_sig}}
+
+\subfigure[Estimated membership parameter by the proposed method]{%
+  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership.eps}
+  \label{LGG_membership}}
+\quad
+\subfigure[Estimated membership parameter by the proposed method in log scale]{%
+  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership_log.eps}
+  \label{LGG_membership_log}}
+
+  \caption{(a) The barplot is divided by 6 substitution patterns. In each division, 
+16 bars show joint probabilities of 16 combinations of the immediate 5' and 3' bases.
+(b, c) We have selected top 100 cancer samples according to the number of mutation.
+The height of bar shows (the logarithm of) the number of mutations for each sample,
+and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
+The low grade glioma specific signature detected by the proposed method is the signature 2.
+We can see that the mutations corresponding to signature 2 is mostly from the sample with an extremely high mutation rate.
+}
+
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+\centering
+
+\subfigure[The result for the original data ($K = 2$)]{%
+  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_original_signature_K2.eps}
+  \label{LGG_roriginal_sigK2_sig}}
+
+\subfigure[The result for the data without the hyper-mutated sample ($K = 2$)]{%
+  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K2.eps}
+  \label{LGG_removed_sigK2_sig}}
+
+\subfigure[The result for the data without the hyper-mutated sample ($K = 3$)]{%
+  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K3.eps}
+  \label{LGG_removed_sigK3_sig}}
+
+\subfigure[The result for the data without the hyper-mutated sample ($K = 4$)]{%
+  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K4.eps}
+  \label{LGG_removed_sigK4_sig}}
+
+\subfigure[The result for the data without the hyper-mutated sample ($K = 5$)]{%
+  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K5.eps}
+  \label{LGG_removed_sigK5_sig}}
+
+  \caption{(a) The signatures obtained for the original data. The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
+  The second signature is the low grade glioma signature. 
+  (b, c, d, e) The signature obtained for the data without the hyper-mutated case for several number of mutation signatures ($K$, including the background signature).
+  Although the signature related to deamination of 5-methyl-cytosine remained, low grade glioma specific signature could not be observed.}
+
+\end{figure}
+
+\clearpage
+
+\begin{figure}
+\centering
+
+\subfigure[The signature 2 detected in kidney clear cell carcinomas]{%
+  \includegraphics[height=3cm,width=4.8cm,clip]{Kidney-Clear-Cell_oxidation_signature.eps}
+  \label{LGG_membership}}
+\quad
+\subfigure[The signature 1 detected in lung adenocarcinomas]{%
+  \includegraphics[height=3cm,width=4.8cm,clip]{Lung-Adeno_oxidation_signature.eps}
+  \label{LGG_membership_log}}
+\quad
+\subfigure[The signature 1 detected in melanomas]{%
+  \includegraphics[height=3cm,width=4.8cm,clip]{Melanoma_oxidation_signature.eps}
+  \label{LGG_membership_log}}
+
+\subfigure[Estimated membership parameter for kidney clear cell carcinomas]{%
+  \includegraphics[height=4cm,width=12cm,clip]{Kidney-Clear-Cell_oxidation_membership.eps}
+  \label{LGG_nature2013_sig}}
+
+\subfigure[Estimated membership parameter for lung adenocarcinomas]{%
+  \includegraphics[height=4cm,width=12cm,clip]{Lung-Adeno_oxidation_membership.eps}
+  \label{LGG_membership}}
+\quad
+\subfigure[Estimated membership parameter for melanomas]{%
+  \includegraphics[height=4cm,width=12cm,clip]{Melanoma_oxidation_membership.eps}
+  \label{LGG_membership_log}}
+
+  \caption{(a, b, c) The putative oxidative artifact signatures estimated for each cancer.
+  We can observe consistent abundance of the base C at the $-2 $ position.
+  (d, e, f) For each cancer type, 
+We have selected top 100 cancer samples according to the number of mutation.
+The height of bar shows the number of mutations for each sample,
+and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
+We can see that the signature corresponding to putative oxidative artifacts concentrates on a small number of samples.}
+
+\end{figure}
+
 \end{document}

### 2015-03-14 figure/script/AlexandrovEtAl_Glioma-Low-Grade_removed.R

@@ -1,15 +1,35 @@
 library(pmsignature)
 
+# dummy signature estimation for generating class
+inputFile <- system.file("extdata/Nik_Zainal_2012.mutationPositionFormat.txt", package="pmsignature");
+G <- readMPFile(inputFile, numBases = 3, type = "full");
+Param <- getPMSignature(G, K = 2);
+
+# reading the signature from Alexandrov et al. Nature 2013
+nature2013_sig_raw <- read.table("../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
+nature2013_sig <- t(nature2013_sig_raw[,4:30]);
+colnames(nature2013_sig) <- nature2013_sig_raw[,3];
+
+Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
+Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
+
+visPMSignature(Param, 1)
+ggsave("../../supp/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
+
+
+##########
+
 G <- readMPFile("../../AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
+
 
 ##########
 # membership plot
 load("../../AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
 visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
-ggsave("../../supp/Glioma-Low-Grade.membership.eps", height = 4, width = 15, pointsize = 9);
+ggsave("../../supp/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
 
 visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
-ggsave("../../supp/Glioma-Low-Grade.membership_log.eps", height = 4, width = 15, pointsize = 9);
+ggsave("../../supp/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
 ##########
 
 
@@ -39,12 +59,35 @@ for (l1 in 1:length(Params)) {
    
   for (l2 in 1:l1) {
     aInd <- which(alignOrder[[l1]] == l2);
[…]

### 2015-03-14 figure/script/AlexandrovEtAl_oxidation.R

@@ -37,18 +37,18 @@ for (i in 1:length(oxidation_types)) {
   inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
   inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
         
-  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
+  # G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
   load(inputRdata);
-  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
-  ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 15, pointsize = 9);
+  # visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
+  # ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
 
   oxInd <- checkOxidationInd(resultForSave[[1]])
   
   tempMar <- par("mar");
-  par(mar = 0.4 * tempMar);
+  par(mar = 0.2 * tempMar);
   
-  visPMSignature(resultForSave[[1]], oxInd, charSize = 0.7);
-  dev.copy2eps(file=paste("../../supp/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 2.4, width = 3.75);    
+  visPMSignature(resultForSave[[1]], oxInd, charSize = 0.8);
+  dev.copy2eps(file=paste("../../supp/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 3.0, width = 5);    
 
   par(.pardefault);
   

### 2015-03-14 manuscript/plos_template.tex
modified manuscript

@@
 Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at the $-2$ position,
-and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
-In fact, nonsense mutations in TP53 at R231X, which are shown to be enriched in cancers having Pol $\epsilon$,
-and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
+and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}, Supplementary Figure 7(b, c)).
 In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
@@
 The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, 
-can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 7 (a)). 
+can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 8 (a)). 
 This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
-Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate (Supplementary Figure 7 (b,c)), 
-and the signature 2 disappeared when we re-analyze the data removing this sample (Supplementary Figure 8).
+Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate (Supplementary Figure 8 (b,c)), 
+and the signature 2 disappeared when we re-analyze the data removing this sample (Supplementary Figure 9).
 We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
@@
 The signature 4 (C $>$ A at CpCpG) observed in kidney clear cell carcinomas, lung adenocarcinomas and melanomas 
-shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 9 (a,b,c)).
-However, this signature is confide to only a few samples for every cancer type (see Supplementary Figure 9 (d, e, f)).
+shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 10 (a,b,c)).
+However, this signature is confide to only a few samples for every cancer type (see Supplementary Figure 10 (d, e, f)).
 This seems to correspond to the ``signatures R2''  detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas
@@
 The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
-and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}).
+and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}, Supplementary Figure 7(a)).
 This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
@@
 The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
+Observed consistent strand specificities (Figure \ref{UV_sig_list}) matches to the previous results \cite{pmid23318258}.   
+Also, higher frequencies of T base at the $\pm 2$ position are consistently observed (Figure \ref{UV_sig_list}, Supplementary Figure 7(d)).
 The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
 and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear,
-and the depletion of C base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_sig_list}).
-the signature 12 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas,
+and the abundance of G base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_sig_list}, Supplementary Figure 7(e)).
 
@@
 though the source of this signature is still not clear. 
-The fact that very strong strains specificity is shown and 
-the evidence of transcription-coupled nucleotide excision repair is implied is consistent to the previous studies \cite{pmid25362482, pmid23318258}.
-
-In addition to the above signatures, a number of prominent mutation signatures observed in one cancer type 
-had common characteristics with those observed in the previous study:
-the signature 21 (T $>$ [CG] at CpTpT) observed in oesophagus cancers,
-
-and the signature 19 (T $>$ C at GpTpN) in stomach cancers.
-
-There results indicate that the proposed method can capture not only many of the prominent mutation signatures observed in the previous study,
-but also novel characteristics for $\pm 2$ sites.
+In this signature, we can observe very strong strains specificity implying the evidence of transcription-coupled nucleotide excision repair as shown in \cite{pmid25362482, pmid23318258}.
+The unknown signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively,
+coincides well with the those detected in the same cancer types in the previous study \cite{pmid23318258}.
+Although several interesting characteristics were observed at the $\pm$ positions for these signatures, 
+their validity are difficult to confirm at this point since these are not reproduced in multiple cancer types.
+
+In summary, the propose method can capture many of the prominent mutation signatures observed in the previous study.
+Also, the propose method detected several novel characteristics at the $\pm 2$ position that is consistent across multiple cancer types,
+suggesting the existence of more hidden characteristics at the distant position from the mutated site.
 

### 2015-03-14 supp/supp.tex
modified manuscript

@@
 
-% \subfigure[POLE1 signatures in each cancer type]{%
-%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-%   \label{POLE1_sig_list}}
+\subfigure[APOBEC signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=14.5cm,height=6cm]{APOBEC_two5prime.eps}
+   \label{APOBEC_two5prime}}
 
-% \subfigure[POLE2 signature in each cancer type]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-%   \label{POLE2_sig_list}}
+\subfigure[POLE1 signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{POLE1_two5prime.eps}
+  \label{POLE1_two5prime}}
+\quad
+\subfigure[POLE2 signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{POLE2_two5prime.eps}
+  \label{POLE2_two5prime}}
 
-% \subfigure[UV signature in each cancer type]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-%   \label{UV_sig_list}}
+\subfigure[UV signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{UV_two5prime.eps}
+  \label{UV_two5prime}}
+\quad
+\subfigure[The signature 11 intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{LMST_two5prime.eps}
+  \label{LMST_two5prime}}
 
-% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-%   \label{LMST_sig_list}}
+\caption{The estimated frequencies of bases at two 5' to the mutated site for each cancer type.
+The heights of bar show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
+The error bars show standard errors estimated by bootstrap.}
+\label{two5prime}
 
-% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
-% \label{nature2013_sig_others}
-% \end{figure*}
+\end{figure*}
 

### 2015-03-14 figure/script/AlexandrovEtAl_figures_dir.R

@@ -354,9 +354,9 @@ for (i in 1:length(importantSigs)) {
     theme(text = element_text(size=20)) +
     theme(axis.text.x= element_text(angle=45,hjust=1));
 
-  outputName <- paste("../../manuscript/", names(importantSigs)[i], "_two5prime.eps", sep="");
+  outputName <- paste("../../supp/", names(importantSigs)[i], "_two5prime.eps", sep="");
   
-  ggsave(outputName, width=length(types) * 1.2 + 3, height=8, units="in");
+  ggsave(outputName, width=length(types) * 1.0 + 2.5, height = 6, units="in");
 
   
 }

### 2015-03-14 figure/script/AlexandrovEtAl_oxidation.R

@@ -37,10 +37,10 @@ for (i in 1:length(oxidation_types)) {
   inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
   inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
         
-  # G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
+  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
   load(inputRdata);
-  # visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
-  # ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
+  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
+  ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
 
   oxInd <- checkOxidationInd(resultForSave[[1]])
   

### 2015-03-18 manuscript/plos_template.tex
method overview figure

@@
 
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-\section*{Methods}
+% Results and Discussion can be combined.
+\section*{Results}
 
@@
 
-\subsection*{Mathematical representation of mutation signatures}
-
-In this subsection, 
-we show how to mathematically describe mutations collected from genome sequencing studies and mutational signatures.
-First, let $\Delta^S = \{ (t_1, \cdots, t_S) |\ t_s \geq 0 (\forall s = 1, \cdots, S), \sum_{s=1}^S t_s = 1 \}$ denote S-dimensional simplex,
-which is used to represent nonnegative vectors summing to $1$ throughout the paper.
-
-Suppose each somatic mutation has $L$ mutation features, $\bm{m} = (m_1, m_2, \cdots, m_L)$, 
-where each $m_l$ can take $M_l$ discrete values. Also, we set $\bm{M} = (M_1, \cdots, M_L)$.
-Let $\bm{x}_{i, j} = (x_{i, j, 1}, \cdots, x_{i, j, L}), (i = 1, \cdots, I, j = 1, \cdots, J_i)$ denote the mutation feature vector for the $j$-th mutation of $i$-th cancer genome,
-where $x_{i, j, l} \in \{1, \cdots, M_l \}$, $I$ is the number of available cancer genomes and $J_i$ is the number of mutations in the $i$-th cancer genome.
-When taking account of 6 substitution patterns and $\pm 2$ sites,
-$\bm{M} = (6, 4, 4, 4, 4)$.
-See Table \ref{tab_rep} for other representation example.
-
-Suppose that there are $K$ mutation signatures,
-Let  $\bm{f}_{k, l} = (f_{k, l, 1}, \cdots, f_{k, l, M_l }) \in \Delta^{M_l}$ denote 
-the multinomial distribution parameter of $k$-th mutation signature and the $l$-th mutation feature, 
-then the probabilistic distribution of $k$-th signature corresponds to $\{ \bm{f}_{k, l} \}_{l=1, \cdots, L}$.
-
-Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
+\subsection*{Overview of the proposed method}
+
+The proposed method has mainly two novelties compared to existing approaches.
+The first point is the assumption of independence, where we can identify mutation signatures as multiple multinomial distribution parameters on multiple contextual factors.
 We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
 which is reminiscent of sequencing logos \cite{pmid2172928}. 
-
-\subsection*{Non-independent model as a special case of the proposed method}
-
-The previous approaches, 
-where each mutation signature is not a set multiple multinomial distributions but one high-dimensional multinomial distribution,
-can be considered as a special case of the above framework.
-
-Suppose each mutation feature vector (such as substitution patterns, adjacent bases, and so on) 
-is re-labeled as a single-valued mutation feature by some lexicographical order (see Table \ref{tab_rep}),
-then each mutation signature becomes one multinomial distribution. 
-We call this representation and mutational process modeling as ``full representation,'' and ``full model.''
-In this case, when taking account of 6 substitution patterns and $\pm 2$ sites,
-$\bm{M} = (1536)$.
-On the other hand, the usual representation, where each mutation feature has its corresponding value and individual multinomial distribution,
-is called as ``independent representation,'' and ``independent model.''
-
-The full representation model potentially represent complicated mutational processes
-(e.g., a situation where C $>$ A is frequent at ApCpG sites and C $>$ T is frequent at TpCpA sites) with one signature.
-However, when many mutation contextual factors are taken into account and the number of free parameters get huge,
-estimated results tend to be unstable and unreliable.
-Furthermore, there is a great risk that we over-interpreting the seemingly complex estimated results.
-
-\subsection*{Generative model of mutation features for each cancer genome}
-
-Each cancer genome naturally has multiple mutation signatures,
-because we are living exposed to a variety of carcinogens. 
-Also, the strength of each mutation signature varies among cancer samples, 
-depending on lifestyles, genetic difference and so on.
-We represent the distribution of mutation signatures for the $i$-the cancer genome 
-by $\bm{q}_i = (q_{i,1}, q_{i,2}, \cdots, q_{i, K}) \in \Delta^K,  (i = 1, \cdots, I)$.
-
-We adopt a two step model for a generative model of mutations.
-First, one of the contributing mutation signature is chosen for each mutation 
-depending on the signature distribution parameter of the corresponding cancer genome $\{ \bm{q}_i \}$.
+This representation will help us to immediately identify which elements are strongly working on each contextual factors.
+In this paper, we focus on substitution patterns, flanking bases (up to the $\pm 2$ positions) and transcribed directions as contextual factors of mutation patterns.
+
+The second novel point is that we adopt  the ``mixed membership model'' for the mutation signature problem.
+Briefly, the mixed membership model is basically clustering method, but allowing that each individual can have multiple clusters.
+The mixed membership model is widely used in many different fields such as population structure classification in population genetics and document classification problems in machine learning.
+This model is also shown to be closely related to nonnegative matrix factorization.
+
+Suppose there are in total $K$ source of mutations (or mutation signatures) in the cohort of interest.
+For each cancer genome, somatic mutations are generated by one of the $K$ sources,
+and the ratio of the amount of contributing sources varies across cancer genomes depending on the ages, lifestyle, races and so on.
+We call this parameter as ``membership parameters.''
+
+Here, by convention of the mixed membership model, we assume that each mutation for each cancer genome is generated by a two step model:
+First, one of the contributing mutation signature is chosen depending on the membership parameter of that cancer genome.
 Then, according to the probability distribution of the selected mutation signature, 
 mutation features such as substitution patterns and flanking base pairs are generated.
-
-The detailed description of generative process of $\{ \bm{x}_{i,j} \}$ is as follows:
-For the $j$-th mutation in the $i$-th cancer genome, 
-\begin{enumerate}
-\item
-Generate $z_{i,j} \sim \text{Multinomial} (\bm{q}_m)$, where $z_{i,j} \in \{1,\cdots,K \}$ is the underlying mutation signature causing that mutation.
-\item 
-For each $l (= 1, \cdots, L)$, generate the values of each mutation feature $x_{i,j,l} \sim \text{Multinomial}(\bm{f}_{z_{i,j},l})$.
-\end{enumerate}
-
-\subsection*{Relationship with mixed-membership models}
-
-The two step generative model described in the previous subsection has close relationships with
-the mixed-membership models that have been adopted in many other applications, 
-such as document classification and population structure inference problems. 
-In this subsection, we show the relationships between the proposed method and mixed-membership models,
-slightly abusing notations to contrast the relationships with the proposed method.
-
-In topic models \cite{Hofmann:1999,Blei:2003},
-which is a form of mixed-membership models frequently used in document classification problems,
-each document is assumed to have $K$ different ``topics'' in varying proportions ($\bm{q}_i \in \Delta^K$),
-where each topic is characterized by a word frequency (a multinomial distribution on a set of words $W$ ($\bm{f}_k \in \Delta^W$).
-And each word is assumed to be generated by one of $K$ multinomial distributions (topics).
-The detailed generative process of the $j$-th word in the $i$-th document $x_{i,j}$ is:  
-\begin{enumerate}
-\item
-Generate the underlying topic for the $j$-th word, $z_{i,j} \sim \text{Multinomial} (\bm{q}_i)$., where $z_{i,j} \in \{1,\cdots,K \}$.
-\item
-Generate $x_{i,j} \sim \text{Multinomial} (\bm{f}_{z_{i,j}})$, where $x_{i,j} \in \{1, \cdots, W \}$.
-\end{enumerate}
-Actually, the proposed model in case of $L = 1$ or ``full representation''  is mostly the same as topic models.
-
-On the other hand, in population structure inference problems \cite{pmid10835412, pmid19648217}, 
-each individual is assumed to be an admixture of $K$ ancestries in varying proportions, 
-where each ancestry is characterized by the allele frequency at each SNP locus,
-and each SNP genotype of an individual are assumed to be generated by the two step model:
-first, one of the ancestries are chosen by the admixture ratio of ancestries given for each individual,
-and then the SNP genotype is generated according to the allele frequency of the selected ancestry at that locus.
-The relationships among the mutation signature model, topic models and population structure models are summarized in the Table \ref{tab_pop}.
-
-Furthermore, close relationships between mixed-membership models and nonnegative matrix factorization,
-which has been successfully used in the previous studies for mutational signature problems \cite{pmid22608084, pmid23318258, pmid23945592},
-have been pointed out \cite{ding2008equivalence}.
-In fact, the proposed method can be seen as non-negative matrix factorization with additional restrictions.
-See Supplementary Materials for the detail of the relationships between the proposed approach and nonnegative matrix factorization.
-
-\subsection*{Estimating parameters}
-
-The parameters $\{ \bm{f}_{k, l} \}$ and $\{ \bm{q}_i \}$
-are not given and have to be estimated from the available mutation data $\{ \bm{x}_{i,j} \}$.
-On estimating the parameters for mixed-membership models, a number of approaches have been proposed.
-
-After realizing the relationships with mixed-membership models, 
-a number of  past parameter estimation techniques proposed for mixed-membership models can be tailored for the proposed model.
-EM-algorithm (or its variant, called tempered EM algorithm) have been adopted in classical topic models 
-for document classification (\cite{Hofmann:1999}) and population structure estimation (\cite{tang2005estimation}).
-In the population structure estimation problem, \cite{pmid19648217} has proposed 
-a fast block relaxation scheme using sequential quadratic programming for block updates
-with a quasi-Newton acceleration of convergence \cite{zhou2011quasi}, demonstrating a great improvement over EM-algorithm.
-Similar techniques are used in \cite{taddy2011estimation} for the document classification.
-For the estimation of Bayesian mixed-membership models, 
-(collapsed) Gibbs sampling \cite{pmid10835412,pmid14872004} and 
-variational method \cite{Blei:2003,teh2006collapsed,raj2014variational} have been proposed.
-
-In this paper, we adopt a relatively simple approach that uses EM-algorithm.
-However, there is a great possibility that we can devise far more efficient approach based on past experiences.
-Let $g_{i, \bm{m}}$ denote the number of the $i$-th sample's mutations with the mutation feature vector $\bm{m}$.
-
-Introducing the auxiliary variables $\theta_{i, k, \bm{m}}$, 
-we update these auxiliary variables in the E-step as
-
-\begin{equation*}
-\theta_{i, k, \bm{m}} = \frac{  q_{i,k} \prod_{l=1}^L f_{k,l,m_l} }{ \sum_{k^{\prime} = 1}^K q_{i, k^{\prime} } \prod_{l=1}^L f_{k^{\prime}, l, m_l } }.
-\end{equation*}
-
-Then, in the M-step, we update the parameters $\{ f_{k, \bm{m} } \}$ and $\{ q_{i, k} \}$ as
-
-\begin{equation*}
-f_{k, l, p} = \frac{ \sum_{\bm{m} : m_l = p} g_{i, \bm{m}} \theta_{i, k, \bm{m}} }{ \sum_{p^{\prime} } 
-\sum_{\bm{m} : m_l = p^{\prime}}  g_{i, \bm{m}}\theta_{i, k, \bm{m}} },
-\end{equation*}
-\begin{equation*}
-q_{i, k} = \frac{ \sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k, \bm{m} } }{  \sum_{k^{\prime} }\sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k^{\prime}, \bm{m} } }.
-\end{equation*}
-
-In addition, we use SQUAREM \cite{varadhan2008simple},
-which is a general framework for accelerating the convergence of any fixed-point iteration such as EM algorithm. 
-Furthermore, to reduce problems with convergence to local minima, 
-we performe EM-algorithm several times (10 times in this paper) changing the initial points,
-and adopt the estimate with the maximum log-likelihood.
-For the derivation of EM-algorithm, please see Supplementary Material.
-
-\subsection*{Adding background signatures}
-
-There may be a possibility that the intrinsic composition of the genome sequence influences the estimated mutation signatures.
-For example, the number of observed C $>$ T transitions at CpG sites may increase at promoter regions,
-just because CpG dinucleotides are more frequent in those regions.
-In the previous research \cite{pmid23628380}, this background problem was dealt 
-by explicitly incorporating mutation ``opportunity'' coefficients into the model.
-
-Here, to offset the influences of intrinsic sequence composition, we add background signatures 
-$\{ \bm{f}_{0, \bm{m}} \} \in \Delta^{M_1 \times  \cdots \times M_L}$.
-For example, when obtaining the background signatures in case of considering substitution patterns and up to $\pm 2$ bases, 
-we first calculate the frequencies of 5-mers where complement sequences are taken when the central bases are A or G,
-and then the frequencies divided by 3 was give on each mutation feature vectors $\bm{m}$
-considering the alternated bases are equally likely.
-Since we mainly deal with mutation data from exome sequencing,
-background signatures were calculated on entire exonic regions.
-
-\subsection*{Estimating standard errors}
-
-The standard errors for the parameter estimates are calculated using bootstrap, 
-where somatic mutations are resampled according to the empirical distribution of the original data $\{ \bm{x}_{i,j} \}$ for each cancer genome.
-For each bootstrap resample, we performed the re-estimation using parameters obtained for the original data as initial points,
-and calculated sample standard errors of the inferred mutational signatures as estimates of parameter standard errors.
-In this paper, we performed 100 bootstrap resampling.
-
-\subsection*{Selecting the number of signatures}
-
-Determining the number of mutation signatures $K$ is an important and challenging task. 
-One approach is to utilize some statistical information criteria such as AIC \cite{akaike1974new}, BIC \cite{schwarz1978estimating}.
-In the population structure problems, for example, 
-the Bayesian deviance \cite{pmid10835412}, 
-and cross-validation \cite{alexander2011enhancements} have been suggested.
-One previous study on mutation signature problems \cite{pmid23628380} utilized BIC.
-The problem of using these statistical information criteria is that most of them are based on the likelihood,
-where slight deviations between the specified probabilistic models and the reality sometimes lead to 
-larger number of mutation signatures for compensating those deviations,
-and produce results with a risk of over-interpretation.
-
-In this paper, instead of utilizing statistical information criteria,
-we adopt following strategies:
-
-\begin{itemize}
-\item
-After calculating the likelihood and standard errors of parameters for a range of $K$,
-the value of $K$ is determined at the point where the likelihood is sufficiently high, 
-and the standard errors are sufficiently low \cite{pmid23318258}.
-
-\item
-When, for $k_1$-th and $k_2$-th mutation signatures, 
-we could detect strong correlations between the estimated membership parameters 
-for each cancer genome
-($(q_{1,k_1}, q_{2,k_1}, \cdots, q_{I,k_1})$ and $(q_{1,k_2}, q_{2,k_2}, \cdots, q_{I,k_2})$),
-and the two mutation signatures ($\{ f_{k_1, \bm{m} } \}$ and $\{ f_{k_2, \bm{m} } \}$ ) show similar patterns,
-then it is likely that an excess amount of $K$ forced to split one mutation signature into two.
-We stop increasing $K$ before these pairs of mutation signatures are observed.
-
-\end{itemize}
-
-The strategies listed above are probably not exhaustive, 
-and we should add other practical strategies observing various situations.
-In addition, it would be nice is we could devise automated and practical approaches for choosing $K$,
-which is a possible future challenges.
-
-% Results and Discussion can be combined.
-\section*{Results}
+Although the mutation signatures and membership parameters are unknown,
+we can efficiently estimate these parameters by an EM-algorithm.
+
+To account for the intrinsic sequential context, we add background signatures.
+See the Method section for the mathematical details, 
+detailed relationships with mixed membership models and nonnegative matrix factorizations, how to choose the number of mutation signatures and so on.
 
@@
 Finally, we have applied the proposed method for the somatic mutation data of 30 cancer types \cite{pmid23945592}.
-For each cancer type, the proposed method was applied separately to see the robustness of the estimated signatures among different cancer types.
-By changing the number of mutation signatures and calculating values of the log-likelihood and bootstrap errors,
+The proposed method was applied on each cancer type separately to see the robustness of the estimated signatures across different cancer types.
+By changing the number of mutation signatures and examining values of the log-likelihood and bootstrap errors as well as correlation of membership parameters,
 we have chosen the number of mutation signatures for each cancer type.
-The estimated mutation signatures across cancer types were clustered by Frobenius Distance.
+Finally, similar mutation signatures observed across cancer types were clustered by Frobenius Distance.
  % (see Supplementary Material for the values of the log-likelihoods and bootstrap errors for each cancer type).
@@
 In total, 27 mutation signatures were extracted.
-By comparing the composition of nucleotides and memberships of signatures with the result of the previous study,
+By comparing the composition of nucleotides and cancer types exhibiting the signatures with the result of the previous study,
 we could associate many of the detected signatures with known mutational processes.
 
-The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) were observed in colorectal and uterine cancers,
-are associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
-The proposed method used two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
+The signature 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) observed in colorectal and uterine cancers
+are arguably associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
+The proposed method uses two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
 while was represented by one signature in the previous study.
-Interestingly, signature 1 showed transcription biases whereas signature 8 did not.
+Interestingly, the signature 1 shows transcription biases whereas the signature 8 does not.
 This phenomena are consistent for both colorectal and uterus cancers (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
@@
 whereas, there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
-Although replication strand biases is different from the transcription strand biases, these two biases may have strong correlation,
-considering the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
-Therefore, the transcription strand bias observed in the signature 1 may actually come from replication strand biases.
-
-Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at the $-2$ position,
-and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}, Supplementary Figure 7(b, c)).
-In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
-In fact, this mutation is C $>$ T at TpTpCpGpA, the most likely patten with the signature 8.
+Although the replication strand bias is different from the transcription strand bias, these two biases may have strong correlation
+through the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
+Therefore, the transcription strand bias observed in the signature 1 may actually come from the replication strand biases.
+
+Furthermore, both the signature 1 and 8 signatures shows strong abundance of the T base at the $-2$ position,
+and the signature 1 also shows a little abundance of the T base at the $+2$ position.
+ (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}, Supplementary Figure 7(b, c)).
+In the previous study of Pol $\epsilon$  \cite{pmid25228659}, 
+a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
+In fact, the pattern of this mutation is C $>$ T at TpTpCpGpA, that is the most likely patten with the signature 8.
 This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.
 
-The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, 
+The signature 2 (C $>$ A at [CT]pCpT) observed in low grade gliomas
 can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 8 (a)). 
-This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
-Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate (Supplementary Figure 8 (b,c)), 
-and the signature 2 disappeared when we re-analyze the data removing this sample (Supplementary Figure 9).
+This seems to be because the corresponding signature in the previous study shows very complex patterns (C $>$ A at NpCpT or C $>$ T at GpTpN).
+Actually, the mutation related to this signature is mostly coming from one sample with an extremely high mutation rate (Supplementary Figure 8 (b,c)), 
+and the signature 2 disappeared when we remove this sample (Supplementary Figure 9).
 We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
 but cannot be decomposed due to the lack of the number of samples having these signatures.
-Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.
+Therefore, we need more number of samples to validate the authenticity of the signature 2 and the corresponding signature in the previous study.
 
@@
 shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 10 (a,b,c)).
-However, this signature is confide to only a few samples for every cancer type (see Supplementary Figure 10 (d, e, f)).
-This seems to correspond to the ``signatures R2''  detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas
-(see supplementary figures of the previous study \cite{pmid23318258})
+However, for each cancer type, only a few samples support this signature (see Supplementary Figure 10 (d, e, f)).
+Also, this signature seems to correspond to the ``signatures R2''  detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas (see supplementary figures of the previous study \cite{pmid23318258})
 that could not be validated in the previous study,
-meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
-Actually, the pattern of this signature largely matches to that of the putative artifact by oxidation of DNA during acoustic shearing suggested in \cite{pmid23303777},
-we guess that this signature is the same as what is observed in that study.
-Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.
+meaning that significant ratios of somatic mutations corresponding to that signature could not be validated by re-sequencing or visual inspection of BAM files using genomic viewers.
+Actually, the pattern of this signature largely matches to that of the putative artifact suggested in \cite{pmid23303777}, 
+that is caused by oxidation of DNA during acoustic shearings, 
+and we guess that this signature is the same as what is observed in that study.
+Although this signature may be from the artifacts, 
+knowing higher resolution of the artifact signature is very helpful to accurately remove false positive mutations.
 
 The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
-The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
+The composition of nucleotides of estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
 and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}, Supplementary Figure 7(a)).
 This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
-and implies the robustness of the proposed approach.
-The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
+and indicates the robustness of the proposed approach.
+The estimated transcribed strand specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
 The Signature 15 and 16 may also be delated to the APOBEC signatures although the estimated forms are slightly different,
-and are not merged to the cluster of signature 13 in the current criteria.
+and are not merged to the cluster of signature 13 by the specified criteria.
 
@@
 The signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas matches the pattern of that detected in the same cancer type in the previous study.
-The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
+The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and arguably related to deamination of 5-methyl-cytosine.
 The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, and are probably associated with a chemotherapy drug, temozolomide.
@@
 
-The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, 
+The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancers has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, 
 though the source of this signature is still not clear. 
-In this signature, we can observe very strong strains specificity implying the evidence of transcription-coupled nucleotide excision repair as shown in \cite{pmid25362482, pmid23318258}.
+In this signature, we can observe a very strong strand specificity as shown in \cite{pmid25362482, pmid23318258},
+ indicating the evidence of transcription-coupled nucleotide excision repairs.
+The signature 20 (T $>$ C at [AC]pTpN) observed in thyroid cancers also shows a very strong strand specificity 
+and a sign of transcription-coupled nucleotide excision repairs.
+Actually, this signature does not match to any of signatures in the previous paper \cite{pmid23318258}.
+This may be because the mutation ratio of thyroid cancer is low and the signature is too week to detect for the previous methods.
+%% download the TCGA data and check the visualization of the mutation position for this data.
 The unknown signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively,
 coincides well with the those detected in the same cancer types in the previous study \cite{pmid23318258}.
-Although several interesting characteristics were observed at the $\pm$ positions for these signatures, 
-their validity are difficult to confirm at this point since these are not reproduced in multiple cancer types.
+We could observe several interesting characteristics at the $\pm 2$ positions for these signatures,
+which should be validated by more number of samples from different cohorts and cancer types.
 
@@
 
-In this paper, we proposed a novel framework for extracting mutation signatures from a set of somatic point substitutions.
-We have shown close relationships with mixed membership models, 
-that have been deeply investigated in statistical genetics and machine learning fields, as well as nonnegative matrix factorization.
-We have demonstrated that the proposed method could capture meaningful characteristics of mutational patterns.
-
-There is a lot of room for improvement by learning from past experiences and knowledge.
-First, introducing certain prior distributions or penalty terms can lead more efficiency in terms of both accuracy and interpretation,
+In this paper, we proposed a novel method for extracting mutation signatures from a set of somatic point substitutions.
+The proposed model utilizes the framework of mixed membership models, 
+that have been deeply investigated in statistical genetics and machine learning fields.
+We believe that there is a great possibility of improvement by learning the past experiences from those fields (see Method section for detail).
+
+Even though the proposed method impose an independence assumptions among each contextual factor,
+the proposed method could capture many of the previously detected mutation signatures.
+Also, we have shown that the proposed method with the independence assumption is gives more robust estimates compared to that without the independence assumption.
+In addition, the estimated mutation signatures with the independence assumption can be relatively easily visualized in an interpretable manner.
+As we have discussed, the independence assumption may not be completely consistent with the nature of DNA damages and repairs.
+Still, we have demonstrated that the model with the independence assumption is ``practically'' very helpful.
+
+We have just focused on somatic substitutions in this paper.
+However, there are a variety of mutations in cancer genomes, 
+such as insertions, deletions, double nucleotides substitutions, structural variations and copy number alterations.
+Our framework can potentially treat those mutations by considering appropriate mutation features.
+For example, deletions can be represented by features such as the lengths of deletions, adjacent bases, and so on.
+However, detailed investigation on what mutation features are practical is a future problem.
+
+Finally, mutation ratios have been shown to be closely related to a number of genomic and epigenetic factors,
+such as GC contents, repeat sequences, chromatin accessibility and modifications, and replication timing and so on \cite{pmid22820252, pmid21953857, pmid23422670, pmid23770567}.
+Also, a recent study has shown that the epigenetic information of the cell type of origin of the corresponding tumors is the most predictive \cite{pmid25693567}.
+As a number of epigenetic data from various type of cells accumulate,
+one of the next main topic will be to integrating those epigenetic factors into the mutation signature 
+and to understand how these epigenetic factors influence on DNA damage and repair mechanisms.
+
+% In this paper, we have demonstrated that the independent representations could obtain more robust estimates than the full representations.
+% Also, even in the independent representations, more complex mutational process could be represented by utilizing multiple mutation signatures.
+% However, we still do not know about the exact mechanisms of mutational processes and 
+% there might be those that should be represented by the full representations or somewhere between the full representations and the independent representations.
+% We should keep exploring appropriate representations of mutation signature with expertise from biology and chemistry.
+
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Methods}
+
+\subsection*{Mathematical representation of mutation signatures}
+
+In this subsection, 
+we show how to mathematically describe mutations collected from genome sequencing studies and mutational signatures.
+First, let $\Delta^S = \{ (t_1, \cdots, t_S) |\ t_s \geq 0 (\forall s = 1, \cdots, S), \sum_{s=1}^S t_s = 1 \}$ denote S-dimensional simplex,
+which is used to represent nonnegative vectors summing to $1$ throughout the paper.
+
+Suppose each somatic mutation has $L$ mutation features, $\bm{m} = (m_1, m_2, \cdots, m_L)$, 
+where each $m_l$ can take $M_l$ discrete values. Also, we set $\bm{M} = (M_1, \cdots, M_L)$.
+Let $\bm{x}_{i, j} = (x_{i, j, 1}, \cdots, x_{i, j, L}), (i = 1, \cdots, I, j = 1, \cdots, J_i)$ denote the mutation feature vector for the $j$-th mutation of $i$-th cancer genome,
+where $x_{i, j, l} \in \{1, \cdots, M_l \}$, $I$ is the number of available cancer genomes and $J_i$ is the number of mutations in the $i$-th cancer genome.
+When taking account of 6 substitution patterns and $\pm 2$ sites,
+$\bm{M} = (6, 4, 4, 4, 4)$.
+See Table \ref{tab_rep} for other representation example.
+
+Suppose that there are $K$ mutation signatures,
+Let  $\bm{f}_{k, l} = (f_{k, l, 1}, \cdots, f_{k, l, M_l }) \in \Delta^{M_l}$ denote 
+the multinomial distribution parameter of $k$-th mutation signature and the $l$-th mutation feature, 
+then the probabilistic distribution of $k$-th signature corresponds to $\{ \bm{f}_{k, l} \}_{l=1, \cdots, L}$.
+
+Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
+We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
+which is reminiscent of sequencing logos \cite{pmid2172928}.
+
+\subsection*{Non-independent model as a special case of the proposed method}
+
+The previous approaches, 
+where each mutation signature is not a set multiple multinomial distributions but one high-dimensional multinomial distribution,
+can be considered as a special case of the above framework.
+
+Suppose each mutation feature vector (such as substitution patterns, adjacent bases, and so on) 
+is re-labeled as a single-valued mutation feature by some lexicographical order (see Table \ref{tab_rep}),
+then each mutation signature becomes one multinomial distribution. 
+We call this representation and mutational process modeling as ``full representation,'' and ``full model.''
+In this case, when taking account of 6 substitution patterns and $\pm 2$ sites,
+$\bm{M} = (1536)$.
+On the other hand, the usual representation, where each mutation feature has its corresponding value and individual multinomial distribution,
+is called as ``independent representation,'' and ``independent model.''
+
+The full representation model potentially represent complicated mutational processes
+(e.g., a situation where C $>$ A is frequent at ApCpG sites and C $>$ T is frequent at TpCpA sites) with one signature.
+However, when many mutation contextual factors are taken into account and the number of free parameters get huge,
+estimated results tend to be unstable and unreliable.
+Furthermore, there is a great risk that we over-interpreting the seemingly complex estimated results.
+
+\subsection*{Generative model of mutation features for each cancer genome}
+
+Each cancer genome naturally has multiple mutation signatures,
+because we are living exposed to a variety of carcinogens. 
+Also, the strength of each mutation signature varies among cancer samples, 
+depending on lifestyles, genetic difference and so on.
+We represent the distribution of mutation signatures for the $i$-the cancer genome 
+by $\bm{q}_i = (q_{i,1}, q_{i,2}, \cdots, q_{i, K}) \in \Delta^K,  (i = 1, \cdots, I)$.
+
+We adopt a two step model for a generative model of mutations.
+First, one of the contributing mutation signature is chosen for each mutation 
+depending on the signature distribution parameter of the corresponding cancer genome $\{ \bm{q}_i \}$.
+Then, according to the probability distribution of the selected mutation signature, 
+mutation features such as substitution patterns and flanking base pairs are generated.
+
+The detailed description of generative process of $\{ \bm{x}_{i,j} \}$ is as follows:
+For the $j$-th mutation in the $i$-th cancer genome, 
+\begin{enumerate}
+\item
+Generate $z_{i,j} \sim \text{Multinomial} (\bm{q}_m)$, where $z_{i,j} \in \{1,\cdots,K \}$ is the underlying mutation signature causing that mutation.
+\item 
+For each $l (= 1, \cdots, L)$, generate the values of each mutation feature $x_{i,j,l} \sim \text{Multinomial}(\bm{f}_{z_{i,j},l})$.
+\end{enumerate}
+
+\subsection*{Relationship with mixed-membership models}
+
+The two step generative model described in the previous subsection has close relationships with
+the mixed-membership models that have been adopted in many other applications, 
+such as document classification and population structure inference problems. 
+In this subsection, we show the relationships between the proposed method and mixed-membership models,
+slightly abusing notations to contrast the relationships with the proposed method.
+
+In topic models \cite{Hofmann:1999,Blei:2003},
+which is a form of mixed-membership models frequently used in document classification problems,
+each document is assumed to have $K$ different ``topics'' in varying proportions ($\bm{q}_i \in \Delta^K$),
+where each topic is characterized by a word frequency (a multinomial distribution on a set of words $W$ ($\bm{f}_k \in \Delta^W$).
+And each word is assumed to be generated by one of $K$ multinomial distributions (topics).
+The detailed generative process of the $j$-th word in the $i$-th document $x_{i,j}$ is:  
+\begin{enumerate}
+\item
+Generate the underlying topic for the $j$-th word, $z_{i,j} \sim \text{Multinomial} (\bm{q}_i)$., where $z_{i,j} \in \{1,\cdots,K \}$.
+\item
+Generate $x_{i,j} \sim \text{Multinomial} (\bm{f}_{z_{i,j}})$, where $x_{i,j} \in \{1, \cdots, W \}$.
+\end{enumerate}
+Actually, the proposed model in case of $L = 1$ or ``full representation''  is mostly the same as topic models.
+
+On the other hand, in population structure inference problems \cite{pmid10835412, pmid19648217}, 
+each individual is assumed to be an admixture of $K$ ancestries in varying proportions, 
+where each ancestry is characterized by the allele frequency at each SNP locus,
+and each SNP genotype of an individual are assumed to be generated by the two step model:
+first, one of the ancestries are chosen by the admixture ratio of ancestries given for each individual,
+and then the SNP genotype is generated according to the allele frequency of the selected ancestry at that locus.
+The relationships among the mutation signature model, topic models and population structure models are summarized in the Table \ref{tab_pop}.
+
+Furthermore, close relationships between mixed-membership models and nonnegative matrix factorization,
+which has been successfully used in the previous studies for mutational signature problems \cite{pmid22608084, pmid23318258, pmid23945592},
+have been pointed out \cite{ding2008equivalence}.
+In fact, the proposed method can be seen as non-negative matrix factorization with additional restrictions.
+See Supplementary Materials for the detail of the relationships between the proposed approach and nonnegative matrix factorization.
+
+\subsection*{Estimating parameters}
+
+The parameters $\{ \bm{f}_{k, l} \}$ and $\{ \bm{q}_i \}$
+are not given and have to be estimated from the available mutation data $\{ \bm{x}_{i,j} \}$.
+On estimating the parameters for mixed-membership models, a number of approaches have been proposed.
+
+In this paper, we adopt a relatively simple approach that uses EM-algorithm.
+However, there is a great possibility that we can devise far more efficient approach based on past experiences.
+Let $g_{i, \bm{m}}$ denote the number of the $i$-th sample's mutations with the mutation feature vector $\bm{m}$.
+
+Introducing the auxiliary variables $\theta_{i, k, \bm{m}}$, 
+we update these auxiliary variables in the E-step as
+
+\begin{equation*}
+\theta_{i, k, \bm{m}} = \frac{  q_{i,k} \prod_{l=1}^L f_{k,l,m_l} }{ \sum_{k^{\prime} = 1}^K q_{i, k^{\prime} } \prod_{l=1}^L f_{k^{\prime}, l, m_l } }.
+\end{equation*}
+
+Then, in the M-step, we update the parameters $\{ f_{k, \bm{m} } \}$ and $\{ q_{i, k} \}$ as
+
+\begin{equation*}
+f_{k, l, p} = \frac{ \sum_{\bm{m} : m_l = p} g_{i, \bm{m}} \theta_{i, k, \bm{m}} }{ \sum_{p^{\prime} } 
+\sum_{\bm{m} : m_l = p^{\prime}}  g_{i, \bm{m}}\theta_{i, k, \bm{m}} },
+\end{equation*}
+\begin{equation*}
+q_{i, k} = \frac{ \sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k, \bm{m} } }{  \sum_{k^{\prime} }\sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k^{\prime}, \bm{m} } }.
+\end{equation*}
+
+In addition, we use SQUAREM \cite{varadhan2008simple},
+which is a general framework for accelerating the convergence of any fixed-point iteration such as EM algorithm. 
+Furthermore, to reduce problems with convergence to local minima, 
+we performe EM-algorithm several times (10 times in this paper) changing the initial points,
+and adopt the estimate with the maximum log-likelihood.
+For the derivation of EM-algorithm, please see Supplementary Material.
+
+\subsection*{Adding background signatures}
+
+There may be a possibility that the intrinsic composition of the genome sequence influences the estimated mutation signatures.
+For example, the number of observed C $>$ T transitions at CpG sites may increase at promoter regions,
+just because CpG dinucleotides are more frequent in those regions.
+In the previous research \cite{pmid23628380}, this background problem was dealt 
+by explicitly incorporating mutation ``opportunity'' coefficients into the model.
+
+Here, to offset the influences of intrinsic sequence composition, we add background signatures 
+$\{ \bm{f}_{0, \bm{m}} \} \in \Delta^{M_1 \times  \cdots \times M_L}$.
+For example, when obtaining the background signatures in case of considering substitution patterns and up to $\pm 2$ bases, 
+we first calculate the frequencies of 5-mers where complement sequences are taken when the central bases are A or G,
+and then the frequencies divided by 3 was give on each mutation feature vectors $\bm{m}$
+considering the alternated bases are equally likely.
+Since we mainly deal with mutation data from exome sequencing,
+background signatures were calculated on entire exonic regions.
+
+\subsection*{Estimating standard errors}
+
+The standard errors for the parameter estimates are calculated using bootstrap, 
+where somatic mutations are resampled according to the empirical distribution of the original data $\{ \bm{x}_{i,j} \}$ for each cancer genome.
+For each bootstrap resample, we performed the re-estimation using parameters obtained for the original data as initial points,
+and calculated sample standard errors of the inferred mutational signatures as estimates of parameter standard errors.
+In this paper, we performed 100 bootstrap resampling.
+
+\subsection*{Selecting the number of signatures}
+
+Determining the number of mutation signatures $K$ is an important and challenging task. 
+One approach is to utilize some statistical information criteria such as AIC \cite{akaike1974new}, BIC \cite{schwarz1978estimating}.
+In the population structure problems, for example, 
+the Bayesian deviance \cite{pmid10835412}, 
+and cross-validation \cite{alexander2011enhancements} have been suggested.
+One previous study on mutation signature problems \cite{pmid23628380} utilized BIC.
+The problem of using these statistical information criteria is that most of them are based on the likelihood,
+where slight deviations between the specified probabilistic models and the reality sometimes lead to 
+larger number of mutation signatures for compensating those deviations,
+and produce results with a risk of over-interpretation.
+
+In this paper, instead of utilizing statistical information criteria,
+we adopt following strategies:
+
+\begin{itemize}
+\item
+After calculating the likelihood and standard errors of parameters for a range of $K$,
+the value of $K$ is determined at the point where the likelihood is sufficiently high, 
+and the standard errors are sufficiently low \cite{pmid23318258}.
+
+\item
+When, for $k_1$-th and $k_2$-th mutation signatures, 
+we could detect strong correlations between the estimated membership parameters 
+for each cancer genome
+($(q_{1,k_1}, q_{2,k_1}, \cdots, q_{I,k_1})$ and $(q_{1,k_2}, q_{2,k_2}, \cdots, q_{I,k_2})$),
+and the two mutation signatures ($\{ f_{k_1, \bm{m} } \}$ and $\{ f_{k_2, \bm{m} } \}$ ) show similar patterns,
+then it is likely that an excess amount of $K$ forced to split one mutation signature into two.
+We stop increasing $K$ before these pairs of mutation signatures are observed.
+
+\end{itemize}
+
+The strategies listed above are probably not exhaustive, 
+and we should add other practical strategies observing various situations.
+In addition, it would be nice is we could devise automated and practical approaches for choosing $K$,
+which is a possible future challenges.
+
+\subsection*{Future works for the methodology}
+
+Since the mixed membership model has been widely studied in the field machine learning and population genetics,
+there are many rooms for improvement in the current form of the proposed method by learning from past experiences and knowledges.
+
+First, a number of large scale cancer genome sequencing projects have been launched all over the world,
+and we need to improve the parameter estimation so that those large scale data can be handled.
+A number of  past parameter estimation techniques proposed for mixed-membership models can be tailored for the proposed model.
+EM-algorithm (or its variant, called tempered EM algorithm) have been adopted in classical topic models 
+for document classification (\cite{Hofmann:1999}) and population structure estimation (\cite{tang2005estimation}).
+In the population structure estimation problem, \cite{pmid19648217} has proposed 
+a fast block relaxation scheme using sequential quadratic programming for block updates
+with a quasi-Newton acceleration of convergence \cite{zhou2011quasi}, demonstrating a great improvement over EM-algorithm.
+Similar techniques are used in \cite{taddy2011estimation} for the document classification.
+For the estimation of Bayesian mixed-membership models, 
+(collapsed) Gibbs sampling \cite{pmid10835412,pmid14872004} and 
+variational method \cite{Blei:2003,teh2006collapsed,raj2014variational} have been proposed. 
+Evaluate those techniques in the problem of mutation signature extraction problems, and incorporate the efficient method is an important future work.
+
+Second, introducing certain prior distributions or penalty terms can lead more efficiency in terms of both accuracy and interpretation,
 considering the success in machine learning and statistical genetics communities.
@@
 avoiding over fine-tuning for frequent groups of words with multiple similar topics.
-Second, utilizing other metrics than Euclidean distance and Kullback-Leibler Divergence might be worth investigating.
+
+Tihrd, for avoiding the problem of determining the number of latent variables, 
+introducing Hierarchical dirichlet processes \cite{teh2006hierarchical}, 
+which is a natural extension of topic models to nonparametric Bayesian frameworks, might be helpful.
+
+Finally, utilizing other metrics than Euclidean distance and Kullback-Leibler Divergence might be worth investigating.
 For nonnegative matrix factorization, 
@@
 has been successfully utilized in many research fields (reviewed in \cite{fevotte2011algorithms}).
-Third, for avoiding the problem of determining the number of latent variables, 
-introducing Hierarchical dirichlet processes \cite{teh2006hierarchical}, 
-which is a natural extension of topic models to nonparametric Bayesian frameworks, might be helpful.
-
-In this paper, we have demonstrated that the independent representations could obtain more robust estimates than the full representations.
-Also, even in the independent representations, more complex mutational process could be represented by utilizing multiple mutation signatures.
-However, we still do not know about the exact mechanisms of mutational processes and 
-there might be those that should be represented by the full representations or somewhere between the full representations and the independent representations.
-We should keep exploring appropriate representations of mutation signature with expertise from biology and chemistry.
-
-We have just focused on somatic substitutions in this paper.
-However, there are a variety of mutations in cancer genomes, 
-such as insertions, deletions, double nucleotides substitutions, structural variations and copy number alterations.
-Our framework can potentially treat those mutations by considering appropriate mutation features.
-For example, for deletions, potential mutation features can be the lengths of deletions, adjacent bases, and so on.
-However, detailed investigation on what mutation features are practical is a future problem.
 

### 2015-03-18 figure/script/AlexandrovEtAl_figures_dir.R

@@ -179,10 +179,10 @@ colnames(sig2type) <- type2sigNum[,1];
 myCol <- colorRampPalette(c("#F8F8FF", "#F8F8FF", "#F8F8FF", "#6B8E23"));
 
 #' Export Width 800, Height 800
-corrplot(sig2type, is.corr=FALSE, bg="#F8F8FF", col=myCol(200), tl.col="black", tl.cex=1.2, tl.srt=90, cl.pos="n");
+corrplot(sig2type, is.corr=FALSE, bg="#F8F8FF", col=myCol(200), tl.col="black", tl.cex=1.0, tl.srt=90, cl.pos="n");
 
 outputName <- "../../manuscript/corrplot.eps"
-dev.copy2eps(file=outputName, height = 15, width = 15, pointsize = 18);
+dev.copy2eps(file=outputName, height = 15, width = 15, pointsize = 12);
 par(.pardefault);
 
 ##########
@@ -308,6 +308,38 @@ for (i in 1:length(importantSigs)) {
 par(.pardefault);
 
 
+###
+
+importantSigs <- c(APOBEC_ind, UV_ind, LMST_ind);
+names(importantSigs) <- c("APOBEC", "UV", "LMST");
+
+par(mar=c(0, 0, 0, 0));
+par(xaxs = "i", yaxs = "i");
+
+for (i in 1:length(importantSigs)) {
+  
+
+  F <- Fs[[as.numeric(importantSigs[i])]];
+    
+  plot.new();
+  plot.window(xlim=c(-0.3, 6.3), ylim=c(-0.3, 3.8));
+  
+  # the setting of charSize is temporary... why the size changes by the number of rows.....
+  visPMS_ind(F, numBases = 5, trDir = TRUE, charSize = 0);
+       
+  
+  outputName <- paste("../methodOverview/signatureList_", names(importantSigs)[i], ".eps", sep="");
+  dev.copy2eps(file=outputName, height = 2.4, width = 3.75, pointsize = 1e-10);
+
+}
+
+par(.pardefault);
+
+
+
+###
+
+##########
 
 ##########
 

### 2015-03-18 figure/script/AlexandrovEtAl_function.R

@@ -48,7 +48,7 @@ visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale 
     for(w in 1:4) {
       endx <- startx + A[l,w]
       polygon(c(startx, endx, endx, startx), c(0, 0, sizes[l], sizes[l]), col = baseCol[w], border=F);
-      if (endx - startx > 1 / 4 & sizes[l] > 0.5) {
+      if (endx - startx > 1 / 4 & sizes[l] > 0.5 & charSize > 0) {
         text(0.5 * (endx + startx), 0.5 * sizes[l], num2base[w], col="white", cex=charSize)
       }
       startx <- endx;
@@ -63,7 +63,7 @@ visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale 
     for(ww in 1:4) {
       endy <- starty + B[w,ww];
       polygon(c(startx, endx, endx, startx), c(starty, starty, endy, endy), col=baseCol[ww], border=F);
-      if ((endy - starty > 1 / 4) & (endx - startx > 1 / 4)) {
+      if ((endy - starty > 1 / 4) & (endx - startx > 1 / 4) & charSize > 0) {
         text(0.5 * (endx + startx), 0.5 * (endy + starty), num2base[ww], col="white", cex=charSize)
       }
       starty <- endy;
@@ -105,7 +105,7 @@ visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale 
     starty <- 2;
     endy <- starty + v3[1];
     polygon(c(startx, endx, endx, startx), c(starty, starty, endy, endy), col=baseCol[5], border=F);
-    if (endy - starty > 1 / 8) {
+    if (endy - starty > 1 / 8 & charSize > 0) {
       text(0.5 * (startx + endx), 0.5 * (starty + endy), "+", col="white", cex=charSize)
     }
     
@@ -114,7 +114,7 @@ visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale 
     starty <- 2;
     endy <- starty + v3[2];
     polygon(c(startx, endx, endx, startx), c(starty, starty, endy, endy), col=baseCol[6], border=F);
-    if (endy - starty > 1 / 8) {
+    if (endy - starty > 1 / 8 & charSize > 0) {
       text(0.5 * (startx + endx), 0.5 * (starty + endy), "-", col="white", cex=charSize)
     }
     

### 2015-03-19 manuscript/plos_template.tex
a number of modifications such as addition of method overview figures and many corrections of manuscript

@@
 we could not treat many contextual factors all together because of instability of estimation results.
-Furthermore, interpretation of mutations signatures of huge dimensional vectors is often troublesome.
+Furthermore, interpretation of mutations signatures with huge dimensional vectors is often troublesome.
 
 In this paper, we propose a novel approach based on hierarchical probabilistic modeling.
-The proposed approach reduces the number of parameters by the independence assumption on each factor 
+The proposed approach reduces the number of parameters by an independence assumption on each factor 
 so that we can obtain more robust and interpretable estimates.
@@
 Recognizing these relationships will help us to develop a grasp of mutation signature extraction problems,
-and will allow us to utilize a lot of techniques accumulated on other fields to further improve the statistical methods.
-
-Finally, we have prepared an R package of the proposed approach (probabilistic mutation signature, pmsignature),
+and will allow us to utilize a number of techniques accumulated on other fields to further improve the statistical methods.
+
+We have prepared an R package of the proposed approach (probabilistic mutation signature, pmsignature),
 which is available at \url{https://github.com/friend1ws/pmsignature}.
@@
 
-It is known that the pattern of somatic mutations depends on cancer types and even individuals within the same caner type.
-For example, C $>$ A mutations are frequent in lung caner
-whereas C $>$ T and CC $>$ TT mutations are frequent in skin cancer,
-and their patterns are usually associated to some kinds of carcinogens.
+The pattern of somatic mutations has been known to depend on cancer types and even individuals within the same caner type.
+For example, C $>$ A mutations are frequent in lung caners
+whereas C $>$ T and CC $>$ TT mutations are frequent in skin cancers,
+and their patterns are usually associated to some kinds of carcinogens such as tobacco smoking and ultraviolets.
 However, since the cost and throughput of sequencing technologies were limited, 
-high-resolution analysis of mutation patterns was not possible.
+high-resolution analyses of mutation patterns were not possible.
 
@@
 and there is a great possibility of detecting novel mutation patterns leading to identification of novel carcinogens
-and obtaining better classification of cancer genomes based on mutation patterns differences.
-On the other hand, there is an increasing need for novel statistical method for mutation patterns analysis 
-from vast amount of somatic mutation data.
+and better classification of cancer genomes.
+On the other hand, there is an increasing need for novel statistical method for analyzing mutation patterns
+from vast amounts of somatic mutation data.
 
@@
 a small portions of mutations at some specific sites in cancer genes (called ``driver mutations'') 
-confer growth activities to the cells having them over other cells,
-allowing such as autonomous proliferation and tissue invasion,
-and contribute to oncogenesis \cite{stratton2009cancer}.
-The main goal in cancer genome study has been to find the driver mutations to understand the mechanism of cancer development.
+confer growth activities to the cells over other cells,
+causing autonomous proliferation, tissue invasion, and contribute to oncogenesis \cite{stratton2009cancer}.
+The main goal in cancer genome study has been to find driver mutations to understand the mechanism of cancer development.
 On the other hand,  even passenger mutations can also give us an important information, 
@@
 For example, C $>$ A mutations at CpG sites are abundant in lung cancers with smoking history,
-and these are caused by benzo(a)pyrene included in tobacco smoke \cite{pmid12379884}.
+and these are caused by benzo(a)pyrene included in tobacco smokes \cite{pmid12379884}.
 Also, C $>$ T and CC $>$ TT  mutations are abundant in ultraviolet-light-associated skin cancers, 
@@
 Due to limited sequencing throughput, 
-they mostly aggregated mutations collected from each individuals over the same cancer types focusing on a few cancer genes such as TP53
+they mostly aggregated mutations collected from each individual over the same cancer type focusing on a few cancer genes such as TP53
 where high mutation frequencies could be expected,
 and compared mutation patten profiles among different cancer types.
-However, many of those mutations in those cancer genes are considered to be driver mutations adding selective activities to the cells,
-which can give biased mutation profiles from those purely reflect driving forces of mutations.
-Furthermore, the mutation pattern profile not for each cancer type but for each individual could not be explored in the above approach.
+However, since many of the mutations in cancer genes are considered to be driver mutations adding selective activities to the cells,
+the resultant mutation profiles are considered to biased from those purely reflect driving forces of mutations.
+Furthermore, the mutation patterns not for each cancer type but for each individual could not be explored in the above approach.
 
@@
 A study using 21 breast cancer samples identified the association between C $>$ [AGT] mutations at TpC sites, 
-which is later proved to be caused by APOBEC proteins, and the novel phenomenon called {\it kataegis} \cite{pmid22608084}. 
-Moreover, the recent landmark study revealed a landscape of mutation signatures using from 7,034 primary cancer samples of 30 different classes \cite{pmid23945592}. 
-Furthermore, it is expected that detection of novel mutation signatures and associated mutagens can lead to identification of novel mutagens and prevention of cancer.
+which is later proved to be caused by APOBEC protein family, and a novel phenomenon called {\it kataegis} \cite{pmid22608084}. 
+Moreover, a landmark study using from 7,034 primary cancer samples of 30 different classes revealed an overview of mutation signatures in most cancer types \cite{pmid23945592}. 
+Now, there are great hopes that detection of novel mutation signatures and associated mutagens can lead to identification of novel mutagens and prevention of cancer.
 
 At the same time, for extracting prominent mutation signatures from vast amounts of somatic mutation data,
-there is an increasing need for novel latent variable modeling approaches.
+there is an increasing need for novel statistical approaches.
 Currently, only a few approaches have been proposed. 
@@
 
-One of the major problems of the previous approach is that 
+One of the major problems of the previous approaches is that 
 the number of parameters in the model grows exponentially 
@@
 despite there is a potential need to observe two base 5' and 3' positions to the mutated sites \cite{pmid9683596}.
-Also, it is often very hard to interpret and extract important characteristics from the estimate with high-dimensional parameters.
-Furthermore, considering that latent variable modeling have enormous number of researches in statistics and machine learning, 
-there is a great possibility that we can further elaborate statistical methods for mutational signature extraction problems.
-
-In this paper, we present a novel probabilistic approach.
-To avoid the problem of the exponential increase of the number of parameters,
-we assumed independence among contextual factors, which can give more interpretable and robust estimation.
-We demonstrate that independence assumption can robustly capture known mutational signatures with additional contextual informations.
-
-In addition, even though it has not been widely discussed, 
-the problem of identifying mutation signatures is closely related to "mixed-membership models'' in other fields, 
-such as ``population admixture'' models \cite{pmid10835412} in the analysis of population structure, 
-and the ``latent dirichlet allocation'' (LDA) model for document clustering \cite{Blei:2003}.
-In this paper, we explicitly show the relationships between the proposed model and the mixed-membership models and discuss the relationships,
-which we believe will be highly helpful for future elaboration of the statistical method for mutation signature extraction problems.
+Also, it is often very hard to interpret the estimates with high-dimensional parameters.
+
+In this paper, we present a novel probabilistic approach based on ``mixed membership models,'' that is widely used in other fields such as population genetics and machine learning.
+Even though it has not been widely discussed, there is close relationships between the mutation signature problems and mixed membership models,
+and we believe that being aware of those relationships will be very helpful for future elaboration of the statistical methods.
+Furthermore, to avoid the problem of the exponential increase of the number of parameters,
+we assume an independence among contextual factors, which can give more interpretable and robust estimation.
+We demonstrate that the independence assumption can robustly capture known mutational signatures with additional contextual informations.
+
+% In addition, even though it has not been widely discussed, 
+% the problem of identifying mutation signatures is closely related to "mixed-membership models,''
+% that is widely used in other field such as ``population admixture'' models \cite{pmid10835412} in the analysis of population structure, 
+% and the ``latent dirichlet allocation'' (LDA) model for document clustering \cite{Blei:2003}.
+% In this paper, we explicitly show the relationships between the proposed model and the mixed-membership models and discuss the relationships,
+% which we believe will be highly helpful for future elaboration of the statistical method for mutation signature extraction problems.
 
@@
 For example, if we want to take account of up to $n$ bases 5' and 3' to the mutated site 
-(which we call $-n$ site and $+n$ site, respectively, in this paper) as well as 6 substitution patterns, 
+(which we call $-n$ position and $+n$ position, respectively, in this paper) as well as 6 substitution patterns, 
 the number of parameters becomes $6 \times 4^{2n} - 1$. 
@@
 (where about 30,000 mutations are expected in entire genomic regions).
-Therefore, considering even $\pm 2$ sites and substitution patterns (1535 parameters)
+Therefore, considering even $\pm 2$ positions and substitution patterns (1535 parameters)
 is often difficult because of instability of parameter estimation.
 
-On the other hand, observing the mutation signatures observed in the previous studies, 
-many mutation signatures had clear characteristics such as:
+On the other hand, observing the mutation signatures detected in the previous studies, 
+many mutation signatures have clear characteristics such as:
 \begin{itemize}
@@
 A certain value of mutation features is often dominant in many known mutation signatures. 
-For example, the base T is highly dominant at the $-1$ site in the APOBEC signature, 
+For example, the base T is highly dominant at the $-1$ position in the APOBEC signature, 
 and C $>$ T substitution is highly dominant in the ultraviolet signature \cite{pmid23945592} (see Figure \ref{APOBEC_nature2013_sig}).
@@
 Proportional relationships among mutation features are often observed in many signatures as shown in the APOBEC signature, 
-where the frequencies of bases at the $+1$ site are largely proportional across 3 major substitution patterns C$>$T, C$>$G and C$>$A,
-and show consistent tendencies (A and T bases are more frequent than C and G bases).
+where the frequencies of bases at the $+1$ position are largely proportional across 3 major substitution patterns C$>$T, C$>$G and C$>$A,
+showing consistent tendencies  (A and T bases are more frequent than C and G bases).
 
@@
 In this paper, we propose to represent mutation signatures with independent multiple multi-nominal distributions
-on each mutation feature. 
-By doing this, we can reduce the number of parameters from exponential to linear with respect to the number of mutation features. 
-When we consider up to $\pm n$ sites as well as 6 substitution patterns, 
+on each mutation feature,
+and thus, we can reduce the number of parameters from exponential to linear with respect to the number of mutation features. 
+When we consider up to $\pm n$ positions as well as 6 substitution patterns, 
 the required number of parameters becomes $5 + 6n$. 
-For example, when considering $\pm 2$ sites and substitution patterns,
-the number of parameters becomes 17, which is far less compared to 1535 in the previous approach.
+For example, when considering $\pm 2$ positions and substitution patterns,
+the number of parameters becomes 17, which is far less compared to 1535 in the previous approach without the independence assumption.
 Furthermore, independence assumption enables us to come up with fairly interpretable representation for mutation signatures.
@@
 has been quite successfully utilized 
-even though it is never likely that transcription factors independently choose the bases of their binding sites. 
-This is probably because PSWM can capture important characteristics of transcription factor binding sites 
+even though it is never likely that transcription factors independently choose the nucleotides of their binding sites. 
+The success of the PSWM may be because it can capture important characteristics of transcription factor binding sites 
 in spite of the independence assumption,
@@
 As we demonstrate in later sections, 
-the proposed approach can robustly extract most of previously-collected mutation signatures even independence assumption is imposed.
+the proposed approach can robustly extract most of previously-collected mutation signatures even independence assumption is imposed,
+indicating the practical validity of the independence assumption.
 
@@
 on just two pattern TpCpT $>$ TpApT and TpCpG $>$ TpTpG, 
-and the frequencies of substitution patterns and $+1$ bases are not proportional unlike the APOBEC signature.
-However,  by using multiple mutation signatures, we can represent this phenomenon even with independence assumption.
+and the frequencies of substitution patterns and bases at the $+1$ positions are not proportional unlike the APOBEC signature.
+However, by using multiple mutation signatures, we can represent this phenomenon even under the independence assumption.
 \end{enumerate}
@@
 The proposed method has mainly two novelties compared to existing approaches.
-The first point is the assumption of independence, where we can identify mutation signatures as multiple multinomial distribution parameters on multiple contextual factors.
+The first point is, as noted above, the assumption of independence, where we can identify mutation signatures as multiple multinomial distribution parameters on multiple contextual factors.
 We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
-which is reminiscent of sequencing logos \cite{pmid2172928}. 
-This representation will help us to immediately identify which elements are strongly working on each contextual factors.
+which is somewhat reminiscent of sequencing logos \cite{pmid2172928}. 
+This representation will help us to immediately identify which elements are strongly expressed on each mutation factors.
 In this paper, we focus on substitution patterns, flanking bases (up to the $\pm 2$ positions) and transcribed directions as contextual factors of mutation patterns.
 
-The second novel point is that we adopt  the ``mixed membership model'' for the mutation signature problem.
-Briefly, the mixed membership model is basically clustering method, but allowing that each individual can have multiple clusters.
-The mixed membership model is widely used in many different fields such as population structure classification in population genetics and document classification problems in machine learning.
-This model is also shown to be closely related to nonnegative matrix factorization.
-
-Suppose there are in total $K$ source of mutations (or mutation signatures) in the cohort of interest.
+The second point is that we adopt the framework of ``mixed membership models'' for the mutation signature problem.
+The mixed memberships model are basically clustering methods, but allowing that each individual can have multiple clusters,
+and widely used in many different fields such as population structure classification in population genetics \cite{pmid10835412} and document classification problems in machine learning \cite{Blei:2003}.
+This models are also shown to be closely related to nonnegative matrix factorization \cite{ding2008equivalence}.
+
+The generative process of somatic mutations proposed proposed in this paper is summarized in the Figure \ref{methodOverview}.
+Suppose there are in total $K$ sources of somatic mutations (or mutation signatures) in the cohort of interest.
 For each cancer genome, somatic mutations are generated by one of the $K$ sources,
@@
 We call this parameter as ``membership parameters.''
-
-Here, by convention of the mixed membership model, we assume that each mutation for each cancer genome is generated by a two step model:
-First, one of the contributing mutation signature is chosen depending on the membership parameter of that cancer genome.
+Here, by convention of the mixed membership models, we assume that each mutation for each cancer genome is generated by a two step model.
+First, one of the contributing mutation signatures is chosen depending on the membership parameter of the corresponding cancer genome.
 Then, according to the probability distribution of the selected mutation signature, 
-mutation features such as substitution patterns and flanking base pairs are generated.
+mutation features such as substitution patterns and flanking bases are generated.
 Although the mutation signatures and membership parameters are unknown,
@@
 
-To account for the intrinsic sequential context, we add background signatures.
+To account for the intrinsic sequential context, we add background signatures calculated by compositions of short nucleotides sequences in the human genome.
 See the Method section for the mathematical details, 
-detailed relationships with mixed membership models and nonnegative matrix factorizations, how to choose the number of mutation signatures and so on.
+detailed relationships with mixed membership models and nonnegative matrix factorizations, parameter estimation details, how to choose the number of mutation signatures and so on.
 
@@
 First, we would like to investigate whether the proposed approach can extract ``true'' mutation signatures or not.
-Since we can not know the true mutation signatures in real biological data, we resorted to simulation studies.
-
-Here, we generated a set of mutations changing the number of cancer genomes (10, 25, 50, 100), 
+Since we cannot know the true mutation signatures in real biological data, we resorted to simulation studies.
+
+Here, we assumed substitution pattens and up to $\pm 2$ flanking bases as mutation features,
+and generated a set of mutations changing the number of cancer genomes (10, 25, 50, 100), 
 and the number of mutations for each cancer genome (10, 25, 50, 100, 250, 500, 1000).
 The number of mutation signatures were set to 5 including background mutation ratio.
-The mutation feature parameters and signature distribution parameters were generated by Dirichlet distribution,
+The mutation feature parameters and membership parameters were generated by Dirichlet distribution,
 
@@
 where the $\alpha$ and $\gamma$ represent the amounts of dispersion 
-for the mutation features parameters and signature distribution parameters, respectively.
-When these values are smaller, only a fewer components can have larger probability masses.
-and the rest will have much smaller masses.  
-When these are large, all the components tend to have evenly-distributed probability masses.
-
-As Figure \ref{sim_facet} shows, we could estimate the mutation signatures very accurately overall.
-(see Supplementary Figure 1 for an example).
+for the mutation signature parameters and membership parameters, respectively.
+When these values are smaller, only a fewer components can have larger probability masses and the rest will have much smaller masses.
+On the other hand, when these are large, all the components tend to have evenly-distributed probability masses.
+
+As Figure \ref{sim_facet} shows, we could estimate the mutation signatures very accurately overall (see Supplementary Figure 1 for an example).
 In most cases, the log-likelihood stopped increasing at five mutation signatures, 
@@
 As expected, as we increase the number of either cancer genomes or mutations, the accuracy of the estimates increased.
-Also, as we increase the value of $\alpha$, then the accuracy of the estimated signatures got worse.
+Also, as we increase the value of $\alpha$, then the accuracy of the estimated signatures became worse.
 However, even we increase the value of $\gamma$, the accuracy of the estimate signatures did not changed much.
 Therefore, the dispersion of mutation feature parameters influences the accuracy of mutation signatures
-more sensitively than signature distribution parameters.
+more sensitively than that of membership parameters.
 
@@
 
-In this section, we compare the ``full model'' and the ``independent model'' by examining mutation signatures obtained by the two approaches,
-and investigate the robustness by downsampling experiments.
+In this subsection, we compare  the ``independent model,'' where the independence assumption are imposed among mutation features
+and the ``full model,'' where no independence assumption are imposed (see Method section for the mathematical detail)
+by examining mutation signatures obtained by the two approaches and investigating the robustness by downsampling experiments.
 The dataset used here is a list of 14717 somatic substitutions 
@@
 
-We performed the proposed method considering substitution patterns, up to $\pm 2$ sites and strand directions as mutation features.
-For the full model, we assign one integer for each combination of these features, and thus $L = 1, \bm{M} = (3072)$.
-For the independent model, $L = 6, \bm{M} = (6, 4, 4, 4, 4, 2)$.
+We performed the proposed method considering substitution patterns, up to $\pm 2$ positions and strand directions as mutation features.
+For the full model, the number of parameters per mutation signature is 3071, where as it is 18 for the independent model.
 
 First, performing on various number of mutation signatures $K$ (including a background signature) on the independent model, 
-we observed how the detected mutation signatures change (See Supplementary Figure 3).
-For $K = 2$, a mutation signature which seem to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription direction) was observed (Supplementary Figure 3 (a)).
+we observed how the detected mutation signatures change (see Supplementary Figure 3).
+For $K = 2$, a mutation signature which seem to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription strand specificity) was observed (Supplementary Figure 3 (a)).
 For $K = 3$, an additional mutation signature corresponding APOBEC enzyme (C $>$ [AGT] at TpCpN sites) was observed
 in addition to the AA mutation signature (Figure \ref{UTUC:APOBEC_ind5_sig}, \ref{UTUC:AA_ind5_sig}, Supplementary Figure 3 (b)).
-For $K = 4$, an additional signature (T $>$ A at NpTpN sites with strong transcription direction) 
+For $K = 4$, an additional signature (T $>$ A at NpTpN sites with strong strand specificity) 
 that is somewhat similar to the AA signature was observed (Supplementary Figure 3 (c)). 
@@
 
-For the independent model, we could observed the depletion of G base at the $-2$ site,
+For the independent model, we could observed the depletion of G base at the $-2$ position,
 which is consistent to the previous study \cite{pmid23318258} and the result in the next subsection.
-On the other hand, for the full representation model, this tendency was rather mild (Figure \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}, Supplementary Figure 6).
-Inferred AA mutation signature had no clear characteristics at two bases 5' and 3' to the mutated site compared to the APOBEC mutation signature.
+On the other hand, for the full model, this tendency was rather mild (Figure \ref{UTUC:APOBEC_full5_sig}, \ref{UTUC:AA_full5_sig}, Supplementary Figure 6).
+Inferred AA mutation signature had no clear characteristics at the $-2$ position compared to the APOBEC mutation signature.
 
@@
 and compared obtained signatures with the gold standard.
-For measuring the deviations of the mutation signature, the cosine similarity was used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
+For measuring the deviations of the mutation signature, the cosine similarity was used on the dimensional vector space full model,
 so that the comparison between the full model and the independent model become possible.
@@
 
+% For measuring the deviations of the mutation signature, the cosine similarity was used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,
+
 As the Figure \ref{UTUC:APOBEC_downsampling} and \ref{UTUC:AA_downsampling} shows, 
-the independent representation could recover the original signatures even when, e.g, about 90 \% of the data was removed.
+the independent model could mostly recover the original signatures even when, e.g, about 90 \% of the data was removed,
+whereas the full model could not recover, especially, the APOBEC signature.
 In this experiment, the AA signature was more robust than the APOBEC signature.
@@
 are arguably associated with deregulated activity of the error-prone polymerase Pol $\epsilon$.
-The proposed method uses two signatures for representing the signature for the Pol $\epsilon$ dysfunction, 
-while was represented by one signature in the previous study.
+In the proposed method, the signature for the Pol $\epsilon$ dysfunction was represented by two signatures
+while it was represented by one signature in the previous study.
 Interestingly, the signature 1 shows transcription biases whereas the signature 8 does not.
-This phenomena are consistent for both colorectal and uterus cancers (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
-C $>$A at TpCpT mutations were shown to be enriched in leading strands of replication forks.
-whereas, there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
-Although the replication strand bias is different from the transcription strand bias, these two biases may have strong correlation
-through the fact that replication origins prefer transcription start sites \cite{pmid23187890}.
-Therefore, the transcription strand bias observed in the signature 1 may actually come from the replication strand biases.
-
-Furthermore, both the signature 1 and 8 signatures shows strong abundance of the T base at the $-2$ position,
-and the signature 1 also shows a little abundance of the T base at the $+2$ position.
- (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}, Supplementary Figure 7(b, c)).
+This phenomena are consistent for both colorectal and uterus cancers (Supplementary Figures 7(c, d)).
+C $>$A at TpCpT mutations were shown to be enriched in leading strands of replication forks,
+whereas there is no mention about the replication strand-specificity of C $>$ T at TpCpG mutations \cite{pmid25228659}.
+Although the replication strand bias is a different phenomenon from the transcription strand bias, these two biases may have strong correlation
+through the fact that replication origins prefer transcription start sites \cite{pmid23187890},
+and the transcription strand bias observed in the signature 1 may actually come from the replication strand biases.
+
+Furthermore, both the signature 1 and 8 signatures show higher frequency of the T base at the $-2$ position,
+and the signature 1 also shows a little higher frequency of the T base at the $+2$ position.
+ (Figures \ref{POLE1_two5prime} and \ref{POLE2_two5prime}, Supplementary Figure 7(c, d)).
 In the previous study of Pol $\epsilon$  \cite{pmid25228659}, 
-a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
-In fact, the pattern of this mutation is C $>$ T at TpTpCpGpA, that is the most likely patten with the signature 8.
-This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.
-
-The signature 2 (C $>$ A at [CT]pCpT) observed in low grade gliomas
+a nonsense mutation R23X of TP53 was shown to be enriched in cancers with Pol $\epsilon$ defects.
+In fact, the pattern of this mutation is C $>$ T at TpTpCpGpA, that is the most consistent patten with the signature 8.
+Therefore, observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.
+
+The signature 2 (C $>$ A at [CT]pCpT) observed solely in low grade gliomas
 can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 8 (a)). 
@@
 The signature 4 (C $>$ A at CpCpG) observed in kidney clear cell carcinomas, lung adenocarcinomas and melanomas 
-shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 10 (a,b,c)).
+shows remarkably high ratio of the base C at the $-2$ position (see Supplementary Figure 10 (a,b,c)).
 However, for each cancer type, only a few samples support this signature (see Supplementary Figure 10 (d, e, f)).
@@
 that could not be validated in the previous study,
-meaning that significant ratios of somatic mutations corresponding to that signature could not be validated by re-sequencing or visual inspection of BAM files using genomic viewers.
-Actually, the pattern of this signature largely matches to that of the putative artifact suggested in \cite{pmid23303777}, 
-that is caused by oxidation of DNA during acoustic shearings, 
+meaning that most of somatic mutations corresponding to that signature could not be validated by re-sequencing or visual inspection of BAM files using genomic viewers.
+Actually, the pattern of this signature largely matches to that of the putative artifact caused by oxidation of DNA during acoustic shearings \cite{pmid23303777}, 
 and we guess that this signature is the same as what is observed in that study.
-Although this signature may be from the artifacts, 
-knowing higher resolution of the artifact signature is very helpful to accurately remove false positive mutations.
+Although this signature may come from the artifacts, 
+collecting artifact signatures with a higher resolution will be very helpful to accurately remove false positive mutations.
 
 The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
-The composition of nucleotides of estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
-and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}, Supplementary Figure 7(a)).
+The composition of nucleotides for estimated signatures in 12 different cancer types is highly consistent (Supplementary Figure 7(a)) except for the one in B-cell lymphoma,
+and almost all of them showed slightly higher abundance of the A and T bases and depletion of the G base at the $-2$ position (Figure \ref{APOBEC_two5prime}, Supplementary Figure 7(a)).
 This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
@@
 The estimated transcribed strand specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
-The Signature 15 and 16 may also be delated to the APOBEC signatures although the estimated forms are slightly different,
+The Signature 15 and 16 may also be related to the APOBEC signatures although the estimated forms are slightly different,
 and are not merged to the cluster of signature 13 by the specified criteria.
@@
 The signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
-The estimated signature in each cancer type shows higher mutation prevalence on the template strand, which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
+The estimated signature in each cancer type shows higher mutation prevalence on the template strand (Supplementary Figure 7(b)), 
+which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
 The signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas matches the pattern of that detected in the same cancer type in the previous study.
@@
 The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, and are probably associated with a chemotherapy drug, temozolomide.
-The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
-Observed consistent strand specificities (Figure \ref{UV_sig_list}) matches to the previous results \cite{pmid23318258}.   
-Also, higher frequencies of T base at the $\pm 2$ position are consistently observed (Figure \ref{UV_sig_list}, Supplementary Figure 7(d)).
+The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light,
+and consistent strand specificities among the two cancer types (Supplementary Figure 7(e)) matches to the previous results \cite{pmid23318258}.   
+In addition, higher frequencies of the T base at the $\pm 2$ position are consistently observed (Figure \ref{UV_two5prime}, Supplementary Figure 7(e)).
 The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
 and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear,
-and the abundance of G base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_sig_list}, Supplementary Figure 7(e)).
+and the abundance of G base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_two5prime}, Supplementary Figure 7(f)).
 
@@
 and a sign of transcription-coupled nucleotide excision repairs.
-Actually, this signature does not match to any of signatures in the previous paper \cite{pmid23318258}.
+Actually, this signature does not match to any of signatures in the previous paper \cite{pmid23318258} and possibly a novel signature.
 This may be because the mutation ratio of thyroid cancer is low and the signature is too week to detect for the previous methods.
 %% download the TCGA data and check the visualization of the mutation position for this data.
+
 The unknown signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively,
@@
 In summary, the propose method can capture many of the prominent mutation signatures observed in the previous study.
-Also, the propose method detected several novel characteristics at the $\pm 2$ position that is consistent across multiple cancer types,
+Also, the propose method detected several novel characteristics at the $\pm 2$ position that are consistent across multiple cancer types,
 suggesting the existence of more hidden characteristics at the distant position from the mutated site.
@@
 In this paper, we proposed a novel method for extracting mutation signatures from a set of somatic point substitutions.
-The proposed model utilizes the framework of mixed membership models, 
+The proposed model utilizes the framework of mixed membership models
 that have been deeply investigated in statistical genetics and machine learning fields.
@@
 
-Even though the proposed method impose an independence assumptions among each contextual factor,
+Even though the proposed method impose an independence assumption among contextual factors,
 the proposed method could capture many of the previously detected mutation signatures.
-Also, we have shown that the proposed method with the independence assumption is gives more robust estimates compared to that without the independence assumption.
+Also, we have shown that the proposed method with the independence assumption gives more robust estimates compared to that without the independence assumption.
 In addition, the estimated mutation signatures with the independence assumption can be relatively easily visualized in an interpretable manner.
-As we have discussed, the independence assumption may not be completely consistent with the nature of DNA damages and repairs.
-Still, we have demonstrated that the model with the independence assumption is ``practically'' very helpful.
+The independence assumption may not be completely consistent with the nature of DNA damages and repairs.
+Still, we would like to argue that the model with the independence assumption is ``practically'' very helpful.
 
@@
 However, there are a variety of mutations in cancer genomes, 
-such as insertions, deletions, double nucleotides substitutions, structural variations and copy number alterations.
+such as insertions, deletions, double nucleotides substitutions, structural variations and copy number alterations \cite{meyerson2010advances, helleday2014mechanisms}.
 Our framework can potentially treat those mutations by considering appropriate mutation features.
@@
 Finally, mutation ratios have been shown to be closely related to a number of genomic and epigenetic factors,
-such as GC contents, repeat sequences, chromatin accessibility and modifications, and replication timing and so on \cite{pmid22820252, pmid21953857, pmid23422670, pmid23770567}.
-Also, a recent study has shown that the epigenetic information of the cell type of origin of the corresponding tumors is the most predictive \cite{pmid25693567}.
-As a number of epigenetic data from various type of cells accumulate,
-one of the next main topic will be to integrating those epigenetic factors into the mutation signature 
+such as GC contents, repeat sequences, chromatin accessibility and modifications, replication timing and so on \cite{pmid22820252, pmid21953857, pmid23422670, pmid23770567}.
+Also, a recent study has shown that the epigenetic information of the cell type of origin of the corresponding tumors is the most predictive \cite{pmid25693567} for regional mutation ratios.
+Nowadays, a number of epigenetic data from various type of cells have been accumulating,
+and one of the next main topic will be to integrate those epigenetic factors into the mutation signature 
 and to understand how these epigenetic factors influence on DNA damage and repair mechanisms.
@@
 
-Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
-We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
-which is reminiscent of sequencing logos \cite{pmid2172928}.
+% Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
+% We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
+% which is reminiscent of sequencing logos \cite{pmid2172928}.
 
@@
 We call this representation and mutational process modeling as ``full representation,'' and ``full model.''
-In this case, when taking account of 6 substitution patterns and $\pm 2$ sites,
+In this case, when taking account of 6 substitution patterns and $\pm 2$ positions,
 $\bm{M} = (1536)$.
-On the other hand, the usual representation, where each mutation feature has its corresponding value and individual multinomial distribution,
+On the other hand, the representation where each mutation feature has its corresponding value and individual multinomial distribution
 is called as ``independent representation,'' and ``independent model.''
@@
 (e.g., a situation where C $>$ A is frequent at ApCpG sites and C $>$ T is frequent at TpCpA sites) with one signature.
-However, when many mutation contextual factors are taken into account and the number of free parameters get huge,
+However, when many mutation contextual factors are taken into account and the number of free parameters becomes huge,
 estimated results tend to be unstable and unreliable.
@@
 EM-algorithm (or its variant, called tempered EM algorithm) have been adopted in classical topic models 
-for document classification (\cite{Hofmann:1999}) and population structure estimation (\cite{tang2005estimation}).
+for document classification \cite{Hofmann:1999} and population structure estimation \cite{tang2005estimation}.
 In the population structure estimation problem, \cite{pmid19648217} has proposed 
@@
 Finally, utilizing other metrics than Euclidean distance and Kullback-Leibler Divergence might be worth investigating.
-For nonnegative matrix factorization, 
-the metric called $\beta$ divergence \cite{basu1998robust, eguchi2001robustifying}, 
-has been successfully utilized in many research fields (reviewed in \cite{fevotte2011algorithms}).
+For nonnegative matrix factorization, the metric called $\beta$ divergence \cite{basu1998robust, eguchi2001robustifying}, 
+has been successfully utilized in many research fields (reviewed in \cite{fevotte2011algorithms}),
+and evaluating these metrics on mutation signature problems will be an interesting future tasks.
 
@@
 
-The first author would like to thank to Dr. Daichi Mochihashi for helpful discussion and comments on earlier version on the proposed method.
-Many of the contents in this paper is deeply influenced by what was going on at Matthew Stephens Laboratory,
+The first author would like to thank to Dr. Daichi Mochihashi for helpful discussion and comments on the earlier version on the proposed method.
+He also thanks to Dr. Hiromichi Suzuki for helpful feedbacks for R packages.
+This paper is a product deeply influenced by what was going on at Matthew Stephens Laboratory,
 when the first author stayed at the University of Chicago as a visiting scholar.
-The first author would like to thank to the members in Matthew Stephens and John Novembre Laboratory, especially Dr. John Novembre,
-Dr. Jacob Degner for helpful discussion and comments.
+The first author would like to thank to Dr. John Novembre,
+Dr. Jacob Degner and all the members in Matthew Stephens and John Novembre Laboratory for helpful discussion and comments.
 
@@
 
+\begin{figure*}
+\includegraphics[width=450pt]{methodOverview.pdf}
+\caption{Suppose there are three types of mutation sources (mutation signatures) such as ultraviolet, tobacco smoking chemicals and transcription coupled repairs.
+Each cancer genome has ratios showing which types of mutation sources are contributing to its mutations (membership parameters).
+The generative model of the pattern of each mutation is:
+first, one of the mutation signatures is chosen according to the membership parameter.
+Second, according to the selected mutation signature, mutation features such as substitution pattens and flanking bases are generated 
+by the corresponding multinomial distributions for that mutation signature.
+}
+\label{methodOverview}
+\end{figure*}
+
+\clearpage
+
 \begin{figure*}[b]
@@
 where the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
+First, the mutation signatures were estimated with respect to each cancer type,
+then the similar signatures (the sum of the square distances between the parameters is below 0.6) emerged across different cancer types were merged.
 }
@@
 
+% \begin{figure*}[b]
+
 \begin{figure*}[b]
 
-\subfigure[APOBEC signatures obtained in each cancer type]{%
-  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
-  \label{APOBEC_sig_list}}
-
-\subfigure[Smoking signature  in each cancer type]{%
-  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
-  \label{LUNG_sig_list}}
-
-\subfigure[POLE1 signatures in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-  \label{POLE1_sig_list}}
+\subfigure[APOBEC signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=14.5cm,height=6cm]{APOBEC_two5prime.eps}
+   \label{APOBEC_two5prime}}
+
+\subfigure[POLE1 signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{POLE1_two5prime.eps}
+  \label{POLE1_two5prime}}
 \quad
-\subfigure[POLE2 signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-  \label{POLE2_sig_list}}
-
-\subfigure[UV signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-  \label{UV_sig_list}}
+\subfigure[POLE2 signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{POLE2_two5prime.eps}
+  \label{POLE2_two5prime}}
+
+\subfigure[UV signature intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{UV_two5prime.eps}
+  \label{UV_two5prime}}
 \quad
-\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-  \label{LMST_sig_list}}
-
-% \subfigure[Signature intensities at two 5' to the mutated site]{%
-% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
- % \label{APOBEC_twofive}}
-
-\caption{(a) APOBEC signatures obtained in 11 cancer types, 
-(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
-(f) the frequencies of bases at two 5' to the mutated site. 
-}
-\label{nature2013_apobec}
+\subfigure[The signature 11 intensities at two 5' to the mutated site]{%
+  \includegraphics[width=4.5cm,height=6cm]{LMST_two5prime.eps}
+  \label{LMST_two5prime}}
+
+\caption{The estimated frequencies of bases at two 5' to the mutated site for each cancer type.
+The heights of bar show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
+The error bars show standard errors estimated by bootstrap.}
+\label{two5prime}
+
 \end{figure*}

### 2015-03-19 supp/supp.tex
a number of modifications such as addition of method overview figures and many corrections of manuscript

--- 2015-03-14 supp/supp.tex
@@
 For solving NMF, 
-the previous study \cite{pmid23318258} used the following updating rule \cite{Lee00algorithmsfor}:
+the previous study (Lee et al. 2000) used the following updating rule:
 \begin{equation*}
@@
 
-On the other hand, there is another type of updating rule \cite{Lee00algorithmsfor}:
+On the other hand, there is another type of updating rule:
 \begin{equation*}
@@
 
-% \subfigure[POLE1 signatures in each cancer type]{%
-%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-%   \label{POLE1_sig_list}}
+\subfigure[APOBEC signatures obtained in each cancer type]{%
+  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
+  \label{APOBEC_sig_list}}
 
-% \subfigure[POLE2 signature in each cancer type]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-%   \label{POLE2_sig_list}}
+\subfigure[Smoking signature  in each cancer type]{%
+  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
+  \label{LUNG_sig_list}}
 
-% \subfigure[UV signature in each cancer type]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-%   \label{UV_sig_list}}
+\subfigure[POLE1 signatures in each cancer type]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
+  \label{POLE1_sig_list}}
+\quad
+\subfigure[POLE2 signature in each cancer type]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
+  \label{POLE2_sig_list}}
 
-% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-%   \label{LMST_sig_list}}
+\subfigure[UV signature in each cancer type]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
+  \label{UV_sig_list}}
+\quad
+\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
+  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
+  \label{LMST_sig_list}}
 
-% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
-% \label{nature2013_sig_others}
-% \end{figure*}
+% \subfigure[Signature intensities at two 5' to the mutated site]{%
+% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
+ % \label{APOBEC_twofive}}
+
+\caption{(a) APOBEC signatures obtained in 11 cancer types, 
+(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
+(f) the frequencies of bases at two 5' to the mutated site. 
+}
+\label{nature2013_apobec}
+\end{figure*}
 
@@
 
-\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in \cite{pmid23945592})]{%
+\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in Alexandrov et al. (2013)]{%
   \includegraphics[height=3cm,width=12cm,clip]{example96_Glioma-Low-Grade.eps}

### 2015-03-19 figure/script/AlexandrovEtAl_figures_dir.R

@@ -295,7 +295,7 @@ for (i in 1:length(importantSigs)) {
   
   
   
-  outputName <- paste("../../manuscript/signatureList_", names(importantSigs)[i], ".eps", sep="");
+  outputName <- paste("../../supp/signatureList_", names(importantSigs)[i], ".eps", sep="");
   
   if (length(types) > 4) {
     dev.copy2eps(file=outputName, height = ceiling(length(types) / 4) * 2.4, width = 15);
@@ -309,28 +309,41 @@ par(.pardefault);
 
 
 ###
+# for method overview figure
 
-importantSigs <- c(APOBEC_ind, UV_ind, LMST_ind);
-names(importantSigs) <- c("APOBEC", "UV", "LMST");
+for (k in 1:length(Fs)) {
+  if (Fs[[k]][1,5] > 0.6 & Fs[[k]][3,1] > 0.25 & Fs[[k]][6,2] > 0.7) {
+    TCR_ind <- k;
+  }
+}
+  
+  
+importantSigs2 <- c(LUNG1_ind, UV_ind, TCR_ind);
+names(importantSigs2) <- c("LUNG1", "UV", "TCR");
 
 par(mar=c(0, 0, 0, 0));
 par(xaxs = "i", yaxs = "i");
 
-for (i in 1:length(importantSigs)) {
+for (i in 1:length(importantSigs2)) {
   
 
-  F <- Fs[[as.numeric(importantSigs[i])]];
+  F <- Fs[[as.numeric(importantSigs2[i])]];
     
   plot.new();
   plot.window(xlim=c(-0.3, 6.3), ylim=c(-0.3, 3.8));
   
   # the setting of charSize is temporary... why the size changes by the number of rows.....
   visPMS_ind(F, numBases = 5, trDir = TRUE, charSize = 0);
-       
-  
-  outputName <- paste("../methodOverview/signatureList_", names(importantSigs)[i], ".eps", sep="");
+  outputName <- paste("../methodOverview/signatureList_", names(importantSigs2)[i], "_nonBase.eps", sep="");
   dev.copy2eps(file=outputName, height = 2.4, width = 3.75, pointsize = 1e-10);
+  
+  plot.new();
+  plot.window(xlim=c(-0.3, 6.3), ylim=c(-0.3, 3.8));
 
+  # the setting of charSize is temporary... why the size changes by the number of rows.....
+  visPMS_ind(F, numBases = 5, trDir = TRUE, charSize = 0.6 + 0.1333);
[…]

### 2015-03-19 supp/supp.tex
modification of the table of mutation signature nuns…

@@
 \bibliography{document}
+
+\clearpage
+
+\begin{table}
+\caption{The number of mutation signatures selected for each cancer type.
+}
+\begin{center}
+\begin{tabular}{|c|c|} \hline
+cancer type & \# mutation signature \\ \hline
+AML & 2 \\ \hline
+ALL & 3 \\ \hline
+Bladder & 3 \\ \hline
+Breast & 4 \\ \hline
+Cervix & 3 \\ \hline
+CLL & 3 \\ \hline
+Colorectum & 5 \\ \hline
+Esophageal & 4 \\ \hline
+Glioblastoma & 3 \\ \hline
+Glioma-Low-Grade & 3 \\ \hline
+Head-and-Neck & 5 \\ \hline
+Kidney-Chromophobe & 3 \\ \hline
+Kidney-Clear-Cell & 3 \\ \hline
+Kidney-Papillary & 3 \\ \hline
+Liver & 3 \\ \hline
+Lung-Adeno & 4 \\ \hline
+Lung-Small-Cell & 4 \\ \hline
+Lung-Squamous & 3 \\ \hline
+Lymphoma-B-Cell & 4 \\ \hline
+Medulloblastoma & 3 \\ \hline
+Melanoma & 4 \\ \hline
+Myeloma & 3 \\ \hline
+Neuroblastoma & 3 \\ \hline
+Ovary & 3 \\ \hline
+Pancreas & 4 \\ \hline
+Pilocytic-Astrocytoma & 4 \\ \hline
+Prostate & 2 \\ \hline
+Stomach & 5 \\ \hline
+Thyroid & 4 \\ \hline
+Uterus & 4 \\ \hline
+\end{tabular}
+\end{center}
+\label{tab_rep}
+\end{table}
 

### 2015-05-18 manuscript20150320/plos_template.tex
backup manuscript as of 20150320

--- 2015-03-19 manuscript/plos_template.tex
@@
 By changing the number of mutation signatures and examining values of the log-likelihood and bootstrap errors as well as correlation of membership parameters,
-we have chosen the number of mutation signatures for each cancer type.
+we have chosen the number of mutation signatures for each cancer type
+(see Supplementary Table 1 for the number of mutation signatures chosen for each cancer type).
 Finally, similar mutation signatures observed across cancer types were clustered by Frobenius Distance.

### 2015-05-18 manuscript20150515/plos_template.tex
backup manuscript as of 20150515

% Template for PLoS
% Version 2.0 July 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% Be advised that this is merely a template 
% designed to facilitate accurate translation of manuscript content 
% into our production files. 
%
% This template contains extensive comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template 
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication and enters production, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% DO NOT ADD EXTRA PACKAGES TO THIS TEMPLATE unless absolutely necessary.
% Packages included in this template are intentionally
% limited and basic in order to reduce the possibility
% of issues during our production process.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text. For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
%
% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
% For bolding characters in mathmode, please use $\mathbf{}$ 
%
% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
%
% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
%
% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
%
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\usepackage{bm}
\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
% \DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
% \usepackage{setspace} 
% \doublespacing

\usepackage{graphicx}
\usepackage{subfigure}

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}

%% Include all macros below. Please limit the use of macros.

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{A simple model-based approach to inferring and visualizing cancer mutation signatures}
}
% Insert Author names, affiliations and corresponding author email.
\\
Yuichi Shiraishi$^{1,\ast}$, 
Georg Tremmel$^{1}$, 
Satoru Miyano$^{1}$, 
Matthew Stephens$^{2,3, \ast}$
\\
\bf{1} Laboratory of DNA Information Analysis, Human Genome Center, Institute of Medical Science, The University of Tokyo, Tokyo, Japan
\\
\bf{2} Dept. of Human Genetics, University of Chicago, Chicago, Illinois, United States of America
\\
\bf{3} Dept. of Statistics, University of Chicago, Chicago, Illinois, United States of America
% \bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: yshira@hgc.jp; mstephens@uchicago.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

Recent advances in sequencing technologies have enabled
the production of
massive amounts of data on somatic mutations from cancer genomes. These data have led to the detection of characteristic patterns of somatic mutations or ``mutation signatures'' at an unprecedented resolution, with the
potential for new insights into
the causes and mechanisms of tumorigenesis.

Here we present new methods for modelling, identifying and visualizing such mutation signatures. Our methods
greatly simplify mutation signature models compared with existing approaches, reducing the number of parameters by orders of magnitude even while increasing the contextual factors (e.g.~number of flanking bases) that are accounted for. This improves both sensitivity and robustness of inferred signatures. We also provide a new intuitive way to visualize the signatures, analogous to the use of sequence logos to visualize transcription factor binding sites.

We illustrate our new method on somatic mutation data from urothelial carcinoma of the upper urinary tract, and a
larger dataset from 30 diverse cancer types.
The results illustrate several important features
of our methods, including the ability of our new visualization
tool to clearly highlight the key features of each signature,
the improved robustness of signature inferences from small sample sizes, and more detailed inference of signature characteristics such as strand biases and sequence context effects at the base two positions 5' to the mutated site.

%This paragraph is nice, but could be removed if short on words?
Our work draws on ideas from several tools that are widely used in other contexts. The model for each signature is connected with position weight matrix models for transcription factor binding, and the overall framework is based on probabilistic models that are closely
connected with ``mixed-membership models'' which are widely used in population genetic admixture analysis, and in machine learning for document clustering. We argue that recognizing these relationships should help improve
understanding of mutation signature extraction problems,
and suggests ways to further improve the statistical methods.

Our methods are implemented in an R package 
pmsignature (\url{https://github.com/friend1ws/pmsignature})
and a web application available at \url{https://friend1ws.shinyapps.io/pmsignature_shiny/}.

% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}

Somatic (non-inherited) mutations are acquired throughout our lives in cells throughout our body. These mutations
can be caused, for example, by DNA replication errors or exposure to environmental mutagens such as tobacco smoke. 
Some of these mutations can lead to cancer.

Different cancers, and even different instances of the same
cancer, can show different distinctive patterns of somatic mutations. These distinctive patterns 
have become known as ``mutation signatures".
For example, C $>$ A mutations are frequent in lung caners
whereas C $>$ T and CC $>$ TT mutations are frequent in skin cancers. Each mutation signature may be associated with a specific kind of carcinogen, such as tobacco smoke or ultraviolet light. Identifying mutation signatures
therefore has the potential to identify new carcinogens, and
yield new insights into the mechanisms and causes of cancer,

In this paper, we introduce new statistical tools
for tackling this important problem. These tools provide more robust and interpretable mutation signatures compared to previous approaches, as we demonstrate by applying them to large-scale cancer genomic data.

\section*{Introduction}

Cancer is a genomic disease. 
As we lead a life, DNA within our cells acquires random somatic mutations,
mainly caused by DNA replication errors and exposures to mutagens such as chemical substances, radioactivities and inflammatory reactions.  
Although most mutations are harmless (called ``passenger mutations''), 
a small portions of mutations at some specific sites in cancer genes (``driver mutations'') 
%confer growth activities to the cells over other cells,
affect cell growth,
causing autonomous proliferation, tissue invasion, and contributing to oncogenesis \cite{stratton2009cancer}.
Cancer genome studies typically focus on identifying driver mutations, 
to help understand the mechanism of cancer development.
However, passenger mutations can also yield important information, 
because they often show patterns (``mutation signatures'') which 
can provide insights into the forces that cause somatic mutations.
%Classical analyses of mutation patterns have revealed a number of relationships between mutation patterns and carcinogens.
For example, classical studies of mutation patterns
revealed that C $>$ A mutations are abundant in lung cancers in patients with smoking history,
and these are caused by benzo(a)pyrene included in tobacco smokes \cite{pmid12379884}.
Also, C $>$ T and CC $>$ TT  mutations are abundant in ultraviolet-light-associated skin cancers, 
and these are caused by pyrimidine dimers as a result of ultraviolet radiation \cite{pmid15748635}.

%There were several problems in these classical studies \cite{pmid12379884, pmid15748635}.
The potential for classical studies to yield insights into somatic mutation processes
were limited in several ways.
Due to limited sequencing throughputs, most classical studies 
focused on a few cancer genes, such as TP53,
where high mutation frequencies could be expected. They then
contrasted mutation pattern profiles among different cancer types,
aggregating mutations across multiple individuals within the same cancer type to
yield sufficient mutations for analysis.
However, since many of the mutations in cancer genes are driver mutations causing cell proliferation,
the resultant mutation profiles are a biased representation of the underlying mutation process.
Furthermore, the paucity of mutation data made it effectively impossible to assess variation in mutation patterns among individuals.

Recent advances in high-throughput sequencing provide
new opportunities to investigate sample-by-sample mutation signatures in an unbiased way
using genome-wide somatic mutation data. For example, a large-scale
study using 21 breast cancer samples identified an association of C $>$ [AGT] mutations at TpC sites, 
%Yuichi - is the change association -> excess ok?
%Matthew - actually, what they showed is the relationship between the C>[AGT] mutation at TpC sites and kataegis.
which was later proved to be caused by APOBEC protein family \cite{pmid23389445, pmid23852168, pmid23852170}, 
and a novel phenomenon called {\it kataegis} \cite{pmid22608084}. 
Moreover, a landmark study of 7,034 primary cancer samples, representing 30 different cancer classes, has provided the first large-scale overview of mutation signatures across a large number of cancer types \cite{pmid23945592}. This has lead to great hopes that detection of novel mutation signatures and associated mutagens can lead to identification of novel mutagens and prevention of cancer.

To make the most of these opportunities requires the development
of efficient and effective statistical methods for analyzing
mutation signatures in vast amounts of somatic mutation data.
Current statistical approaches \cite{pmid23318258,pmid23628380}, 
%based on nonnegative matrix factorization (NMF, \cite{pmid10548103}) and Poisson models
are an excellent starting point, 
and have helped generate the new insights noted above.
However, we argue here that these existing methods 
have two important limitations, caused by the fact that they
use an unconstrained model for each ``mutation signature."
First, although using an unconstrained model might appear to be a good thing in terms of flexibility, 
in practice it can actually reduce flexibility, because
the price of using an unconstrained model is that one must limit 
the domain of mutation signatures considered. 
For example, most recent analyses of mutation signatures
consider only the immediate flanking 5' and 3' bases of each substitution to be part of
the signature, even though it is known that more distal bases -- and particularly the
next flanking base on each side -- can contain important contextual information
\cite{pmid9683596}. These recent analyses take this approach because, in the unconstrained model,
incorporating the more distal bases into the signature very substantially increases the number of parameters, making estimated mutation signatures unstable.
Secondly, and just as important, the unconstrained model means
that each signature is a probability distribution in a high-dimensional
parameter space, which can make signatures difficult to interpret.

In this paper, we present a novel probabilistic approach 
to mutation signature modelling that addresses these limitations.
In brief, we first simplify the modelling of mutation signatures by decomposing them into separate ``mutation features".
For example, the substitution type is one feature; flanking bases are each another feature. We then exploit this 
decomposition by using a probabilistic model for signatures that assumes independence across features. This approach substantially reduces the number of parameters associated with
each signature, greatly facilitating the incorporation of additional relevant sequence context. 
For example, our approach can
incorporate both the two bases
3' and 5' of the substitution, and transcription strand biases using only 18 parameters
per signature, compared with 3071 parameters per signature with current approaches. We demonstrate the benefits of this simplification in data analyses. These benefits include: more stable estimation of signatures from smaller samples, 
refinement of the detail and resolution of many mutation signatures, and, possibly, identification of novel signatures.

Assuming independence among features in a signature may initially seem unnatural. However, its use here is analogous  ``position weight matrix models" which have been highly
successful 
for modelling transcription factor binding motifs.
Indeed, an important second contribution of
our paper is to provide intuitive visual representations for mutation signatures,
analogous to the ``sequence logos" used for visualizing binding motifs.
Finally, we also highlight the close connection
between mutation signature models and the ``mixed-membership models'',
also known as``admixture models" \cite{pmid10835412} or ``latent Dirichlet allocation''
models \cite{Blei:2003} that are widely used in population genetics 
and document clustering applications.
These connections should be helpful for future elaboration of 
computational and statistical methods for cancer mutation signature detection.

%The second point is that we adopt the framework of ``mixed membership models'' for the mutation signature problem. The mixed memberships model are basically clustering methods, but allowing that each individual can have multiple clusters, and widely used in many different fields such as population structure classification in population genetics \cite{pmid10835412} and document classification problems in machine learning \cite{Blei:2003}. This models are also shown to be closely related to nonnegative matrix factorization \cite{ding2008equivalence}.

% In addition, even though it has not been widely discussed, 
% the problem of identifying mutation signatures is closely related to "mixed-membership models,''
% that is widely used in other field such as ``population admixture'' models \cite{pmid10835412} in the analysis of population structure, 
% and the ``latent dirichlet allocation'' (LDA) model for document clustering \cite{Blei:2003}.
% In this paper, we explicitly show the relationships between the proposed model and the mixed-membership models and discuss the relationships,
% which we believe will be highly helpful for future elaboration of the statistical method for mutation signature extraction problems.

Software implementing the proposed methods is available in an R package 
{\bf pmsignature} ({\bf p}robabilistic {\bf m}utation signature),
at \url{https://github.com/friend1ws/pmsignature}).
The core part of the estimation process is implemented in C++ by way of the Rcpp package \cite{eddelbuettel2011rcpp},
which enables handling millions of somatic mutations from thousands of cancer genomes using a standard desktop computer.
In addition, a web-based application of our method is available at 
\url{https://friend1ws.shinyapps.io/pmsignature_shiny/}

% Results and Discussion can be combined.
\section*{Methods Overview}

\subsection*{New model for mutation signatures}

The term ``mutation signature'' is used to describe a characteristic mutational pattern observed in cancer genomes. Such patterns are often related to carcinogens (e.g., frequent C $>$ A mutations in lung cancers with smoking histories).

What constitutes a mutational pattern varies among papers.
The simplest approach is to consider 6 possible mutation patterns, corresponding to
6 possible substitution patterns (C$>$A, C$>$G, C$>$T, T$>$A, T$>$C, T$>$G; the original base is often fixed to C or T to remove redundancy of taking complementary strands).
However, in practice we know that DNA context of a substitution is often important, and so it is common to go the next level of complexity, and include the immediate 5' and 3' flanking bases in the mutation pattern.
This results in 96 (6 $\times$ 4 $\times$ 4) patterns. 
Further incorporating the strand (plus or minus) of each substitution extends this to 192 patterns \cite{pmid23945592, pmid23318258}.

% In this paper, we call the factors composing mutation patterns (such as substitution patterns, adjacent bases and transcriptional strand) 
% as ``mutation features.''

Mathematically, mutation signatures have previously been characterized using an unconstrained distribution over mutation patterns \cite{pmid23318258,pmid23628380}. Thus, if the number of mutation patterns considered is $M$ then each
mutation signature is characterized by a probability vector
of length $M$ (which must sum to 1, so $M-1$ parameters). 
A problem with this approach is that
it requires a large number of parameters per mutation signature. As noted above,
even accounting only for immediately flanking bases 
gives $M=96$.  Furthermore, $M$ increases exponentially if we try to account for additional context: to take account of up to $n$ bases 5' and 3' to the mutated site 
(henceforth referred to as the $-n$ position and the $+n$ position, respectively) gives $M=6 \times 4^{2n}$. 
Having a large number of parameters per signature causes
two important problems: i) estimates of signature parameters can become statistically unstable; ii) signatures can become
difficult to interpret.

% On the other hand, the mutation rates are mostly below 10 per Mb \cite{pmid23945592},
% (where about 30,000 mutations are expected in entire genomic regions).
% Therefore, considering even $\pm 2$ positions and substitution patterns (1535 parameters)
% can cause instability in parameter estimation.

% On the other hand, observing the mutation signatures detected in the previous studies, 
% many mutation signatures have clear characteristics such as:
% \begin{itemize}
% \item 
% A certain value of mutation features is often dominant in many known mutation signatures. 
% For example, the base T is highly dominant at the $-1$ position in the APOBEC signature, 
% and C $>$ T substitution is highly dominant in the ultraviolet signature \cite{pmid23945592} (see Figure \ref{APOBEC_nature2013_sig}).

% \item
% Proportional relationships among mutation features are often observed in many signatures as shown in the APOBEC signature, 
% where the frequencies of bases at the $+1$ position are largely proportional across 3 major substitution patterns C$>$T, C$>$G and C$>$A,
% showing consistent tendencies  (A and T bases are more frequent than C and G bases).

% \end{itemize}

% These characteristics imply that the parameter spaces of mutation signatures can be degenerated to lower dimensional spaces in many cases, 
% and we can potentially reduce the number of parameters by imposing additional restrictions.

The first contribution of this paper is to suggest
a more parsimonious approach to modelling 
mutation signatures, with the benefit of producing both more stable estimates and more easily-interpretable signatures.
In brief, we substantially reduce the number of parameters
per signature by breaking each mutation pattern into 
``features", and assuming independence
across mutation features. For example, consider
the case where a mutation pattern is defined by the substitution and its two flanking bases. We break this into three features (substitution,
3' base, 5' base), and characterize each mutation
signature by a probability distribution for each feature
(which, by our independence assumption, are multiplied
together to define a distribution on mutation patterns). Since the number of possible values for each feature is 6, 4, and 4 respectively this requires $5+3+3=11$ parameters instead of $96-1=95$ parameters. Furthermore, extending this model to account for $\pm n$ neighboring bases requires only $5 + 6n$ parameters instead of $6 \times 4^{2n} - 1$. For example,
 considering $\pm 2$ positions requires 17 parameters
 instead of 1535. Finally, incorporating
transcription strand as an additional feature adds just one parameter, instead of doubling the number of parameters.

Since the aim of a mutation signature is, in some ways, to capture dependencies among features, the independence assumption may seem counter-intuitive. However, the
idea is exactly analogous to the use of a ``position weight matrix" (PWM) to represent motifs in sequence data. In this
analogy, a motif is analogous to a mutation signature, and each location in the motif is analogous to a ``feature". Just as we use a probability vector for each feature, a PWM defines a probability vector for each location in a motif,
and these probabilities at each location can be multiplied together to yield a probability
distribution on sequences.
Even though a PWM cannot capture complex 
higher-order dependencies, some of which likely do exist in practice, it has been a highly successful tool for motif
analysis - likely because it can capture the most important characteristics of transcription factor binding sites 
(that some locations will show strong preference for a particular base, whereas others will not), and also
because it can be represented in an easily-interpretable way via sequencing logos \cite{pmid2172928}.
For similar reasons -- in addition to the empirical
demonstrations we present later -- we believe our mutation signature representation will prove useful for mutation signature analysis.

Figure \ref{mutSig_example} illustrates the way that our new representation of signatures can simply capture
a previously-identified signature \cite{pmid25362482, pmid23318258} and provides
an easily-interpretable visualization of the signature 
that is analogous to  sequencing logos \cite{pmid2172928}.
We particularly note how the key elements of this mutation signature are more immediately visually apparent than with visualizations of the full vector of probabilities used by existing approaches.

% Let's consider putting this in the discussion:
% Although most mutation signatures show proportional relationships among mutation features, 
% a few signatures seem to violate proportional relationships and cannot be explained by independent models. 
% For example, the pol $\epsilon$ mutation signature \cite{pmid23945592} (see Figure \ref{POLE_nature2013_sig}) puts strong probability masses 
% on just two pattern TpCpT $>$ TpApT and TpCpG $>$ TpTpG, 
% and the frequencies of substitution patterns and bases at the $+1$ positions are not proportional unlike the APOBEC signature.
% However, by using multiple mutation signatures, we can represent this phenomenon even under the independence assumption.

%In this paper, we focus on substitution patterns, flanking bases (up to the $\pm 2$ positions) and transcribed directions as contextual factors of mutation patterns.

\subsection*{Mathematical specification, and connection with mixed-membership models}
% yshira20150421 I think the title of the section should change (because this is not "detail").

Suppose each somatic mutation has $L$ mutation features, $\bm{m} = (m_1, m_2, \cdots, m_L)$, 
where each $m_l$ can take $M_l$ discrete values. Also, let $\bm{M} := (M_1, \cdots, M_L)$.
For example, when taking account of 6 substitution patterns and $\pm 2$ flanking sites,
$\bm{M} = (6, 4, 4, 4, 4)$.
See Supplementary Table 1 for other examples.

Suppose we have observed mutations in $I$ sampled cancer genomes, and let $J_i$ denote the number of observed mutations in the $i$-th cancer genome.
Further, let
$\bm{x}_{i, j} = (x_{i, j, 1}, \cdots, x_{i, j, L}), (i = 1, \cdots, I, j = 1, \cdots, J_i)$ denote the observed mutation feature vector for the $j$-th mutation of $i$-th cancer genome,
where $x_{i, j, l} \in \{1, \cdots, M_l \}$.

Our model assumes that each mutation arose from one of $K$ possible mutation signatures. Each cancer sample has its own characteristic proportion of mutations of each signature type (which might depend on lifestyle, genetic differences, etc.).
We let $q_{i,k}$ denote the proportion of signature $k$ in sample $i$, so $\bm{q}_i = (q_{i,1}, q_{i,2}, \cdots, q_{i, K}) \in \Delta^K,  (i = 1, \cdots, I)$ where $\Delta^S = \{ (t_1, \cdots, t_S) : t_s \geq 0 \, \forall s, \sum_{s=1}^S t_s = 1 \}$ denotes the S-dimensional simplex of non-negative vectors summing to 1.
Further, each mutation signature is characterized by parameter vectors $\bm{F}_k := (\bm{f}_{k,1},\dots,\bm{f}_{k,L})$, where \bm{$f}_{k,l}$ is a probability vector for
the $l$-th feature in the $k$-th signature.
That is,  $\bm{f}_{k, l} = (f_{k, l, 1}, \cdots, f_{k, l, M_l }) \in \Delta^{M_l}$.

Our generative model for the observed mutations $\{ \bm{x}_{i,j} \}$ in each
cancer sample can now be described as a two-step process.
\begin{enumerate}
\item
Generate $z_{i,j} \sim \text{Multinomial} (\bm{q}_m)$, where $z_{i,j} \in \{1,\cdots,K \}$ denotes the (unobserved) underlying mutation signature that caused the $j$-th mutation in the $i$-th sample.
\item 
For each $l (= 1, \cdots, L)$, generate $x_{i,j,l} \sim \text{Multinomial}(\bm{f}_{z_{i,j},l})$.
Thus, 
\begin{equation}
\Pr(x_{i,j,l} = m | z_{i,j}=k) = f_{k,l,m} 
\end{equation}
\end{enumerate}

This generative model is summarized in Figure \ref{methodOverview}.
The key parameters in this model
are the membership proportions for each sample, $\bm{q}_i$, and the mutation signature parameters, $\bm{F}_k$. 
We estimate these parameters by maximizing likelihood using an EM algorithm. 
A simulation study demonstrates that the estimation method can reproduce the mutation signature very accurately 
provided enough mutations and samples are available (see Supplementary Material and Supplementary Figure 1). 
See Methods for more details, how to select $K$, etc.

% yshira20150421 I tried to mention about background signature very shortly in the main text as above. 
% But I feel the necessity of further discussion.
The intrinsic composition of genome sequence, if unaccounted for, can undesirably influence estimated mutation signatures.
For example, since the di-nucleotide CpG is underrepresented in most genomic regions (other than promoters), 
a signature with substitutions from a C base can have weaker signals of G base at the $+1$ position.
In previous work \cite{pmid23628380}, this background problem was dealt by explicitly incorporating ``mutation opportunity'' coefficients into the model.
Here, to reduce the influences of intrinsic sequence composition on our signature estimates, we introduce a special ``background signature"
$\{ \bm{F}_{0} \} \in \Delta^{M_1 \times \cdots \times M_L}$, 
which is designed to capture biases in intrinsic genome sequence composition and is calculated from the composition of consecutive nucleotides of the human genome sequence 
(See Methods for the detail).

This model is essentially a ``mixed-membership model", also known as an ``admixture model" \cite{pmid10835412} or ``Latent Dirichlet Allocation" \cite{Blei:2003}. For example,
the membership proportions for each sample are analogous to admixture proportions in an admixture model; the mutation signatures are analogous to 
populations, and the mutation signature-specific parameters are analogous to population-specific allele frequencies
(see Relationship with mixed-membership models in Methods for further discussion).

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Robustness experiments using cancer genomes from urothelial carcinoma of the upper urinary tract}

Here we compare our new ``independent model'' for 
mutation signatures, which assumes independence
among mutation features,
with the ``full model'', which corresponds to existing approaches. We compare mutation signatures obtained by the two approaches and investigate the robustness of each approach by downsampling experiments.

The data consist of 14717 somatic substitutions 
collected from a study of 26 urothelial carcinomas of the upper urinary tract (UCUT) \cite{pmid23926200}.
The original study identified a novel mutation signature in these data: 
T $>$ A substitutions at CpTpG sites with a strong transcription strand specificity caused by aristolochic acids (AA).

We consider a mutation pattern to consist of the substitution pattern, the $\pm 2$ flanking bases, and the transcription strand direction. Thus each signature is characterized by 18 parameters in our independent model, and by 3071 parameters in the full model. After analyzing the data with various number of mutation signatures $K$,
we selected $K=3$ signatures for these analyses (Supplementary Material, Supplementary Figure 2, 3 and 4).

The inferred APOBEC signature under the independent model
shows a clear depletion of G base at the $-2$ position,
which is consistent to the previous study \cite{pmid23318258} and results in the next subsection.
In contrast, for the full model, this tendency is rather mild (Figure \ref{UCUT:APOBEC_full5_sig}, \ref{UCUT:AA_full5_sig} and Supplementary Figure 5).
The inferred AA mutation signature has no clear characteristics at the $-2$ position.
These results suggest that our independent model has the potential
to identify signatures in more detail and with less data than existing approaches based on the full model.

To investigate this further we performed downsampling experiments.
Using the mutation signatures obtained using all 14717 substitutions as a gold standard, 
we assessed performance of the proposed method on down-sampled data consisting of $r\%$ of the original data, where  $r=(1\%, 2.5\%, 5\%, 10\%, 25\%, 50\%)$.
To measure robustness we used the cosine similarity on the 
full dimensional vector space, which allows comparison between the full model and the independent model.
We repeated each down-sampling experiment 100 times for each model.

% For measuring the deviations of the mutation signature, the cosine similarity was used on the $\prod_{l=1}^L M_l$ dimensional vector space $f_{i, \bm{m}} = \prod_{l=1}^L f_{k,l,m_l}$,

The results (Figures \ref{UCUT:APOBEC_downsampling} and \ref{UCUT:AA_downsampling})
confirm that the results of the independent model 
are substantially more robust to reductions in data size than the full model. Indeed, mutation signatures inferred using the independent model with only 10\% of the data 
remain highly similar to the signatures inferred from the full data; by comparison the full model shows a much larger drop-off in similarity, especially in the APOBEC signature where even using 50\% of the data gives a substantial drop-off in similarity. Both methods found the AA signature 
easier to recover than the APOBEC signature.
We believe that this is because the number of T $>$ A substitutions at GpTpC sites are far more frequent in this dataset.

\subsection*{Application to somatic mutation data of 30 cancer types}

To provide a more comprehensive practical illustration of
our method, we applied it to somatic mutation data from 30 cancer types \cite{pmid23945592}.
We applied the method to each cancer type separately to assess similarity of estimated signatures across cancer types.
For each cancer type we selected the number of signatures $K$ by fitting the model with increasing $K$ and examining the log-likelihood, bootstrap errors, and correlation of membership parameters.The selected values of $K$ are given in 
Supplementary Table 2.
Also, we simply removed somatic mutations located in an intergenic region to include transcription strand biases as mutation features.
Finally, we merged similar mutation signatures across different cancer types 
(when their Frobenius Distance were $<$ $0.6$). 
% XXX WHAT KIND OF CLUSTERING?
 % (see Supplementary Material for the values of the log-likelihoods and bootstrap errors for each cancer type).

Figures \ref{nature2013_sig_summary} and \ref{nature2013_sig_member} summarise the results.
In total, we identified 27 mutation signatures.
Many of these signatures show reassuring similarities with
signatures identified in previous studies. However
signatures from our independence model, because of its ability to effectively
and parsimoniously deal with both $\pm2$ flanking
base context and strand bias, are often more refined, highlighting additional details or features not previously evident. By comparing the composition of nucleotides and cancer types exhibiting the signatures with results of previous studies, we were able to associate many of the detected signatures with known mutational processes.
In addition, as we reviewed these signatures
and compared them with previous work, we noticed connections
that, while not directly related to our new model, appear
novel and noteworthy. The remainder of this section
provides a comprehensive discussion of these findings.

Signatures 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) observed in colorectal and uterine cancers
appear likely to be associated with deregulated activity of the error-prone polymerase Pol $\epsilon$. 
In previous analyses of these data \cite{pmid23945592}, 
the signature for Pol $\epsilon$ dysfunction was represented by a single signature (their ``signature 10'').
In contrast our new approach uses two signatures.
Since these signatures are highly correlated,
and appear connected by
a single biological mechansim, we certainly do not argue
that inferring them as a single signature is ``wrong". However, splitting them into two signatures does help highlight certain features. Specifically,
signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Supplementary Figures 6(c, d)).
This strand bias may be connected with the enrichment of 
C $>$A at TpCpT mutations in leading strands of replication forks
observed by \cite{pmid25228659}. Although replication strand bias is different from transcription strand bias, these two biases may be connected through the fact that replication origins prefer transcription start sites \cite{pmid23187890}.

These signatures also illustrate the ability of our model
to help highlight sequence context effects beyond the immediate flanking bases. Specifically, both signatures 1 and 8  show an elevated frequency of the T base at position $-2$,
and signature 1 also shows slightly elevated frequency of the T base at position $+2$
 (Figures \ref{POLE1_two5prime} and \ref{POLE2_two5prime}, Supplementary Figure 6(c, d)).
A previous study of Pol $\epsilon$  \cite{pmid25228659} 
found that a nonsense mutation R23X of TP53 is enriched in cancers with Pol $\epsilon$ defects.
In fact, the pattern of this mutation is C $>$ T at TpTpCpGpA, closely matching signature 8.
This illustrates that the inclusion of $\pm 2$ bases 
into signatures may be helpful for identifying underlying mechanisms.

%XX Is the point here that looking $\pm 2$ bases can be useful and our method is particularly helpful here? XX

%XX this is important that it is drive by a single sample, but maybe just because yuichi looked more carefully and not new method XX
Signature 2 (C $>$ A at [CT]pCpT) is observed solely in low grade gliomas, and
appears related to, but slightly different from, the signature previously detected in the same cancer types (Supplementary Figure 7 (a)). Indeed, the corresponding signature in the previous study shows very complex patterns (C $>$ A at NpCpT or C $>$ T at GpTpN). Further investigation
revealed that this signature is driven by 
a single sample with an extremely high mutation rate (Supplementary Figure 7 (b,c)), and signature 2 disappeared when we removed this sample (Supplementary Figure 8).
It may be that the complex low-grade-glioma specific signature detected in the previous study is driven by the same single sample.
We suggest that these signatures should be treated with caution until validated in additional samples. 
% XXX Need to discuss this wording and what it means XXX

%XX Also important, making a connection, but not the new method XX
Signature 4 (C $>$ A at CpCpG) observed in kidney clear cell carcinomas, lung adenocarcinomas and melanomas 
seems to correspond to the ``signature R2''  detected in the same cancer types (plus lung squamous carcinomas) in 
\cite{pmid23318258} (see their Supplementary Figures).
Again our analysis highlights additional contextual information, with a strikingly elevated frequency of base C at the $-2$ position (Supplementary Figure 9 (a,b,c)).
However, for each cancer type, only a few samples support this signature (see Supplementary Figure 9 (d, e, f)),
and the corresponding signature 
could not be validated in the previous study: most somatic mutations corresponding to that signature could not be validated by re-sequencing or visual inspection of BAM files using genomic viewers. Again, further investigation
yielded a potential explanation for this finding:
this signature largely matches that of a putative artifact caused by oxidation of DNA during acoustic shearing \cite{pmid23303777}, 
and we conclude that this signature, and the corresponding signature in previous work, are likely artefactual.
Although not of direct biological interest,
identifying artefactual signatures could be helpful in removing false positive mutations.

%XX highlights robustness of approach XX
Signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and is surely related to the activity of the APOBEC family. The 12 inferred signatures are highly consistent among cancer types except for B-cell lymphoma (Supplementary Figure 6(a)), highlighting the robustness
of our approach.  Almost all of them show enrichment of A and T and depletion of G base at the $-2$ position (Figure \ref{APOBEC_two5prime}, Supplementary Figure 6(a)), 
consistent with the UCUT data above and previous analyses \cite{pmid23318258}.
The estimated transcribed strand specificities varied among cancer types, suggesting that there is not consistent strand-specificity in APOBEC signatures (and the observed variation may be due to estimation errors).
%Yuichi is that true?
Signatures 15 and 16 may also be related to APOBEC, although the estimated signatures are sufficiently different from 13 that they
were not merged into a single 
cluster by our specified clustering criteria.

Signatures 10, 11, 12, 19 and 21 provide further examples
of our method refining previously-observed signatures, highlighting strand biases and/or sequence context effects,
particularly 2 bases upstream of the substitution.
Signature 10 (C $>$T at [CT]pCpC) was observed in head and neck cancers and melanomas, and probably relates to ultraviolet light.
Consistent strand specificities among the two cancer types (Supplementary Figure 6(e)) matches previous results \cite{pmid23318258}, but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figure \ref{UV_two5prime}, Supplementary Figure 6(e)). 
Signature 11 (C $>$ T at GpCp[CG]) appears in small-cell lung cancers and stomach cancers,
and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear. Again our analysis highlights elevated abundance of G at the $-2$ position (Figure \ref{LMST_two5prime}, Supplementary Figure 6(f)).
Signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively, agree well with those detected in the same cancer types in the previous study \cite{pmid23318258}. However our analysis again refines these signatures,
highlighting a strand bias in all three, and sequence context effects at the -2 position in Signatures 12 and 21.
%although the causes of these signatures remain unknown.
%which should be validated by analysis of more samples from different cohorts and cancer types.

One signature, Signature 20, appears not to match any signatures in the previous analysis \cite{pmid23318258} and represents a potentially novel signature. 
This signature (T $>$ C at [AC]pTpN) is observed in thyroid cancers, and shows a very strong strand specificity, which could be due to transcription-coupled nucleotide excision repairs.
This signature may have been too weak for previous
methods to detect, perhaps because the mutation ratio of thyroid cancer is low, possibly reflecting improved sensitivity of our more parsimonious model.
%% download the TCGA data and check the visualization of the mutation position for this data.

The remaining signatures largely recapitulate previous results.
Signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
The estimated signature in each cancer type shows higher mutation prevalence on the template strand (Supplementary Figure 6(b)), 
which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
Signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas matches the pattern detected in the same cancer type in the previous study.
Signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and arguably relates to deamination of 5-methyl-cytosine.
Signature 9 (C $>$ T at NpCp[CT]) was observed in melanomas and glioblastomas, and is probably associated with a chemotherapy drug, temozolomide.
Signature 18 (T $>$ C at ApTp[AG]) observed in liver cancers has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, 
though the source of this signature is still not clear. 
In this signature, we observe a very strong strand specificity as shown in \cite{pmid25362482, pmid23318258},
suggesting a possible role for transcription-coupled nucleotide excision repairs.

%In summary, our method and analysis captures and refines 
%several previously observed mutation signatures, particularly
%detecting several novel characteristics at the $\pm 2$ position that are consistent across multiple cancer types.
%suggesting the existence of more hidden characteristics at the distant position from the mutated site.

\section*{Discussion}

In this paper, we presented new methods for inferring and visualizing mutation signatures from multiple cancer samples.
The new methods exploit simpler, more parsimonious, models
for mutation signatures than existing methods.
This improves stability of statistical estimation, and easily allows
a wider range of contextual factors (e.g. more flanking bases) to be
incorporated into the analysis. In addition, we provide a new
intuitive way to visualize the inferred signatures.

We have also emphasised the connection between mutation
signature detection, and the use of mixed-membership models
in other fields, particularly admixture analysis and document clustering.  This connection naturally raises the possibility of improving the proposed approach by learning from experiences in those other fields. For example, in admixture analysis, \cite{pmid12930761} found that the use of a correlated
prior on allele frequencies improved sensitivity to detect
population clusters; this suggests that it might be fruitful to consider a correlated prior distribution on signatures, to allow that some signatures - perhaps in different cancers - may be similar to one another (though not identical).
More generally, introducing certain prior distributions or penalty terms, such as sparsity-promoting penalties \cite{hoyer2004non, engelhardt2010analysis} and determinantal point process priors \cite{kulesza2012determinantal, kwok2012priors}
could improve both accuracy and interpretation.
Further, as the scale of cancer genome data becomes large,
more sophisticated computational approaches for estimating parameters may become necessary. We can potentially borrow a number of computational techniques such as those using EM-algorithm \cite{Hofmann:1999, tang2005estimation},
sequential quadratic programming \cite{zhou2011quasi}, Gibbs sampling \cite{pmid10835412,pmid14872004} 
and variational methods \cite{Blei:2003,teh2006collapsed,raj2014variational}.
Finally, to address the problem of determining the number of signatures, it may be fruitful to extend the framework to  
the Hierarchical Dirichlet processes \cite{teh2006hierarchical}.
% XX I think we should try to summarize the future work part from the methods more succinctly and put it here. XX
% yshira20150421 tried to summarize briefly as above

Although we have focused on point substitution mutations in this paper, 
many other types of mutations occur in cancer genomes, 
including insertions, deletions, double nucleotides substitutions, structural variations and copy number alterations \cite{meyerson2010advances, helleday2014mechanisms}.
Our framework could incorporate these additional mutation types, by 
summarizing them using appropriate mutation features.
In some cases, choice of appropriate features may need investigation. 
For example, longer deletions could be represented by the length of deletion and the adjacent bases; 
for short deletions (a few bases) it may be fruitful to include the actual deleted bases as part of the features.
%However, detailed investigation on what mutation features are practical is a future problem.

We have detected a number of mutation signatures having transcription strand biases, 
which are naturally considered to be associated with transcription activities.
Therefore, to further understand the effect of transcription activities on mutational mechanisms, 
our framework will be helpful: by allowing for including gene expression or RNA polymerase II occupancies 
to mutation features, the relationships of strand biases and transcription activities will be clarified.

Although we believe our new methods already provide
useful gains compared with existing approaches,
the methods are perhaps even more important for their future potential to incorporate other contextual data, including epigenetic data,
into mutation signature analysis. This is important, because
local mutation densities are closely related to a number of genomic and epigenetic factors, such as GC content, repeat sequences, chromatin accessibility and modifications, and replication timing \cite{pmid22820252, pmid21953857, pmid23422670, pmid23770567}.
A recent study found that epigenetic information in the cell types of origin of the corresponding tumors is the most predictive \cite{pmid25693567} for local mutation densities. A growing range of epigenetic data from many cell types are now available, and it will be interesting to integrate these epigenetic factors into mutation signature analysis to help understand how these epigenetic factors influence DNA damage and repair mechanisms. 
Our work here provides a straightforward way to do this: epigenetic
data can be simply added as features
to the mutation signature. 
We believe that 
the value and impact of our work, and 
specifically our proposed approach to modelling mutation signatures via independent features, will grow as more and more features
are incorporated into the analysis.

% In this paper, we have demonstrated that the independent representations could obtain more robust estimates than the full representations.
% Also, even in the independent representations, more complex mutational process could be represented by utilizing multiple mutation signatures.
% However, we still do not know about the exact mechanisms of mutational processes and 
% there might be those that should be represented by the full representations or somewhere between the full representations and the independent representations.
% We should keep exploring appropriate representations of mutation signature with expertise from biology and chemistry.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Methods}

% Therefore, each probabilistic mutation signature, in general, consists of multiple multinomial distribution parameters.
% We give a way of visualizing probabilistic mutation signature (see Figure \ref{mutSig_example}), 
% which is reminiscent of sequencing logos \cite{pmid2172928}.

\subsection*{Parameter Estimation}

The parameters $\{ \bm{f}_{k, l} \}$ and $\{ \bm{q}_i \}$
must be estimated from the available mutation data $\{ \bm{x}_{i,j} \}$. 
Here we adopt a simple approach that uses an EM-algorithm to maximise the likelihood.

Let $g_{i, \bm{m}}$ denote the number of mutations in the $i$-th sample that have mutation feature vector $\bm{m}$.
In the E step of the EM algorithm, we calculate values of auxiliary variables $\theta_{i, k, \bm{m}}$ defined as
\begin{equation}
\theta_{i, k, \bm{m}} = \frac{  q_{i,k} \prod_{l=1}^L f_{k,l,m_l} }{ \sum_{k^{\prime} = 1}^K q_{i, k^{\prime} } \prod_{l=1}^L f_{k^{\prime}, l, m_l } }.
\end{equation}
Then, in the M-step, we update the parameters $\{ \bm{f}_{k, l} \}$ and $\{ q_{i, k} \}$ as
\begin{equation}
f_{k, l, p} = \frac{ \sum_{\bm{m} : m_l = p} g_{i, \bm{m}} \theta_{i, k, \bm{m}} }{ \sum_{p^{\prime} } 
\sum_{\bm{m} : m_l = p^{\prime}}  g_{i, \bm{m}}\theta_{i, k, \bm{m}} },
\end{equation}
\begin{equation}
q_{i, k} = \frac{ \sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k, \bm{m} } }{  \sum_{k^{\prime} }\sum_{\bm{m}} g_{i, \bm{m}} \theta_{i, k^{\prime}, \bm{m} } }.
\end{equation}

We use the R package SQUAREM \cite{varadhan2008simple} to accelerate
convergence of this EM algorithm (SQUAREM implements a general approach to accelerate the convergence of any fixed-point iterative scheme such as an EM algorithm).
To address potential problems with convergence to local minima,
we apply the EM algorithm several times (10 times in this paper) using different initial points,
and use the estimate with the largest log-likelihood.
See Supplementary Material for the derivation of the above updating procedures.

\subsection*{Background signatures}

% The intrinsic composition of genome sequence may 
% influence the estimated mutation signatures.
% For example, the number of observed C $>$ T transitions at CpG sites may increase at promoter regions,
% just because CpG dinucleotides are more frequent in those regions.
% In previous work \cite{pmid23628380}, this background problem was dealt by explicitly incorporating mutation ``opportunity'' coefficients into the model.

% XX This is not quite clear to me. Also may need including in Main text? XX

% Here, to offset the influences of intrinsic sequence composition, we add background signatures 
% $\{ \bm{f}_{0, \bm{m}} \} \in \Delta^{M_1 \times \cdots \times M_L}$.
% For example, when obtaining the background signatures in case of considering substitution patterns and up to $\pm 2$ flanking bases, 
Here, we describe how the background mutation signature is 
obtained in the case where mutation features are the substitution patterns, the $\pm 2$ flanking bases, and the transcription strand.
Since the majority of the data used in this paper is exome sequencing data, and since we consider transcription strand as a mutation feature,
we use the exonic regions of the human genome reference sequence to obtain the background mutation signature.
First we calculate the frequencies of 5-mers with transcription strands of the corresponding exon,
where we take complement sequences and flip the strand for those whose central bases are A or G.
Then, assuming alternated bases are equally likely from each central base C and T, 
the frequency of each mutation feature is derived directly from those of the 5-mers and transcription strands.
Finally, the probability of each mutation feature is derived by normalizing each frequency to sum to one.

\subsection*{Estimating standard errors}

We use the non-parametric bootstrap \cite{efron1994introduction} to calculate standard errors for parameter estimates. This involves resampling somatic mutations from the empirical distribution of the original data $\{ \bm{x}_{i,j} \}$ for each cancer genome. For each of 100 such bootstrap samples, we re-fitted the model, using parameters obtained for the original data as initial points.
We then used sample standard errors of the inferred mutational signatures as estimates of parameter standard errors.

\subsection*{Selecting the number of signatures}

Determining an appropriate number of mutation signatures $K$ is a challenging task. 
One approach is to utilize some statistical information criteria such as AIC \cite{akaike1974new} or BIC \cite{schwarz1978estimating}.
In the population structure problems, for example, 
the Bayesian deviance \cite{pmid10835412} 
and cross-validation \cite{alexander2011enhancements} have been suggested.
One previous study on mutation signature problems \cite{pmid23628380} utilized BIC.
On the other hand, the problem of using these statistical information criteria is that most of them are based on the likelihood,
where slight deviations between the specified probabilistic models and the reality sometimes leads to additional
(possibly spurious) mutation signatures being selected to
compensate for those deviations.

In this paper, instead of utilizing a statistical information criteria, we adopt the following strategy:
\begin{itemize}
\item
After calculating the likelihood and standard errors of parameters for a range of $K$,
the value of $K$ is determined at the point where the likelihood is sufficiently high, 
and the standard errors are sufficiently low \cite{pmid23318258}.

\item
When, for $k_1$-th and $k_2$-th mutation signatures, 
we could detect strong correlations between the estimated membership parameters 
for each cancer genome
($(q_{1,k_1}, q_{2,k_1}, \cdots, q_{I,k_1})$ and $(q_{1,k_2}, q_{2,k_2}, \cdots, q_{I,k_2})$),
and the two mutation signatures ($\{ \bm{F}_{k_1} \}$ and $\{ \bm{F}_{k_2} \}$) show similar patterns,
then this suggests that the method may have split one mutation signature into two. We choose $K$ to be small enough that such pairs of mutation signatures do not occur.
\end{itemize}

These strategies are not claimed as optimal, but appeared
to  provide satisfactory results in our applications here.
The development of automated and practical approaches for choosing $K$ is a possible area for future development.

\subsection*{Existing methods as a special case}

Previous approaches to mutation signature modelling in \cite{pmid23945592, pmid23318258} are a special case of our framework. Specifically, they correspond to combining
all possible combinations of mutation features into a single ``meta-feature'', which takes $M_1 \times M_2 \times \dots \times M_L$ possible values.
Thus, instead of having $L$ features with $\bm{M} = (M_1,\dots,M_L)$,
existing approaches have one feature with $\bm{M} = (M_1 \times \dots \times M_L)$ (see Supplementary Table 1). 
The resulting model allows for arbitrary distributions on the $M_1 \times \dots \times M_L$
feature space, and we call the resulting model the ``full model''.
The full model can represent complicated dependencies in a single
signature. For example,
a situation where C $>$ A is frequent at ApCpG sites and C $>$ T is frequent at TpCpA sites could be represented with one signature.
This may be desirable in some settings and not in others.
However, when many mutation contextual factors are taken into account and the number of free parameters becomes huge,
estimated results can be unstable and unreliable.
Furthermore, there is a risk of over-interpreting the complex features of estimated signatures.

\subsection*{Relationship with mixed-membership models}

Our model is closely related to
mixed-membership models that have been adopted in other applications, 
such as document classification and population structure inference problems. 
In this subsection, we outline these relationships,
slightly abusing notation to contrast the relationships.

In the topic model \cite{Hofmann:1999,Blei:2003},
which are a form of mixed-membership models frequently used in document classification problems,
each document is assumed to have $K$ different ``topics'' in varying proportions ($\bm{q}_i \in \Delta^K$),
where each topic is characterized by a word frequency (a multinomial distribution on a set of words $W$ ($\bm{f}_k \in \Delta^W$).
And each word is assumed to be generated by one of $K$ multinomial distributions (topics).
The detailed generative process of the $j$-th word in the $i$-th document $x_{i,j}$ is:  
\begin{enumerate}
\item
Generate the underlying topic for the $j$-th word: $z_{i,j} \sim \text{Multinomial} (\bm{q}_i)$, where $z_{i,j} \in \{1,\cdots,K \}$.
\item
Generate $x_{i,j} \sim \text{Multinomial} (\bm{f}_{z_{i,j}})$, where $x_{i,j} \in \{1, \cdots, W \}$.
\end{enumerate}
Actually our ``full model'' ($L=1$) is essentially the same as a topic model.

On the other hand, in population structure inference problems \cite{pmid10835412, pmid19648217}, 
each individual is assumed to be an admixture of $K$ ancestries in varying proportions, 
where each ancestry is characterized by the allele frequency at each SNP locus.
Each SNP genotype of an individual is assumed to be generated by the two step model:
first, an ancestry (``population") is chosen according to the admixture proportion for each individual,
and then the SNP genotype is generated according to the allele frequency of the selected ancestry at that locus.
The relationships among the mutation signature models, topic models and population structure models are summarized in Table \ref{tab_pop}.

As pointed out by \cite{ding2008equivalence}, there is a  close relationships between mixed-membership models and nonnegative matrix factorization,
which has been successfully used in the previous studies for mutational signature problems \cite{pmid22608084, pmid23318258, pmid23945592}.
In fact, the proposed method can be seen as non-negative matrix factorization with additional restrictions.
See Supplementary Materials for details of the relationship between the proposed approach and nonnegative matrix factorization.

% \subsection*{Future works for the methodology}

% Since the mixed membership model has been widely studied in the field machine learning and population genetics,
% there are many rooms for improvement in the current form of the proposed method by learning from past experiences and knowledge.

% First, a number of large scale cancer genome sequencing projects have been launched all over the world,
% and we need to improve the parameter estimation so that those large scale data can be handled.
% A number of  past parameter estimation techniques proposed for mixed-membership models can be tailored for the proposed model.
% EM-algorithm (or its variant, called tempered EM algorithm) have been adopted in classical topic models 
% for document classification \cite{Hofmann:1999} and population structure estimation \cite{tang2005estimation}.
% In the population structure estimation problem, \cite{pmid19648217} has proposed 
% a fast block relaxation scheme using sequential quadratic programming for block updates
% with a quasi-Newton acceleration of convergence \cite{zhou2011quasi}, demonstrating a great improvement over EM-algorithm.
% Similar techniques are used in \cite{taddy2011estimation} for the document classification.
% For the estimation of Bayesian mixed-membership models, 
% (collapsed) Gibbs sampling \cite{pmid10835412,pmid14872004} and 
% variational method \cite{Blei:2003,teh2006collapsed,raj2014variational} have been proposed. 
% Evaluate those techniques in the problem of mutation signature extraction problems, and incorporate the efficient method is an important future work.

% Second, introducing certain prior distributions or penalty terms can lead more efficiency in terms of both accuracy and interpretation,
% considering the success in machine learning and statistical genetics communities.
% \cite{hoyer2004non, engelhardt2010analysis}.
% In the documentation classification problem, 
% adopting determinantal point process priors \cite{kulesza2012determinantal, kwok2012priors}
% is demonstrated to be helpful for obtaining more diverse and intuitively interpretable topics,
% avoiding over fine-tuning for frequent groups of words with multiple similar topics.

% Third, for avoiding the problem of determining the number of latent variables, 
% introducing Hierarchical dirichlet processes \cite{teh2006hierarchical}, 
% which is a natural extension of topic models to nonparametric Bayesian frameworks, might be helpful.

% Finally, utilizing other metrics than Euclidean distance and Kullback-Leibler Divergence might be worth investigating.
% For nonnegative matrix factorization, the metric called $\beta$ divergence \cite{basu1998robust, eguchi2001robustifying}, 
% has been successfully utilized in many research fields (reviewed in \cite{fevotte2011algorithms}),
% and evaluating these metrics on mutation signature problems will be an interesting future tasks.

% Do NOT remove this, even if you are not including acknowledgments.

\section*{Acknowledgments}

We thank to Dr. Daichi Mochihashi for helpful discussion and comments on an earlier version of the method;
Dr. Hiromichi Suzuki for helpful feedback on the R packages;
Dr. Ziyue Gao for helpful comments on the manuscript.

This project was deeply influenced by the broader intellectual and academic environment surrounding Matthew Stephens's Laboratory,
when the first author stayed at the University of Chicago as a visiting scholar. The first author thanks Dr. John Novembre,
Dr. Jacob Degner and all the members of Matthew Stephens and John Novembre Laboratories for helpful discussion and comments.

% \section*{References}

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2009.bst
% style file and paste the contents of your .bbl file
% here.
%

% \bibliographystyle{plos2009}

\bibliography{plos_template}

% \section*{Figure Legends}
% This section is for figure legends only, do not include
% graphics in your manuscript file.
%
%\begin{figure}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure caption.  
%}
%\label{Figure_label}
%\end{figure}

\newpage

 \section*{Tables}
% 
% See introductory notes if you wish to include sideways tables.
%
% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements. Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
%
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\begin{table}[!ht]
\caption{
{\bf Relationships among mutation signature model, topic models, and population structure models.}
}
\begin{center}
\begin{tabular}{|l|l|l|l|} \hline
problem & $\bm{x}_{i,j}$ & $\bm{f}_{k}$ & $\bm{q}_{i}$  \\ \hline
mutation signature model & the $j$-th mutation  & the feature dist.  & the signature dist.  \\
& in the $i$-th cancer genome & for the $k$-th signature & for the $i$-th cancer genome \\ \hline
topic model &  the $j$-th word & the word dist. & the topic dist. \\ 
& in the $i$-th document & for the $k$-th topic & for the $i$-th document \\ \hline
population structure model & the $j$-th locus genotype & the allele freq.  & the admixture dist. \\
& of the individual $i$  & for the ancestry $k$ & for the individual $i$ \\
\hline 
\end{tabular}
\end{center}
\label{tab_pop}
\end{table}

\clearpage

\begin{figure*}[b]

\subfigure[Example mutation signature representation for the unconstrained model]{%
   \includegraphics[width=15cm,height=4cm]{example3076_TCR.eps}
  \label{TCR_ful}
}

\subfigure[Example mutation signature representation for the independent model]{%
   \includegraphics[width=8cm,height=5.32cm]{example18_TCR.eps}
  \label{TCR_ind}
}
\quad
\subfigure[Mutation signature parameter values for the signature in (b)]{%
    \includegraphics[width=175pt]{exampleTable.pdf}
    \label{TCR_ind_table}
}

\caption{Examples of visualizations and parameter values for the mutation signatures of the unconstrained (full) model and our independent model, 
where substitution patterns, two 5' and 3' bases and transcription strand direction are considered as mutation features. 
(a) The barplots are divided by 6 substitution patterns and transcription strand direction.
In each division, 256 bars show joint probabilities of up to two base 5' and 3' bases with substitution patterns and transcription strands
(ApApNpApA, ApApNpApC, ApApNpApG, ApApNpApT, $\cdots$, TpTpNpTpT).
(b, c) An example mutation signature representation and parameter values from our independent model, where mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).
In the bottom five rectangles, the width of each box represents the frequencies of bases (A, C, G and T) at the substitution and flanking site. 
To highlight the most informative flanking sites, the heights of flanking site boxes are scaled by ($1 + 0.5 \times \log \sum_{n = A,C,G,T} f_n^2$, where $f_n$ is the parameter for each base), which can be interpreted as ($1 - 0.5 \times \text{R\'enyi entropy}$) \cite{rrnyi1961measures}. This is analogous to the information content scaling used in sequencing logos.
In the top rectangle, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
In the upper right, the height of the $+$ box represents the frequencies of mutations in the coding strand (or the plus strand, the sense strand and the untranscribed strand)
whose nucleotide sequences directly corresponds to mRNA, 
whereas the height of $-$ box represents those in the template strand (or the minus strand, the antisense strand, the transcribed strand and the noncoding strand)
whose sequences are copied during the synthesis of mRNA.
}
\label{mutSig_example}

\end{figure*}

\clearpage

\begin{figure*}
\includegraphics[width=450pt]{methodOverview.pdf}
\caption{An overview of the generative model of somatic mutations proposed in this paper. Suppose there are three types of mutation sources (mutation signatures) such as ultraviolet, tobacco smoking chemicals and transcription coupled repairs.
Each cancer genome has ratios showing which types of mutation sources are contributing to its mutations (membership parameters).
The generative model of the pattern of each mutation is:
first, one of the mutation signatures is chosen according to the membership parameter.
Second, according to the selected mutation signature, mutation features such as substitution pattens and flanking bases are generated 
by the corresponding multinomial distributions for that mutation signature.
}
\label{methodOverview}
\end{figure*}

\clearpage

\begin{figure*}[ht]
\centering

\subfigure[APOBEC signature for the independent model]{%
  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_APOBEC_ind.eps}
  \label{UCUT:APOBEC_ind5_sig}}
\quad
\subfigure[AA signature for the independent model]{%
  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_AA_ind.eps}
  \label{UCUT:AA_ind5_sig}}

  \subfigure[APOBEC signature for the full model]{%
  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_APOBEC_full.eps}
  \label{UCUT:APOBEC_full5_sig}}
\quad
\subfigure[AA signature for the full model]{%
  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_AA_full.eps}
  \label{UCUT:AA_full5_sig}}

\subfigure[APOBEC signature stability]{%
  \includegraphics[height=5.5cm,width=7.5cm,clip]{UCUT_downsampling_APOBEC.eps}
  \label{UCUT:APOBEC_downsampling}}
\quad
\subfigure[AA signature stability]{%
  \includegraphics[height=5.5cm,width=7.5cm,clip]{UCUT_downsampling_AA.eps}
  \label{UCUT:AA_downsampling}}
%
\caption{The mutation signatures for the UCUT data, and the results of downsampling experiments. 
3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.}
\label{UCUT}
\end{figure*}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
\caption{The summary of mutation signatures obtained in a reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method,
where the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
First, mutation signatures were estimated separately in each cancer type, and then similar signatures were merged (see text).
}
\label{nature2013_sig_summary}
\end{figure*}

\clearpage

\begin{figure*}[b]
\centering
\includegraphics[width=15cm,height=15cm]{corrplot.eps}
\caption{The summary of membership of each mutation signature across 30 cancer types obtained using the proposed method.}
\label{nature2013_sig_member}
\end{figure*}

\clearpage

% \begin{figure*}[b]

\begin{figure*}[b]

\subfigure[APOBEC signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=14.5cm,height=6cm]{APOBEC_two5prime.eps}
   \label{APOBEC_two5prime}}

\subfigure[POLE1 signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{POLE1_two5prime.eps}
  \label{POLE1_two5prime}}
\quad
\subfigure[POLE2 signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{POLE2_two5prime.eps}
  \label{POLE2_two5prime}}

\subfigure[UV signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{UV_two5prime.eps}
  \label{UV_two5prime}}
\quad
\subfigure[The signature 11 intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{LMST_two5prime.eps}
  \label{LMST_two5prime}}

\caption{The estimated frequencies of bases at two 5' to the mutated site for each cancer type.
The bar heights show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
The error bars show bootstrapped standard errors.}
\label{two5prime}

\end{figure*}

% \begin{figure*}[b]

% \subfigure[POLE1 signatures in each cancer type]{%
%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
%   \label{POLE1_sig_list}}

% \subfigure[POLE2 signature in each cancer type]{%
%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
%   \label{POLE2_sig_list}}

% \subfigure[UV signature in each cancer type]{%
%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
%   \label{UV_sig_list}}

% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
%   \label{LMST_sig_list}}

% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
% \label{nature2013_sig_others}
% \end{figure*}

% \section*{Supporting Information Legends}
%
% Please enter your Supporting Information captions below in the following format:
%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
% 
%\begin{description}
%\item {\bf}
%\item {\bf}
%\end{description}

\end{document}

### 2015-05-25 backup/supp20150518/supp.tex
organized directory

--- 2015-03-19 supp/supp.tex
@@
 \usepackage{setspace} 
+\usepackage{subfigure}
 
@@
 \vspace*{0.25cm}
+
+\section*{Experiments on synthetic data}
+
+We begin by illustrating the ability of our method to extract ``true'' mutation signatures on simulated data. In these simulations a mutation pattern
+is defined by a substitution and the $\pm 2$ flanking bases.
+We generated a set of mutations changing the number of cancer genomes ($I=10, 25, 50, 100$),  and the number of mutations for each cancer genome ($J=10, 25, 50, 100, 250, 500, 1000$).
+The number of mutation signatures were set to $K=5$ including background mutation ratio.
+The mutation feature parameters and membership parameters were generated by Dirichlet distribution,
+\begin{equation}
+\bm{f}_{k,l} \sim \text{Dir} (\alpha \bm{1} ),\ k = 1, \cdots, K,\  l = 1, \cdots, L.
+\end{equation}
+ \begin{equation}
+ \bm{q}_{i, k} \sim \text{Dir} (\gamma \bm{1} ),\ i = 1, \cdots, I,
+ \end{equation}
+where $\alpha$ and $\gamma$ control the amount of dispersion 
+for the mutation signature parameters and membership parameters, respectively.
+For example, when $\gamma$ is small, most samples will have most of their
+mutations coming from a single signature (but not the same signature for each sample). When $\gamma$ is large most samples will have mutations coming from all signatures in roughly equal proportions.
+
+As Supplementary Figure 1(a) shows, we estimate the mutation signatures very accurately overall (see Supplementary Figure 1(b) for an example).
+As expected, accuracy improves as we increase the number of cancer genomes or mutations.
+Also, as the mutation feature dispersion decreases ($\alpha$ increases), accuracy decreases.
+This is expected: as $\alpha$ increases the signatures will become more ``fuzzy", with
+probability mass spread over a large number of mutation patterns, and so harder to infer
+(In other words, larger $\alpha$ corresponds to a less informative motif).
+In contrast, accuracy was relatively insensitive to 
+individual membership dispersion ($\gamma$), and estimates remain accurate even when most individuals have mutations in many different signatures.
+
+In most cases, the log-likelihood stopped increasing at $K=5$ mutation signatures, 
+whereas the standard error of the estimated parameters started increasing past $K=5$ (see Supplementary Figure 1(c)). We conclude that examining
+the trade-off between likelihood and standard-errors may be a helpful guide to selecting an appropriate number of mutation signatures.
+
+\section*{Experiment on UCUT data with the various number of mutation signatures}
+
+We analyzed the data using our new model, with increasing number of mutation signatures $K=2,3,\dots$. 
+Results are shown in Supplementary Figure 2.
+As expected, the likelihood increased with $K$. Bootstrap-errors started to increase at $K = 5$ (Supplementary Figure 3).
+
+With $K = 2$ (Supplementary Figure 2 (a)) we observed a mutation signature that appears to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription strand specificity).
+Increasing to  $K=3$ introduced an additional mutation signature corresponding to the APOBEC enzyme (C $>$ [AGT] at TpCpN sites) (Figure \ref{UTUC:APOBEC_ind5_sig}, \ref{UTUC:AA_ind5_sig}, Supplementary Figure 2 (b)).
+Increasing to $K = 4$ introduced an additional signature (T $>$ A at NpTpN sites with strong strand specificity) 
+that is somewhat similar to the AA signature 
+(Supplementary Figure 2 (c)). 
+The membership in the AA signature and this ``AA-like" signature were correlated ($R = 0.77$; Supplementary Figure 4). 
+This additional signature may be just making up for the residual of the AA signature 
+which the original AA signature could not explain due to a slight deviance of the probabilistic model.
+
+The strong correlation among estimated membership parameters started to be shown at $K = 4$.
+Considering all these factors together, $K = 3$ seems to be a reasonable choice in terms of the interpretability,
+and we adopted $K = 3$.
 
@@
 
+\begin{table}[!ht]
+\caption{
+{\bf Example of representation for mutation patterns (substitution patterns and one 5' and 3' bases)}
+In the independent representation, the elements of vector show substitution patterns, 5' adjacent bases and 3' adjacent bases, respectively.
+For substitution pattens, 1 to 6 values are assigned to C$>$A, C$>$G, C$>$T, T$>$A, T$>$C and T$>$G in this order.
+For 5' and 3' adjacent bases, 1 to 4 values are assigned to A, C, G and T.
+Note that the original base is fixed to C or T to remove the redundancy of complement sequences.
+}
+\begin{center}
+\begin{tabular}{|c|c|c|} \hline
+mutation pattern & full model & independent model \\ \hline
+$L$ & 1  & 3  \\
+$\bm{M}$ & (96)  & (6, 4, 4) \\  \hline
+ApCpA $\to$ ApCpA & (1)  & (1, 1, 1)  \\ 
+ApCpC $\to$ ApApC & (2)  & (1, 1, 2)  \\ 
+ApCpG $\to$ ApApG & (3)  & (1, 1, 3)  \\
+ApCpT $\to$ ApApT & (4)  & (1, 1, 4)  \\
+CpCpA $\to$ CpApA & (5)  & (1, 2, 1) \\
+$\cdots$ & $\cdots$ & $\cdots$ \\
+ApCpA $\to$ ApGpA & (17)  & (2, 1, 1) \\
+$\cdots$ & $\cdots$ & $\cdots$ \\
+TpTpT $\to$ TpGpT & (96)  & (6, 4, 4) \\ \hline
+\end{tabular}
+\end{center}
+\label{tab_rep}
+\end{table}
+
+\clearpage
+
 \begin{table}
@@
 \begin{figure}
-\centering
-\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
-\caption{An example of the relationship between true and estimated mutation signatures in the simulation.
+
+\subfigure[The accuracy of the proposed approach for the simulated data when changing the number of samples, mutations, 
+and the amounts of dispersion parameters ($\alpha$ and $\gamma$) for the mutation features and signature distribution parameters.]{%
+\includegraphics[width=16cm,height=6cm]{simulation_result.eps}
+\label{simulation_facet}
+}
+
+\subfigure[An example of the relationship between true and estimated mutation signatures in the simulation.
 In this example, the numbers of cancer genomes and mutations for each cancer genome are 25 and 100, respectively,
 and the parameters $\alpha$ and $\gamma$ are both set to 1. 
-From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations}
+From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations]{%
+\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
 \label{example_signature.pd}
-\end{figure}
+}
 
-\clearpage
+\subfigure[The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters 
+for several numbers of signatures $K$ in the trial of the above figure.]{%
+\includegraphics[width=12cm,height=4cm]{simulation_stat.eps}
+\label{example_summary.pd}
+}
 
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in the trial of the Supplementary Figure 1.}
-\label{example_summary.pd}
+\caption{The summary of simulation study.}
+
 \end{figure}

### 2015-05-25 paper/TextS1/TextS1.tex
organized directory

--- 2015-03-19 supp/supp.tex
@@
 \vspace*{1.0cm}
-{\LARGE \bf Supplementary material}
+{\LARGE \bf Experiments on synthetic data}
 \vspace*{0.25cm}
 
-\section*{Derivation of EM algorithm}
+We begin by illustrating the ability of our method to extract ``true'' mutation signatures on simulated data. In these simulations a mutation pattern
+is defined by a substitution and the $\pm 2$ flanking bases.
+We generated a set of mutations changing the number of cancer genomes ($I=10, 25, 50, 100$),  and the number of mutations for each cancer genome ($J=10, 25, 50, 100, 250, 500, 1000$).
+The number of mutation signatures were set to $K=5$ including background mutation ratio.
+The mutation feature parameters and membership parameters were generated by Dirichlet distribution,
+\begin{equation}
+\bm{f}_{k,l} \sim \text{Dir} (\alpha \bm{1} ),\ k = 1, \cdots, K,\  l = 1, \cdots, L.
+\end{equation}
+ \begin{equation}
+ \bm{q}_{i, k} \sim \text{Dir} (\gamma \bm{1} ),\ i = 1, \cdots, I,
+ \end{equation}
+where $\alpha$ and $\gamma$ control the amount of dispersion 
+for the mutation signature parameters and membership parameters, respectively.
+For example, when $\gamma$ is small, most samples will have most of their
+mutations coming from a single signature (but not the same signature for each sample). When $\gamma$ is large most samples will have mutations coming from all signatures in roughly equal proportions.
 
-The complete log-likelihood including missing data $\{ \bm{z}_i \}$ for the proposed model is
-\begin{equation*}
-\sum_{i = 1}^I \sum_{j=1}^J \sum_{k=1}^K I(z_{i,j} = k) \Bigl( \sum_{l = 1}^L \log f_{k, l, x_{i,j,l}} + \log q_{i,k} \Bigr).
-\end{equation*}
-Here, we introduce the variable for conditional probability for $z_{i,j}$ given the parameters and the mutation features $\bm{x}_{i,j}$,
-\begin{equation*}
-\theta_{i,k, \bm{m}} = \Pr \bigl( z_{i,j} = k \big| \bm{x}_{i, j} = \bm{m}, \{ \bm{f}_{k, l} \}, \{  \bm{q}_i \} \bigr)
-\end{equation*}
-Note that this conditional probability just depends on the value of mutation feature $\bm{m} = (m_1, \cdots, m_L)$, not on the index $j$.
-Then, the expected complete log-likelihood augmented by Lagrange multipliers is calculated as
-\begin{equation*}
-\sum_{i=1}^I \sum_{\bm{m}} g_{i, \bm{m} } \sum_{k=1}^K \theta_{i, k, \bm{m}}   \Bigl( \sum_{l = 1}^L \log f_{k, l, m_l} + \log q_{i,k} \Bigr)
-+ \sum_{k=1}^K \sum_{l=1}^L \tau_{k,l} \bigl( 1 - \sum_{p=1}^{M_l} f_{k,l,p} \bigr)
-+ \sum_{i=1}^I \rho_i \bigl( 1 - \sum_{k=1}^K q_{i,k} \bigr).
-\end{equation*}
-Differentiating it leads to following stationary equations: 
-\begin{align*}
-\sum_{i=1}^I \sum_{\bm{m}: m_l = p} g_{i, \bm{m}} \theta_{i,k,\bm{m}} - \tau_{k,l} f_{k,l,p} & = 0,\ (p=1, \cdots, M_l, k=1, \cdots, K,\ l=1, \cdots, L)., \\
-\sum_{\bm{m}} g_{i,\bm{m}} \theta_{i,k,\bm{m}} - \rho_i q_{i, k} & = 0,\ (k = 1, \cdots, K, i = 1, \cdots, I).
-\end{align*}
-Then, by eliminating Lagrange multipliers, updating rules can be obtained.
+As Fig. 1(a) shows, we estimate the mutation signatures very accurately overall (see Fig. 1(b) for an example).
+As expected, accuracy improves as we increase the number of cancer genomes or mutations.
+Also, as the mutation feature dispersion decreases ($\alpha$ increases), accuracy decreases.
+This is expected: as $\alpha$ increases the signatures will become more ``fuzzy", with
+probability mass spread over a large number of mutation patterns, and so harder to infer
+(In other words, larger $\alpha$ corresponds to a less informative motif).
+In contrast, accuracy was relatively insensitive to 
+individual membership dispersion ($\gamma$), and estimates remain accurate even when most individuals have mutations in many different signatures.
 
-\section*{Relationship with nonnegative matrix factorization}
-
-First, for ease of explanation, let assume that the full representation'' representation ($L = 1$) is used.
-Suppose that each $\bm{m}$ has unique appropriate index from 1 to  $| \bm{M} | = \prod_{l=1}^L M_l$ (the number of possible mutation patterns),
-so that $\bm{m}$ can be indices of matrices.
-
-Let $G = \{ g_{i, \bm{m}} \}$ denote the $I \times | \bm{M} | $ matrix, 
-where $g_{i, \bm{m}}$ is the number of mutations whose mutation patters are equal to $\bm{m}$ in the $i$-th cancer genome.
-Nonnegative matrix factorization aims to find low rank decomposition, 
-$G \sim \tilde{Q} F$,
-where $\tilde{Q} = \{ \tilde{q}_{i,k} \}$ and $F = \{ f_{k, \bm{m}} \}$ are nonnegative matrix, 
-and row vectors of $F$ are often restricted to be sum to one. 
-We used the notation $\tilde{Q}$ instead of $Q$
-to represent that the row vectors of $\tilde{Q}$ are not normalized to sum to one in general.
-
-For solving NMF, 
-the previous study (Lee et al. 2000) used the following updating rule:
-\begin{equation*}
-f_{k,m} \gets f_{k,m} \frac{ ( \tilde{Q}^T G)_{k,m} }{ ( \tilde{Q}^T \tilde{Q} F  )_{k,m} },\  \
-\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ (G F^T)_{i,k} }{ ( \tilde{Q} F F^T )_{i,k} },
-\end{equation*}
-that reduces the {\it Euclidean distance} $|| G - \tilde{Q} F ||$. 
-Therefore, the optimization problem for the existing approach is
-\begin{equation}
-\begin{aligned}
-& \text{minimize}
-& & ||G - \tilde{Q} F || \\
-& \text{subject to}
-& & \sum_{ \bm{m} } f_{k, \bm{m}} =  1, \; k = 1, \cdots, K \\ &
-& & f_{k, \bm{m}} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\end{equation}
-
-On the other hand, there is another type of updating rule:
-\begin{equation*}
-f_{k,m} \gets f_{k,m} \frac{ \sum_i \tilde{q}_{i,k} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_i \tilde{q}_{i,k} },
-\end{equation*}
-\begin{equation*}
-\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ \sum_m f_{k,m} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_m f_{k,m} }.
-\end{equation*}
-that reduces the Kullback-Liebler Divergence:
-\begin{equation*}
-KL(G ||  \tilde{Q} F) = \sum_{i,m} \Bigl( g_{i,m} \log  \frac{ g_{i,m} }{ (\tilde{Q} F)_{i,m} } - g_{i,m} + ( \tilde{Q} f )_{i,m} \Bigr).
-\end{equation*}
-
-In general cases including the independent representation, 
-there is restrictions $f_{k, \bm{m}} = \prod_l f_{k, l, m_l}$ by smaller set of parameters.
-Let us consider the following optimization problem with the Kullback-Liebler Divergence and the restrictions on $F$:
-\begin{equation}
-\begin{aligned}
-& \text{minimize}
-& & KL(G ||  \tilde{Q} F) \\
-& \text{subject to}
-& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\label{KL}
-\end{equation}
-
-In fact, this is equivalent to the proposed method, whose optimization problem can be written as:
-\begin{equation}
-\begin{aligned}
-& \text{maximize}
-& & L(Q, F | G) \bigl(= \sum_{i,m}  g_{i,m} \log (Q F)_{i,m} \bigr)  \\
-& \text{subject to}
-& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \sum_k  q_{i, k} =  1, \; i = 1, \cdots, I \\ &
-& & q_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\label{likelihood}
-\end{equation}
-
-\begin{prop}
-When $(Q, F) = (Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}), 
-then $(\tilde{Q}, F) = (R^* Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}).
-On the other hand, when $(\tilde{Q}, F) = (\tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}),
-then $(Q, F) = (R^{*-1} \tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}),
-where $R^* = \text{diag} (r_1^*, \cdots, r_I^*), r_i^* = \sum_{\bm{m}} g_{i,\bm{m}}, i = 1, \cdots, I$.
-\end{prop}
-\proof
-This is because
-\begin{equation*}
-KL(G ||  \tilde{Q} F) = -  \sum_i \Bigl( ( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i \Bigr)  - L(Q, F | G) + \text{(constant value)},
-\end{equation*}
-where $Q$ is row-normalized matrix for $\tilde{Q}$, $\tilde{r}_i = \sum_k q_{i, k}$ for each $i$, 
-and $( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i$ takes its maximum at $\tilde{r}_i = r_i^*$.
-\hfill $\square$
+In most cases, the log-likelihood stopped increasing at $K=5$ mutation signatures, 
+whereas the standard error of the estimated parameters started increasing past $K=5$ (see Fig. 1(c)). We conclude that examining
+the trade-off between likelihood and standard-errors may be a helpful guide to selecting an appropriate number of mutation signatures.
 
@@
 
-\clearpage
+\begin{figure}
 
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
-\caption{An example of the relationship between true and estimated mutation signatures in the simulation.
+\subfigure[The accuracy of the proposed approach for the simulated data when changing the number of samples, mutations, 
+and the amounts of dispersion parameters ($\alpha$ and $\gamma$) for the mutation features and signature distribution parameters.]{%
+\includegraphics[width=16cm,height=6cm]{simulation_result.eps}
+\label{simulation_facet}
+}
+
+\subfigure[An example of the relationship between true and estimated mutation signatures in the simulation.
 In this example, the numbers of cancer genomes and mutations for each cancer genome are 25 and 100, respectively,
 and the parameters $\alpha$ and $\gamma$ are both set to 1. 
-From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations}
+From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations]{%
+\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
 \label{example_signature.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in the trial of the Supplementary Figure 1.}
-\label{example_summary.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[$K = 2$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K2.eps}
-  \label{UTUC:sigK2_sig}}
-
-\subfigure[$K = 3$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K3.eps}
-  \label{UTUC:sigK3_sig}}
-
-\subfigure[$K = 4$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K4.eps}
-  \label{UTUC:sigK4_sig}}
-
-\caption{The result of estimated mutation signatures and membership parameters for UTUC data 
-when changing the number of mutation signatures $K$.}
-\label{UTUC_multiK}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{UTUC_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in UTUC data.}
-\label{UTUC_stat}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm,height=10cm]{UTUC_AA_AAlike_cor.eps}
-\caption{The relationships between the two membership parameters, AA (the first signature in the Supplementary Figure 3) and AA\_like (the second signature in the Supplementary Figure 3) signatures.}
-\label{UTUT_AA_AAlike}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=4cm,height=8cm]{UTUC_APOBEC_twoFivePrime.eps}
-\caption{The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
-\label{UTUC_APOBEC_twoFivePrime}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}
-
-\subfigure[APOBEC signatures obtained in each cancer type]{%
-  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
-  \label{APOBEC_sig_list}}
-
-\subfigure[Smoking signature  in each cancer type]{%
-  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
-  \label{LUNG_sig_list}}
-
-\subfigure[POLE1 signatures in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-  \label{POLE1_sig_list}}
-\quad
-\subfigure[POLE2 signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-  \label{POLE2_sig_list}}
-
-\subfigure[UV signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-  \label{UV_sig_list}}
-\quad
-\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-  \label{LMST_sig_list}}
-
-% \subfigure[Signature intensities at two 5' to the mutated site]{%
-% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
- % \label{APOBEC_twofive}}
-
-\caption{(a) APOBEC signatures obtained in 11 cancer types, 
-(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
-(f) the frequencies of bases at two 5' to the mutated site. 
-}
-\label{nature2013_apobec}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in Alexandrov et al. (2013)]{%
-  \includegraphics[height=3cm,width=12cm,clip]{example96_Glioma-Low-Grade.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter by the proposed method]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter by the proposed method in log scale]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership_log.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a) The barplot is divided by 6 substitution patterns. In each division, 
-16 bars show joint probabilities of 16 combinations of the immediate 5' and 3' bases.
-(b, c) We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows (the logarithm of) the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-The low grade glioma specific signature detected by the proposed method is the signature 2.
-We can see that the mutations corresponding to signature 2 is mostly from the sample with an extremely high mutation rate.
 }
 
-\end{figure}
+\subfigure[The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters 
+for several numbers of signatures $K$ in the trial of the above figure.]{%
+\includegraphics[width=12cm,height=4cm]{simulation_stat.eps}
+\label{example_summary.pd}
+}
 
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The result for the original data ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_original_signature_K2.eps}
-  \label{LGG_roriginal_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K2.eps}
-  \label{LGG_removed_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 3$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K3.eps}
-  \label{LGG_removed_sigK3_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 4$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K4.eps}
-  \label{LGG_removed_sigK4_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 5$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K5.eps}
-  \label{LGG_removed_sigK5_sig}}
-
-  \caption{(a) The signatures obtained for the original data. The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
-  The second signature is the low grade glioma signature. 
-  (b, c, d, e) The signature obtained for the data without the hyper-mutated case for several number of mutation signatures ($K$, including the background signature).
-  Although the signature related to deamination of 5-methyl-cytosine remained, low grade glioma specific signature could not be observed.}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The signature 2 detected in kidney clear cell carcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Kidney-Clear-Cell_oxidation_signature.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[The signature 1 detected in lung adenocarcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Lung-Adeno_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-\quad
-\subfigure[The signature 1 detected in melanomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Melanoma_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-
-\subfigure[Estimated membership parameter for kidney clear cell carcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Kidney-Clear-Cell_oxidation_membership.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter for lung adenocarcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Lung-Adeno_oxidation_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter for melanomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Melanoma_oxidation_membership.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a, b, c) The putative oxidative artifact signatures estimated for each cancer.
-  We can observe consistent abundance of the base C at the $-2 $ position.
-  (d, e, f) For each cancer type, 
-We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-We can see that the signature corresponding to putative oxidative artifacts concentrates on a small number of samples.}
+\caption{The summary of simulation study.}
 

### 2015-05-25 paper/TextS2/TextS2.tex
organized directory

--- 2015-03-19 supp/supp.tex
@@
 \vspace*{1.0cm}
-{\LARGE \bf Supplementary material}
+{\LARGE \bf Experiment on UCUT data with the various number of mutation signatures}
 \vspace*{0.25cm}
 
-\section*{Derivation of EM algorithm}
+We analyzed the data using our new model, with increasing number of mutation signatures $K=2,3,\dots$. 
+Results are shown in Figure 1.
+As expected, the likelihood increased with $K$. Bootstrap-errors started to increase at $K = 5$ (Figure 2).
 
-The complete log-likelihood including missing data $\{ \bm{z}_i \}$ for the proposed model is
-\begin{equation*}
-\sum_{i = 1}^I \sum_{j=1}^J \sum_{k=1}^K I(z_{i,j} = k) \Bigl( \sum_{l = 1}^L \log f_{k, l, x_{i,j,l}} + \log q_{i,k} \Bigr).
-\end{equation*}
-Here, we introduce the variable for conditional probability for $z_{i,j}$ given the parameters and the mutation features $\bm{x}_{i,j}$,
-\begin{equation*}
-\theta_{i,k, \bm{m}} = \Pr \bigl( z_{i,j} = k \big| \bm{x}_{i, j} = \bm{m}, \{ \bm{f}_{k, l} \}, \{  \bm{q}_i \} \bigr)
-\end{equation*}
-Note that this conditional probability just depends on the value of mutation feature $\bm{m} = (m_1, \cdots, m_L)$, not on the index $j$.
-Then, the expected complete log-likelihood augmented by Lagrange multipliers is calculated as
-\begin{equation*}
-\sum_{i=1}^I \sum_{\bm{m}} g_{i, \bm{m} } \sum_{k=1}^K \theta_{i, k, \bm{m}}   \Bigl( \sum_{l = 1}^L \log f_{k, l, m_l} + \log q_{i,k} \Bigr)
-+ \sum_{k=1}^K \sum_{l=1}^L \tau_{k,l} \bigl( 1 - \sum_{p=1}^{M_l} f_{k,l,p} \bigr)
-+ \sum_{i=1}^I \rho_i \bigl( 1 - \sum_{k=1}^K q_{i,k} \bigr).
-\end{equation*}
-Differentiating it leads to following stationary equations: 
-\begin{align*}
-\sum_{i=1}^I \sum_{\bm{m}: m_l = p} g_{i, \bm{m}} \theta_{i,k,\bm{m}} - \tau_{k,l} f_{k,l,p} & = 0,\ (p=1, \cdots, M_l, k=1, \cdots, K,\ l=1, \cdots, L)., \\
-\sum_{\bm{m}} g_{i,\bm{m}} \theta_{i,k,\bm{m}} - \rho_i q_{i, k} & = 0,\ (k = 1, \cdots, K, i = 1, \cdots, I).
-\end{align*}
-Then, by eliminating Lagrange multipliers, updating rules can be obtained.
+With $K = 2$ (Figure 1(a)) we observed a mutation signature that appears to correspond to AA  (T $>$ A substitutions at CpTpG sites with strong transcription strand specificity).
+Increasing to  $K=3$ introduced an additional mutation signature corresponding to the APOBEC enzyme (C $>$ [AGT] at TpCpN sites) (Figure 1(b)).
+Increasing to $K = 4$ introduced an additional signature (T $>$ A at NpTpN sites with strong strand specificity) 
+that is somewhat similar to the AA signature 
+(Figure 1(c)). 
+The membership in the AA signature and this ``AA-like" signature were correlated ($R = 0.77$; Figure 3). 
+This additional signature may be just making up for the residual of the AA signature 
+which the original AA signature could not explain due to a slight deviance of the probabilistic model.
 
-\section*{Relationship with nonnegative matrix factorization}
+The strong correlation among estimated membership parameters started to be shown at $K = 4$.
+Considering all these factors together, $K = 3$ seems to be a reasonable choice in terms of the interpretability,
+and we adopted $K = 3$.
 
-First, for ease of explanation, let assume that the full representation'' representation ($L = 1$) is used.
-Suppose that each $\bm{m}$ has unique appropriate index from 1 to  $| \bm{M} | = \prod_{l=1}^L M_l$ (the number of possible mutation patterns),
-so that $\bm{m}$ can be indices of matrices.
-
-Let $G = \{ g_{i, \bm{m}} \}$ denote the $I \times | \bm{M} | $ matrix, 
-where $g_{i, \bm{m}}$ is the number of mutations whose mutation patters are equal to $\bm{m}$ in the $i$-th cancer genome.
-Nonnegative matrix factorization aims to find low rank decomposition, 
-$G \sim \tilde{Q} F$,
-where $\tilde{Q} = \{ \tilde{q}_{i,k} \}$ and $F = \{ f_{k, \bm{m}} \}$ are nonnegative matrix, 
-and row vectors of $F$ are often restricted to be sum to one. 
-We used the notation $\tilde{Q}$ instead of $Q$
-to represent that the row vectors of $\tilde{Q}$ are not normalized to sum to one in general.
-
-For solving NMF, 
-the previous study (Lee et al. 2000) used the following updating rule:
-\begin{equation*}
-f_{k,m} \gets f_{k,m} \frac{ ( \tilde{Q}^T G)_{k,m} }{ ( \tilde{Q}^T \tilde{Q} F  )_{k,m} },\  \
-\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ (G F^T)_{i,k} }{ ( \tilde{Q} F F^T )_{i,k} },
-\end{equation*}
-that reduces the {\it Euclidean distance} $|| G - \tilde{Q} F ||$. 
-Therefore, the optimization problem for the existing approach is
-\begin{equation}
-\begin{aligned}
-& \text{minimize}
-& & ||G - \tilde{Q} F || \\
-& \text{subject to}
-& & \sum_{ \bm{m} } f_{k, \bm{m}} =  1, \; k = 1, \cdots, K \\ &
-& & f_{k, \bm{m}} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\end{equation}
-
-On the other hand, there is another type of updating rule:
-\begin{equation*}
-f_{k,m} \gets f_{k,m} \frac{ \sum_i \tilde{q}_{i,k} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_i \tilde{q}_{i,k} },
-\end{equation*}
-\begin{equation*}
-\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ \sum_m f_{k,m} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_m f_{k,m} }.
-\end{equation*}
-that reduces the Kullback-Liebler Divergence:
-\begin{equation*}
-KL(G ||  \tilde{Q} F) = \sum_{i,m} \Bigl( g_{i,m} \log  \frac{ g_{i,m} }{ (\tilde{Q} F)_{i,m} } - g_{i,m} + ( \tilde{Q} f )_{i,m} \Bigr).
-\end{equation*}
-
-In general cases including the independent representation, 
-there is restrictions $f_{k, \bm{m}} = \prod_l f_{k, l, m_l}$ by smaller set of parameters.
-Let us consider the following optimization problem with the Kullback-Liebler Divergence and the restrictions on $F$:
-\begin{equation}
-\begin{aligned}
-& \text{minimize}
-& & KL(G ||  \tilde{Q} F) \\
-& \text{subject to}
-& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\label{KL}
-\end{equation}
-
-In fact, this is equivalent to the proposed method, whose optimization problem can be written as:
-\begin{equation}
-\begin{aligned}
-& \text{maximize}
-& & L(Q, F | G) \bigl(= \sum_{i,m}  g_{i,m} \log (Q F)_{i,m} \bigr)  \\
-& \text{subject to}
-& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \sum_k  q_{i, k} =  1, \; i = 1, \cdots, I \\ &
-& & q_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\label{likelihood}
-\end{equation}
-
-\begin{prop}
-When $(Q, F) = (Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}), 
-then $(\tilde{Q}, F) = (R^* Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}).
-On the other hand, when $(\tilde{Q}, F) = (\tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}),
-then $(Q, F) = (R^{*-1} \tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}),
-where $R^* = \text{diag} (r_1^*, \cdots, r_I^*), r_i^* = \sum_{\bm{m}} g_{i,\bm{m}}, i = 1, \cdots, I$.
-\end{prop}
-\proof
-This is because
-\begin{equation*}
-KL(G ||  \tilde{Q} F) = -  \sum_i \Bigl( ( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i \Bigr)  - L(Q, F | G) + \text{(constant value)},
-\end{equation*}
-where $Q$ is row-normalized matrix for $\tilde{Q}$, $\tilde{r}_i = \sum_k q_{i, k}$ for each $i$, 
-and $( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i$ takes its maximum at $\tilde{r}_i = r_i^*$.
-\hfill $\square$
-
-% \bibliographystyle{natbib}
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
 \bibliographystyle{plain}
@@
 
-\clearpage
-
 \begin{figure}
-\centering
-\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
-\caption{An example of the relationship between true and estimated mutation signatures in the simulation.
-In this example, the numbers of cancer genomes and mutations for each cancer genome are 25 and 100, respectively,
-and the parameters $\alpha$ and $\gamma$ are both set to 1. 
-From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations}
-\label{example_signature.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in the trial of the Supplementary Figure 1.}
-\label{example_summary.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
 
@@
 
-\clearpage
-
 \begin{figure}
@@
 \includegraphics[width=10cm,height=10cm]{UTUC_AA_AAlike_cor.eps}
-\caption{The relationships between the two membership parameters, AA (the first signature in the Supplementary Figure 3) and AA\_like (the second signature in the Supplementary Figure 3) signatures.}
+\caption{The relationships between the two membership parameters, AA (the first signature in the Figure 1(c)) and AA\_like (the second signature in the Figure 1(c)) signatures.}
 \label{UTUT_AA_AAlike}
@@
 
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=4cm,height=8cm]{UTUC_APOBEC_twoFivePrime.eps}
-\caption{The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
-\label{UTUC_APOBEC_twoFivePrime}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}
-
-\subfigure[APOBEC signatures obtained in each cancer type]{%
-  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
-  \label{APOBEC_sig_list}}
-
-\subfigure[Smoking signature  in each cancer type]{%
-  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
-  \label{LUNG_sig_list}}
-
-\subfigure[POLE1 signatures in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-  \label{POLE1_sig_list}}
-\quad
-\subfigure[POLE2 signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-  \label{POLE2_sig_list}}
-
-\subfigure[UV signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-  \label{UV_sig_list}}
-\quad
-\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-  \label{LMST_sig_list}}
-
-% \subfigure[Signature intensities at two 5' to the mutated site]{%
-% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
- % \label{APOBEC_twofive}}
-
-\caption{(a) APOBEC signatures obtained in 11 cancer types, 
-(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
-(f) the frequencies of bases at two 5' to the mutated site. 
-}
-\label{nature2013_apobec}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in Alexandrov et al. (2013)]{%
-  \includegraphics[height=3cm,width=12cm,clip]{example96_Glioma-Low-Grade.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter by the proposed method]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter by the proposed method in log scale]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership_log.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a) The barplot is divided by 6 substitution patterns. In each division, 
-16 bars show joint probabilities of 16 combinations of the immediate 5' and 3' bases.
-(b, c) We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows (the logarithm of) the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-The low grade glioma specific signature detected by the proposed method is the signature 2.
-We can see that the mutations corresponding to signature 2 is mostly from the sample with an extremely high mutation rate.
-}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The result for the original data ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_original_signature_K2.eps}
-  \label{LGG_roriginal_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K2.eps}
-  \label{LGG_removed_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 3$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K3.eps}
-  \label{LGG_removed_sigK3_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 4$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K4.eps}
-  \label{LGG_removed_sigK4_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 5$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K5.eps}
-  \label{LGG_removed_sigK5_sig}}
-
-  \caption{(a) The signatures obtained for the original data. The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
-  The second signature is the low grade glioma signature. 
-  (b, c, d, e) The signature obtained for the data without the hyper-mutated case for several number of mutation signatures ($K$, including the background signature).
-  Although the signature related to deamination of 5-methyl-cytosine remained, low grade glioma specific signature could not be observed.}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The signature 2 detected in kidney clear cell carcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Kidney-Clear-Cell_oxidation_signature.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[The signature 1 detected in lung adenocarcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Lung-Adeno_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-\quad
-\subfigure[The signature 1 detected in melanomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Melanoma_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-
-\subfigure[Estimated membership parameter for kidney clear cell carcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Kidney-Clear-Cell_oxidation_membership.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter for lung adenocarcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Lung-Adeno_oxidation_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter for melanomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Melanoma_oxidation_membership.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a, b, c) The putative oxidative artifact signatures estimated for each cancer.
-  We can observe consistent abundance of the base C at the $-2 $ position.
-  (d, e, f) For each cancer type, 
-We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-We can see that the signature corresponding to putative oxidative artifacts concentrates on a small number of samples.}
-
-\end{figure}
-
 \end{document}

### 2015-05-25 paper/TextS3/TextS3.tex
organized directory

--- 2015-03-19 supp/supp.tex
@@
 \vspace*{1.0cm}
-{\LARGE \bf Supplementary material}
+{\LARGE \bf Derivation of EM algorithm}
 \vspace*{0.25cm}
-
-\section*{Derivation of EM algorithm}
 
@@
 
-\section*{Relationship with nonnegative matrix factorization}
-
-First, for ease of explanation, let assume that the full representation'' representation ($L = 1$) is used.
-Suppose that each $\bm{m}$ has unique appropriate index from 1 to  $| \bm{M} | = \prod_{l=1}^L M_l$ (the number of possible mutation patterns),
-so that $\bm{m}$ can be indices of matrices.
-
-Let $G = \{ g_{i, \bm{m}} \}$ denote the $I \times | \bm{M} | $ matrix, 
-where $g_{i, \bm{m}}$ is the number of mutations whose mutation patters are equal to $\bm{m}$ in the $i$-th cancer genome.
-Nonnegative matrix factorization aims to find low rank decomposition, 
-$G \sim \tilde{Q} F$,
-where $\tilde{Q} = \{ \tilde{q}_{i,k} \}$ and $F = \{ f_{k, \bm{m}} \}$ are nonnegative matrix, 
-and row vectors of $F$ are often restricted to be sum to one. 
-We used the notation $\tilde{Q}$ instead of $Q$
-to represent that the row vectors of $\tilde{Q}$ are not normalized to sum to one in general.
-
-For solving NMF, 
-the previous study (Lee et al. 2000) used the following updating rule:
-\begin{equation*}
-f_{k,m} \gets f_{k,m} \frac{ ( \tilde{Q}^T G)_{k,m} }{ ( \tilde{Q}^T \tilde{Q} F  )_{k,m} },\  \
-\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ (G F^T)_{i,k} }{ ( \tilde{Q} F F^T )_{i,k} },
-\end{equation*}
-that reduces the {\it Euclidean distance} $|| G - \tilde{Q} F ||$. 
-Therefore, the optimization problem for the existing approach is
-\begin{equation}
-\begin{aligned}
-& \text{minimize}
-& & ||G - \tilde{Q} F || \\
-& \text{subject to}
-& & \sum_{ \bm{m} } f_{k, \bm{m}} =  1, \; k = 1, \cdots, K \\ &
-& & f_{k, \bm{m}} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\end{equation}
-
-On the other hand, there is another type of updating rule:
-\begin{equation*}
-f_{k,m} \gets f_{k,m} \frac{ \sum_i \tilde{q}_{i,k} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_i \tilde{q}_{i,k} },
-\end{equation*}
-\begin{equation*}
-\tilde{q}_{i,k} \gets \tilde{q}_{i,k}  \frac{ \sum_m f_{k,m} g_{i,m} / ( \tilde{Q} F)_{i,m} }{ \sum_m f_{k,m} }.
-\end{equation*}
-that reduces the Kullback-Liebler Divergence:
-\begin{equation*}
-KL(G ||  \tilde{Q} F) = \sum_{i,m} \Bigl( g_{i,m} \log  \frac{ g_{i,m} }{ (\tilde{Q} F)_{i,m} } - g_{i,m} + ( \tilde{Q} f )_{i,m} \Bigr).
-\end{equation*}
-
-In general cases including the independent representation, 
-there is restrictions $f_{k, \bm{m}} = \prod_l f_{k, l, m_l}$ by smaller set of parameters.
-Let us consider the following optimization problem with the Kullback-Liebler Divergence and the restrictions on $F$:
-\begin{equation}
-\begin{aligned}
-& \text{minimize}
-& & KL(G ||  \tilde{Q} F) \\
-& \text{subject to}
-& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \tilde{q}_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\label{KL}
-\end{equation}
-
-In fact, this is equivalent to the proposed method, whose optimization problem can be written as:
-\begin{equation}
-\begin{aligned}
-& \text{maximize}
-& & L(Q, F | G) \bigl(= \sum_{i,m}  g_{i,m} \log (Q F)_{i,m} \bigr)  \\
-& \text{subject to}
-& & f_{k, \bm{m}} = \prod_l f_{k, l, m_l}, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & f_{k, l, p} \geq 0, \; k = 1, \cdots, K,\ \bm{m} \in M \\ & 
-& & \sum_k  q_{i, k} =  1, \; i = 1, \cdots, I \\ &
-& & q_{i, k} \geq 0, \; i = 1, \cdots, I,\ k = 1, \cdots, K.
-\end{aligned}
-\label{likelihood}
-\end{equation}
-
-\begin{prop}
-When $(Q, F) = (Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}), 
-then $(\tilde{Q}, F) = (R^* Q^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}).
-On the other hand, when $(\tilde{Q}, F) = (\tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{KL}),
-then $(Q, F) = (R^{*-1} \tilde{Q}^*, F^*)$ is an optimal solution of the optimization problem (\ref{likelihood}),
-where $R^* = \text{diag} (r_1^*, \cdots, r_I^*), r_i^* = \sum_{\bm{m}} g_{i,\bm{m}}, i = 1, \cdots, I$.
-\end{prop}
-\proof
-This is because
-\begin{equation*}
-KL(G ||  \tilde{Q} F) = -  \sum_i \Bigl( ( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i \Bigr)  - L(Q, F | G) + \text{(constant value)},
-\end{equation*}
-where $Q$ is row-normalized matrix for $\tilde{Q}$, $\tilde{r}_i = \sum_k q_{i, k}$ for each $i$, 
-and $( \sum_{m} g_{i, m} ) \log \tilde{r}_i - \tilde{r}_i$ takes its maximum at $\tilde{r}_i = r_i^*$.
-\hfill $\square$
-
-% \bibliographystyle{natbib}
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
-\bibliographystyle{plain}
-\bibliography{document}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
-\caption{An example of the relationship between true and estimated mutation signatures in the simulation.
-In this example, the numbers of cancer genomes and mutations for each cancer genome are 25 and 100, respectively,
-and the parameters $\alpha$ and $\gamma$ are both set to 1. 
-From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations}
-\label{example_signature.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in the trial of the Supplementary Figure 1.}
-\label{example_summary.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[$K = 2$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K2.eps}
-  \label{UTUC:sigK2_sig}}
-
-\subfigure[$K = 3$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K3.eps}
-  \label{UTUC:sigK3_sig}}
-
-\subfigure[$K = 4$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K4.eps}
-  \label{UTUC:sigK4_sig}}
-
-\caption{The result of estimated mutation signatures and membership parameters for UTUC data 
-when changing the number of mutation signatures $K$.}
-\label{UTUC_multiK}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{UTUC_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in UTUC data.}
-\label{UTUC_stat}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm,height=10cm]{UTUC_AA_AAlike_cor.eps}
-\caption{The relationships between the two membership parameters, AA (the first signature in the Supplementary Figure 3) and AA\_like (the second signature in the Supplementary Figure 3) signatures.}
-\label{UTUT_AA_AAlike}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=4cm,height=8cm]{UTUC_APOBEC_twoFivePrime.eps}
-\caption{The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
-\label{UTUC_APOBEC_twoFivePrime}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}
-
-\subfigure[APOBEC signatures obtained in each cancer type]{%
-  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
-  \label{APOBEC_sig_list}}
-
-\subfigure[Smoking signature  in each cancer type]{%
-  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
-  \label{LUNG_sig_list}}
-
-\subfigure[POLE1 signatures in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-  \label{POLE1_sig_list}}
-\quad
-\subfigure[POLE2 signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-  \label{POLE2_sig_list}}
-
-\subfigure[UV signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-  \label{UV_sig_list}}
-\quad
-\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-  \label{LMST_sig_list}}
-
-% \subfigure[Signature intensities at two 5' to the mutated site]{%
-% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
- % \label{APOBEC_twofive}}
-
-\caption{(a) APOBEC signatures obtained in 11 cancer types, 
-(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
-(f) the frequencies of bases at two 5' to the mutated site. 
-}
-\label{nature2013_apobec}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in Alexandrov et al. (2013)]{%
-  \includegraphics[height=3cm,width=12cm,clip]{example96_Glioma-Low-Grade.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter by the proposed method]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter by the proposed method in log scale]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership_log.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a) The barplot is divided by 6 substitution patterns. In each division, 
-16 bars show joint probabilities of 16 combinations of the immediate 5' and 3' bases.
-(b, c) We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows (the logarithm of) the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-The low grade glioma specific signature detected by the proposed method is the signature 2.
-We can see that the mutations corresponding to signature 2 is mostly from the sample with an extremely high mutation rate.
-}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The result for the original data ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_original_signature_K2.eps}
-  \label{LGG_roriginal_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K2.eps}
-  \label{LGG_removed_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 3$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K3.eps}
-  \label{LGG_removed_sigK3_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 4$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K4.eps}
-  \label{LGG_removed_sigK4_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 5$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K5.eps}
-  \label{LGG_removed_sigK5_sig}}
-
-  \caption{(a) The signatures obtained for the original data. The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
-  The second signature is the low grade glioma signature. 
-  (b, c, d, e) The signature obtained for the data without the hyper-mutated case for several number of mutation signatures ($K$, including the background signature).
-  Although the signature related to deamination of 5-methyl-cytosine remained, low grade glioma specific signature could not be observed.}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The signature 2 detected in kidney clear cell carcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Kidney-Clear-Cell_oxidation_signature.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[The signature 1 detected in lung adenocarcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Lung-Adeno_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-\quad
-\subfigure[The signature 1 detected in melanomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Melanoma_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-
-\subfigure[Estimated membership parameter for kidney clear cell carcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Kidney-Clear-Cell_oxidation_membership.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter for lung adenocarcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Lung-Adeno_oxidation_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter for melanomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Melanoma_oxidation_membership.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a, b, c) The putative oxidative artifact signatures estimated for each cancer.
-  We can observe consistent abundance of the base C at the $-2 $ position.
-  (d, e, f) For each cancer type, 
-We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-We can see that the signature corresponding to putative oxidative artifacts concentrates on a small number of samples.}
-
-\end{figure}
-
 \end{document}

### 2015-05-25 paper/TextS4/TextS4.tex
organized directory

--- 2015-03-19 supp/supp.tex
@@
 \vspace*{1.0cm}
-{\LARGE \bf Supplementary material}
+{\LARGE \bf Relationship with nonnegative matrix factorization}
 \vspace*{0.25cm}
-
-\section*{Derivation of EM algorithm}
-
-The complete log-likelihood including missing data $\{ \bm{z}_i \}$ for the proposed model is
-\begin{equation*}
-\sum_{i = 1}^I \sum_{j=1}^J \sum_{k=1}^K I(z_{i,j} = k) \Bigl( \sum_{l = 1}^L \log f_{k, l, x_{i,j,l}} + \log q_{i,k} \Bigr).
-\end{equation*}
-Here, we introduce the variable for conditional probability for $z_{i,j}$ given the parameters and the mutation features $\bm{x}_{i,j}$,
-\begin{equation*}
-\theta_{i,k, \bm{m}} = \Pr \bigl( z_{i,j} = k \big| \bm{x}_{i, j} = \bm{m}, \{ \bm{f}_{k, l} \}, \{  \bm{q}_i \} \bigr)
-\end{equation*}
-Note that this conditional probability just depends on the value of mutation feature $\bm{m} = (m_1, \cdots, m_L)$, not on the index $j$.
-Then, the expected complete log-likelihood augmented by Lagrange multipliers is calculated as
-\begin{equation*}
-\sum_{i=1}^I \sum_{\bm{m}} g_{i, \bm{m} } \sum_{k=1}^K \theta_{i, k, \bm{m}}   \Bigl( \sum_{l = 1}^L \log f_{k, l, m_l} + \log q_{i,k} \Bigr)
-+ \sum_{k=1}^K \sum_{l=1}^L \tau_{k,l} \bigl( 1 - \sum_{p=1}^{M_l} f_{k,l,p} \bigr)
-+ \sum_{i=1}^I \rho_i \bigl( 1 - \sum_{k=1}^K q_{i,k} \bigr).
-\end{equation*}
-Differentiating it leads to following stationary equations: 
-\begin{align*}
-\sum_{i=1}^I \sum_{\bm{m}: m_l = p} g_{i, \bm{m}} \theta_{i,k,\bm{m}} - \tau_{k,l} f_{k,l,p} & = 0,\ (p=1, \cdots, M_l, k=1, \cdots, K,\ l=1, \cdots, L)., \\
-\sum_{\bm{m}} g_{i,\bm{m}} \theta_{i,k,\bm{m}} - \rho_i q_{i, k} & = 0,\ (k = 1, \cdots, K, i = 1, \cdots, I).
-\end{align*}
-Then, by eliminating Lagrange multipliers, updating rules can be obtained.
-
-\section*{Relationship with nonnegative matrix factorization}
 
@@
 
-% \bibliographystyle{natbib}
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
-\bibliographystyle{plain}
-\bibliography{document}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5.5cm]{simulation_signature_example.eps}
-\caption{An example of the relationship between true and estimated mutation signatures in the simulation.
-In this example, the numbers of cancer genomes and mutations for each cancer genome are 25 and 100, respectively,
-and the parameters $\alpha$ and $\gamma$ are both set to 1. 
-From this figure, we can see that fairly accurate estimation is possible even with moderate numbers of cancer genomes and mutations}
-\label{example_signature.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{simulation_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in the trial of the Supplementary Figure 1.}
-\label{example_summary.pd}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[$K = 2$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K2.eps}
-  \label{UTUC:sigK2_sig}}
-
-\subfigure[$K = 3$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K3.eps}
-  \label{UTUC:sigK3_sig}}
-
-\subfigure[$K = 4$]{%
-  \includegraphics[height=6cm,width=15cm,clip]{UTUC_signature_K4.eps}
-  \label{UTUC:sigK4_sig}}
-
-\caption{The result of estimated mutation signatures and membership parameters for UTUC data 
-when changing the number of mutation signatures $K$.}
-\label{UTUC_multiK}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=15cm,height=5cm]{UTUC_stat.eps}
-\caption{The log-likelihood, bootstrap-errors and maximum correlation values among estimated membership parameters for several numbers of signatures $K$ in UTUC data.}
-\label{UTUC_stat}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm,height=10cm]{UTUC_AA_AAlike_cor.eps}
-\caption{The relationships between the two membership parameters, AA (the first signature in the Supplementary Figure 3) and AA\_like (the second signature in the Supplementary Figure 3) signatures.}
-\label{UTUT_AA_AAlike}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=4cm,height=8cm]{UTUC_APOBEC_twoFivePrime.eps}
-\caption{The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
-\label{UTUC_APOBEC_twoFivePrime}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}
-
-\subfigure[APOBEC signatures obtained in each cancer type]{%
-  \includegraphics[width=15cm,height=7.5cm]{signatureList_APOBEC.eps}
-  \label{APOBEC_sig_list}}
-
-\subfigure[Smoking signature  in each cancer type]{%
-  \includegraphics[width=15cm,height=2.5cm]{signatureList_LUNG.eps}
-  \label{LUNG_sig_list}}
-
-\subfigure[POLE1 signatures in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-  \label{POLE1_sig_list}}
-\quad
-\subfigure[POLE2 signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-  \label{POLE2_sig_list}}
-
-\subfigure[UV signature in each cancer type]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-  \label{UV_sig_list}}
-\quad
-\subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-  \label{LMST_sig_list}}
-
-% \subfigure[Signature intensities at two 5' to the mutated site]{%
-% \includegraphics[width=8cm,height=5cm]{APOBEC_two5prime.eps}
- % \label{APOBEC_twofive}}
-
-\caption{(a) APOBEC signatures obtained in 11 cancer types, 
-(b, c, d, e) Several signatures having prominent characteristics at 5'  or 3' to the mutated sites,
-(f) the frequencies of bases at two 5' to the mutated site. 
-}
-\label{nature2013_apobec}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The low grade glioma specific signature in the previous study (the ``Signature 14'' in Alexandrov et al. (2013)]{%
-  \includegraphics[height=3cm,width=12cm,clip]{example96_Glioma-Low-Grade.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter by the proposed method]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter by the proposed method in log scale]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Glioma-Low-Grade_membership_log.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a) The barplot is divided by 6 substitution patterns. In each division, 
-16 bars show joint probabilities of 16 combinations of the immediate 5' and 3' bases.
-(b, c) We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows (the logarithm of) the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-The low grade glioma specific signature detected by the proposed method is the signature 2.
-We can see that the mutations corresponding to signature 2 is mostly from the sample with an extremely high mutation rate.
-}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The result for the original data ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_original_signature_K2.eps}
-  \label{LGG_roriginal_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 2$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K2.eps}
-  \label{LGG_removed_sigK2_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 3$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K3.eps}
-  \label{LGG_removed_sigK3_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 4$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K4.eps}
-  \label{LGG_removed_sigK4_sig}}
-
-\subfigure[The result for the data without the hyper-mutated sample ($K = 5$)]{%
-  \includegraphics[height=2.5cm,width=15cm,clip]{Glioma-Low-Grade_removed_signature_K5.eps}
-  \label{LGG_removed_sigK5_sig}}
-
-  \caption{(a) The signatures obtained for the original data. The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
-  The second signature is the low grade glioma signature. 
-  (b, c, d, e) The signature obtained for the data without the hyper-mutated case for several number of mutation signatures ($K$, including the background signature).
-  Although the signature related to deamination of 5-methyl-cytosine remained, low grade glioma specific signature could not be observed.}
-
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-\centering
-
-\subfigure[The signature 2 detected in kidney clear cell carcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Kidney-Clear-Cell_oxidation_signature.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[The signature 1 detected in lung adenocarcinomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Lung-Adeno_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-\quad
-\subfigure[The signature 1 detected in melanomas]{%
-  \includegraphics[height=3cm,width=4.8cm,clip]{Melanoma_oxidation_signature.eps}
-  \label{LGG_membership_log}}
-
-\subfigure[Estimated membership parameter for kidney clear cell carcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Kidney-Clear-Cell_oxidation_membership.eps}
-  \label{LGG_nature2013_sig}}
-
-\subfigure[Estimated membership parameter for lung adenocarcinomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Lung-Adeno_oxidation_membership.eps}
-  \label{LGG_membership}}
-\quad
-\subfigure[Estimated membership parameter for melanomas]{%
-  \includegraphics[height=4cm,width=12cm,clip]{Melanoma_oxidation_membership.eps}
-  \label{LGG_membership_log}}
-
-  \caption{(a, b, c) The putative oxidative artifact signatures estimated for each cancer.
-  We can observe consistent abundance of the base C at the $-2 $ position.
-  (d, e, f) For each cancer type, 
-We have selected top 100 cancer samples according to the number of mutation.
-The height of bar shows the number of mutations for each sample,
-and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
-We can see that the signature corresponding to putative oxidative artifacts concentrates on a small number of samples.}
-
-\end{figure}
-
 \end{document}

### 2015-05-25 paper/manuscript/plos_template.tex
organized directory

--- 2015-05-18 manuscript20150515/plos_template.tex
@@
 % Template for PLoS
-% Version 2.0 July 2014
+% Version 3.1 February 2015
 %
@@
 %
-% Be advised that this is merely a template 
-% designed to facilitate accurate translation of manuscript content 
-% into our production files. 
-%
-% This template contains extensive comments intended 
+% This template contains comments intended 
 % to minimize problems and delays during our production 
-% process. Please follow the template 
+% process. Please follow the template instructions
 % whenever possible.
@@
 %
-% Once your paper is accepted for publication and enters production, 
+% Once your paper is accepted for publication, 
 % PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
@@
 %
-% DO NOT ADD EXTRA PACKAGES TO THIS TEMPLATE unless absolutely necessary.
-% Packages included in this template are intentionally
-% limited and basic in order to reduce the possibility
-% of issues during our production process.
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
 %
@@
 % -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
 %
@@
 % - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig." instead of "Figure".
 % See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
@@
 % Tables should be cell-based and may not contain:
-% - tabs/spacing/line breaks within cells to alter layout
+% - tabs/spacing/line breaks within cells to alter layout or alignment
 % - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
@@
 %
-% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
 %
@@
 %
-% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text.
-For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
-%
-% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
-% For bolding characters in mathmode, please use $\mathbf{}$ 
-%
-% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
-%
-% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
-%
-% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
-%
+% Please be sure to include all portions of an equation in the math environment.
+%
+% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
 %
 % % % % % % % % % % % % % % % % % % % % % % % % 
-
-\documentclass[10pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% fixltx2e package for \textsubscript
+\usepackage{fixltx2e}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
 
@@
 \usepackage{cite}
-\usepackage{bm}
-\usepackage{hyperref}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
 
 % line numbers
-\usepackage{lineno}
+\usepackage[right]{lineno}
 
@@
 \usepackage{microtype}
-% \DisableLigatures[f]{encoding = *, family = * }
+\DisableLigatures[f]{encoding = *, family = * }
 
 % rotating package for sideways tables
-%\usepackage{rotating}
-
-% If you wish to include algorithms, please use one of the packages below.
-Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
-%\usepackage{algorithmic}
-%\usepackage{algorithmicx}
-
-% Use doublespacing - comment out for single spacing
-% \usepackage{setspace} 
-% \doublespacing
-
+\usepackage{rotating}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% added by yshira( for using subfigure... this will be not necessary at the time of submission)
 \usepackage{graphicx}
 \usepackage{subfigure}
+\usepackage{bm}
+\usepackage{comment}
 
 % Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
 % Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
 
 % Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2009}
+\bibliographystyle{plos2015}
 
@@
 
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
 \pagestyle{myheadings}
-
-%% Include all macros below. Please limit the use of macros.
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
 
@@
 \begin{document}
-
-% Title must be 150 characters or less
+\vspace*{0.35in}
+
+% Title must be 250 characters or less.
+% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
 \begin{flushleft}
 {\Large
-\textbf{A simple model-based approach to inferring and visualizing cancer mutation signatures}
+\textbf\newline{A simple model-based approach to inferring and visualizing cancer mutation signatures}
 }
-% Insert Author names, affiliations and corresponding author email.
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
 \\
-Yuichi Shiraishi$^{1,\ast}$, 
-Georg Tremmel$^{1}$, 
-Satoru Miyano$^{1}$, 
-Matthew Stephens$^{2,3, \ast}$
+Yuichi Shiraishi\textsuperscript{1, *},
+Georg Tremmel\textsuperscript{1},
+Satoru Miyano\textsuperscript{1},
+Matthew Stephens\textsuperscript{2,3,*},
 \\
+\bigskip
 \bf{1} Laboratory of DNA Information Analysis, Human Genome Center, Institute of Medical Science, The University of Tokyo, Tokyo, Japan
@@
 \bf{3} Dept. of Statistics, University of Chicago, Chicago, Illinois, United States of America
-% \bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
 \\
-$\ast$ E-mail: yshira@hgc.jp; mstephens@uchicago.edu
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+% \Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
+
+% Current address notes
+% \textcurrency a Insert current address of first author with an address update
+% \textcurrency b Insert current address of second author with an address update
+% \textcurrency c Insert current address of third author with an address update
+
+% Deceased author note
+% \dag Deceased
+
+% Group/Consortium Author Note
+% \textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* yshira@hgc.jp; mstephens@uchicago.edu
+
 \end{flushleft}
 
-% Please keep the abstract between 250 and 300 words
+% Please keep the abstract below 300 words
 \section*{Abstract}
@@
 Here we present new methods for modelling, identifying and visualizing such mutation signatures. Our methods
-greatly simplify mutation signature models compared with existing approaches, reducing the number of parameters by orders of magnitude even while increasing the contextual factors (e.g.~number of flanking bases) that are accounted for.
+greatly simplify mutation signature models compared with existing approaches, reducing the number of parameters by orders of magnitude even while increasing the contextual factors (e.g.
+the number of flanking bases) that are accounted for.
 This improves both sensitivity and robustness of inferred signatures.
@@
 %This paragraph is nice, but could be removed if short on words?
-Our work draws on ideas from several tools that are widely used in other contexts.
+% Our work draws on ideas from several tools that are widely used in other contexts.
 The model for each signature is connected with position weight matrix models for transcription factor binding, and the overall framework is based on probabilistic models that are closely
+The overall framework of our work is based on probabilistic models that are closely
 connected with ``mixed-membership models'' which are widely used in population genetic admixture analysis, and in machine learning for document clustering.
@@
 
+% Please keep the Author Summary between 150 and 200 words
 % Use first person. PLOS ONE authors please skip this step. 
@@
 These tools provide more robust and interpretable mutation signatures compared to previous approaches, as we demonstrate by applying them to large-scale cancer genomic data.
+
+\linenumbers
 
@@
 
-Figure \ref{mutSig_example} illustrates the way that our new representation of signatures can simply capture
+Fig. 1 illustrates the way that our new representation of signatures can simply capture
 a previously-identified signature \cite{pmid25362482, pmid23318258} and provides
@@
 We particularly note how the key elements of this mutation signature are more immediately visually apparent than with visualizations of the full vector of probabilities used by existing approaches.
+
+\begin{figure}[h]
+
+\caption{{\bf Examples of visualizations and parameter values for the mutation signatures of the unconstrained (full) model and our independent model, 
+where substitution patterns, two 5' and 3' bases and transcription strand direction are considered as mutation features.}
+(A) The barplots are divided by 6 substitution patterns and transcription strand direction.
+In each division, 256 bars show joint probabilities of up to two base 5' and 3' bases with substitution patterns and transcription strands
+(ApApNpApA, ApApNpApC, ApApNpApG, ApApNpApT, $\cdots$, TpTpNpTpT).
+(B, C) An example mutation signature representation and parameter values from our independent model, where mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
+are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).
+In the bottom five rectangles, the width of each box represents the frequencies of bases (A, C, G and T) at the substitution and flanking site. 
+To highlight the most informative flanking sites, the heights of flanking site boxes are scaled by ($1 + 0.5 \times \log \sum_{n = A,C,G,T} f_n^2$, where $f_n$ is the parameter for each base), which can be interpreted as ($1 - 0.5 \times \text{R\'enyi entropy}$) \cite{rrnyi1961measures}.
+This is analogous to the information content scaling used in sequencing logos.
+In the top rectangle, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
+In the upper right, the height of the $+$ box represents the frequencies of mutations in the coding strand (or the plus strand, the sense strand and the untranscribed strand)
+whose nucleotide sequences directly corresponds to mRNA, 
+whereas the height of $-$ box represents those in the template strand (or the minus strand, the antisense strand, the transcribed strand and the noncoding strand)
+whose sequences are copied during the synthesis of mRNA.
+}
+\label{mutSig_example}
+
+\end{figure}
 
@@
 $\bm{M} = (6, 4, 4, 4, 4)$.
-See Supplementary Table 1 for other examples.
+See S1 Table for other examples.
 
@@
 
-This generative model is summarized in Figure \ref{methodOverview}.
+This generative model is summarized in Fig. \ref{methodOverview}.
 The key parameters in this model
@@
 A simulation study demonstrates that the estimation method can reproduce the mutation signature very accurately 
-provided enough mutations and samples are available (see Supplementary Material and Supplementary Figure 1). 
+provided enough mutations and samples are available (see S1 Text). 
 See Methods for more details, how to select $K$, etc.
+
+\begin{figure}[h]
+% \includegraphics[width=450pt]{methodOverview.pdf}
+\caption{{\bf An overview of the generative model of somatic mutations proposed in this paper.} 
+Suppose there are three types of mutation sources (mutation signatures) such as ultraviolet, tobacco smoking chemicals and transcription coupled repairs.
+Each cancer genome has ratios showing which types of mutation sources are contributing to its mutations (membership parameters).
+The generative model of the pattern of each mutation is:
+first, one of the mutation signatures is chosen according to the membership parameter.
+Second, according to the selected mutation signature, mutation features such as substitution pattens and flanking bases are generated 
+by the corresponding multinomial distributions for that mutation signature.
+}
+\label{methodOverview}
+\end{figure}
 
@@
 After analyzing the data with various number of mutation signatures $K$,
-we selected $K=3$ signatures for these analyses (Supplementary Material, Supplementary Figure 2, 3 and 4).
+we selected $K=3$ signatures for these analyses (see S2 Text).
 
@@
 shows a clear depletion of G base at the $-2$ position,
-which is consistent to the previous study \cite{pmid23318258} and results in the next subsection.
-In contrast, for the full model, this tendency is rather mild (Figure \ref{UCUT:APOBEC_full5_sig}, \ref{UCUT:AA_full5_sig} and Supplementary Figure 5).
+which is consistent to the previous study \cite{pmid23318258} and results in the next subsection ((Figs. 3A and 3B)).
+In contrast, for the full model, this tendency is rather mild (Figs. 3C, 3D and S1).
 The inferred AA mutation signature has no clear characteristics at the $-2$ position.
@@
 to identify signatures in more detail and with less data than existing approaches based on the full model.
+
+\begin{figure}[h]
+\caption{{\bf The mutation signatures for the UCUT data, and the results of downsampling experiments.}
+3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.
+(A) APOBEC signature for the independent model.
+(B) AA signature for the independent model.
+(C) APOBEC signature for the full model.
+(D) AA signature for the full model.
+(E) APOBEC signature stability.
+(F) AA signature stability.}
+\label{UCUT}
+\end{figure}
 
@@
 
-The results (Figures \ref{UCUT:APOBEC_downsampling} and \ref{UCUT:AA_downsampling})
+The results (Figs. 3E and 3F)
 confirm that the results of the independent model 
@@
 We applied the method to each cancer type separately to assess similarity of estimated signatures across cancer types.
-For each cancer type we selected the number of signatures $K$ by fitting the model with increasing $K$ and examining the log-likelihood, bootstrap errors, and correlation of membership parameters.The selected values of $K$ are given in
-Supplementary Table 2.
+For each cancer type we selected the number of signatures $K$ by fitting the model with increasing $K$ and examining the log-likelihood, bootstrap errors, and correlation of membership parameters.The selected values of $K$ are given in S2 Table.
 Also, we simply removed somatic mutations located in an intergenic region to include transcription strand biases as mutation features.
@@
 
-Figures \ref{nature2013_sig_summary} and \ref{nature2013_sig_member} summarise the results.
+Figs. \ref{nature2013_sig_summary} and \ref{nature2013_sig_member} summarise the results.
 In total, we identified 27 mutation signatures.
@@
 
+\begin{figure}[h]
+\centering
+% \includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
+\caption{{\bf The summary of mutation signatures obtained in a reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method.}
+Here, the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
+First, mutation signatures were estimated separately in each cancer type, and then similar signatures were merged (see text).
+}
+\label{nature2013_sig_summary}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+% \includegraphics[width=15cm,height=15cm]{corrplot.eps}
+\caption{{\bf The summary of membership of each mutation signature across 30 cancer types obtained using the proposed method.}}
+\label{nature2013_sig_member}
+\end{figure}
+
 Signatures 1 and 8 (C $>$ A at TpCpT and C $>$ T at TpCpG, respectively) observed in colorectal and uterine cancers
@@
 that inferring them as a single signature is ``wrong". However, splitting them into two signatures does help highlight certain features. Specifically,
-signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Supplementary Figures 6(c, d)).
+signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Figs. S2C S2D).
 This strand bias may be connected with the enrichment of 
@@
 and signature 1 also shows slightly elevated frequency of the T base at position $+2$
- (Figures \ref{POLE1_two5prime} and \ref{POLE2_two5prime}, Supplementary Figure 6(c, d)).
+ (Figs. 6B, 6C ,S6C and S6D).
 A previous study of Pol $\epsilon$  \cite{pmid25228659} 
@@
 
+\begin{figure}[h]
+% \subfigure[APOBEC signature intensities at two 5' to the mutated site]{%
+%   \includegraphics[width=14.5cm,height=6cm]{APOBEC_two5prime.eps}
+%    \label{APOBEC_two5prime}}
+% \subfigure[POLE1 signature intensities at two 5' to the mutated site]{%
+%   \includegraphics[width=4.5cm,height=6cm]{POLE1_two5prime.eps}
+%   \label{POLE1_two5prime}}
+% \quad
+% \subfigure[POLE2 signature intensities at two 5' to the mutated site]{%
+%   \includegraphics[width=4.5cm,height=6cm]{POLE2_two5prime.eps}
+%   \label{POLE2_two5prime}}
+% \subfigure[UV signature intensities at two 5' to the mutated site]{%
+%   \includegraphics[width=4.5cm,height=6cm]{UV_two5prime.eps}
+%   \label{UV_two5prime}}
+% \quad
+% \subfigure[The signature 11 intensities at two 5' to the mutated site]{%
+%   \includegraphics[width=4.5cm,height=6cm]{LMST_two5prime.eps}
+%   \label{LMST_two5prime}}
+\caption{{\bf The estimated frequencies of bases at two 5' to the mutated site for each cancer type.}
+The bar heights show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
+The error bars show bootstrapped standard errors.
+(A) The signature 13 (APOBEC signature)  intensities at two 5' to the mutated site.
+(B) The signature 1 (the first POL $\epsilon$ signature) intensities at two 5' to the mutated site.
+(C) The signature 8 (the second POL $\epsilon$ signature) intensities at two 5' to the mutated site.
+(D) The signature 10 (ultraviolet signature) intensities at two 5' to the mutated site.
+(E) The signature 11 intensities at two 5' to the mutated site.
+}
+\label{two5prime}
+\end{figure}
+
 %XX Is the point here that looking $\pm 2$ bases can be useful and our method is particularly helpful here? XX
@@
 Signature 2 (C $>$ A at [CT]pCpT) is observed solely in low grade gliomas, and
-appears related to, but slightly different from, the signature previously detected in the same cancer types (Supplementary Figure 7 (a)).
+appears related to, but slightly different from, the signature previously detected in the same cancer types (``signature 14'' \cite{pmid23318258}).
 Indeed, the corresponding signature in the previous study shows very complex patterns (C $>$ A at NpCpT or C $>$ T at GpTpN).
@@
 revealed that this signature is driven by 
-a single sample with an extremely high mutation rate (Supplementary Figure 7 (b,c)), and signature 2 disappeared when we removed this sample (Supplementary Figure 8).
+a single sample with an extremely high mutation rate (see Figs. S3A and S3B), and signature 2 disappeared when we removed this sample (see Fig. S4).
 It may be that the complex low-grade-glioma specific signature detected in the previous study is driven by the same single sample.
@@
 \cite{pmid23318258} (see their Supplementary Figures).
-Again our analysis highlights additional contextual information, with a strikingly elevated frequency of base C at the $-2$ position (Supplementary Figure 9 (a,b,c)).
-However, for each cancer type, only a few samples support this signature (see Supplementary Figure 9 (d, e, f)),
+Again our analysis highlights additional contextual information, with a strikingly elevated frequency of base C at the $-2$ position (Figs. S5A, S5B and S5C).
+However, for each cancer type, only a few samples support this signature (see Figs. S5D, S5E and S5F),
 and the corresponding signature 
@@
 Signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and is surely related to the activity of the APOBEC family.
-The 12 inferred signatures are highly consistent among cancer types except for B-cell lymphoma (Supplementary Figure 6(a)), highlighting the robustness
-of our approach.  Almost all of them show enrichment of A and T and depletion of G base at the $-2$ position (Figure \ref{APOBEC_two5prime}, Supplementary Figure 6(a)), 
+The 12 inferred signatures are highly consistent among cancer types except for B-cell lymphoma (see Fig.
+S2A), highlighting the robustness
+of our approach.  Almost all of them show enrichment of A and T and depletion of G base at the $-2$ position (Fig. 6A and S2A), 
 consistent with the UCUT data above and previous analyses \cite{pmid23318258}.
@@
 Signature 10 (C $>$T at [CT]pCpC) was observed in head and neck cancers and melanomas, and probably relates to ultraviolet light.
-Consistent strand specificities among the two cancer types (Supplementary Figure 6(e)) matches previous results \cite{pmid23318258}, but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figure \ref{UV_two5prime}, Supplementary Figure 6(e)).
+Consistent strand specificities among the two cancer types (Fig. S2E) matches previous results \cite{pmid23318258}, 
+but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figs. 6D and S2E). 
 Signature 11 (C $>$ T at GpCp[CG]) appears in small-cell lung cancers and stomach cancers,
-and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear.
-Again our analysis highlights elevated abundance of G at the $-2$ position (Figure \ref{LMST_two5prime}, Supplementary Figure 6(f)).
+and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear. Again our analysis highlights elevated abundance of G at the $-2$ position (Figs. 6E and S2F).
 Signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively, agree well with those detected in the same cancer types in the previous study \cite{pmid23318258}.
@@
 Signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
-The estimated signature in each cancer type shows higher mutation prevalence on the template strand (Supplementary Figure 6(b)), 
+The estimated signature in each cancer type shows higher mutation prevalence on the template strand (Fig. S2B), 
 which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
@@
 and use the estimate with the largest log-likelihood.
-See Supplementary Material for the derivation of the above updating procedures.
+See S3 for the derivation of the above updating procedures.
 
@@
 Thus, instead of having $L$ features with $\bm{M} = (M_1,\dots,M_L)$,
-existing approaches have one feature with $\bm{M} = (M_1 \times \dots \times M_L)$ (see Supplementary Table 1). 
+existing approaches have one feature with $\bm{M} = (M_1 \times \dots \times M_L)$ (see S1 Table). 
 The resulting model allows for arbitrary distributions on the $M_1 \times \dots \times M_L$
@@
 
-As pointed out by \cite{ding2008equivalence}, there is a  close relationships between mixed-membership models and nonnegative matrix factorization,
-which has been successfully used in the previous studies for mutational signature problems \cite{pmid22608084, pmid23318258, pmid23945592}.
-In fact, the proposed method can be seen as non-negative matrix factorization with additional restrictions.
-See Supplementary Materials for details of the relationship between the proposed approach and nonnegative matrix factorization.
-
-% \subsection*{Future works for the methodology}
-
-% Since the mixed membership model has been widely studied in the field machine learning and population genetics,
-% there are many rooms for improvement in the current form of the proposed method by learning from past experiences and knowledge.
-
-% First, a number of large scale cancer genome sequencing projects have been launched all over the world,
-% and we need to improve the parameter estimation so that those large scale data can be handled.
-% A number of  past parameter estimation techniques proposed for mixed-membership models can be tailored for the proposed model.
-% EM-algorithm (or its variant, called tempered EM algorithm) have been adopted in classical topic models 
-% for document classification \cite{Hofmann:1999} and population structure estimation \cite{tang2005estimation}.
-% In the population structure estimation problem, \cite{pmid19648217} has proposed 
-% a fast block relaxation scheme using sequential quadratic programming for block updates
-% with a quasi-Newton acceleration of convergence \cite{zhou2011quasi}, demonstrating a great improvement over EM-algorithm.
-% Similar techniques are used in \cite{taddy2011estimation} for the document classification.
-% For the estimation of Bayesian mixed-membership models, 
-% (collapsed) Gibbs sampling \cite{pmid10835412,pmid14872004} and 
-% variational method \cite{Blei:2003,teh2006collapsed,raj2014variational} have been proposed. 
-% Evaluate those techniques in the problem of mutation signature extraction problems, and incorporate the efficient method is an important future work.
-
-% Second, introducing certain prior distributions or penalty terms can lead more efficiency in terms of both accuracy and interpretation,
-% considering the success in machine learning and statistical genetics communities.
-% \cite{hoyer2004non, engelhardt2010analysis}.
-% In the documentation classification problem, 
-% adopting determinantal point process priors \cite{kulesza2012determinantal, kwok2012priors}
-% is demonstrated to be helpful for obtaining more diverse and intuitively interpretable topics,
-% avoiding over fine-tuning for frequent groups of words with multiple similar topics.
-
-% Third, for avoiding the problem of determining the number of latent variables, 
-% introducing Hierarchical dirichlet processes \cite{teh2006hierarchical}, 
-% which is a natural extension of topic models to nonparametric Bayesian frameworks, might be helpful.
-
-% Finally, utilizing other metrics than Euclidean distance and Kullback-Leibler Divergence might be worth investigating.
-% For nonnegative matrix factorization, the metric called $\beta$ divergence \cite{basu1998robust, eguchi2001robustifying}, 
-% has been successfully utilized in many research fields (reviewed in \cite{fevotte2011algorithms}),
-% and evaluating these metrics on mutation signature problems will be an interesting future tasks.
-
-% Do NOT remove this, even if you are not including acknowledgments.
-
-\section*{Acknowledgments}
-
-We thank to Dr. Daichi Mochihashi for helpful discussion and comments on an earlier version of the method;
-Dr. Hiromichi Suzuki for helpful feedback on the R packages;
-Dr. Ziyue Gao for helpful comments on the manuscript.
-
-This project was deeply influenced by the broader intellectual and academic environment surrounding Matthew Stephens's Laboratory,
-when the first author stayed at the University of Chicago as a visiting scholar. The first author thanks Dr. John Novembre,
-Dr. Jacob Degner and all the members of Matthew Stephens and John Novembre Laboratories for helpful discussion and comments.
-
-% \section*{References}
-
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% OR
-%
-% Compile your BiBTeX database using our plos2009.bst
-% style file and paste the contents of your .bbl file
-% here.
-%
-
-% \bibliographystyle{plos2009}
-
-\bibliography{plos_template}
-
-% \section*{Figure Legends}
-% This section is for figure legends only, do not include
-% graphics in your manuscript file.
-%
-%\begin{figure}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure caption.  
-%}
-%\label{Figure_label}
-%\end{figure}
-
-\newpage
-
- \section*{Tables}
-% 
-% See introductory notes if you wish to include sideways tables.
-%
-% NOTE: Please look over our table guidelines at http://www.plosone.org/static/figureGuidelines#tables to make sure that your tables meet our requirements.
-Certain types of spacing, cell merging, and other formatting tricks may have unintended results and will be returned for revision.
-%
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
-
 \begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
 \caption{
@@
 }
-\begin{center}
 \begin{tabular}{|l|l|l|l|} \hline
@@
 \end{tabular}
-\end{center}
 \label{tab_pop}
+\end{adjustwidth}
 \end{table}
 
-\clearpage
-
-\begin{figure*}[b]
-
-\subfigure[Example mutation signature representation for the unconstrained model]{%
-   \includegraphics[width=15cm,height=4cm]{example3076_TCR.eps}
-  \label{TCR_ful}
-}
-
-\subfigure[Example mutation signature representation for the independent model]{%
-   \includegraphics[width=8cm,height=5.32cm]{example18_TCR.eps}
-  \label{TCR_ind}
-}
-\quad
-\subfigure[Mutation signature parameter values for the signature in (b)]{%
-    \includegraphics[width=175pt]{exampleTable.pdf}
-    \label{TCR_ind_table}
-}
-
-\caption{Examples of visualizations and parameter values for the mutation signatures of the unconstrained (full) model and our independent model, 
-where substitution patterns, two 5' and 3' bases and transcription strand direction are considered as mutation features. 
-(a) The barplots are divided by 6 substitution patterns and transcription strand direction.
-In each division, 256 bars show joint probabilities of up to two base 5' and 3' bases with substitution patterns and transcription strands
-(ApApNpApA, ApApNpApC, ApApNpApG, ApApNpApT, $\cdots$, TpTpNpTpT).
-(b, c) An example mutation signature representation and parameter values from our independent model, where mutation features (substitution patterns, two 5' and 3' bases and strand direction) 
-are assumed to be independent ($L=6$, $\bm{M} = (6, 4, 4, 4, 4, 2)$).
-In the bottom five rectangles, the width of each box represents the frequencies of bases (A, C, G and T) at the substitution and flanking site. 
-To highlight the most informative flanking sites, the heights of flanking site boxes are scaled by ($1 + 0.5 \times \log \sum_{n = A,C,G,T} f_n^2$, where $f_n$ is the parameter for each base), which can be interpreted as ($1 - 0.5 \times \text{R\'enyi entropy}$) \cite{rrnyi1961measures}.
-This is analogous to the information content scaling used in sequencing logos.
-In the top rectangle, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
-In the upper right, the height of the $+$ box represents the frequencies of mutations in the coding strand (or the plus strand, the sense strand and the untranscribed strand)
-whose nucleotide sequences directly corresponds to mRNA, 
-whereas the height of $-$ box represents those in the template strand (or the minus strand, the antisense strand, the transcribed strand and the noncoding strand)
-whose sequences are copied during the synthesis of mRNA.
-}
-\label{mutSig_example}
-
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}
-\includegraphics[width=450pt]{methodOverview.pdf}
-\caption{An overview of the generative model of somatic mutations proposed in this paper.
-Suppose there are three types of mutation sources (mutation signatures) such as ultraviolet, tobacco smoking chemicals and transcription coupled repairs.
-Each cancer genome has ratios showing which types of mutation sources are contributing to its mutations (membership parameters).
-The generative model of the pattern of each mutation is:
-first, one of the mutation signatures is chosen according to the membership parameter.
-Second, according to the selected mutation signature, mutation features such as substitution pattens and flanking bases are generated 
-by the corresponding multinomial distributions for that mutation signature.
-}
-\label{methodOverview}
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}[ht]
-\centering
-
-\subfigure[APOBEC signature for the independent model]{%
-  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_APOBEC_ind.eps}
-  \label{UCUT:APOBEC_ind5_sig}}
-\quad
-\subfigure[AA signature for the independent model]{%
-  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_AA_ind.eps}
-  \label{UCUT:AA_ind5_sig}}
-
-  \subfigure[APOBEC signature for the full model]{%
-  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_APOBEC_full.eps}
-  \label{UCUT:APOBEC_full5_sig}}
-\quad
-\subfigure[AA signature for the full model]{%
-  \includegraphics[height=5cm,width=7.5cm,clip]{UCUT_AA_full.eps}
-  \label{UCUT:AA_full5_sig}}
-
-\subfigure[APOBEC signature stability]{%
-  \includegraphics[height=5.5cm,width=7.5cm,clip]{UCUT_downsampling_APOBEC.eps}
-  \label{UCUT:APOBEC_downsampling}}
-\quad
-\subfigure[AA signature stability]{%
-  \includegraphics[height=5.5cm,width=7.5cm,clip]{UCUT_downsampling_AA.eps}
-  \label{UCUT:AA_downsampling}}
-%
-\caption{The mutation signatures for the UCUT data, and the results of downsampling experiments. 
-3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.}
-\label{UCUT}
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}[b]
-\centering
-\includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
-\caption{The summary of mutation signatures obtained in a reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method,
-where the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
-First, mutation signatures were estimated separately in each cancer type, and then similar signatures were merged (see text).
-}
-\label{nature2013_sig_summary}
-\end{figure*}
-
-\clearpage
-
-\begin{figure*}[b]
-\centering
-\includegraphics[width=15cm,height=15cm]{corrplot.eps}
-\caption{The summary of membership of each mutation signature across 30 cancer types obtained using the proposed method.}
-\label{nature2013_sig_member}
-\end{figure*}
-
-\clearpage
-
-% \begin{figure*}[b]
-
-\begin{figure*}[b]
-
-\subfigure[APOBEC signature intensities at two 5' to the mutated site]{%
-  \includegraphics[width=14.5cm,height=6cm]{APOBEC_two5prime.eps}
-   \label{APOBEC_two5prime}}
-
-\subfigure[POLE1 signature intensities at two 5' to the mutated site]{%
-  \includegraphics[width=4.5cm,height=6cm]{POLE1_two5prime.eps}
-  \label{POLE1_two5prime}}
-\quad
-\subfigure[POLE2 signature intensities at two 5' to the mutated site]{%
-  \includegraphics[width=4.5cm,height=6cm]{POLE2_two5prime.eps}
-  \label{POLE2_two5prime}}
-
-\subfigure[UV signature intensities at two 5' to the mutated site]{%
-  \includegraphics[width=4.5cm,height=6cm]{UV_two5prime.eps}
-  \label{UV_two5prime}}
-\quad
-\subfigure[The signature 11 intensities at two 5' to the mutated site]{%
-  \includegraphics[width=4.5cm,height=6cm]{LMST_two5prime.eps}
-  \label{LMST_two5prime}}
-
-\caption{The estimated frequencies of bases at two 5' to the mutated site for each cancer type.
-The bar heights show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
-The error bars show bootstrapped standard errors.}
-\label{two5prime}
-
-\end{figure*}
-
-% \begin{figure*}[b]
-
-% \subfigure[POLE1 signatures in each cancer type]{%
-%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
-%   \label{POLE1_sig_list}}
-
-% \subfigure[POLE2 signature in each cancer type]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
-%   \label{POLE2_sig_list}}
-
-% \subfigure[UV signature in each cancer type]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
-%   \label{UV_sig_list}}
-
-% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
-%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
-%   \label{LMST_sig_list}}
-
-% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
-% \label{nature2013_sig_others}
-% \end{figure*}
-
-% \section*{Supporting Information Legends}
-%
-% Please enter your Supporting Information captions below in the following format:
-%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
-% 
-%\begin{description}
-%\item {\bf}
-%\item {\bf}
-%\end{description}
+As pointed out by \cite{ding2008equivalence}, there is a  close relationships between mixed-membership models and nonnegative matrix factorization,
+which has been successfully used in the previous studies for mutational signature problems \cite{pmid22608084, pmid23318258, pmid23945592}.
+In fact, the proposed method can be seen as non-negative matrix factorization with additional restrictions.
+See S4 Text  for details of the relationship between the proposed approach and nonnegative matrix factorization.
+
+\section*{Supporting Information}
+\subsection*{S1 Text}
+\label{simulation}
+{\bf Experiments on synthetic data.}
+
+\subsection*{S2 Text}
+\label{UCUT_supp}
+{\bf Experiment on UCUT data with the various number of mutation signatures.}
+
+\subsection*{S3 Text}
+\label{EM_math}
+{\bf Derivation of EM algorithm.}
+
+\subsection*{S4 Text}
+\label{NMF_math}
+{\bf Relationship with nonnegative matrix factorization.}
+
+\subsection*{S1 Table}
+\label{mfexample}
+{\bf Example of representation for mutation patterns (substitution patterns and one 5' and 3' bases)}
+In the independent representation, the elements of vector show substitution patterns, 5' adjacent bases and 3' adjacent bases, respectively.
+For substitution pattens, 1 to 6 values are assigned to C$>$A, C$>$G, C$>$T, T$>$A, T$>$C and T$>$G in this order.
+For 5' and 3' adjacent bases, 1 to 4 values are assigned to A, C, G and T.
+Note that the original base is fixed to C or T to remove the redundancy of complement sequences.
+
+\subsection*{S2 Table}
+\label{nature2013_sigNum}
+{\bf The number of mutation signatures selected for each cancer type.}
+
+\subsection*{S1 Figure}
+\label{UCUT_APOBEC_twoFivePrime}
+{\bf The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
+
+\subsection*{S2 Figure}
+\label{nature2013_sig_list}
+{\bf The list of several signatures extracted in each cancer type in the Alexandrov et al. (2013) data.}
+(A) APOBEC signatures obtained in each cancer type.
+(B) Smoking signature in each cancer type.
+(C) The first Pol $\epsilon$ signature (signature 1)  in each cancer type.
+(D) The second Pol $\epsilon$ signature (signature 8) in each cancer type.
+(E) The ultraviolet signature in each cancer type.
+(F) Unknown signature obtained in lung small cell carcinomas and stomach cancers.
+
+\subsection*{S3 Figure}
+\label{LGG_membership}
+{\bf The estimated membership parameters of low grade gliomas.}
+(A, B) Estimated membership parameter by the proposed method in normal and log scale.
+We have selected top 100 cancer samples according to the number of mutation.
+The height of bar shows (the logarithm of) the number of mutations for each sample,
+and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
+The low grade glioma specific signature detected by the proposed method is the signature 2.
+We can see that the mutations corresponding to signature 2 is mostly from the sample with an extremely high mutation rate.
+
+\subsection*{S4 Figure}
+\label{LGG_signatures}
+{\bf The signatures obtained for the original data and for the data without the hyper-mutated case.}
+(A) The result for the original data ($K = 3$). The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
+  The second signature is the low grade glioma signature.
+(B, C, D, E) The result for the data without the hyper-mutated sample for $K = 2, 3, 4$ and $5$, respectively.
+Although the signature related to deamination of 5-methyl-cytosine remained, low grade glioma specific signature could not be observed.
+
+\subsection*{S5 Figure}
+\label{oxidation}
+{\bf The putative oxidative artifact signatures and membership parameters estimated for each cancer.}
+(A) The second signature detected in kidney clear cell carcinomas.
+(B) The first signature detected in lung adenocarcinomas.
+(C) The first signature detected in melanomas
+(D, E, F) Estimated membership parameter for kidney clear cell carcinomas, lung adenocarcinomas and melanomas, respectively.
+For each cancer type, 
+We have selected top 100 cancer samples according to the number of mutation.
+The height of bar shows the number of mutations for each sample,
+and the ratio of colored division shows the ratio of estimated membership parameters for each signature and sample.
+We can see that the signature corresponding to putative oxidative artifacts concentrates on a small number of samples.
+
+\section*{Acknowledgments}
+
+We thank to Dr. Daichi Mochihashi for helpful discussion and comments on an earlier version of the method;
+Dr. Hiromichi Suzuki for helpful feedback on the R packages;
+Dr. Ziyue Gao for helpful comments on the manuscript.
+
+This project was deeply influenced by the broader intellectual and academic environment surrounding Matthew Stephens's Laboratory,
+when the first author stayed at the University of Chicago as a visiting scholar. The first author thanks Dr. John Novembre,
+Dr. Jacob Degner and all the members of Matthew Stephens and John Novembre Laboratories for helpful discussion and comments.
+
+\nolinenumbers
+
+% \section*{References}
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% OR
+%
+% Compile your BiBTeX database using our plos2009.bst
+% style file and paste the contents of your .bbl file
+% here.
+%
+
+% \bibliographystyle{plos2009}
+
+\bibliography{plos_template}
 

### 2015-05-25 submission20150525/plos_template.tex
organized directory

--- 2015-05-25 paper/manuscript/plos_template.tex
@@
 Our methods are implemented in an R package 
-pmsignature (\url{https://github.com/friend1ws/pmsignature})
+{\bf pmsignature} (\url{https://github.com/friend1ws/pmsignature})
 and a web application available at \url{https://friend1ws.shinyapps.io/pmsignature_shiny/}.
@@
 {\bf pmsignature} ({\bf p}robabilistic {\bf m}utation signature),
-at \url{https://github.com/friend1ws/pmsignature}).
+at \url{https://github.com/friend1ws/pmsignature}.
 The core part of the estimation process is implemented in C++ by way of the Rcpp package \cite{eddelbuettel2011rcpp},
@@
 In addition, a web-based application of our method is available at 
-\url{https://friend1ws.shinyapps.io/pmsignature_shiny/}
+\url{https://friend1ws.shinyapps.io/pmsignature_shiny/}.
 
 % Results and Discussion can be combined.
-\section*{Methods Overview}
+% \section*{Methods Overview}
+\section{Result}
 
@@
 (A) The barplots are divided by 6 substitution patterns and transcription strand direction.
-In each division, 256 bars show joint probabilities of up to two base 5' and 3' bases with substitution patterns and transcription strands
+In each division, 256 bars show joint probabilities of up to two base 5' and 3' bases
 (ApApNpApA, ApApNpApC, ApApNpApG, ApApNpApT, $\cdots$, TpTpNpTpT).
@@
 
-\subsection*{Mathematical specification, and connection with mixed-membership models}
+% \subsection*{Mathematical specification, and connection with mixed-membership models}
+\subsection*{An overview of mathematical specification of mutation signatures and the generative model of somatic mutations}
+
 % yshira20150421 I think the title of the section should change (because this is not "detail").
@@
 This generative model is summarized in Fig. \ref{methodOverview}.
+This model is essentially a ``mixed-membership model", also known as an ``admixture model" \cite{pmid10835412} or ``latent Dirichlet allocation" \cite{Blei:2003}. For example,
+the membership proportions for each sample are analogous to admixture proportions in an admixture model; the mutation signatures are analogous to 
+populations, and the mutation signature-specific parameters are analogous to population-specific allele frequencies.
+
 The key parameters in this model
@@
 provided enough mutations and samples are available (see S1 Text). 
-See Methods for more details, how to select $K$, etc.
+See Methods for more detailed models, parameter estimation, further discussion on relationships with mixed membership models, how to select $K$, etc.
 
@@
 first, one of the mutation signatures is chosen according to the membership parameter.
-Second, according to the selected mutation signature, mutation features such as substitution pattens and flanking bases are generated 
-by the corresponding multinomial distributions for that mutation signature.
+Second, each mutation feature such as substitution patterns and flanking bases is generated 
+by the corresponding multinomial distributions for the selected mutation signature.
 }
@@
 
-This model is essentially a ``mixed-membership model", also known as an ``admixture model" \cite{pmid10835412} or ``Latent Dirichlet Allocation" \cite{Blei:2003}. For example,
-the membership proportions for each sample are analogous to admixture proportions in an admixture model; the mutation signatures are analogous to 
-populations, and the mutation signature-specific parameters are analogous to population-specific allele frequencies
-(see Relationship with mixed-membership models in Methods for further discussion).
-
 % Results and Discussion can be combined.
-\section*{Results}
+% \section*{Results}
 
@@
 with the ``full model'', which corresponds to existing approaches.
-We compare mutation signatures obtained by the two approaches and investigate the robustness of each approach by downsampling experiments.
+We compare mutation signatures obtained by the two approaches and investigate the robustness of each approach by down-sampling experiments.
 
@@
 shows a clear depletion of G base at the $-2$ position,
-which is consistent to the previous study \cite{pmid23318258} and results in the next subsection ((Figs. 3A and 3B)).
+which is consistent to the previous study \cite{pmid23318258} and results in the next subsection (Figs. 3A and 3B).
 In contrast, for the full model, this tendency is rather mild (Figs. 3C, 3D and S1).
@@
 \begin{figure}[h]
-\caption{{\bf The mutation signatures for the UCUT data, and the results of downsampling experiments.}
+\caption{{\bf The mutation signatures for the UCUT data, and the results of down-sampling experiments.}
 3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.
-(A) APOBEC signature for the independent model.
-(B) AA signature for the independent model.
-(C) APOBEC signature for the full model.
-(D) AA signature for the full model.
-(E) APOBEC signature stability.
-(F) AA signature stability.}
+(A, B) APOBEC and AA signature for the independent model.
+(C, D) APOBEC and AA signature for the full model.
+(E, F) APOBEC and AA signature stability (the mean cosine similarity for each donw-sampling ratio).}
 \label{UCUT}
@@
 
-To investigate this further we performed downsampling experiments.
+To investigate this further we performed down-sampling experiments.
 Using the mutation signatures obtained using all 14717 substitutions as a gold standard, 
@@
 % \includegraphics[width=16cm,height=17.5cm]{AlexandrovEtAl_mergedSignature.eps}
-\caption{{\bf The summary of mutation signatures obtained in a reanalysis of the data of the previous study \cite{pmid23945592} using the proposed method.}
+\caption{{\bf The summary of mutation signatures across 30 cancer types \cite{pmid23945592} obtained using the proposed method.}
 Here, the substitution patterns and two 5' and 3' bases from the mutated sites are taken into account as mutation features.
@@
 The error bars show bootstrapped standard errors.
-(A) The signature 13 (APOBEC signature)  intensities at two 5' to the mutated site.
-(B) The signature 1 (the first POL $\epsilon$ signature) intensities at two 5' to the mutated site.
-(C) The signature 8 (the second POL $\epsilon$ signature) intensities at two 5' to the mutated site.
-(D) The signature 10 (ultraviolet signature) intensities at two 5' to the mutated site.
-(E) The signature 11 intensities at two 5' to the mutated site.
+(A, B, C, D, E) The intensities of signature 13 (APOBEC signature), signature 1 (the first POL $\epsilon$ signature),
+signature 8 (the second POL $\epsilon$ signature), signature 10 (ultraviolet signature) and signature 11, respectively, at two 5' to the mutated site.
+intensities at two 5' to the mutated site.
 }
@@
 Signature 2 (C $>$ A at [CT]pCpT) is observed solely in low grade gliomas, and
-appears related to, but slightly different from, the signature previously detected in the same cancer types (``signature 14'' \cite{pmid23318258}).
-Indeed, the corresponding signature in the previous study shows very complex patterns (C $>$ A at NpCpT or C $>$ T at GpTpN).
-Further investigation
+appears related to, but slightly different from, the signature previously detected in the same cancer types (``signature 14'', \cite{pmid23318258}). 
+Indeed, the corresponding signature in the previous study shows very complex patterns (C $>$ A at NpCpT or C $>$ T at GpTpN). Further investigation
 revealed that this signature is driven by 
@@
 Therefore, to further understand the effect of transcription activities on mutational mechanisms, 
-our framework will be helpful: by allowing for including gene expression or RNA polymerase II occupancies 
-to mutation features, the relationships of strand biases and transcription activities will be clarified.
+we can include gene expression or RNA polymerase II occupancies to mutation features, 
+so that the relationships of strand biases and transcription activities will be clarified.
 
@@
 and use the estimate with the largest log-likelihood.
-See S3 for the derivation of the above updating procedures.
+See S3 Text for the derivation of the above updating procedures.
 
@@
 \label{mfexample}
-{\bf Example of representation for mutation patterns (substitution patterns and one 5' and 3' bases)}
+{\bf Example of representation for mutation patterns (substitution patterns and one 5' and 3' bases).}
 In the independent representation, the elements of vector show substitution patterns, 5' adjacent bases and 3' adjacent bases, respectively.
@@
 \label{nature2013_sigNum}
-{\bf The number of mutation signatures selected for each cancer type.}
+{\bf The number of mutation signatures selected for each cancer type for each cancer type in the Alexandrov et al. (2013) data.}
 
@@
 \label{UCUT_APOBEC_twoFivePrime}
-{\bf The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UTUC data using the independent and full models.}
+{\bf The frequencies of bases at two 5' to the mutated site for the APOBEC mutation signatures obtained in UCUT data using the independent and full models.}
 
@@
 {\bf The list of several signatures extracted in each cancer type in the Alexandrov et al. (2013) data.}
-(A) APOBEC signatures obtained in each cancer type.
+(A) APOBEC signatures (signature 13) obtained in each cancer type.
 (B) Smoking signature in each cancer type.
@@
 (D) The second Pol $\epsilon$ signature (signature 8) in each cancer type.
-(E) The ultraviolet signature in each cancer type.
-(F) Unknown signature obtained in lung small cell carcinomas and stomach cancers.
+(E) The ultraviolet signature i(signature 10) n each cancer type.
+(F) Unknown signature (signature 11) obtained in lung small cell carcinomas and stomach cancers.
 
@@
 {\bf The signatures obtained for the original data and for the data without the hyper-mutated case.}
-(A) The result for the original data ($K = 3$). The first (from the left) signature seems to be deamination of 5-methyl-cytosine. 
+(A) The result for the original data ($K = 3$). The first (from the left) signature seems to be one from deamination of 5-methyl-cytosine. 
 The second signature is the low grade glioma signature.
@@
 We thank to Dr. Daichi Mochihashi for helpful discussion and comments on an earlier version of the method;
-Dr. Hiromichi Suzuki for helpful feedback on the R packages;
-Dr. Ziyue Gao for helpful comments on the manuscript.
+Dr. Hiromichi Suzuki for many helpful feedback on the R packages;
+Dr. Ziyue Gao for reading the manuscript and providing many helpful comments.
 

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_Glioma-Low-Grade_removed.R

@@ -1,0 +1,95 @@
+library(pmsignature)
+
+# dummy signature estimation for generating class
+inputFile <- system.file("extdata/Nik_Zainal_2012.mutationPositionFormat.txt", package="pmsignature");
+G <- readMPFile(inputFile, numBases = 3, type = "full");
+Param <- getPMSignature(G, K = 2);
+
+# reading the signature from Alexandrov et al. Nature 2013
+nature2013_sig_raw <- read.table("../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
+nature2013_sig <- t(nature2013_sig_raw[,4:30]);
+colnames(nature2013_sig) <- nature2013_sig_raw[,3];
+
+Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
+Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
+
+visPMSignature(Param, 1, isScale = TRUE)
+ggsave("../../supp/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
+
+
+##########
+
+G <- readMPFile("../../AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
+
+
+##########
+# membership plot
+load("../../AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
+visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
+ggsave("../../supp/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
+
+visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
+ggsave("../../supp/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
+##########
+
+
+##########
+# remove the sample with maximum number of mutation and check the signatures
+mutNum <- getMutNum(G);
+mutMaxInd <- which(mutNum[,2] == max(mutNum[,2]));
+G@countData[3, G@countData[2,] == mutMaxInd] <- 0;
+BG <- readBGFile(G);
+
[…]

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_checkThyroidSig.R

@@ -1,0 +1,29 @@
+
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+matchMutNum <- matrix(0, length(type2sigNum[,1]), 2);
+rownames(matchMutNum) <- type2sigNum[,1];
+
+for (i in 1:length(type2sigNum[,1])) {
+  
+  type <-  type2sigNum[i, 1];
+  print(type);
+  inputFile <- paste("../../AlexandrovEtAl/result/MPFormat/", type, ".mp.txt.gz", sep="");
+  G <- readMPFile(inputFile, numBases = 5, trDir = TRUE);
+
+  FVList <- G@featureVectorList;
+
+  mutInds_plus <- which(FVList[1,] == 5 & FVList[2, ] == 2 & FVList[3,] %in% c(1, 2) & FVList[6,] == 1);
+  mutInds_minus <- which(FVList[1,] == 5 & FVList[2, ] == 2 & FVList[3,] %in% c(1, 2) & FVList[6,] == 2);
+
+  countData <- G@countData;
+  countData[3, countData[1, ] %in% mutInds_plus];
+  countData[3, countData[1, ] %in% mutInds_minus];
+
+  matchMutNum[i, 1] <- sum( countData[3, countData[1, ] %in% mutInds_plus] );
+  matchMutNum[i, 2] <- sum( countData[3, countData[1, ] %in% mutInds_minus] );
+  
+}
+
+
+

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_figures.R

@@ -1,0 +1,505 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("AlexandrovEtAl_function.R")
+
+
+##########
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], fdim));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
+  # indList <- which(closeSig[i,] > 0);
[…]

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_figures.R.bk.R

@@ -1,0 +1,425 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+##########
+# functions
+
+
+visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2) {
+  
+  if (is.na(baseCol)) {
+    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
+    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
+  }
+  
+  centerBase <- (1 + numBases) / 2;
+  
+  v1 <- vF[1,1:6];
+  V2 <- vF[2:(numBases),1:4];
+  A <- matrix(0, numBases, 4);
+  B <- matrix(0, 4, 4);
+  
+  if (trDir == TRUE) {
+    v3 <- vF[(numBases + 1),1:2];
+  }
+  
+  for (l in 1:numBases) {
+    if (l < centerBase) {
+      A[l, ] <- V2[l, ];
+    } else if (l > centerBase) {
+      A[l, ] <- V2[l - 1, ];
+    }
+  }
+  A[centerBase,2] <- sum(v1[1:3]);
+  A[centerBase,4] <- sum(v1[4:6]);
+  
+  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
+  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
+  
+  num2base <- c("A", "C", "G", "T");
+  
+  frame();
+  plot.window(xlim=c(-0.25, 1.25 * numBases + 0.25), ylim=c(-0.25, 3.25));
+  
+  startx <- 0;
+  for(l in 1:numBases) {
+    
+    for(w in 1:4) {
+      endx <- startx + A[l,w]
+      polygon(c(startx, endx, endx, startx), c(0, 0, 1, 1), col = baseCol[w], border=F);
+      if (endx - startx > 1 / 4) {
+        text(0.5 * (endx + startx), 0.5, num2base[w], col="white", cex=charSize)
+      }
+      startx <- endx;
+    }
+    startx <- startx + 0.25;
+  }
+  
+  startx <- (centerBase - 1) * 1.25;
+  for (w in 1:4) {
[…]

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_figures_dir.R

@@ -1,0 +1,460 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("AlexandrovEtAl_function.R")
+
+
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4, 2);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
[…]

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_figures_dir_presentation.R

@@ -1,0 +1,416 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("AlexandrovEtAl_function.R")
+
+
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4, 2);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
+  # indList <- which(closeSig[i,] > 0);
[…]

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_function.R

@@ -1,0 +1,306 @@
+
+##########
+# functions for drawing figures for the Alexandrov et al analysis data
+
+
+visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale = TRUE, alpha = 2, charLimit = 0.25) {
+  
+  if (is.na(baseCol)) {
+    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
+    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
+  }
+  
+  centerBase <- (1 + numBases) / 2;
+  
+  v1 <- vF[1,1:6];
+  V2 <- vF[2:(numBases),1:4];
+  A <- matrix(0, numBases, 4);
+  B <- matrix(0, 4, 4);
+  
+  if (trDir == TRUE) {
+    v3 <- vF[(numBases + 1),1:2];
+  }
+  
+  for (l in 1:numBases) {
+    if (l < centerBase) {
+      A[l, ] <- V2[l, ];
+    } else if (l > centerBase) {
+      A[l, ] <- V2[l - 1, ];
+    }
+  }
+  A[centerBase,2] <- sum(v1[1:3]);
+  A[centerBase,4] <- sum(v1[4:6]);
+  
+  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
+  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
+  
+  num2base <- c("A", "C", "G", "T");
+  
+  # frame();
+  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
+  
+  renyi = function(p, tAlpha = alpha) {
+    if (tAlpha == 1) {
+      return(- sum(p * log2(p), na.rm = TRUE));
+    } else {
+      return( log(sum(p^tAlpha)) / (1 - tAlpha));
+    }
+  }
+  
+  # sizes <- 0.5 * (2 - apply(A, MARGIN = 1, FUN = function(p, alpha = scale) {-sum(x * log2(x), na.rm = TRUE)}));
+  # sizes <- sizes ** scale
+  
+  if (scale == FALSE) {
+    sizes <- rep(1, numBases)
+  } else {
[…]

### 2015-05-25 backup/figure20150516/script/AlexandrovEtAl_poleMembership.R

@@ -1,0 +1,75 @@
+library(pmsignature);
+
+##########
+# function for checking the index of the POLE signatures
+checkPoleInd <- function(Param) {
+  
+  maxK <- slot(Param, "signatureNum") - as.integer(slot(Param, "isBackGround"));
+  
+  sigInd1 <- 0;
+  sigInd2 <- 0;
+  for (k in 1:maxK) {
+    
+    F <- Param@signatureFeatureDistribution;
+
+    if (F[k,1,1] > 0.6 & F[k,3,4] > 0.6 & F[k,4,4] > 0.6) {
+      sigInd1 <- k;
+    }
+    
+    if (F[k,1,3] > 0.6 & F[k,3,4] > 0.6 & F[k,4,3] > 0.6) {
+      sigInd2 <- k;
+    }
+    
+  }
+  
+  return(c(sigInd1, sigInd2));
+  
+}
+###################
+
+
+
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+POLE_sample <- read.table("../data/POLE_sample.txt", header=FALSE, stringsAsFactors = FALSE)[[1]];
+
+POLE_types <- c("Colorectum", "Uterus");
+
+mem_pole <- data.frame(membership_sig1 = c(), membership_sig8 = c(), type = c())
+
+##########
+# cololrectum
+K <- type2sigNum[type2sigNum[,1] == "Colorectum", 2];
+inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", "Colorectum", ".", K, ".Rdata", sep="");
+load(inputRdata);
+poleInd_colo <- checkPoleInd(resultForSave[[1]]);
+samples_colo <- unlist(lapply(strsplit(resultForSave[[1]]@sampleList, split="-"), function(x) {paste(x[1], x[2], x[3], sep="-")}));
+mem_colo <- resultForSave[[1]]@sampleSignatureDistribution[samples_colo %in% POLE_sample ,poleInd_colo];
+mem_pole <- cbind(mem_pole, mem_colo[,1], mem_colo[,2], rep("colorectum", nrow(mem_colo)));
+
+
+##########
+# uterus
+K <- type2sigNum[type2sigNum[,1] == "Uterus", 2];
+
+inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", "Uterus", ".mp.txt.gz", sep = "");
[…]

### 2015-05-25 backup/figure20150516/script/UTUC_downsampling.R

@@ -1,0 +1,91 @@
+library(ggplot2);
+library(dplyr);
+
+ratio <- c(0.01, 0.025, 0.05, 0.1, 0.25, 0.5);
+
+Dist <- rep(0, 200 * 6 * 2);
+Type <- rep(0, 200 * 6 * 2);
+Ratio <- rep(0, 200 * 6 * 2);
+Sig <- rep(0, 200 * 6 * 2);
+
+
+# full5: 1: AA 2: APOBEC
+Type[1:1200] <- "full";
+for (i in 1:6) {
+  Inds <- 1:100 + 200 * (i - 1);
+  a <- read.table(paste("../../UTUC/result/cosineDist_full/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
+  Ratio[Inds] <- ratio[i];
+  Ratio[Inds + 100] <-ratio[i];                       
+  Dist[Inds] <- a[,"AA"];
+  Dist[Inds + 100] <- a[,"APOBEC"];
+  Sig[Inds] <- "AA";
+  Sig[Inds + 100] <- "APOBEC";
+}
+
+# ind5: 1: AA 2: APOBEC
+Type[1201:2400] <- "ind";
+for (i in 1:6) {
+  Inds <- 1:100 + 200 * (i - 1) + 1200;
+  a <- read.table(paste("../../UTUC/result/cosineDist_ind/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
+  Ratio[Inds] <- ratio[i];
+  Ratio[Inds + 100] <-ratio[i];                       
+  Dist[Inds] <- a[,"AA"];
+  Dist[Inds + 100] <- a[,"APOBEC"];
+  Sig[Inds] <- "AA";
+  Sig[Inds + 100] <- "APOBEC";
+}
+
+
+UTUC.downsampling <- data.frame(type=Type, ratio=Ratio, distance=Dist, signature=Sig);
+summary_UTUC.downsampling <- UTUC.downsampling %>% group_by(type, ratio, signature) %>% summarise(cosine_similality = mean(distance), seDist = sd(distance));
+
+summary.AA <- summary_UTUC.downsampling %.% filter(signature=="AA");
+summary.AA$seDist_max <- sapply(summary.AA$cosine_similality + summary.AA$seDist, min, 1) - summary.AA$cosine_similality;
[…]

### 2015-05-25 backup/figure20150516/script/UTUC_signature.R

@@ -1,0 +1,78 @@
+library(pmsignature);
+
+source("AlexandrovEtAl_function.R");
+
+load("../../UTUC/result/Param_full/3.Rdata");
+Param_full <- resultForSave[[1]];
+
+if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
+  visPMSignature(Param_full, 1);
+  ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
+  visPMSignature(Param_full, 2);
+  ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5); 
+} else {
+  visPMSignature(Param_full, 2);
+  ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
+  visPMSignature(Param_full, 1);
+  ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5);   
+}
+  
+load("../../UTUC/result/Param_ind/3.Rdata");
+Param_ind <- resultForSave[[1]];
+
+if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
+  visPMSignature(Param_ind, 1, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../../manuscript/UTUC_APOBEC_ind.eps", height = 5, width = 7.5);
+  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../../manuscript/UTUC_AA_ind.eps", height = 5, width = 7.5);
+} else {
+  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../../manuscript/UTUC_APOBEC_ind.eps", height = 5, width = 7.5);
[…]

### 2015-05-25 backup/figure20150516/script/example_figures.R

@@ -1,0 +1,46 @@
+
+####################
+# figure 1
+
+# dummy signature estimation for generating class
+inputFile <- system.file("extdata/Nik_Zainal_2012.mutationPositionFormat.txt", package="pmsignature");
+G <- readMPFile(inputFile, numBases = 3, type = "full");
+Param <- getPMSignature(G, K = 2);
+
+# reading the signature from Alexandrov et al. Nature 2013
+nature2013_sig_raw <- read.table("../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
+nature2013_sig <- t(nature2013_sig_raw[,4:30]);
+colnames(nature2013_sig) <- nature2013_sig_raw[,3];
+
+Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[3,];
+Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[11,];
+
+visPMSignature(Param, 1)
+ggsave("../../manuscript/example96_APOBEC.eps", width=12, height=3, units="in");
+
+visPMSignature(Param, 2)
+ggsave("../../manuscript/example96_POLE.eps", width=12, height=3, units="in");
+####################
+
+
+####################
+# figure 2
+
+inputFile <- system.file("extdata/Nik_Zainal_2012.mutationPositionFormat.txt", package="pmsignature");
+G <- readMPFile(inputFile, numBases = 5, trDir = TRUE);
+Param <- getPMSignature(G, K = 1);
+
+Param@signatureFeatureDistribution[1,1,1:6] <- c(0.201, 0.080, 0.314, 0.000, 0.405, 0.000);
+Param@signatureFeatureDistribution[1,2,1:4] <- c(0.146, 0.257, 0.128, 0.469);
+Param@signatureFeatureDistribution[1,3,1:4] <- c(0.183, 0.389, 0.331, 0.097);
+Param@signatureFeatureDistribution[1,4,1:4] <- c(0.161, 0.107, 0.013, 0.719);
+Param@signatureFeatureDistribution[1,5,1:4] <- c(0.019, 0.265, 0.366, 0.350);
+Param@signatureFeatureDistribution[1,6,1:2] <- c(0.375, 0.625);
+  
+visPMSignature(Param, 1, charSize = 1.2);
+
+outputName <- "../../manuscript/example_signature.eps";
+dev.copy2eps(file=outputName, height = 6.66, width = 10);
+par(.pardefault);
+
+

### 2015-05-25 backup/figure20150516/script/plot_facet_estimation.R

@@ -1,0 +1,160 @@
+library(ggplot2);
+library(dplyr);
+
+
+mutNum <- c(10, 25, 50, 100, 250, 500, 1000);
+sampleNum <- c(10, 25, 50, 100);
+alpha <- c(0.5, 1, 2);
+gamma <- c(0.5, 1, 2);
+
+MutNum <- rep(0, 25200);
+SampleNum <- rep(0, 25200);
+Alpha <- rep(0, 25200);
+Gamma <- rep(0, 25200);
+Dist <- rep(0, 25200);
+
+Ind <- 1:100;
+
+for (i1 in 1:7) {
+  for (i2 in 1:4) {
+    for (i3 in 1:3) {
+      for (i4 in 1:3) {
+        a <- read.table(paste("../result/cosineDist/", mutNum[i1], "_", sampleNum[i2], "_", alpha[i3], "_", gamma[i4], ".txt", sep=""), sep="\t", header=FALSE);
+
+        Dist[Ind] <- rowSums(a) / 4;
+        MutNum[Ind] <- mutNum[i1];
+        SampleNum[Ind] <- sampleNum[i2];
+        Alpha[Ind] <- alpha[i3];
+        Gamma[Ind] <- gamma[i4];
+        
+        Ind <- Ind + 100;
+      }
+    }
+  }
+}
+
+
+
+simulation.result <- data.frame(mutNum=MutNum, sampleNum=SampleNum, alpha=Alpha, gamma=Gamma, distance=Dist);
+summary_simulation.result <- simulation.result  %.% group_by(mutNum, sampleNum, alpha, gamma) %.% summarise(cosine_similality = mean(distance), seDist = sd(distance));
+summary_simulation.result$seDist2 <- sapply(summary_simulation.result$cosine_similality + summary_simulation.result$seDist, min, 1) - summary_simulation.result$cosine_similality;
+
+
+summary_simulation.result.1_1 <- summary_simulation.result %.% filter(alpha == 1 & gamma == 1);
+summary_simulation.result.1_1$seDist2 <- sapply(summary_simulation.result.1_1$cosine_similality + summary_simulation.result.1_1$seDist, min, 1) - summary_simulation.result.1_1$cosine_similality;
+
+
+pd <- position_dodge(0.2);
[…]

### 2015-05-25 figure/script/AlexandrovEtAl_Glioma-Low-Grade_removed.R

@@ -1,95 +1,0 @@
-library(pmsignature)
-
-# dummy signature estimation for generating class
-inputFile <- system.file("extdata/Nik_Zainal_2012.mutationPositionFormat.txt", package="pmsignature");
-G <- readMPFile(inputFile, numBases = 3, type = "full");
-Param <- getPMSignature(G, K = 2);
-
-# reading the signature from Alexandrov et al. Nature 2013
-nature2013_sig_raw <- read.table("../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
-nature2013_sig <- t(nature2013_sig_raw[,4:30]);
-colnames(nature2013_sig) <- nature2013_sig_raw[,3];
-
-Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
-Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
-
-visPMSignature(Param, 1)
-ggsave("../../supp/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
-
-
-##########
-
-G <- readMPFile("../../AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
-
-
-##########
-# membership plot
-load("../../AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
-visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
-ggsave("../../supp/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
-
-visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
-ggsave("../../supp/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
-##########
-
-
-##########
-# remove the sample with maximum number of mutation and check the signatures
-mutNum <- getMutNum(G);
-mutMaxInd <- which(mutNum[,2] == max(mutNum[,2]));
-G@countData[3, G@countData[2,] == mutMaxInd] <- 0;
-BG <- readBGFile(G);
-
[…]

### 2015-05-25 figure/script/AlexandrovEtAl_figures_dir.R

@@ -1,460 +1,0 @@
-
-library(Matrix);
-library(corrplot);
-library(pmsignature);
-
-.pardefault <- par(no.readonly = TRUE);
-
-# functions
-source("AlexandrovEtAl_function.R")
-
-
-#' Here, we focus on the case where substitution patterns 
-#' and two 5' and 3' bases are considered
-fdim <- c(6, 4, 4, 4, 4, 2);
-SigMat <- matrix(0, 0, prod(fdim));
-# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
-eps <- 0.6
-
-#' create the matrix showing the relationships between cancer type 
-#' and the specified numer of mutation signatures
-type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
-
-#' reading and converting the mutation signatures for each cancer type
-for (i in 1:nrow(type2sigNum)) {
-  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
-}
-
-SigV <- rownames(SigMat);
-
-#' calculate cosine distances among mutation signatures
-#' and list up the pairs having smaller distances than the specified threshould
-# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
-closeSig <- diag(nrow(SigMat));
-dist <- c();
-for (i in 1:(nrow(SigMat) - 1)) {
-  for (j in (i+1):nrow(SigMat)) {
-    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
-    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
-    # dist <- c(dist, d); # for checking the histgram of cosine distances
-    if ( d < eps) {
-      closeSig[i,j] <- 1; 
-    }
-  }
-}
-
-
-
-#' clustering the similar mutation signatures
-EXPCLOSESIG <- expm(closeSig);
-removeList <- rep(1, nrow(SigMat));
-curInd <- 1;
-typeVec <- c();
-
-Fs <- list();
-for (i in 1:length(removeList)) {
-  
-  samples <- c();
[…]

### 2015-05-25 figure/script/AlexandrovEtAl_oxidation.R

@@ -1,56 +1,0 @@
-library(pmsignature);
-
-.pardefault <- par(no.readonly = TRUE);
-
-##########
-# function for checking the index of the oxidative signature
-checkOxidationInd <- function(Param) {
-  
-  maxK <- slot(Param, "signatureNum") - as.integer(slot(Param, "isBackGround"));
-  
-  sigInd <- 0;
-  for (k in 1:maxK) {
-    
-    F <- Param@signatureFeatureDistribution;
-    if (F[k,1,1] > 0.6 & F[k,2,2] > 0.5 & F[k,3,2] > 0.5) {
-      sigInd <- k;
-    }
-    
-  }
-  
-  return(sigInd);
-  
-}
-##########
-
-
-
-#' create the matrix showing the relationships between cancer type 
-#' and the specified numer of mutation signatures
-type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
-
-oxidation_types <- c("Kidney-Clear-Cell", "Lung-Adeno", "Melanoma");
-
-for (i in 1:length(oxidation_types)) {
-  
-  K <- type2sigNum[type2sigNum[,1] == oxidation_types[i], 2];
-  inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
-  inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
-        
-  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
-  load(inputRdata);
-  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
-  ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
-
-  oxInd <- checkOxidationInd(resultForSave[[1]])
-  
-  tempMar <- par("mar");
-  par(mar = 0.2 * tempMar);
-  
-  visPMSignature(resultForSave[[1]], oxInd, charSize = 0.8);
-  dev.copy2eps(file=paste("../../supp/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 3.0, width = 5);    
-
-  par(.pardefault);
-  
-}
-

### 2015-05-25 makeFigure/script/AlexandrovEtAl_oxidation.R

@@ -1,0 +1,56 @@
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+##########
+# function for checking the index of the oxidative signature
+checkOxidationInd <- function(Param) {
+  
+  maxK <- slot(Param, "signatureNum") - as.integer(slot(Param, "isBackGround"));
+  
+  sigInd <- 0;
+  for (k in 1:maxK) {
+    
+    F <- Param@signatureFeatureDistribution;
+    if (F[k,1,1] > 0.6 & F[k,2,2] > 0.5 & F[k,3,2] > 0.5) {
+      sigInd <- k;
+    }
+    
+  }
+  
+  return(sigInd);
+  
+}
+##########
+
+
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+oxidation_types <- c("Kidney-Clear-Cell", "Lung-Adeno", "Melanoma");
+
+for (i in 1:length(oxidation_types)) {
+  
+  K <- type2sigNum[type2sigNum[,1] == oxidation_types[i], 2];
+  inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
+  inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
+        
+  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
+  load(inputRdata);
+  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
+  ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
+
+  oxInd <- checkOxidationInd(resultForSave[[1]])
+  
+  tempMar <- par("mar");
+  par(mar = 0.2 * tempMar);
+  
+  visPMSignature(resultForSave[[1]], oxInd, charSize = 0.8, isScale = TRUE);
+  dev.copy2eps(file=paste("../../supp/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 3.0, width = 5);    
+
+  par(.pardefault);
+  
+}
+

### 2015-05-25 makeFigure/script/Fig1.R

@@ -1,0 +1,44 @@
+# script for generating subfigures for Figure 1.
+# data generated in "../../analysis/AlexandrovEtAl/result/" is necessary.
+
+
+if (!file.exists("../result/Fig1")) {
+  dir.create("../result/Fig1")
+}
+
+##########
+# new figure 1(a)
+
+inputFile <- "../../analysis/AlexandrovEtAl/result/MPFormat/Liver.mp.txt.gz";
+G <- readMPFile(inputFile, numBases = 5, trDir = TRUE, type = "full");
+BG_prob <- readBGFile(G);
+Param <- getPMSignature(G, K = 3, BG = BG_prob);
+
+if (sum(Param@signatureFeatureDistribution[1,1,(256 * 10 + 1):(256 * 11)]) > 0.2) {
+  visPMSignature(Param, 1);
+} else {
+  visPMSignature(Param, 2);
+}
+
+ggsave("../result/Fig1/example3076_TCR.eps", width=12, height=5, units="in");
+
+
+# new figure 1(b)
+G <- readMPFile(inputFile, numBases = 5, trDir = TRUE, type = "independent");
+BG_prob <- readBGFile(G);
+Param <- getPMSignature(G, K = 3, BG = BG_prob);
+
+
+if (Param@signatureFeatureDistribution[1, 1, 5] > 0.6) {
+  visPMSignature(Param, 1, isScale = TRUE, alpha = 2); 
+} else {
+  visPMSignature(Param, 2, isScale = TRUE, alpha = 2);
+}
+
+
+outputName <- "../result/Fig1/example18_TCR.eps";
+dev.copy2eps(file=outputName, height = 6.66, width = 10);
+
+
+
+

### 2015-05-25 makeFigure/script/SText1.R

@@ -1,0 +1,25 @@
+# Script for generating figures in SText 1 (simulation study)
+# First, we need the results ("../../analysis/simulation/result"),
+# generated by the script (""../../analysis/simulation/script/runall.sh").
+
+
+if (!file.exists("../result/SText1")) {
+  dir.create("../result/SText1")
+}
+
+# script for generating the cosine distances for varying parameters
+# input: the results of "subscript_SText1/simulation_result.R".
+# output: "../result/SText1/simulation_result.eps"
+source("subscript_SText1/simulation_result.R")
+
+
+# script for generating the statistics on the simulation study 
+# such as likelihood, bootstrap errors, and correlations on signatures,
+# and an example of true and estimated signatures.
+# input: None
+# output: 
+source("subscript_SText1/simulation_example_stat_signature.R")
+
+# * the above procedure is not ideal because the statistic information should have been 
+# included in the simulaton study results. But I forggot to included it...
+# I may fix them iin the future....

### 2015-05-25 makeFigure/script/SText2.R

@@ -1,0 +1,14 @@
+# Script for generating figures in SText2.
+# We need several results ("../../analysis/UCUT/result"),
+# that are generated the script (""../../analysis/UCUT/script/runall.sh").
+
+if (!file.exists("../result/SText2")) {
+  dir.create("../result/SText2")
+}
+
+# You may need to modify the size of device...
+source("subscript_SText2/UCUT_variousK.R")
+
+# generating statistics for the estimation results such as
+# likelihood, bootstrap errors and correlations.
+source("subscript_SText2/UCUT_stat.R")

### 2015-05-25 makeFigure/script/subscript_SText1/simulation_example_stat_signature.R

@@ -1,0 +1,121 @@
+library(pmsignature);
+library(grid);
+
+
+G_sim <- makeSimData(type = "independent", numBases = 5, trDir = FALSE, K = 5, sampleNum = 25, mutationNum = 100, isBG = TRUE);
+brprob <- readBGFile(G_sim[[1]]);
+
+
+value <- c();
+sigNum <- c();
+type <- c();
+
+for (K in 2:8) {
+  
+  Param_est <- getPMSignature(G_sim[[1]], K = K, BG = brprob);
+  Param_boot <- bootPMSignature(G_sim[[1]], Param0 = Param_est, bootNum = 100, BG = brprob);
+  
+  if (K == 5) {
+    Param_est_5 <- Param_est;
+  }
+  
+  sigNum <- c(sigNum, K, K, K);
+  
+  # likelihood
+  value <- c(value, slot(Param_est, "loglikelihood"));
+  type <- c(type, "log-likelihood");
+  
+  # bootstrap standard error
+  bF <- Param_boot[[1]];
+  bFs <- matrix(0, 100, K - 1);
+  for(b in 1:100) {
+    for(k in 1:(K - 1)) {
+      bFs[b, k] <- sum(bF[b,k,,]);
+    }
+  }
+  value <- c(value, sum(bFs) / ((K - 1) * 100));
+  type <- c(type, "bootstrap error");
+  
+  # correlation
+  Q <- slot(Param_est, "sampleSignatureDistribution");
+  value <- c(value, max(cor(Q) - 2 * diag(rep(1, K))));
+  type <- c(type, "max. corr.");
+  
+}
+
+
+type <- factor(type, levels = c("log-likelihood", "bootstrap error", "max. corr."));
+simulation_stat <- data.frame(value = value, type = type, sigNum = sigNum);
+
+
+ggplot(simulation_stat, aes(x = sigNum, y = value)) +
+  geom_line(stat = "identity") + 
+  facet_wrap(~type, scales = "free") +
+  xlab("#signature") +
+  theme_bw() +
+  theme(panel.margin = unit(1, "lines"),
+        plot.title = element_text(size=rel(2)),
+        axis.title = element_text(size = rel(1.2)), 
+        axis.text.x = element_text(size = rel(1.2)),
+        axis.text.y = element_text(size = rel(1.2), angle = 90, hjust = 0.5),
+        legend.title = element_text(size = rel(1.2)),
+        legend.text = element_text(size = rel(1.2)),
+        panel.grid.major.y = element_blank(),
[…]

### 2015-05-25 makeFigure/script/subscript_SText1/simulation_result.R

@@ -1,0 +1,68 @@
+library(ggplot2);
+library(dplyr);
+
+
+mutNum <- c(10, 25, 50, 100, 250, 500, 1000);
+sampleNum <- c(10, 25, 50, 100);
+alpha <- c(0.5, 1, 2);
+gamma <- c(0.5, 1, 2);
+
+MutNum <- rep(0, 25200);
+SampleNum <- rep(0, 25200);
+Alpha <- rep(0, 25200);
+Gamma <- rep(0, 25200);
+Dist <- rep(0, 25200);
+
+Ind <- 1:100;
+
+for (i1 in 1:7) {
+  for (i2 in 1:4) {
+    for (i3 in 1:3) {
+      for (i4 in 1:3) {
+        a <- read.table(paste("../../analysis/simulation/result/cosineDist/", mutNum[i1], "_", sampleNum[i2], "_", alpha[i3], "_", gamma[i4], ".txt", sep=""), sep="\t", header=FALSE);
+
+        Dist[Ind] <- rowSums(a) / 4;
+        MutNum[Ind] <- mutNum[i1];
+        SampleNum[Ind] <- sampleNum[i2];
+        Alpha[Ind] <- alpha[i3];
+        Gamma[Ind] <- gamma[i4];
+        
+        Ind <- Ind + 100;
+      }
+    }
+  }
+}
+
+
+
+simulation.result <- data.frame(mutNum=MutNum, sampleNum=SampleNum, alpha=Alpha, gamma=Gamma, distance=Dist);
+summary_simulation.result <- simulation.result  %>% group_by(mutNum, sampleNum, alpha, gamma) %>% summarise(cosine_similality = mean(distance), seDist = sd(distance));
+summary_simulation.result$seDist2 <- sapply(summary_simulation.result$cosine_similality + summary_simulation.result$seDist, min, 1) - summary_simulation.result$cosine_similality;
+
+
+summary_simulation.result.1_1 <- summary_simulation.result %>% filter(alpha == 1 & gamma == 1);
+summary_simulation.result.1_1$seDist2 <- sapply(summary_simulation.result.1_1$cosine_similality + summary_simulation.result.1_1$seDist, min, 1) - summary_simulation.result.1_1$cosine_similality;
+
+
+pd <- position_dodge(0.2);
[…]

### 2015-05-25 makeFigure/script/subscript_SText2/UCUT_stat.R

@@ -1,0 +1,58 @@
+library(grid);
+library(pmsignature);
+
+value <- c();
+sigNum <- c();
+type <- c();
+
+
+for(K in 2:6) {
+  fileName <- paste("../../analysis/UCUT/result/Param_ind/", K, ".Rdata", sep="");
+  load(fileName);
+  
+  sigNum <- c(sigNum, K, K, K);
+  
+  # likelihood
+  value <- c(value, resultForSave[[1]]@loglikelihood);
+  type <- c(type, "log-likelihood");
+  
+  # bootstrap standard error
+  bF <- resultForSave[[2]][[1]];
+  bFs <- matrix(0, 100, K - 1);
+  for(b in 1:100) {
+      for(k in 1:(K - 1)) {
+        bFs[b, k] <- sum(bF[b,k,,]);
+      }
+  }
+  value <- c(value, sum(bFs) / ((K - 1) * 100));
+  type <- c(type, "bootstrap error");
+  
+  # correlation
+  value <- c(value, max(cor(resultForSave[[1]]@sampleSignatureDistribution) - 2 * diag(rep(1, K))));
+  type <- c(type, "max. corr.");
+}
+
+type <- factor(type, levels = c("log-likelihood", "bootstrap error", "max. corr."));
+UTUC_stat <- data.frame(value = value, type = type, sigNum = sigNum);
+
+
+ggplot(UTUC_stat, aes(x = sigNum, y = value)) +
+  geom_line(stat = "identity") + 
+  facet_wrap(~type, scales = "free") +
+  xlab("#signature") +
+  theme_bw() +
+  theme(panel.margin = unit(1, "lines"),
+        plot.title = element_text(size=rel(2)),
+        axis.title = element_text(size = rel(1.2)), 
+        axis.text.x = element_text(size = rel(1.2)),
+        axis.text.y = element_text(size = rel(1.2), angle = 90, hjust = 0.5),
+        legend.title = element_text(size = rel(1.2)),
+        legend.text = element_text(size = rel(1.2)),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.grid.major.x = element_line(colour="grey60", linetype="dashed"),
+        strip.text = element_text(face="bold", size=rel(1.2)));
+
+ggsave("../result/SText2/UCUT_stat.eps", width=15, height=5, units = "in");
+
+

### 2015-05-25 makeFigure/script/subscript_SText2/UCUT_variousK.R

@@ -1,0 +1,171 @@
+# goal: generating several figures from the results of PMSignature on Hoang et al. (2013) data.
+
+# 2015 3/1 Y.S
+# The size of legend is influenced by the plot device size in RStudio.
+# I trid to fix this probelm, but I couldn't. 
+# When make the figures, please enlarge the device so that the legend sizes in the figures are natural...
+
+.pardefault <- par(no.readonly = TRUE);
+##########
+# functions
+
+library(pmsignature);
+library(RColorBrewer);
+
+PMSSigPlot <- function(K) {
+  
+  fileName <- paste("../..//analysis/UCUT/result/Param_ind/", as.character(K), ".Rdata", sep="");
+  load(fileName);
+
+  # prepare matrix for drawing multiple figures
+  mat <- matrix(c(1:3, rep(K, 3)), 2, 3, byrow = TRUE);
+  if (K <= 3) {
+    for (k in K:3) {
+      mat[1, k] <- 0;
+    }
+  }
+  layout(mat);
+  
+  # memorize the current parameter and change the margin slightly
+  tempMar <- par("mar");
+  par(mar = 0.4 * tempMar);
+ 
+  
+  F <- resultForSave[[1]]@signatureFeatureDistribution;
+  
+  # check the index corresponding each signature types
+  AA_ind <- NULL;
+  AA_like_ind <- NULL;
+  APOBEC_ind <- NULL;
+  for (k in 1:(K - 1)) {
+    # AA
+    if (F[k,1,4] > 0.8 & F[k,3,2] > 0.5 & F[k,4,3] > 0.5 & F[k,6,2] > 0.6) {
+      AA_ind <- k;
+    } else if (sum(F[k,1,1:3]) > 0.8 & F[k,3,4] > 0.7) {
+      APOBEC_ind <- k;
+    } else if (F[k,1,4] > 0.6 & F[k,6,2] > 0.6) {
+      AA_like_ind <- k;
+    }
+  }
+  
+  
+  ##########
+  # here, draw the figure for each signature and set the index and color vector 
+  # for the later membership barplot figrues
+  
+  mainSigInd <- c();
+  remainSigInd <- 1:(K - 1);
+  Q_col <- c();
+  Q_legend <- c();
+
+  if (!is.null(AA_ind)) {
+    visPMSignature(resultForSave[[1]], AA_ind);
[…]

### 2015-05-26 makeFigure/script/Fig1.R
organized several scripts for generating figures for the paper.

@@ -6,9 +6,8 @@ if (!file.exists("../result/Fig1")) {
   dir.create("../result/Fig1")
 }
 
+#' generating the subfigure of Figure 1 (mutation signature example of the full model)
 ##########
-# new figure 1(a)
-
 inputFile <- "../../analysis/AlexandrovEtAl/result/MPFormat/Liver.mp.txt.gz";
 G <- readMPFile(inputFile, numBases = 5, trDir = TRUE, type = "full");
 BG_prob <- readBGFile(G);
@@ -21,9 +20,11 @@ if (sum(Param@signatureFeatureDistribution[1,1,(256 * 10 + 1):(256 * 11)]) > 0.2
 }
 
 ggsave("../result/Fig1/example3076_TCR.eps", width=12, height=5, units="in");
+##########
 
 
-# new figure 1(b)
+#' generating the subfiture of Figure 1 (mutation signature example of the independent model)
+##########
 G <- readMPFile(inputFile, numBases = 5, trDir = TRUE, type = "independent");
 BG_prob <- readBGFile(G);
 Param <- getPMSignature(G, K = 3, BG = BG_prob);
@@ -35,10 +36,9 @@ if (Param@signatureFeatureDistribution[1, 1, 5] > 0.6) {
   visPMSignature(Param, 2, isScale = TRUE, alpha = 2);
 }
 
-
 outputName <- "../result/Fig1/example18_TCR.eps";
 dev.copy2eps(file=outputName, height = 6.66, width = 10);
-
+##########
 
 
 

### 2015-05-26 makeFigure/script/SFig1.R

@@ -1,0 +1,59 @@
+# script for generating subfigures for Supplementary Figure 1.
+# data generated in "../../analysis/UCUT/result/" is necessary.
+
+
+if (!file.exists("../result/SFig1")) {
+  dir.create("../result/SFig1")
+}
+
+library(pmsignature);
+
+
+##########
+load("../../analysis/UCUT/result/Param_full/3.Rdata");
+Param_full <- resultForSave[[1]];
+
+load("../../analysis/UCUT/result/Param_ind/3.Rdata");
+Param_ind <- resultForSave[[1]];
+
+
+
+# 2 base 5' to the mutated site
+## ind model 
+
+if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
+  baseTwoFivePrime <- data.frame(model=rep("ind", 4), base=c("A", "C", "G", "T"), prob = Param_ind@signatureFeatureDistribution[1,2,1:4]);
+} else {
+  baseTwoFivePrime <- data.frame(model=rep("ind", 4), base=c("A", "C", "G", "T"), prob = Param_ind@signatureFeatureDistribution[2,2,1:4]);
+}
+
+## full model
+
+if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
+  tempA <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 0 & 0:3071 %% 16 < 4]);
+  tempC <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 4 & 0:3071 %% 16 < 8]);
+  tempG <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 8 & 0:3071 %% 16 < 12]);
+  tempT <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 12 & 0:3071 %% 16 < 16]);
[…]

### 2015-05-26 makeFigure/script/SFig4.R

@@ -1,3 +1,13 @@
+# script for generating subfigures in the Supplementary Figure 4.
+# data generated in "../../analysis/AlexandrovEtAl/result/" is necessary.
+
+
+
+if (!file.exists("../result/SFig4")) {
+dir.create("../result/SFig4")
+}
+
+
 library(pmsignature)
 
 # dummy signature estimation for generating class
@@ -14,22 +24,22 @@ Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[which(rownames(nature
 Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
 
 visPMSignature(Param, 1, isScale = TRUE)
-ggsave("../../supp/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
+ggsave("../result/SFig4/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
 
 
 ##########
 
-G <- readMPFile("../../AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
+G <- readMPFile("../../analysis/AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
 
 
 ##########
 # membership plot
-load("../../AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
+load("../../analysis/AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
 visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
-ggsave("../../supp/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
+ggsave("../result/SFig4/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
 
 visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
-ggsave("../../supp/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
+ggsave("../result/SFig4/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
 ##########
[…]

### 2015-05-26 makeFigure/script/SFig5.R

@@ -1,3 +1,12 @@
+# script for generating subfigures in the Supplementary Figure 5.
+# data generated in "../../analysis/AlexandrovEtAl/result/" is necessary.
+
+
+if (!file.exists("../result/SFig5")) {
+  dir.create("../result/SFig5")
+}
+
+
 library(pmsignature);
 
 .pardefault <- par(no.readonly = TRUE);
@@ -34,13 +43,13 @@ oxidation_types <- c("Kidney-Clear-Cell", "Lung-Adeno", "Melanoma");
 for (i in 1:length(oxidation_types)) {
   
   K <- type2sigNum[type2sigNum[,1] == oxidation_types[i], 2];
-  inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
-  inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
+  inputMPdata <- paste("../../analysis/AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
+  inputRdata <- paste("../../analysis/AlexandrovEtAl/result/Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
         
   G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
   load(inputRdata);
   visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
-  ggsave(paste("../../supp/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
+  ggsave(paste("../result/SFig5/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
 
   oxInd <- checkOxidationInd(resultForSave[[1]])
   
@@ -48,9 +57,12 @@ for (i in 1:length(oxidation_types)) {
   par(mar = 0.2 * tempMar);
   
   visPMSignature(resultForSave[[1]], oxInd, charSize = 0.8, isScale = TRUE);
-  dev.copy2eps(file=paste("../../supp/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 3.0, width = 5);    
[…]

### 2015-05-26 makeFigure/script/SText2.R

@@ -6,7 +6,8 @@ if (!file.exists("../result/SText2")) {
   dir.create("../result/SText2")
 }
 
-# You may need to modify the size of device...
+#' You may need to modify the size of device...
+#' I'm searching for the better way (e.g., set the size of device in advance)
 source("subscript_SText2/UCUT_variousK.R")
 
 # generating statistics for the estimation results such as

### 2015-05-26 makeFigure/script/oxidation.R

@@ -1,44 +1,0 @@
-library(pmsignature);
-
-
-#' create the matrix showing the relationships between cancer type 
-#' and the specified numer of mutation signatures
-type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
-
-oxidation_types <- c("Kidney-Clear-Cell", "Lung-Adeno", "Melanoma");
-
-for (i in 1:length(oxidation_types)) {
-  
-  K <- type2sigNum[type2sigNum[,1] == oxidation_types[i], 2];
-  inputMPdata <- paste("../../AlexandrovEtAl/result/MPFormat/", oxidation_types[i], ".mp.txt.gz", sep = "");
-  inputRdata <- paste("../../AlexandrovEtAl/result//Param_ind5_dir/", oxidation_types[i], ".", K, ".Rdata", sep="");
-        
-  G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
-  visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
-  ggsave("../../supp/", oxidation_types[i], "_oxidation_membership.eps", height = 4, width = 15, pointsize = 9);
-
-  oxInd <- checkOxidationInd(resultForSave[[1]])
-  visPMSignature(resultForSave[[1]], oxInd);
-  dev.copy2eps(file="../../supp/", oxidation_types[i], "_oxidation_signature.eps", height = 2.4, width = 3.75);    
-
-}
-
-
-checkOxidationInd <- function(Param) {
-  
-  maxK <- slot(Param, "signatureNum") - as.integer(slot(Param, "isBackGround"));
-  
-  sigInd <- 0;
-  for (k in 1:maxK) {
-    
-    F <- Param@signatureFeatureDistribution;
-    if (F[k,1,1] > 0.6 & F[k,2,2] > 0.5 & F[k,3,2] > 0.5) {
-      sigInd <- k;
-    }
-    
-  }
-  
-  return(sigInd);
-  
-}
-

### 2015-05-26 makeFigure/script/AlexandrovEtAl_figures.R.bk.R
organized figures generated by UCUT analysis and so on

@@ -1,425 +1,0 @@
-
-library(Matrix);
-library(corrplot);
-library(pmsignature);
-
-.pardefault <- par(no.readonly = TRUE);
-
-##########
-# functions
-
-
-visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2) {
-  
-  if (is.na(baseCol)) {
-    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
-    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
-  }
-  
-  centerBase <- (1 + numBases) / 2;
-  
-  v1 <- vF[1,1:6];
-  V2 <- vF[2:(numBases),1:4];
-  A <- matrix(0, numBases, 4);
-  B <- matrix(0, 4, 4);
-  
-  if (trDir == TRUE) {
-    v3 <- vF[(numBases + 1),1:2];
-  }
-  
-  for (l in 1:numBases) {
-    if (l < centerBase) {
-      A[l, ] <- V2[l, ];
-    } else if (l > centerBase) {
-      A[l, ] <- V2[l - 1, ];
-    }
-  }
-  A[centerBase,2] <- sum(v1[1:3]);
-  A[centerBase,4] <- sum(v1[4:6]);
-  
-  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
-  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
-  
-  num2base <- c("A", "C", "G", "T");
-  
-  frame();
-  plot.window(xlim=c(-0.25, 1.25 * numBases + 0.25), ylim=c(-0.25, 3.25));
-  
-  startx <- 0;
-  for(l in 1:numBases) {
-    
-    for(w in 1:4) {
-      endx <- startx + A[l,w]
-      polygon(c(startx, endx, endx, startx), c(0, 0, 1, 1), col = baseCol[w], border=F);
-      if (endx - startx > 1 / 4) {
-        text(0.5 * (endx + startx), 0.5, num2base[w], col="white", cex=charSize)
-      }
-      startx <- endx;
-    }
-    startx <- startx + 0.25;
-  }
-  
-  startx <- (centerBase - 1) * 1.25;
-  for (w in 1:4) {
[…]

### 2015-05-26 makeFigure/script/Fig3.R

@@ -1,0 +1,13 @@
+# Script for generating figures in SText 1 (simulation study)
+# First, we need the results ("../../analysis/UCUT/result"),
+
+if (!file.exists("../result/Fig3")) {
+  dir.create("../result/Fig3")
+}
+
+
+# script for generating (full and independent) mutation signatures obtained in UCUT analysis.
+source("subscript_Fig3/UCUT_signature.R")
+
+# script for generating (full and independent) the result of down-sampling analysis.
+source("subscript_Fig3/UCUT_downsampling.R")

### 2015-05-26 makeFigure/script/SFig3_SFig4.R

@@ -1,0 +1,108 @@
+# script for generating subfigures in the Supplementary Figure 4.
+# data generated in "../../analysis/AlexandrovEtAl/result/" is necessary.
+
+
+if (!file.exists("../result/SFig3")) {
+  dir.create("../result/SFig3")
+}
+
+if (!file.exists("../result/SFig4")) {
+dir.create("../result/SFig4")
+}
+
+
+library(pmsignature)
+
+# dummy signature estimation for generating class
+inputFile <- system.file("extdata/Nik_Zainal_2012.mutationPositionFormat.txt", package="pmsignature");
+G <- readMPFile(inputFile, numBases = 3, type = "full");
+Param <- getPMSignature(G, K = 2);
+
+# reading the signature from Alexandrov et al. Nature 2013
+nature2013_sig_raw <- read.table("../data/AlexandrovEtAl_signatures.txt", header=T, sep="\t");
+nature2013_sig <- t(nature2013_sig_raw[,4:30]);
+colnames(nature2013_sig) <- nature2013_sig_raw[,3];
+
+Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
+Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
+
+visPMSignature(Param, 1, isScale = TRUE)
+ggsave("../result/SFig3/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
+
+
+##########
+
+G <- readMPFile("../../analysis/AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 5, trDir = TRUE);
+
+
+##########
+# membership plot
+load("../../analysis/AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
+visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
+ggsave("../result/SFig3/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
+
+visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
[…]

### 2015-05-26 makeFigure/script/UTUC_signature.R

@@ -1,78 +1,0 @@
-library(pmsignature);
-
-source("AlexandrovEtAl_function.R");
-
-load("../../UTUC/result/Param_full/3.Rdata");
-Param_full <- resultForSave[[1]];
-
-if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
-  visPMSignature(Param_full, 1);
-  ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
-  visPMSignature(Param_full, 2);
-  ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5); 
-} else {
-  visPMSignature(Param_full, 2);
-  ggsave("../../manuscript/UTUC_APOBEC_full.eps", height = 5, width = 7.5);
-  visPMSignature(Param_full, 1);
-  ggsave("../../manuscript/UTUC_AA_full.eps", height = 5, width = 7.5);   
-}
-  
-load("../../UTUC/result/Param_ind/3.Rdata");
-Param_ind <- resultForSave[[1]];
-
-if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
-  visPMSignature(Param_ind, 1, charSize = 1, isScale = TRUE, alpha = 2);
-  dev.copy2eps(file="../../manuscript/UTUC_APOBEC_ind.eps", height = 5, width = 7.5);
-  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
-  dev.copy2eps(file="../../manuscript/UTUC_AA_ind.eps", height = 5, width = 7.5);
-} else {
-  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
-  dev.copy2eps(file="../../manuscript/UTUC_APOBEC_ind.eps", height = 5, width = 7.5);
[…]

### 2015-05-26 makeFigure/script/plot_facet_estimation.R

@@ -1,160 +1,0 @@
-library(ggplot2);
-library(dplyr);
-
-
-mutNum <- c(10, 25, 50, 100, 250, 500, 1000);
-sampleNum <- c(10, 25, 50, 100);
-alpha <- c(0.5, 1, 2);
-gamma <- c(0.5, 1, 2);
-
-MutNum <- rep(0, 25200);
-SampleNum <- rep(0, 25200);
-Alpha <- rep(0, 25200);
-Gamma <- rep(0, 25200);
-Dist <- rep(0, 25200);
-
-Ind <- 1:100;
-
-for (i1 in 1:7) {
-  for (i2 in 1:4) {
-    for (i3 in 1:3) {
-      for (i4 in 1:3) {
-        a <- read.table(paste("../result/cosineDist/", mutNum[i1], "_", sampleNum[i2], "_", alpha[i3], "_", gamma[i4], ".txt", sep=""), sep="\t", header=FALSE);
-
-        Dist[Ind] <- rowSums(a) / 4;
-        MutNum[Ind] <- mutNum[i1];
-        SampleNum[Ind] <- sampleNum[i2];
-        Alpha[Ind] <- alpha[i3];
-        Gamma[Ind] <- gamma[i4];
-        
-        Ind <- Ind + 100;
-      }
-    }
-  }
-}
-
-
-
-simulation.result <- data.frame(mutNum=MutNum, sampleNum=SampleNum, alpha=Alpha, gamma=Gamma, distance=Dist);
-summary_simulation.result <- simulation.result  %.% group_by(mutNum, sampleNum, alpha, gamma) %.% summarise(cosine_similality = mean(distance), seDist = sd(distance));
-summary_simulation.result$seDist2 <- sapply(summary_simulation.result$cosine_similality + summary_simulation.result$seDist, min, 1) - summary_simulation.result$cosine_similality;
-
-
-summary_simulation.result.1_1 <- summary_simulation.result %.% filter(alpha == 1 & gamma == 1);
-summary_simulation.result.1_1$seDist2 <- sapply(summary_simulation.result.1_1$cosine_similality + summary_simulation.result.1_1$seDist, min, 1) - summary_simulation.result.1_1$cosine_similality;
-
-
-pd <- position_dodge(0.2);
[…]

### 2015-05-26 makeFigure/script/subscript_Fig3/UCUT_downsampling.R

@@ -13,7 +13,7 @@ Sig <- rep(0, 200 * 6 * 2);
 Type[1:1200] <- "full";
 for (i in 1:6) {
   Inds <- 1:100 + 200 * (i - 1);
-  a <- read.table(paste("../../UTUC/result/cosineDist_full/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
+  a <- read.table(paste("../../analysis/UCUT/result/cosineDist_full/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
   Ratio[Inds] <- ratio[i];
   Ratio[Inds + 100] <-ratio[i];                       
   Dist[Inds] <- a[,"AA"];
@@ -26,7 +26,7 @@ for (i in 1:6) {
 Type[1201:2400] <- "ind";
 for (i in 1:6) {
   Inds <- 1:100 + 200 * (i - 1) + 1200;
-  a <- read.table(paste("../../UTUC/result/cosineDist_ind/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
+  a <- read.table(paste("../../analysis/UCUT/result/cosineDist_ind/3_", ratio[i], ".txt", sep=""), sep="\t", header=TRUE);
   Ratio[Inds] <- ratio[i];
   Ratio[Inds + 100] <-ratio[i];                       
   Dist[Inds] <- a[,"AA"];
@@ -39,7 +39,7 @@ for (i in 1:6) {
 UTUC.downsampling <- data.frame(type=Type, ratio=Ratio, distance=Dist, signature=Sig);
 summary_UTUC.downsampling <- UTUC.downsampling %>% group_by(type, ratio, signature) %>% summarise(cosine_similality = mean(distance), seDist = sd(distance));
 
-summary.AA <- summary_UTUC.downsampling %.% filter(signature=="AA");
+summary.AA <- summary_UTUC.downsampling %>% filter(signature=="AA");
 summary.AA$seDist_max <- sapply(summary.AA$cosine_similality + summary.AA$seDist, min, 1) - summary.AA$cosine_similality;
 summary.AA$seDist_min <- - sapply(summary.AA$cosine_similality - summary.AA$seDist, max, 0) + summary.AA$cosine_similality;
 
@@ -60,7 +60,7 @@ ggplot(summary.AA, aes(x=ratio, y=cosine_similality, colour=type, group=type, fi
         legend.text = element_text(size = rel(1.5)),
[…]

### 2015-05-26 makeFigure/script/subscript_Fig3/UCUT_signature.R

@@ -1,0 +1,36 @@
+library(pmsignature);
+
+source("AlexandrovEtAl_function.R");
+
+load("../../analysis/UCUT/result/Param_full/3.Rdata");
+Param_full <- resultForSave[[1]];
+
+if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
+  visPMSignature(Param_full, 1);
+  ggsave("../result/Fig3/UCUT_APOBEC_full.eps", height = 5, width = 7.5);
+  visPMSignature(Param_full, 2);
+  ggsave("../result/Fig3/UCUT_AA_full.eps", height = 5, width = 7.5); 
+} else {
+  visPMSignature(Param_full, 2);
+  ggsave("../result/Fig3/UCUT_APOBEC_full.eps", height = 5, width = 7.5);
+  visPMSignature(Param_full, 1);
+  ggsave("../result/Fig3/UCUT_AA_full.eps", height = 5, width = 7.5);   
+}
+  
+load("../../analysis/UCUT/result/Param_ind/3.Rdata");
+Param_ind <- resultForSave[[1]];
+
+if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
+  visPMSignature(Param_ind, 1, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../result/Fig3/UCUT_APOBEC_ind.eps", height = 5, width = 7.5);
+  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../result/Fig3/UCUT_AA_ind.eps", height = 5, width = 7.5);
+} else {
+  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../result/Fig3/UCUT_APOBEC_ind.eps", height = 5, width = 7.5);
[…]

### 2015-05-26 generateFigure/script/Fig2_Fig4_Fig5_Fig6_SFig2.R
organized figure generating script (Alexandrov et al. analysis) and tex data

@@ -1,0 +1,484 @@
+# Script for generating figures in SText 1 (simulation study)
+# First, we need the results ("../../analysis/UCUT/result"),
+
+if (!file.exists("../result/Fig2")) {
+  dir.create("../result/Fig2")
+}
+
+if (!file.exists("../result/Fig4")) {
+  dir.create("../result/Fig4")
+}
+
+if (!file.exists("../result/Fig5")) {
+  dir.create("../result/Fig5")
+}
+
+if (!file.exists("../result/Fig6")) {
+  dir.create("../result/Fig6")
+}
+
+if (!file.exists("../result/SFig2")) {
+  dir.create("../result/SFig2")
+}
+
+
+##########
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("utils.R")
+
+
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4, 2);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
[…]

### 2015-05-26 generateFigure/script/Fig2_Fig4_Fig5_SFig2.R

@@ -1,0 +1,460 @@
+
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("utils.R")
+
+
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4, 2);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
+  # indList <- which(closeSig[i,] > 0);
[…]

### 2015-05-26 generateFigure/script/subscript_Fig3/UCUT_signature.R

@@ -1,0 +1,36 @@
+library(pmsignature);
+
+source("utils.R");
+
+load("../../analysis/UCUT/result/Param_full/3.Rdata");
+Param_full <- resultForSave[[1]];
+
+if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
+  visPMSignature(Param_full, 1);
+  ggsave("../result/Fig3/UCUT_APOBEC_full.eps", height = 5, width = 7.5);
+  visPMSignature(Param_full, 2);
+  ggsave("../result/Fig3/UCUT_AA_full.eps", height = 5, width = 7.5); 
+} else {
+  visPMSignature(Param_full, 2);
+  ggsave("../result/Fig3/UCUT_APOBEC_full.eps", height = 5, width = 7.5);
+  visPMSignature(Param_full, 1);
+  ggsave("../result/Fig3/UCUT_AA_full.eps", height = 5, width = 7.5);   
+}
+  
+load("../../analysis/UCUT/result/Param_ind/3.Rdata");
+Param_ind <- resultForSave[[1]];
+
+if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
+  visPMSignature(Param_ind, 1, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../result/Fig3/UCUT_APOBEC_ind.eps", height = 5, width = 7.5);
+  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../result/Fig3/UCUT_AA_ind.eps", height = 5, width = 7.5);
+} else {
+  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
+  dev.copy2eps(file="../result/Fig3/UCUT_APOBEC_ind.eps", height = 5, width = 7.5);
[…]

### 2015-05-26 generateFigure/script/utils.R

@@ -1,0 +1,306 @@
+
+##########
+# functions for drawing figures for the Alexandrov et al analysis data
+
+
+visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale = TRUE, alpha = 2, charLimit = 0.25) {
+  
+  if (is.na(baseCol)) {
+    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
+    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
+  }
+  
+  centerBase <- (1 + numBases) / 2;
+  
+  v1 <- vF[1,1:6];
+  V2 <- vF[2:(numBases),1:4];
+  A <- matrix(0, numBases, 4);
+  B <- matrix(0, 4, 4);
+  
+  if (trDir == TRUE) {
+    v3 <- vF[(numBases + 1),1:2];
+  }
+  
+  for (l in 1:numBases) {
+    if (l < centerBase) {
+      A[l, ] <- V2[l, ];
+    } else if (l > centerBase) {
+      A[l, ] <- V2[l - 1, ];
+    }
+  }
+  A[centerBase,2] <- sum(v1[1:3]);
+  A[centerBase,4] <- sum(v1[4:6]);
+  
+  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
+  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
+  
+  num2base <- c("A", "C", "G", "T");
+  
+  # frame();
+  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
+  
+  renyi = function(p, tAlpha = alpha) {
+    if (tAlpha == 1) {
+      return(- sum(p * log2(p), na.rm = TRUE));
+    } else {
+      return( log(sum(p^tAlpha)) / (1 - tAlpha));
+    }
+  }
+  
+  # sizes <- 0.5 * (2 - apply(A, MARGIN = 1, FUN = function(p, alpha = scale) {-sum(x * log2(x), na.rm = TRUE)}));
+  # sizes <- sizes ** scale
+  
+  if (scale == FALSE) {
+    sizes <- rep(1, numBases)
+  } else {
[…]

### 2015-05-26 makeFigure/script/AlexandrovEtAl_figures_dir.R

@@ -1,460 +1,0 @@
-
-library(Matrix);
-library(corrplot);
-library(pmsignature);
-
-.pardefault <- par(no.readonly = TRUE);
-
-# functions
-source("AlexandrovEtAl_function.R")
-
-
-#' Here, we focus on the case where substitution patterns 
-#' and two 5' and 3' bases are considered
-fdim <- c(6, 4, 4, 4, 4, 2);
-SigMat <- matrix(0, 0, prod(fdim));
-# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
-eps <- 0.6
-
-#' create the matrix showing the relationships between cancer type 
-#' and the specified numer of mutation signatures
-type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
-
-#' reading and converting the mutation signatures for each cancer type
-for (i in 1:nrow(type2sigNum)) {
-  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
-}
-
-SigV <- rownames(SigMat);
-
-#' calculate cosine distances among mutation signatures
-#' and list up the pairs having smaller distances than the specified threshould
-# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
-closeSig <- diag(nrow(SigMat));
-dist <- c();
-for (i in 1:(nrow(SigMat) - 1)) {
-  for (j in (i+1):nrow(SigMat)) {
-    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
-    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
-    # dist <- c(dist, d); # for checking the histgram of cosine distances
-    if ( d < eps) {
-      closeSig[i,j] <- 1; 
-    }
-  }
-}
-
-
-
-#' clustering the similar mutation signatures
-EXPCLOSESIG <- expm(closeSig);
-removeList <- rep(1, nrow(SigMat));
-curInd <- 1;
-typeVec <- c();
-
-Fs <- list();
-for (i in 1:length(removeList)) {
-  
-  samples <- c();
[…]

### 2015-05-26 makeFigure/script/AlexandrovEtAl_function.R

@@ -1,306 +1,0 @@
-
-##########
-# functions for drawing figures for the Alexandrov et al analysis data
-
-
-visPMS_ind <- function(vF, numBases, baseCol = NA, trDir, charSize = 1.2, scale = TRUE, alpha = 2, charLimit = 0.25) {
-  
-  if (is.na(baseCol)) {
-    gg_color_hue6 <- hcl(h = seq(15, 375, length = 7), l=65, c=100)[1:6]
-    baseCol <- c(gg_color_hue6[3], gg_color_hue6[5], gg_color_hue6[2], gg_color_hue6[1], gg_color_hue6[4], gg_color_hue6[6]);
-  }
-  
-  centerBase <- (1 + numBases) / 2;
-  
-  v1 <- vF[1,1:6];
-  V2 <- vF[2:(numBases),1:4];
-  A <- matrix(0, numBases, 4);
-  B <- matrix(0, 4, 4);
-  
-  if (trDir == TRUE) {
-    v3 <- vF[(numBases + 1),1:2];
-  }
-  
-  for (l in 1:numBases) {
-    if (l < centerBase) {
-      A[l, ] <- V2[l, ];
-    } else if (l > centerBase) {
-      A[l, ] <- V2[l - 1, ];
-    }
-  }
-  A[centerBase,2] <- sum(v1[1:3]);
-  A[centerBase,4] <- sum(v1[4:6]);
-  
-  B[2, c(1, 3, 4)] <- v1[1:3] / sum(v1[1:3]);
-  B[4, c(1, 2, 3)] <- v1[4:6] / sum(v1[4:6]);
-  
-  num2base <- c("A", "C", "G", "T");
-  
-  # frame();
-  # plot.window(xlim=c(-0.05, 1.25 * numBases + 0.05), ylim=c(-0.25, 3.25));
-  
-  renyi = function(p, tAlpha = alpha) {
-    if (tAlpha == 1) {
-      return(- sum(p * log2(p), na.rm = TRUE));
-    } else {
-      return( log(sum(p^tAlpha)) / (1 - tAlpha));
-    }
-  }
-  
-  # sizes <- 0.5 * (2 - apply(A, MARGIN = 1, FUN = function(p, alpha = scale) {-sum(x * log2(x), na.rm = TRUE)}));
-  # sizes <- sizes ** scale
-  
-  if (scale == FALSE) {
-    sizes <- rep(1, numBases)
-  } else {
[…]

### 2015-05-26 makeFigure/script/subscript_Fig3/UCUT_signature.R

@@ -1,36 +1,0 @@
-library(pmsignature);
-
-source("AlexandrovEtAl_function.R");
-
-load("../../analysis/UCUT/result/Param_full/3.Rdata");
-Param_full <- resultForSave[[1]];
-
-if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + sum(Param_full@signatureFeatureDistribution[1,1,(256 * 6 + 1):(256 * 9)]) > 0.6 & sum(Param_full@signatureFeatureDistribution[2,1,(256 * 3 + 1):(256 * 4)]) + sum(Param_full@signatureFeatureDistribution[2,1,(256 * 9 + 1):(256 * 10)]) > 0.6) {
-  visPMSignature(Param_full, 1);
-  ggsave("../result/Fig3/UCUT_APOBEC_full.eps", height = 5, width = 7.5);
-  visPMSignature(Param_full, 2);
-  ggsave("../result/Fig3/UCUT_AA_full.eps", height = 5, width = 7.5); 
-} else {
-  visPMSignature(Param_full, 2);
-  ggsave("../result/Fig3/UCUT_APOBEC_full.eps", height = 5, width = 7.5);
-  visPMSignature(Param_full, 1);
-  ggsave("../result/Fig3/UCUT_AA_full.eps", height = 5, width = 7.5);   
-}
-  
-load("../../analysis/UCUT/result/Param_ind/3.Rdata");
-Param_ind <- resultForSave[[1]];
-
-if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
-  visPMSignature(Param_ind, 1, charSize = 1, isScale = TRUE, alpha = 2);
-  dev.copy2eps(file="../result/Fig3/UCUT_APOBEC_ind.eps", height = 5, width = 7.5);
-  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
-  dev.copy2eps(file="../result/Fig3/UCUT_AA_ind.eps", height = 5, width = 7.5);
-} else {
-  visPMSignature(Param_ind, 2, charSize = 1, isScale = TRUE, alpha = 2);
-  dev.copy2eps(file="../result/Fig3/UCUT_APOBEC_ind.eps", height = 5, width = 7.5);
[…]

### 2015-05-27 bioRxiv20150527/plos_template.tex
bioRxiv version

--- 2015-05-25 submission20150525/plos_template.tex
@@
 % Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
+% \raggedright
+% \setlength{\parindent}{0.5cm}
+% \textwidth 5.25in 
+% \textheight 8.75in
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
 
@@
 % Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
+% \usepackage{lastpage,fancyhdr,graphicx}
+% \usepackage{epstopdf}
 \pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
+% \pagestyle{fancy}
+% \fancyhf{}
+% \lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+% \rfoot{\thepage/\pageref{LastPage}}
+% \renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+% \fancyheadoffset[L]{2.25in}
+% \fancyfootoffset[L]{2.25in}
+% \lfoot{\sf PLOS}
 
@@
 
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
+% \newcommand{\lorem}{{\bf LOREM}}
+% \newcommand{\ipsum}{{\bf IPSUM}}
 
@@
 \begin{document}
-\vspace*{0.35in}
+% \vspace*{0.35in}
 
@@
 {\Large
-\textbf\newline{A simple model-based approach to inferring and visualizing cancer mutation signatures}
+\textbf{A simple model-based approach to inferring and visualizing cancer mutation signatures}
 }
@@
 
-\linenumbers
+% \linenumbers
 
@@
 }
+\includegraphics[width=15cm]{Fig1.eps}
 \label{mutSig_example}
@@
 }
+\includegraphics[width=15cm]{Fig2.eps}
 \label{methodOverview}
@@
 (E, F) APOBEC and AA signature stability (the mean cosine similarity for each donw-sampling ratio).}
+\includegraphics[width=15cm]{Fig3.eps}
 \label{UCUT}
@@
 }
+\includegraphics[width=15cm]{Fig4.eps}
 \label{nature2013_sig_summary}
@@
 \caption{{\bf The summary of membership of each mutation signature across 30 cancer types obtained using the proposed method.}}
+\includegraphics[width=15cm]{Fig5.eps}
 \label{nature2013_sig_member}
@@
 }
+\includegraphics[width=14cm]{Fig6.eps}
 \label{two5prime}
@@
 \begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
 \caption{
@@
 problem & $\bm{x}_{i,j}$ & $\bm{f}_{k}$ & $\bm{q}_{i}$  \\ \hline
-mutation signature model & the $j$-th mutation  & the feature dist.  & the signature dist.  \\
-& in the $i$-th cancer genome & for the $k$-th signature & for the $i$-th cancer genome \\ \hline
+mutation signature & the $j$-th mutation  & the feature dist.  & the signature dist.  \\
+model & in the $i$-th cancer genome & for the $k$-th signature & for the $i$-th cancer genome \\ \hline
 topic model &  the $j$-th word & the word dist. & the topic dist. \\ 
 & in the $i$-th document & for the $k$-th topic & for the $i$-th document \\ \hline
-population structure model & the $j$-th locus genotype & the allele freq.  & the admixture dist. \\
-& of the individual $i$  & for the ancestry $k$ & for the individual $i$ \\
+population structure  & the $j$-th locus genotype & the allele freq.  & the admixture dist. \\
+model & of the individual $i$  & for the ancestry $k$ & for the individual $i$ \\
 \hline 
@@
 \label{tab_pop}
-\end{adjustwidth}
+% \end{adjustwidth}
 \end{table}

### 2015-08-11 generateFigure/script/Fig_revise1.R
supplementary text modification

@@ -41,9 +41,9 @@ for (k in 1:4) {
   }
 }
 
-visPMSignature(Param, POLE1)
+visPMSignature(Param, POLE1, isScale = TRUE)
 ggsave("../result/Fig_revise1/POLE_pmsignature_1.eps", width=5, height=3, units="in")
 
-visPMSignature(Param, POLE2)
+visPMSignature(Param, POLE2, isScale = TRUE)
 ggsave("../result/Fig_revise1/POLE_pmsignature_2.eps", width=5, height=3, units="in")
 

### 2015-08-11 paper/manuscript/plos_template.tex
modify typos

@@
 revealed that C $>$ A mutations are abundant in lung cancers in patients with smoking history,
-and these are caused by benzo(a)pyrene included in tobacco smokes \cite{pmid12379884}.
+and these are caused by benzo(a)pyrene included in tobacco smoke \cite{pmid12379884}.
 Also, C $>$ T and CC $>$ TT  mutations are abundant in ultraviolet-light-associated skin cancers, 
@@
 \begin{figure}[h]
-\caption{{\bf The mutation signatures for the UCUT data, and the results of downsampling experiments.}
+\caption{{\bf The mutation signatures for the UCUT data, and the results of down-sampling experiments.}
 3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.
-(A) APOBEC signature for the independent model.
-(B) AA signature for the independent model.
-(C) APOBEC signature for the full model.
-(D) AA signature for the full model.
-(E) APOBEC signature stability.
-(F) AA signature stability.}
+(A, B) APOBEC and AA signature for the independent model.
+(C, D) APOBEC and AA signature for the full model.
+(E, F) APOBEC and AA signature stability (the mean cosine similarity for each down-sampling ratio).}
 \label{UCUT}
@@
 We applied the method to each cancer type separately to assess similarity of estimated signatures across cancer types.
-For each cancer type we selected the number of signatures $K$ by fitting the model with increasing $K$ and examining the log-likelihood, bootstrap errors, and correlation of membership parameters.The selected values of $K$ are given in S2 Table.
+For each cancer type we selected the number of signatures $K$ by fitting the model with increasing $K$ and examining the log-likelihood, bootstrap errors, and correlation of membership parameters.
+The selected values of $K$ are given in S2 Table.
 Also, we simply removed somatic mutations located in an intergenic region to include transcription strand biases as mutation features.

### 2015-08-11 paper/manuscript/plos_template.tex
add definition of Frobenius distance

@@
 \newcommand{\ipsum}{{\bf IPSUM}}
+
+\DeclareMathOperator{\Tr}{Tr}
 
@@
 Finally, we merged similar mutation signatures across different cancer types 
-(when their Frobenius Distance were $<$ $0.6$). 
+(when their Frobenius Distance were $<$ $0.6$, where Frobenius Distance between two matrices ($\bm{F}_1$, $\bm{F}_2$) is $\sqrt{ \Tr  (\bm{F}_1 -  \bm{F}_2 )  (\bm{F}_1 -  \bm{F}_2 )^t }$).
 % XXX WHAT KIND OF CLUSTERING?

### 2015-08-11 generateFigure/script/Fig_for_tremmel.R
add rebutall modified by MS

@@ -1,0 +1,101 @@
+# Script for generating data for tremmel.
+
+
+if (!file.exists("../result/Fig_tremmel")) {
+  dir.create("../result/Fig_tremmel")
+}
+
+##########
+library(Matrix);
+library(corrplot);
+library(pmsignature);
+
+.pardefault <- par(no.readonly = TRUE);
+
+# functions
+source("utils.R")
+
+
+#' Here, we focus on the case where substitution patterns 
+#' and two 5' and 3' bases are considered
+fdim <- c(6, 4, 4, 4, 4, 2);
+SigMat <- matrix(0, 0, prod(fdim));
+# eps <- 2 - 2 * cos(pi / 5); # this value can be tuned.. where is the best value??
+eps <- 0.6
+
+#' create the matrix showing the relationships between cancer type 
+#' and the specified numer of mutation signatures
+type2sigNum <- read.table("../data/AlexandrovEtAl_sigNum.txt", sep=",", header=FALSE);
+
+#' reading and converting the mutation signatures for each cancer type
+for (i in 1:nrow(type2sigNum)) {
+  SigMat <- rbind(SigMat, getSigMat(type2sigNum[i,1], type2sigNum[i,2], trDir = TRUE));
+}
+
+SigV <- rownames(SigMat);
+
+#' calculate cosine distances among mutation signatures
+#' and list up the pairs having smaller distances than the specified threshould
+# closeSig <- Matrix(0, nrow(SigMat), nrow(SigMat));
+closeSig <- diag(nrow(SigMat));
+dist <- c();
+for (i in 1:(nrow(SigMat) - 1)) {
+  for (j in (i+1):nrow(SigMat)) {
+    d <- sqrt(sum( (SigMat[i,] - SigMat[j,])^2 ));
+    # d <- 2 - 2 * innerProd(SigMat[i,], SigMat[j,]);
+    # dist <- c(dist, d); # for checking the histgram of cosine distances
+    if ( d < eps) {
+      closeSig[i,j] <- 1; 
+    }
+  }
+}
+
+
+
+#' clustering the similar mutation signatures
+EXPCLOSESIG <- expm(closeSig);
+removeList <- rep(1, nrow(SigMat));
+curInd <- 1;
+typeVec <- c();
+
+Fs <- list();
+for (i in 1:length(removeList)) {
+  
+  samples <- c();
+  # indList <- which(closeSig[i,] > 0);
[…]

### 2015-08-17 generateFigure/script/Fig2_Fig4_Fig5_Fig6_SFig3.R
modify supp Fig file names

@@ -17,8 +17,8 @@ if (!file.exists("../result/Fig6")) {
   dir.create("../result/Fig6")
 }
 
-if (!file.exists("../result/SFig2")) {
-  dir.create("../result/SFig2")
+if (!file.exists("../result/SFig3")) {
+  dir.create("../result/SFig3")
 }
 
 
@@ -319,7 +319,7 @@ for (i in 1:length(importantSigs)) {
   
   
   
-  outputName <- paste("../result/SFig2/signatureList_", names(importantSigs)[i], ".eps", sep="");
+  outputName <- paste("../result/SFig3/signatureList_", names(importantSigs)[i], ".eps", sep="");
   
   if (length(types) > 4) {
     dev.copy2eps(file=outputName, height = ceiling(length(types) / 4) * 2.4, width = 15);

### 2015-08-17 generateFigure/script/SFig4_SFig5.R

@@ -2,12 +2,12 @@
 # data generated in "../../analysis/AlexandrovEtAl/result/" is necessary.
 
 
-if (!file.exists("../result/SFig3")) {
-  dir.create("../result/SFig3")
+if (!file.exists("../result/SFig4")) {
+  dir.create("../result/SFig4")
 }
 
-if (!file.exists("../result/SFig4")) {
-dir.create("../result/SFig4")
+if (!file.exists("../result/SFig5")) {
+dir.create("../result/SFig5")
 }
 
 
@@ -27,7 +27,7 @@ Param@signatureFeatureDistribution[1,1,] <- nature2013_sig[which(rownames(nature
 Param@signatureFeatureDistribution[2,1,] <- nature2013_sig[which(rownames(nature2013_sig) == "Signature.14"),];
 
 visPMSignature(Param, 1, isScale = TRUE)
-ggsave("../result/SFig3/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
+ggsave("../result/SFig4/example96_Glioma-Low-Grade.eps", width=12, height=3, units="in");
 
 
 ##########
@@ -39,10 +39,10 @@ G <- readMPFile("../../analysis/AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.
 # membership plot
 load("../../analysis/AlexandrovEtAl/result/Param_ind5_dir/Glioma-Low-Grade.3.Rdata");
 visMembership(G, resultForSave[[1]], ylog = FALSE, colourBrewer = "Set2", toSample = 100);
-ggsave("../result/SFig3/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
+ggsave("../result/SFig4/Glioma-Low-Grade_membership.eps", height = 4, width = 12, pointsize = 9);
 
 visMembership(G, resultForSave[[1]], ylog = TRUE, colourBrewer = "Set2", toSample = 100);
-ggsave("../result/SFig3/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
+ggsave("../result/SFig4/Glioma-Low-Grade_membership_log.eps", height = 4, width = 12, pointsize = 9);
 ##########
 
 
@@ -75,7 +75,7 @@ for (l1 in 1:length(Params)) {
     visPMSignature(Params[[l1]], aInd, charSize = 1, isScale = TRUE);
   }
   
[…]

### 2015-08-17 generateFigure/script/SFig6.R

@@ -2,8 +2,8 @@
 # data generated in "../../analysis/AlexandrovEtAl/result/" is necessary.
 
 
-if (!file.exists("../result/SFig5")) {
-  dir.create("../result/SFig5")
+if (!file.exists("../result/SFig6")) {
+  dir.create("../result/SFig6")
 }
 
 
@@ -49,7 +49,7 @@ for (i in 1:length(oxidation_types)) {
   G <- readMPFile(inputMPdata, numBases = 5, trDir = TRUE);
   load(inputRdata);
   visMembership(G, resultForSave[[1]], toSample = 100, colourBrewer = "Set2");
-  ggsave(paste("../result/SFig5/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
+  ggsave(paste("../result/SFig6/", oxidation_types[i], "_oxidation_membership.eps", sep=""), height = 4, width = 12, pointsize = 9);
 
   oxInd <- checkOxidationInd(resultForSave[[1]])
   
@@ -57,7 +57,7 @@ for (i in 1:length(oxidation_types)) {
   par(mar = 0.2 * tempMar);
   
   visPMSignature(resultForSave[[1]], oxInd, charSize = 0.8, isScale = TRUE);
-  dev.copy2eps(file=paste("../result/SFig5/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 3.0, width = 5);    
+  dev.copy2eps(file=paste("../result/SFig6/", oxidation_types[i], "_oxidation_signature.eps", sep=""), height = 3.0, width = 5);
 
   par(.pardefault);
   

### 2015-08-17 submission20150818/plos_template.tex
add submission20150818

--- 2015-08-11 paper/manuscript/plos_template.tex
@@
 signatures from our independence model, because of its ability to effectively
-and parsimoniously deal with both $\pm2$ flanking
+and parsimoniously deal with both $\pm 2$ flanking
 base context and strand bias, are often more refined, highlighting additional details or features not previously evident.
@@
 In previous analyses of these data \cite{pmid23945592}, 
-the signature for Pol $\epsilon$ dysfunction was represented by a single signature (their ``signature 10'').
+the signature for Pol $\epsilon$ dysfunction was represented by a single signature (their ``signature 10'', see Fig. S2).
 In contrast our new approach uses two signatures.
@@
 that inferring them as a single signature is ``wrong". However, splitting them into two signatures does help highlight certain features. Specifically,
-signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Figs. S2C S2D).
+signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Figs. S3C and S3D).
 This strand bias may be connected with the enrichment of 
@@
 and signature 1 also shows slightly elevated frequency of the T base at position $+2$
- (Figs. 6B, 6C ,S6C and S6D).
+ (Figs. 6B, 6C ,S3C and S3D).
 A previous study of Pol $\epsilon$  \cite{pmid25228659} 
@@
 revealed that this signature is driven by 
-a single sample with an extremely high mutation rate (see Figs. S3A and S3B), and signature 2 disappeared when we removed this sample (see Fig. S4).
+a single sample with an extremely high mutation rate (see Figs. S4A and S4B), and signature 2 disappeared when we removed this sample (see Fig. S5).
 It may be that the complex low-grade-glioma specific signature detected in the previous study is driven by the same single sample.
@@
 \cite{pmid23318258} (see their Supplementary Figures).
-Again our analysis highlights additional contextual information, with a strikingly elevated frequency of base C at the $-2$ position (Figs. S5A, S5B and S5C).
-However, for each cancer type, only a few samples support this signature (see Figs. S5D, S5E and S5F),
+Again our analysis highlights additional contextual information, with a strikingly elevated frequency of base C at the $-2$ position (Figs. S6A, S6B and S6C).
+However, for each cancer type, only a few samples support this signature (see Figs. S6D, S6E and S6F),
 and the corresponding signature 
@@
 The 12 inferred signatures are highly consistent among cancer types except for B-cell lymphoma (see Fig.
-S2A), highlighting the robustness
-of our approach.  Almost all of them show enrichment of A and T and depletion of G base at the $-2$ position (Fig. 6A and S2A), 
+S3A), highlighting the robustness
+of our approach.  Almost all of them show enrichment of A and T and depletion of G base at the $-2$ position (Fig. 6A and S3A), 
 consistent with the UCUT data above and previous analyses \cite{pmid23318258}.
@@
 Signature 10 (C $>$T at [CT]pCpC) was observed in head and neck cancers and melanomas, and probably relates to ultraviolet light.
-Consistent strand specificities among the two cancer types (Fig. S2E) matches previous results \cite{pmid23318258}, 
-but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figs. 6D and S2E). 
+Consistent strand specificities among the two cancer types (Fig. S3E) matches previous results \cite{pmid23318258}, 
+but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figs. 6D and S3E). 
 Signature 11 (C $>$ T at GpCp[CG]) appears in small-cell lung cancers and stomach cancers,
-and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear. Again our analysis highlights elevated abundance of G at the $-2$ position (Figs. 6E and S2F).
+and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear. Again our analysis highlights elevated abundance of G at the $-2$ position (Figs. 6E and S3F).
 Signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively, agree well with those detected in the same cancer types in the previous study \cite{pmid23318258}.
@@
 Signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
-The estimated signature in each cancer type shows higher mutation prevalence on the template strand (Fig. S2B), 
+The estimated signature in each cancer type shows higher mutation prevalence on the template strand (Fig. S3B), 
 which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
@@
 \subsection*{S2 Figure}
+\label{nature2013_POLE}
+{\bf The pol epsilon signature (Signature 10) derived in the Alexandrov et al., (2013).}
+The barplots are divided by 6 substitution pattern. 
+In each division, 16 bars show joint probabilities of substitution pattern, 5' and 3' bases (ApNpA, ApNpC, ApNpG, ApNpT, $\cdots$, TpNpT).
+
+\subsection*{S3 Figure}
 \label{nature2013_sig_list}
@@
 
-\subsection*{S3 Figure}
+\subsection*{S4 Figure}
 \label{LGG_membership}
@@
 
-\subsection*{S4 Figure}
+\subsection*{S5 Figure}
 \label{LGG_signatures}
@@
 
-\subsection*{S5 Figure}
+\subsection*{S6 Figure}
 \label{oxidation}

### 2015-08-21 generateFigure/script/Fig_for_tremmel.R
modify supp table

@@ -95,7 +95,10 @@ for (i in 1:length(removeList)) {
 newOrder <- sortSignature(Fs)
 
 
-
+num <- 1
 for (i in newOrder) {
   write.table(Fs[[i]], file = paste("../result/Fig_tremmel/sig_", i, ".txt", sep = ""), sep="\t", col.names = FALSE, row.names = FALSE)
+  num <- num + 1
 }
+
+

### 2015-08-21 generateFigure/script/Fig_for_tremmel.R
modify supp fig 1

@@ -97,7 +97,7 @@ newOrder <- sortSignature(Fs)
 
 num <- 1
 for (i in newOrder) {
-  write.table(Fs[[i]], file = paste("../result/Fig_tremmel/sig_", i, ".txt", sep = ""), sep="\t", col.names = FALSE, row.names = FALSE)
+  write.table(Fs[[i]], file = paste("../result/Fig_tremmel/sig_", num, ".txt", sep = ""), sep="\t", col.names = FALSE, row.names = FALSE)
   num <- num + 1
 }
 

### 2015-08-21 generateFigure/script/SFig1.R

@@ -22,9 +22,9 @@ Param_ind <- resultForSave[[1]];
 ## ind model 
 
 if (sum(Param_ind@signatureFeatureDistribution[1,1,1:3]) > 0.75 & Param_ind@signatureFeatureDistribution[2,1,4] > 0.75) {
-  baseTwoFivePrime <- data.frame(model=rep("ind", 4), base=c("A", "C", "G", "T"), prob = Param_ind@signatureFeatureDistribution[1,2,1:4]);
+  baseTwoFivePrime <- data.frame(model=rep("ind", 4), base=c("A", "C", "G", "T"), frequency = Param_ind@signatureFeatureDistribution[1,2,1:4]);
 } else {
-  baseTwoFivePrime <- data.frame(model=rep("ind", 4), base=c("A", "C", "G", "T"), prob = Param_ind@signatureFeatureDistribution[2,2,1:4]);
+  baseTwoFivePrime <- data.frame(model=rep("ind", 4), base=c("A", "C", "G", "T"), frequency = Param_ind@signatureFeatureDistribution[2,2,1:4]);
 }
 
 ## full model
@@ -34,21 +34,21 @@ if (sum(Param_full@signatureFeatureDistribution[1,1,(256 * 0 + 1):(256 * 3)]) + 
   tempC <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 4 & 0:3071 %% 16 < 8]);
   tempG <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 8 & 0:3071 %% 16 < 12]);
   tempT <- sum(Param_full@signatureFeatureDistribution[1,1,0:3071 %% 16 >= 12 & 0:3071 %% 16 < 16]);
-  temp <- data.frame(model=rep("full", 4), base=c("A", "C", "G", "T"), prob=c(tempA, tempC, tempG, tempT));
+  temp <- data.frame(model=rep("full", 4), base=c("A", "C", "G", "T"), frequency=c(tempA, tempC, tempG, tempT));
   baseTwoFivePrime <- rbind(baseTwoFivePrime, temp);
 } else {
   tempA <- sum(Param_full@signatureFeatureDistribution[2,1,0:3071 %% 16 >= 0 & 0:3071 %% 16 < 4]);
   tempC <- sum(Param_full@signatureFeatureDistribution[2,1,0:3071 %% 16 >= 4 & 0:3071 %% 16 < 8]);
[…]

### 2015-09-14 analysis/simulation/script/runall_epi.sh
epigenetic data addition simulation

@@ -1,0 +1,50 @@
+#! /bin/sh
+#$ -S /bin/sh
+#$ -cwd
+
+export R_LIBS=/home/yshira/.R
+export R_PATH=/home/yshira/local/bin
+
+if [ ! -d ../result ]
+then
+    mkdir ../result
+fi
+
+if [ ! -d ../result/cosineDist_epi ]
+then
+    mkdir ../result/cosineDist_epi
+fi
+
+if [ ! -d log ]
+then
+    mkdir log
+fi
+
+SCRIPTDIR=`pwd`
+cd ../result
+RESULTDIR=`pwd`
+cd ../script/subscript
+SUBSCRIPTDIR=`pwd`
+
+
+for FEATNUM in 0 1 2 3 4 5; do
+    # for SAMPLENUM in 10 25 50 100; do
+    for SAMPLENUM in 10; do
+        # for PARAM_ALPHA in 0.5 1 2; do
+	for PARAM_ALPHA in 1; do
+            # for PARAM_GAMMA in 0.5 1 2; do
+	    for PARAM_GAMMA in 1; do
+
+                OUTPUTFILE=${RESULTDIR}/cosineDist_epi/${FEATNUM}_${SAMPLENUM}_${PARAM_ALPHA}_${PARAM_GAMMA}.txt
+
+                # echo "qsub -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}"
+                # qsub -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}
+	
+		echo "sh repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}"
+		sh repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}
+
+            done;
+        done;
+    done;
+done
+

### 2015-09-14 analysis/simulation/script/subscript/repeatSimulation_epi.R

@@ -1,0 +1,21 @@
+library(pmsignature)
+
+mutationNum <- 25;
+featureNum <- as.integer(commandArgs()[5]);
+sampleNum <- as.integer(commandArgs()[6]);
+alpha <- as.numeric(commandArgs()[7]);
+gamma <- as.numeric(commandArgs()[8]);
+outputFile <- commandArgs()[9];
+
+fdim <- c(6, 4, 4, 4, 4, rep(2, featureNum));
+
+Dists <- matrix(0, 100, 4);
+
+for (i in 1:100) {
+  G_sim <- makeSimData(type = "custom", K = 4, fdim = fdim, sampleNum = sampleNum, mutationNum = mutationNum, param_alpha = alpha, param_gamma = gamma);
+  Param_est <- getPMSignature(G_sim[[1]], K = 4);
+  Dists[i, ] <- getCosDistance(Param_est@signatureFeatureDistribution, G_sim[[2]], Param_est@possibleFeatures);
+}
+
+
+write.table(Dists, file=outputFile, quote=FALSE, row.names=FALSE, col.names=FALSE, sep="\t")

### 2015-09-14 analysis/simulation/script/subscript/repeatSimulation_epi.sh

@@ -1,0 +1,16 @@
+#! /bin/sh
+#$ -S /bin/sh
+#$ -cwd
+
+# export R_LIBS=/home/yshira/.R
+# export R_PATH=/home/yshira/local/bin
+export R_PATH=/usr/bin
+
+FEATNUM=$1
+SAMPLENUM=$2
+ALPHA=$3
+GAMMA=$4
+OUTPUTFILE=$5
+
+echo "${R_PATH}/R --vanilla --slave --args ${FEATNUM} ${SAMPLENUM} ${ALPHA} ${GAMMA} ${OUTPUTFILE} < repeatSimulation_epi.R"
+${R_PATH}/R --vanilla --slave --args ${FEATNUM} ${SAMPLENUM} ${ALPHA} ${GAMMA} ${OUTPUTFILE} < repeatSimulation_epi.R

### 2015-09-15 analysis/simulation/script/runall_epi.sh
page

@@ -27,7 +27,7 @@ cd ../script/subscript
 SUBSCRIPTDIR=`pwd`
 
 
-for FEATNUM in 0 1 2 3 4 5; do
+for FEATNUM in 0 1 2 3 4 5 6 7 8; do
     # for SAMPLENUM in 10 25 50 100; do
     for SAMPLENUM in 10; do
         # for PARAM_ALPHA in 0.5 1 2; do

### 2015-09-15 analysis/simulation/script/subscript/repeatSimulation_epi.R

@@ -1,13 +1,13 @@
 library(pmsignature)
 
-mutationNum <- 25;
+mutationNum <- 50;
 featureNum <- as.integer(commandArgs()[5]);
 sampleNum <- as.integer(commandArgs()[6]);
 alpha <- as.numeric(commandArgs()[7]);
 gamma <- as.numeric(commandArgs()[8]);
 outputFile <- commandArgs()[9];
 
-fdim <- c(6, 4, 4, 4, 4, rep(2, featureNum));
+fdim <- c(6, 4, 4, rep(2, featureNum));
 
 Dists <- matrix(0, 100, 4);
 

### 2015-09-16 analysis/simulation/script/runall_epi.sh
intensive epigenetic simulation analysis

@@ -28,20 +28,14 @@ SUBSCRIPTDIR=`pwd`
 
 
 for FEATNUM in 0 1 2 3 4 5 6 7 8; do
-    # for SAMPLENUM in 10 25 50 100; do
-    for SAMPLENUM in 10; do
-        # for PARAM_ALPHA in 0.5 1 2; do
-	for PARAM_ALPHA in 1; do
-            # for PARAM_GAMMA in 0.5 1 2; do
-	    for PARAM_GAMMA in 1; do
+    for SAMPLENUM in 10 25 50 100; do
+        for PARAM_ALPHA in 0.5 1 2; do
+            for PARAM_GAMMA in 0.5 1 2; do
 
                 OUTPUTFILE=${RESULTDIR}/cosineDist_epi/${FEATNUM}_${SAMPLENUM}_${PARAM_ALPHA}_${PARAM_GAMMA}.txt
 
-                # echo "qsub -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}"
-                # qsub -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}
-	
-		echo "sh repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}"
-		sh repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}
+                echo "qsub -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}"
+                qsub -e ${SCRIPTDIR}/log -o ${SCRIPTDIR}/log repeatSimulation_epi.sh ${FEATNUM} ${SAMPLENUM} ${PARAM_ALPHA} ${PARAM_GAMMA} ${OUTPUTFILE}
 
             done;
         done;

### 2015-09-16 analysis/simulation/script/subscript/repeatSimulation_epi.sh

@@ -2,9 +2,8 @@
 #$ -S /bin/sh
 #$ -cwd
 
-# export R_LIBS=/home/yshira/.R
-# export R_PATH=/home/yshira/local/bin
-export R_PATH=/usr/bin
+export R_PATH=/usr/local/package/r/3.2.1/bin
+export R_LIBS=~/.R
 
 FEATNUM=$1
 SAMPLENUM=$2

### 2015-09-16 generateFigure/script/SFig7.R
script for generating (supp figure 7)

@@ -1,0 +1,13 @@
+# Script for generating figures in SFig7 (simulation study)
+# First, we need the results ("../../analysis/simulation/result/cosineDist_epi"),
+# generated by the script (""../../analysis/simulation/script/runall_epi.sh").
+
+
+if (!file.exists("../result/SFig7")) {
+  dir.create("../result/SFig7")
+}
+
+# script for generating the cosine distances for varying parameters
+# output: "../result/SFig7/simulation_result_epi.eps"
+source("subscript_SFig7/simulation_result_epi.R")
+

### 2015-09-16 generateFigure/script/subscript_SFig7/simulation_result_epi.R

@@ -1,0 +1,69 @@
+library(ggplot2);
+library(dplyr);
+
+
+featNum <- c(0, 1, 2, 3, 4, 5, 6, 7, 8);
+sampleNum <- c(10, 25, 50, 100);
+alpha <- c(0.5, 1, 2);
+gamma <- c(0.5, 1, 2);
+
+FeatNum <- rep(0, 21600);
+SampleNum <- rep(0, 21600);
+Alpha <- rep(0, 21600);
+Gamma <- rep(0, 21600);
+Dist <- rep(0, 21600);
+
+Ind <- 1:100;
+
+for (i1 in 1:6) {
+  for (i2 in 1:4) {
+    for (i3 in 1:3) {
+      for (i4 in 1:3) {
+        a <- read.table(paste("../../analysis/simulation/result/cosineDist_epi/", featNum[i1], "_", sampleNum[i2], "_", alpha[i3], "_", gamma[i4], ".txt", sep=""), sep="\t", header=FALSE);
+        
+        Dist[Ind] <- rowSums(a) / 4;
+        FeatNum[Ind] <- featNum[i1];
+        SampleNum[Ind] <- sampleNum[i2];
+        Alpha[Ind] <- alpha[i3];
+        Gamma[Ind] <- gamma[i4];
+        
+        Ind <- Ind + 100;
+      }
+    }
+  }
+}
+
+
+simulation.result <- data.frame(featNum=FeatNum, distance=Dist);
+summary_simulation.result <- simulation.result  %>% group_by(featNum) %>% summarise(cosine_similality = mean(distance), seDist = sd(distance));
+summary_simulation.result$seDist2 <- sapply(summary_simulation.result$cosine_similality + summary_simulation.result$seDist, min, 1) - summary_simulation.result$cosine_similality;
+
+simulation.result <- data.frame(featNum=FeatNum, sampleNum=SampleNum, alpha=Alpha, gamma=Gamma, distance=Dist);
+summary_simulation.result <- simulation.result  %>% group_by(featNum, sampleNum, alpha, gamma) %>% summarise(cosine_similality = mean(distance), seDist = sd(distance));
+summary_simulation.result$seDist2 <- sapply(summary_simulation.result$cosine_similality + summary_simulation.result$seDist, min, 1) - summary_simulation.result$cosine_similality;
+
+
+
+pd <- position_dodge(0.2);
[…]

### 2015-09-16 revise/rebuttal150916.txt
modify rebuttal

#Reviewer 1

The authors have declined to incorporate the changes that I suggested. As a consequence, I can only standby my original comments and leave the final decision to the Editor. This is somewhat disappointing as I feel the additional data would substantially elevate the authors work. I would remark that if no additions are to be made then the authors should qualify the claims in the discussion somewhat. Whilst adding additional features into the model is straightforward, the authors present no evidence to indicate that the resultant analysis would yield anything sensible.

—————
As the reviewer says, performing small analysis incorporating some epigenetic data without careful validation is “straightforward” and not so hard. We have newly included simulation study, the result of which indicate that adding epigenetic features will help improving the accuracy of mutation signature detection as well as gaining novel biological insights.

There are several issues to be addressed for incorporating epigenetic data to mutation signature on real data. First, the choice of normalization methods of epigenetic data will strongly influence the result. Also, there are known relationships between epigenetic features and sequencing depth and errors (replication timing, CG contents) leading the variation in the sensitivity and false positive ratios. So we need to carefully control them.

In spite of the above obstacles, many studies have partly revealed that somatic mutations (mostly structural variations so far) have some relationships with epigenetic features and people will start to try clarifying more comprehensive relationships between mutation signatures and epigenetic data in the not-so-distant future. We believe our proposed model, which gives a natural way for incorporating many additional features, will be of great help in that analysis.
—————

caption for S7 Figure:
The accuracy of the proposed approach for the simulated data when changing the number of features starting from the case where just substation patterns and immediate 5’ and 3’ bases(M = (6, 4, 4), to 5 additional features (M = (6, 4, 4, 2, 2, 2, 2, 2) for each number of samples, dispersion parameters.  The accuracy of the estimated mutation signature improves as the number of additional features increases, indicating that incorporating additional features such as epigenetic data will be potentially beneficial.

### 2015-09-17 generateFigure/script/subscript_SFig7/RFplot_exp.R
kataegis plot test

@@ -1,0 +1,91 @@
+
+getDFForRF <- function(G, sampleInd, typeVec, typeName) {
+  
+  genome <- BSgenome.Hsapiens.UCSC.hg19::BSgenome.Hsapiens.UCSC.hg19
+  targetChr <- paste("chr", c(1:22, "X", "Y"), sep="")
+  seqlen_chr <- GenomeInfoDb::seqlengths(genome)[targetChr];
+  
+  genomeStartInd <- cumsum(as.numeric(seqlen_chr)) - as.numeric(seqlen_chr)
+  names(genomeStartInd) <- targetChr
+  
+  absPos <- genomeStartInd[G@mutationPosition[G@mutationPosition[,3] == sampleInd, 1]] + G@mutationPosition[G@mutationPosition[,3] == sampleInd, 2]
+  mutType <- factor(sapply(G@mutationPosition[G@mutationPosition[,3] == sampleInd, 4], 
+                    function(x) {if (x %in% typeVec) typeName else "other"}), levels = c(typeName, "other"))
+  
+  mutOrder <- order(absPos)
+  absPos_sort <- absPos[mutOrder]
+  mutType_sort <- mutType[mutOrder]
+  
+  dist1 <- abs(absPos_sort - c(0, absPos_sort[1:(length(absPos_sort)-1)]))
+  dist2 <- abs(absPos_sort - c(absPos_sort[2:length(absPos_sort)], 0))
+  dist <- apply(cbind(dist1, dist2), 1, min)
+  
+  return(data.frame(mutind = 1:length(absPos_sort), pos = absPos_sort, type = mutType_sort, log_dist = log10(dist)))
+  
+}
+
+
+
+
+if (!file.exists("../result/Fig_revise2")) {
+  dir.create("../result/Fig_revise2")
+}
+
+##########
+
+G <- readMPFile("../../analysis/AlexandrovEtAl/result/MPFormat/Glioma-Low-Grade.mp.txt.gz", numBases = 3)
+mutNum <- getMutNum(G)
+sampleInd <- which(mutNum[,2] == max(mutNum[,2]))
+
+mutinfo <- getDFForRF(G, sampleInd, c(4, 8, 12, 16), "Np[C>[A]]pT")
+
+ggplot(mutinfo, aes(x = mutind, y = log_dist, colour = type)) + 
+  geom_point() +
+  xlab("ordered mutation number") +
+  ylab("minimum log distance") +
[…]

### 2015-09-18 revise/rebuttal150916.txt
rebutal

@@
 
-—————
 As the reviewer says, performing small analysis incorporating some epigenetic data without careful validation is “straightforward” and not so hard.
@@
 We believe our proposed model, which gives a natural way for incorporating many additional features, will be of great help in that analysis.
-—————
 
@@
 The accuracy of the estimated mutation signature improves as the number of additional features increases, indicating that incorporating additional features such as epigenetic data will be potentially beneficial.
+
+#Reviewer 2
+
+Does this discussion include implicitly PWMs with Markovian structure between the bases, rather than just independent sites?
+
+In this line (L126-130), we did mean Markovian structure.
+However, considering probabilistic model with Markovian structure is definitely one possibility of extention and it is worth mentioning somewhere.
+So, we have included a small discussion in the Discussion section.
+
+---
+Also, we can give mutation signature probabilistic models somewhere between complete independence and non-independence,
+e.g., assuming Markovian structure between adjacent bases as some transcription factor binding site modelings.
+---
+
+Kataegis occurs in some tumours -- regions with extraordinarily high mutation rates. This is not always common, or in the same region. Perhaps worth a mention?
+
+Thanks for the interesting suggestion.
+After this notice, we’ve got interested in checking whether the LGG cancer genome with extremely high mutations has kataegis.
+However, we couldn’t see the characteristic of kataegis from the so called “rain fall plot” from this sample,
+so this extremely high mutations seem to be caused by other phenomena.
+
+Figure R1: Two “rainfall plot”, where mutations aare ordered by the chromosome and coordinates.
+on the x axis from the first variant on the short arm of chromosome 1 to the last variant on the long arm of chromosome X and are colored according to mutation-type.
+The distance between each mutation and the one prior to it (the intermutation distance) is plotted on the vertical axis on a log scale

### 2015-09-18 paper/manuscript/plos_template.tex
rebutal

--- 2015-08-11 paper/manuscript/plos_template.tex
@@
 were limited in several ways.
-Due to limited sequencing throughputs, most classical studies 
+Due to limited sequencing throughput, most classical studies 
 focused on a few cancer genes, such as TP53,
@@
 mutation signatures in vast amounts of somatic mutation data.
-Current statistical approaches \cite{pmid23318258,pmid23628380}, 
+Current statistical approaches \cite{pmid23318258,pmid23628380}
 %based on nonnegative matrix factorization (NMF, \cite{pmid10548103}) and Poisson models
@@
 each signature, greatly facilitating the incorporation of additional relevant sequence context. 
-For example, our approach can
-incorporate both the two bases
-3' and 5' of the substitution, and transcription strand biases using only 18 parameters
+For example, our approach can incorporate the two bases 3' and 5' of the substitution and transcription strand biases using only 18 parameters
 per signature, compared with 3071 parameters per signature with current approaches.
@@
 These benefits include: more stable estimation of signatures from smaller samples,
-refinement of the detail and resolution of many mutation signatures, and, possibly, identification of novel signatures.
-
-Assuming independence among features in a signature may initially seem unnatural. However, its use here is analogous  ``position weight matrix models" which have been highly
+refinement of the detail and resolution of many mutation signatures, and possibly identification of novel signatures.
+
+Assuming independence among features in a signature may initially seem unnatural. However, its use here is analogous to ``position weight matrix models" which have been highly
 successful 
@@
 a more parsimonious approach to modelling 
-mutation signatures, with the benefit of producing both more stable estimates and more easily-interpretable signatures.
+mutation signatures, with the benefit of producing both more stable estimates and more easily interpretable signatures.
 In brief, we substantially reduce the number of parameters
@@
 (that some locations will show strong preference for a particular base, whereas others will not), and also
-because it can be represented in an easily-interpretable way via sequencing logos \cite{pmid2172928}.
+because it can be represented in an easily interpretable way via sequencing logos \cite{pmid2172928}.
 For similar reasons -- in addition to the empirical
@@
 Fig. 1 illustrates the way that our new representation of signatures can simply capture
-a previously-identified signature \cite{pmid25362482, pmid23318258} and provides
-an easily-interpretable visualization of the signature 
+a previously identified signature \cite{pmid25362482, pmid23318258} and provides
+an easily interpretable visualization of the signature 
 that is analogous to  sequencing logos \cite{pmid2172928}.
@@
 In the bottom five rectangles, the width of each box represents the frequencies of bases (A, C, G and T) at the substitution and flanking site. 
-To highlight the most informative flanking sites, the heights of flanking site boxes are scaled by ($1 + 0.5 \times \log \sum_{n = A,C,G,T} f_n^2$, where $f_n$ is the parameter for each base), which can be interpreted as ($1 - 0.5 \times \text{R\'enyi entropy}$) \cite{rrnyi1961measures}.
-This is analogous to the information content scaling used in sequencing logos.
+To highlight the most informative flanking sites, the heights of flanking site boxes are scaled by 
+$1 + 0.5 \times \log \sum_{n = A,C,G,T} f_n^2$, where $f_n$ is the parameter for each base, 
+which can be interpreted as $1 - 0.5 \times \text{R\'enyi entropy}$ \cite{rrnyi1961measures}. This is analogous to the information content scaling used in sequencing logos.
 In the top rectangle, the height of each box represents the conditional frequencies of mutated bases for each original base (C and T).
@@
 
-The data consist of 14717 somatic substitutions 
+The data consist of 14,717 somatic substitutions 
 collected from a study of 26 urothelial carcinomas of the upper urinary tract (UCUT) \cite{pmid23926200}.
@@
 Thus each signature is characterized by 18 parameters in our independent model, and by 3071 parameters in the full model.
-After analyzing the data with various number of mutation signatures $K$,
+After analyzing the data with various numbers of mutation signatures $K$,
 we selected $K=3$ signatures for these analyses (see S2 Text).
@@
 shows a clear depletion of G base at the $-2$ position,
-which is consistent to the previous study \cite{pmid23318258} and results in the next subsection (Figs. 3A and 3B).
+which is consistent with the previous study \cite{pmid23318258} and results in the next subsection (Figs. 3A and 3B).
 In contrast, for the full model, this tendency is rather mild (Figs. 3C, 3D and S1).
@@
 Finally, we merged similar mutation signatures across different cancer types 
-(when their Frobenius Distance were $<$ $0.6$). 
+(when their Frobenius distances were $<$ $0.6$, where the Frobenius distance between two matrices ($\bm{F}_1$, $\bm{F}_2$) is $\sqrt{ \Tr  \bigl( (\bm{F}_1 -  \bm{F}_2 )  (\bm{F}_1 -  \bm{F}_2 )^t \bigr) }$,
+where $\Tr$ means the trace of square matrices.
 % XXX WHAT KIND OF CLUSTERING?
@@
 and appear connected by
-a single biological mechansim, we certainly do not argue
+a single biological mechanism, we certainly do not argue
 that inferring them as a single signature is ``wrong". However, splitting them into two signatures does help highlight certain features. Specifically,
-signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Figs. S2C S2D).
+signature 1 shows a transcription strand bias whereas signature 8 does not, and this is true for both colorectal and uterus cancers (Figs. S3C and S3D).
 This strand bias may be connected with the enrichment of 
@@
 \caption{{\bf The estimated frequencies of bases at two 5' to the mutated site for each cancer type.}
-The bar heights show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
+The bar heights show the estimated frequency for bases A, C, G and T at the $-2$ position (two 5' to the mutated site).
 The error bars show bootstrapped standard errors.
-(A, B, C, D, E) The intensities of signature 13 (APOBEC signature), signature 1 (the first POL $\epsilon$ signature),
-signature 8 (the second POL $\epsilon$ signature), signature 10 (ultraviolet signature) and signature 11, respectively, at two 5' to the mutated site.
-intensities at two 5' to the mutated site.
+(A, B, C, D, E) The intensities of signature 13 (APOBEC signature), signature 1 (the first Pol $\epsilon$ signature),
+signature 8 (the second Pol $\epsilon$ signature), signature 10 (ultraviolet signature) and signature 11, respectively, at the $-2$ position.
 }
@@
 Signatures 10, 11, 12, 19 and 21 provide further examples
-of our method refining previously-observed signatures, highlighting strand biases and/or sequence context effects,
+of our method refining previously observed signatures, highlighting strand biases and/or sequence context effects,
 particularly 2 bases upstream of the substitution.
 Signature 10 (C $>$T at [CT]pCpC) was observed in head and neck cancers and melanomas, and probably relates to ultraviolet light.
-Consistent strand specificities among the two cancer types (Fig. S2E) matches previous results \cite{pmid23318258}, 
-but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figs. 6D and S2E). 
+Consistent strand specificities among the two cancer types (Fig. S3E) matches previous results \cite{pmid23318258}, 
+but our analysis additionally highlights elevated abundance of T at the $-2$ position (Figs. 6D and S3E). 
 Signature 11 (C $>$ T at GpCp[CG]) appears in small-cell lung cancers and stomach cancers,
-and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear. Again our analysis highlights elevated abundance of G at the $-2$ position (Figs. 6E and S2F).
+and seems to be the same as ``signature 15'' in the previous study, whose function remains unclear. Again our analysis highlights elevated abundance of G at the $-2$ position (Figs. 6E and S3F).
 Signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively, agree well with those detected in the same cancer types in the previous study \cite{pmid23318258}.
@@
 
-We have also emphasised the connection between mutation
+We have also emphasized the connection between mutation
 signature detection, and the use of mixed-membership models
@@
 for short deletions (a few bases) it may be fruitful to include the actual deleted bases as part of the features.
+In addition, it may be interesting to devise a probabilistic model of mutation signature somewhere between complete independence and non-independence assumption.
+For example, assuming low-dimensional Markovian structure between adjacent bases as some transcription factor binding site modellings did \cite{pmid16108724}
+may help improving representation power of mutation signatures keeping the number of parameters moderate.
 %However, detailed investigation on what mutation features are practical is a future problem.
@@
 
-Although we believe our new methods already provide
-useful gains compared with existing approaches,
-the methods are perhaps even more important for their future potential to incorporate other contextual data, including epigenetic data,
-into mutation signature analysis. This is important, because
-local mutation densities are closely related to a number of genomic and epigenetic factors, such as GC content, repeat sequences, chromatin accessibility and modifications, and replication timing \cite{pmid22820252, pmid21953857, pmid23422670, pmid23770567}.
+Although we believe our new methods already provide useful gains compared with existing approaches,
+the methods are perhaps even more important for their future potential to incorporate other contextual data, including epigenetic data, into mutation signature analysis. 
+This is important, because local mutation densities are closely related to a number of genomic and epigenetic factors, 
+such as GC content, repeat sequences, chromatin accessibility and modifications, and replication timing \cite{pmid22820252, pmid21953857, pmid23422670, pmid23770567}.
 A recent study found that epigenetic information in the cell types of origin of the corresponding tumors is the most predictive \cite{pmid25693567} for local mutation densities. 
-A growing range of epigenetic data from many cell types are now available, and it will be interesting to integrate these epigenetic factors into mutation signature analysis to help understand how these epigenetic factors influence DNA damage and repair mechanisms.
-Our work here provides a straightforward way to do this: epigenetic
-data can be simply added as features
-to the mutation signature. 
-We believe that 
-the value and impact of our work, and 
-specifically our proposed approach to modelling mutation signatures via independent features, will grow as more and more features
+A growing range of epigenetic data from many cell types are now available, 
+and it will be interesting to integrate these epigenetic factors into mutation signature analysis to help understand how these epigenetic factors influence DNA damage and repair mechanisms. 
+Our work here provides a straightforward way to do this: epigenetic data can be simply added as features to the mutation signature.
+In addition, a simulation study shows that the accuracy of mutation signature detection improves as we increase the number of mutation signatures (see Fig. S7).
+We believe that the value and impact of our work, and specifically our proposed approach to modelling mutation signatures via independent features, will grow as more and more features
 are incorporated into the analysis.
@@
 \label{UCUT_supp}
-{\bf Experiment on UCUT data with the various number of mutation signatures.}
+{\bf Experiment on UCUT data with the various numbers of mutation signatures.}
 
@@
 \subsection*{S2 Figure}
+\label{nature2013_POLE}
+{\bf The Pol epsilon signature (Signature 10) derived in the Alexandrov et al., (2013).}
+The barplots are divided by 6 substitution pattern. 
+In each division, 16 bars show joint probabilities of substitution pattern, 5' and 3' bases (ApNpA, ApNpC, ApNpG, ApNpT, $\cdots$, TpNpT).
+
+\subsection*{S3 Figure}
 \label{nature2013_sig_list}
@@
 
+\subsection*{S7 Figure}
+\label{simu_epi}
+The accuracy of the proposed approach for the simulated data when changing the number of features starting 
+from the case where just substitution patterns and immediate 5' and 3' bases are considered ($\bm{M} = (6, 4, 4)$), to 5 additional features ($\bm{M} = (6, 4, 4, 2, 2, 2, 2, 2)$),
+for each number of samples and dispersion parameters.  
+The accuracy of the estimated mutation signature improves as the number of additional features increases, 
+indicating that incorporating additional features such as epigenetic data will be potentially beneficial.
+
 \section*{Acknowledgments}

### 2015-10-05 paper/manuscript/plos_template.tex
proof paper

@@
 The potential for classical studies to yield insights into somatic mutation processes
-were limited in several ways.
+was limited in several ways.
 Due to limited sequencing throughput, most classical studies 
@@
 For example, our approach can incorporate the two bases 3' and 5' of the substitution and transcription strand biases using only 18 parameters
-per signature, compared with 3071 parameters per signature with current approaches.
+per signature, compared with 3,071 parameters per signature with current approaches.
 We demonstrate the benefits of this simplification in data analyses.
@@
 between mutation signature models and the ``mixed-membership models'',
-also known as``admixture models" \cite{pmid10835412} or ``latent Dirichlet allocation''
+also known as``admixture models" \cite{pmid10835412} or ``Latent Dirichlet Allocation''
 models \cite{Blei:2003} that are widely used in population genetics 
@@
  considering $\pm 2$ positions requires 17 parameters
- instead of 1535. Finally, incorporating
+ instead of 1,535. Finally, incorporating
 transcription strand as an additional feature adds just one parameter, instead of doubling the number of parameters.
@@
 This generative model is summarized in Fig. \ref{methodOverview}.
-This model is essentially a ``mixed-membership model", also known as an ``admixture model" \cite{pmid10835412} or ``latent Dirichlet allocation" \cite{Blei:2003}. For example,
+This model is essentially a ``mixed-membership model", also known as an ``admixture model" \cite{pmid10835412} or ``Latent Dirichlet Allocation" \cite{Blei:2003}. For example,
 the membership proportions for each sample are analogous to admixture proportions in an admixture model; the mutation signatures are analogous to 
@@
 We consider a mutation pattern to consist of the substitution pattern, the $\pm 2$ flanking bases, and the transcription strand direction.
-Thus each signature is characterized by 18 parameters in our independent model, and by 3071 parameters in the full model.
+Thus each signature is characterized by 18 parameters in our independent model, and by 3,071 parameters in the full model.
 After analyzing the data with various numbers of mutation signatures $K$,
@@
 \caption{{\bf The mutation signatures for the UCUT data, and the results of down-sampling experiments.}
-3072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.
+3,072 elements in the full model mutation signatures were shown divided by 6 substitution patterns and strand directions.
 (A, B) APOBEC and AA signature for the independent model.
@@
 To investigate this further we performed down-sampling experiments.
-Using the mutation signatures obtained using all 14717 substitutions as a gold standard, 
+Using the mutation signatures obtained using all 14,717 substitutions as a gold standard, 
 we assessed performance of the proposed method on down-sampled data consisting of $r\%$ of the original data, where  $r=(1\%, 2.5\%, 5\%, 10\%, 25\%, 50\%)$.
@@
 Finally, we merged similar mutation signatures across different cancer types 
-(when their Frobenius distances were $<$ $0.6$, where the Frobenius distance between two matrices ($\bm{F}_1$, $\bm{F}_2$) is $\sqrt{ \Tr  \bigl( (\bm{F}_1 -  \bm{F}_2 )  (\bm{F}_1 -  \bm{F}_2 )^t \bigr) }$,
-where $\Tr$ means the trace of square matrices.
+(when their Frobenius distances were $<$ $0.6$, where the Frobenius distance between two matrices ($\bm{F}_1$, $\bm{F}_2$) is $\sqrt{ \Tr  \bigl( (\bm{F}_1 -  \bm{F}_2 )  (\bm{F}_1 -  \bm{F}_2 )^t \bigr) }$
+($\Tr$ means the trace of square matrices).
 % XXX WHAT KIND OF CLUSTERING?
@@
 
-Figs. \ref{nature2013_sig_summary} and \ref{nature2013_sig_member} summarise the results.
+Figs. \ref{nature2013_sig_summary} and \ref{nature2013_sig_member} summarize the results.
 In total, we identified 27 mutation signatures.
@@
 Signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and is surely related to the activity of the APOBEC family. 
-The 12 inferred signatures are highly consistent among cancer types except for B-cell lymphoma (see Fig.
-S3A), highlighting the robustness
+The 12 inferred signatures were highly consistent among cancer types except for B-cell lymphoma (see Fig. S3A), highlighting the robustness
 of our approach.  Almost all of them show enrichment of A and T and depletion of G base at the $-2$ position (Fig. 6A and S3A), 
@@
 for short deletions (a few bases) it may be fruitful to include the actual deleted bases as part of the features.
-In addition, it may be interesting to devise a probabilistic model of mutation signature somewhere between complete independence and non-independence assumption.
-For example, assuming low-dimensional Markovian structure between adjacent bases as some transcription factor binding site modellings did \cite{pmid16108724}
-may help improving representation power of mutation signatures keeping the number of parameters moderate.
-%However, detailed investigation on what mutation features are practical is a future problem.
+Also, it may be interesting to devise a probabilistic model for mutation signatures somewhere between complete independence and non-independence assumption, 
+for example, using ideas analogous to those in \cite{pmid16108724} that uses a Markovian structure  for transcription  factor binding sites. 
+This may help improve the modelling flexibility of mutation signatures while keeping the number of parameters moderate.
 
@@
 Our work here provides a straightforward way to do this: epigenetic data can be simply added as features to the mutation signature. 
-In addition, a simulation study shows that the accuracy of mutation signature detection improves as we increase the number of mutation signatures (see Fig. S7).
+This has the potential to improve accuracy of signature detection (e.g. Fig S7), and to produce novel biological insights.
 We believe that the value and impact of our work, and specifically our proposed approach to modelling mutation signatures via independent features, will grow as more and more features
@@
 must be estimated from the available mutation data $\{ \bm{x}_{i,j} \}$. 
-Here we adopt a simple approach that uses an EM-algorithm to maximise the likelihood.
+Here we adopt a simple approach that uses an EM-algorithm to maximize the likelihood.
 
